### 2021-02-03 marine/bivalve/bivalve_disturbance/bivalve_disturbance.Rmd

---
title: "Bivalve disturbance"
author: "Gage Clawson (NCEAS, OHI, UCSB)"
date: "12/29/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in libaries and paths}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(raster)
library(data.table)
library(raster)
library(rgdal)
library(here)
library(taxize)

path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/bivalve"

food_crs <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

```

Bivalve specifications:

 - 100 long lines in each km2 
 - each line 4000 meters of rope (13000 feet)
 - each foot of line has 100 bivalves

Which means 130000000 bivalves per km2

We assuming that the rearing time is about 1 year, so no lost bivalves. What is reported is the actual tonnage.

## Estimate number bivalves/tonne production at family level

Using the information in this paper, table 1: https://www.researchgate.net/publication/259436886_Weight_-_Length_relationships_of_25_bivalve_species_Mollusca_Bivalvia_from_the_Algarve_coastsouthern_Portugal

First I need to put the information in table 1 into an excel sheet (do this manually and upload to aurora). I will add a "family_2" column, for the families which do not match our data. It will be the family in our data that is closest in relation (i.e. in the same super family, or the same order)

I will do the same thing with this paper, using table 1 and table 2: http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.825.2317&rep=rep1&type=pdf

We will also use information for oyster species from here (Ostreidae): **be sure to include this as a reference in the supplementary info**

Assume this number for all Ostreidae, regardless of if they are pacific cupped or not (pacific cupped is the most common in our data). 
http://www.ccpo.odu.edu/~klinck/Reprints/PDF/kobayashiAqua1997.pdf
http://www.ccpo.odu.edu/~klinck/Reprints/PDF/kobayashiAqua1997.pdf
Pacific Oyster

Assume live-weight estimates: From Figure 4a

L = 21.4*W^0.377

W = 0.000295836*L^2.563

1. For each species, use their equation and parameter estimates to calculate # bivalves/tonne
2. Average these values at the family level
3. Calculate an average that you can use to gapfill families not included in these studies

```{r}
#### GASPAR (first paper)
## read in the Weight and Length data and wrangle 
gaspar <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/gaspar_bivalve/gaspar_bivalve_WL.csv"))

unique(gaspar$family)

## now lets wrangle a bit, to get family, avg alpga, avg beta

# filter for those that have a "family_2" and get those averages
gaspar_fam_2 <- gaspar %>%
  dplyr::filter(!is.na(family_2)) %>%
  group_by(family_2) %>%
  summarise(alpha = mean(alpha),
            beta = mean(beta)) %>%
  dplyr::rename("family" = "family_2")

# get averages for those that have na for family_2
gaspar_fam <- gaspar %>%
  dplyr::filter(is.na(family_2)) %>%
  group_by(family) %>%
  summarise(alpha = mean(alpha),
            beta = mean(beta))

## bind together and get averages for families
gaspar_final <- rbind(gaspar_fam, gaspar_fam_2) %>%
  group_by(family) %>%
  summarise(alpha = mean(alpha), 
            beta = mean(beta))

#### Park (second paper)

park <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/park_and_oh_bivalve_WL/park_and_oh_bivalve_WL.csv"))

park_final <- park %>%
  group_by(family) %>%
  summarise(alpha = mean(alpha), 
            beta = mean(beta))

## third paper for oysters
koboyashi_final <- data.frame(family = "Ostreidae", alpha = 0.000295836, beta = 2.563)

## combine them all

bivalve_w_l_final <- rbind(park_final, gaspar_final, koboyashi_final) %>%
  group_by(family) %>%
  dplyr::summarise(alpha = mean(alpha), 
            beta = mean(beta))

write.csv(bivalve_w_l_final, file.path("aquaculture/marine/bivalve/bivalve_disturbance/data/weight_length_relationships.csv"), row.names = FALSE)

```

## Convert FAO tonnes of production to number of bivalves
Use FAO mariculture production data for bivalves

1. assign each bivalve to a family
2. merge with paper data by family to add in the number bivalves per tonne production, gapfill missing family data with average (assume an average bivalve length of 4cm (40 mm))
3. Calculate number of bivalves for each taxa/country by dividing tonnes production by number bivalves per tonne 
4. Calculate the number of bivalves for each country (group by country and summarize)
5. Distribute the number of bivalves across the raster based on country and proportion of production for each raster cell
6. Apply model to each raster cell to estimate area:  number bivalves / 130,000,000 bivalves per km2

```{r}
# read in spp specific country bivalve data: 
bivalve_spp <- read_csv(file.path("aquaculture/marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve_taxa_production.csv"))

## read in family information
spp_families <- read_csv(file.path("aquaculture/marine/STEP1_species_groups/int/marine_groups.csv")) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::select(species, family) %>%
  distinct()

## left join
bivalve_spp_family <- bivalve_spp %>%
  left_join(spp_families) %>%
  mutate(family = ifelse(species == "Australian mussel", "Mytilidae", family)) %>% ## fix a mistake
  mutate(family = ifelse(species == "Sea snails", "Haliotidae", family)) ## fix mistake (sea snails from fao likely mean abalone)

summary(bivalve_spp_family) ## check to see there are no NAs in family - there aren't

```

Now compare our families in the weight-length data with our fao data

```{r}
sort(unique(bivalve_spp_family$family))

sort(unique(bivalve_w_l_final$family))

sort(setdiff(bivalve_spp_family$family, bivalve_w_l_final$family))

sort(setdiff(bivalve_w_l_final$family, bivalve_spp_family$family))
```

Now do steps 1-6 listed above 
```{r}
# bivalve_w_l_fix <- bivalve_w_l_final %>%
#   dplyr::filter(!(family %in% c("Donacidae", "Phasianellidae", "Solenidae")))

avg_alpha <- mean(bivalve_w_l_final$alpha)
avg_beta <- mean(bivalve_w_l_final$beta)

# avg_alpha_fix <- mean(bivalve_w_l_fix$alpha)
# avg_beta_fix <- mean(bivalve_w_l_fix$beta)

## join them together 
fao_W_L_explore <- bivalve_spp_family %>%
  left_join(bivalve_w_l_final, by = "family") %>%
  dplyr::mutate(weight = alpha*40**beta) %>% ## assume 40 mm 
  group_by(family) %>%
  summarise(weight = mean(weight),
            number_rows = n(),
            total_tonnes = sum(fao_tonnes_production))

## join them and gapfill the missing rows with the average
fao_W_L <- bivalve_spp_family %>%
  left_join(bivalve_w_l_final, by = "family") %>%
  mutate(alpha = ifelse(is.na(alpha), avg_alpha, alpha),
         beta = ifelse(is.na(beta), avg_beta, beta))

## group by family and country to get tonnes
fao_W_L_summarise <- fao_W_L %>%
  dplyr::group_by(country, iso3c, year, family) %>%
  dplyr::summarise(fao_tonnes_production = sum(fao_tonnes_production),
                   alpha = mean(alpha),
                   beta = mean(beta))

## figure out number of bivalves per country

bivalve_per_country <- fao_W_L_summarise %>%
  dplyr::mutate(bivalve_avg_weight_g = alpha*40**beta) %>%
  dplyr::mutate(bivalve_avg_weight_t = bivalve_avg_weight_g/1000000) %>%
  dplyr::mutate(n_bivalves = fao_tonnes_production/bivalve_avg_weight_t) %>%
  group_by(country, iso3c, year) %>%
  dplyr::summarise(fao_tonnes_production = sum(fao_tonnes_production),
                   total_bivalves = sum(n_bivalves),
                   avg_weight = mean(bivalve_avg_weight_t))

## fix HKG and CHN 
CHN_HKG_biv <- bivalve_per_country %>%
  dplyr::filter((iso3c %in% c("CHN", "HKG")))

sum(CHN_HKG_biv$fao_tonnes_production) # 14646903
chn_bivalves <- sum(CHN_HKG_biv$total_bivalves)

CHN_fix <- data.frame(country = "China", iso3c = "CHN/MAC/HKG", year = 2017, fao_tonnes_production = 14646903, total_bivalves = chn_bivalves)

bivalve_per_country_final <- bivalve_per_country %>%
  dplyr::filter(!(iso3c %in% c("CHN", "HKG"))) %>%
  rbind(CHN_fix) %>%
  dplyr::select(country, iso3c, year, total_bivalves)

# 1. stack with the region raster and covert to data.frame.  merge with the dataframe that will convert the ID_0 values to iso3c. 
# 2.  group-by(iso3c), mutate(country_tonnes = sum(tonnes, na.rm=tRUE)), rowwise(), mutate(prop_cell = tonnes/country_tonnes)
# 3.  Then you merge the country count of bivalves to the raster data and multiply by prop_cell.
# 4. This will give you the estimated number of bivalves in each cell
# 5. Apply space model to raster

## read in production csv
production <- read_csv(file.path(path_prep, "bivalve_farm.csv"))
sum(production$tonnes_production)

iso3c_codes <- read_csv(file.path("_spatial/_output/eez_food_rgns.csv"))

land_eez_region_raster <- raster("/home/shares/food-systems/Food_footprint/dataprep/spatial/land_eez_rgns.tif")

land_eez_region_raster_df <- as.data.frame(land_eez_region_raster, xy = TRUE)

region_raster_iso3c_df <- land_eez_region_raster_df %>%
  left_join(iso3c_codes, by = c("land_eez_rgns" = "ID_0"))

production_cell_iso3c <- production %>%
  left_join(region_raster_iso3c_df, by = c("x", "y")) %>%
  group_by(iso3c) %>%
  mutate(country_tonnes = sum(tonnes_production, na.rm = TRUE)) %>%
  rowwise() %>%
  mutate(prop_cell = tonnes_production/country_tonnes) %>%
  dplyr::mutate(prop_cell = ifelse(prop_cell == "NaN", NA, prop_cell)) %>%
  ungroup()

biv_per_cell <- production_cell_iso3c %>%
  left_join(bivalve_per_country_final, by = "iso3c") %>%
  mutate(bivalves_per_cell = total_bivalves*prop_cell)

sum(biv_per_cell$bivalves_per_cell, na.rm = TRUE) # 3.801619e+12

sum(bivalve_per_country_final$total_bivalves) # 3.801619e+12

## apply our disturbance model
biv_disturbance <- biv_per_cell %>%
  dplyr::mutate(disturbance_km2 = bivalves_per_cell/130000000) %>%
  dplyr::select(x, y, disturbance_km2)

sum(biv_disturbance$disturbance_km2, na.rm = TRUE) 
# 29243.22 km2 disturbance
# 17435623 tonnes production bivalve
# tonnes/disturbance = 596.2279

## compare to salmon: 
# 70.82 km2 disturbance
# 2746086 tonnes produciton salmon
# tonnes/disturbance = 38775.57

# 412 times bigger disturbance for bivalves
# ~ 6 times bigger production for bivalves

## compare to tuna:
#  2.817567 km2 disturbance
# 37072.56 tonnes production tuna
# tonnes/disturbance = 13157.65

# 10378.89 times bigger disturbance for bivalves
# 470 times bigger production for bivalves

## Now multiply the disturbance by 0.1, since bivalve aquaculture likely actually creates more habitat than it disturbs.

biv_disturbance <- biv_disturbance %>%
  dplyr::mutate(disturbance_km2 = disturbance_km2*0.1)

sum(biv_disturbance$disturbance_km2, na.rm = TRUE) # 2924.322
# 2924.322 km2 disturbance
# 17435623 tonnes production bivalve
# tonnes/disturbance = 5962.279

## compare to salmon: 
# 70.82 km2 disturbance
# 2746086 tonnes produciton salmon
# tonnes/disturbance = 38775.57

# 41.29232 times bigger disturbance for bivalves
# ~ 6 times bigger production for bivalves

## compare to tuna:
#  2.817567 km2 disturbance
# 37072.56 tonnes production tuna
# tonnes/disturbance = 13157.65

# 1037.889 times bigger disturbance for bivalves
# 470 times bigger production for bivalves
write.csv(biv_disturbance, file.path(path_prep, "bivalve_disturbance.csv"), row.names = FALSE)

# final_layer_dist <- read_csv(file.path(path_prep, "bivalve_disturbance.csv"))

final_rast <- rasterFromXYZ(biv_disturbance, crs = food_crs)
crs(final_rast)
plot(final_rast)
cellStats(final_rast, "sum") # 2924.322
final_rast

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 2924.322
final_rast
plot(log(final_rast+1))

writeRaster(final_rast, file.path(path_final, "marine_bivalve_aquaculture_meat_disturbance.tif"), overwrite = TRUE)

```