### 2021-02-03 marine/crustaceans/crustacean_nutrient/step1_nutrient.Rmd

---
title: "crustacean_nutrient_pollution"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/22/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    includes: 
toc: true
---

This markdown uses models and species specific parameters from the paper *Aquaculture Production is a Large, Spatially Concentrated Source of Nutrients in Chinese Freshwater and Coastal Seas* (2020 Wang et al).

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/crustaceans"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

production <- read_csv(file.path(path_prep, "crustaceans_farm.csv"))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

crustacean N/P pollution values

N: 57 kg/tonne of growth
P: 9.2 kg/tonne of growth

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P value
3. We report in total excess nutrients (N + P): Sum the N and P kg.
4. We report in tonnes, so divide by 1000. We will also so individual N and P rasters for reporting purposes.

```{r}

crustacean_pollution <- production %>%
  dplyr::mutate(prod_growth = tonnes_production*0.95) %>%
  dplyr::mutate(prod_growth_N_poll = prod_growth*57,
                prod_growth_P_poll = prod_growth*9.2) %>% ## in kg pollution now
  # dplyr::mutate(PO4eq_kg = prod_growth_N_poll*0.42 + prod_growth_P_poll*3.07) %>% Don't do this anymore (01/14/2021)
  dplyr::mutate(total_N_P_kg = prod_growth_N_poll + prod_growth_P_poll) %>%
  dplyr::mutate(total_N_P_t = total_N_P_kg/1000,
                N_poll_t = prod_growth_N_poll/1000, 
                P_poll_t = prod_growth_P_poll/1000) %>%
  dplyr::select(x, y, total_N_P_t, N_poll_t, P_poll_t)

sum(crustacean_pollution$total_N_P_t) # 25264.11
sum(production$tonnes_production) # 401719.1

## save to dataprep folder with the correct naming conventions: 
write.csv(crustacean_pollution, file.path(path_prep, "crustacean_nutrient.csv"))

## Save total N P raster
total_df <- crustacean_pollution %>% 
  dplyr::select(x, y, total_N_P_t)

final_rast <- rasterFromXYZ(total_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 25264.11

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 25264.11
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_final, "marine_crustaceans_aquaculture_meat_nutrient.tif"), overwrite = T)

## Save total N raster
total_df <- crustacean_pollution %>% 
  dplyr::select(x, y, N_poll_t)

final_rast <- rasterFromXYZ(total_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 21753.09

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 21753.09
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_crustaceans_aquaculture_meat_nutrientN.tif"), overwrite = T)

## Save total P raster
total_df <- crustacean_pollution %>% 
  dplyr::select(x, y, P_poll_t)

final_rast <- rasterFromXYZ(total_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 3511.025

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 3511.025
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_crustaceans_aquaculture_meat_nutrientP.tif"), overwrite = T)
```