### 2021-02-03 marine/marine_fish_general/finfish_nutrient/step1_nutrient.Rmd

---
title: "marine_fish_general_nutrient_pollution"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: 12/22/2020
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    includes: 
toc: true
---

This markdown uses models and species specific parameters from the paper *Hindcasts and Future Projections of Global Inland and Coastal Nitrogen and Phosphorus Loads Due to Finfish Aquaculture* (2013 Bouwman et al)

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/marine_fish_general"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

production <- read_csv(file.path(path_prep, "marine_fish_general_farm.csv"))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

Because we have multiple values and species for nutrient pollution for general marine fish, we will have to do this analysis in a series of steps. Workflow will follow these rules:

1. Apply the general pelagic species pollution numbers to areas with those species
2. Any remaining species apply the regional values for Asia and Europe
3. Any remaining species or areas that don't fall in the above steps, take the average of the Asia and Europe nutrient pollution values.

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P value
3. We report in total excess nutrient: add N and P together.
4. We report in tonnes, so divide by 1000. Save N tonnes, P tonnes, and N + P tonnes layers.

**Numbers for each category:** 
Other finfish - pelagics:  Amberjacks, nei, Cobia, Japanese amberjacks
N = 152, P = 25	P assumes

Other finfish - coastal, Europe: 
N = 80, P =	16

Other finfish - coastal, Asia:
N = 112, P = 18

```{r}
fao <- read.csv(here("aquaculture/data/fao_mariculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

## Read in fao data and get a species list for salmon
species_list <- read.csv(here("aquaculture/marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

# the list of species we are including
finfish_list <- species_list$species[species_list$aq_group=="marine_fish_general"]
finfish_list

## write out shrimp taxa and production for Juliette
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

## set up lists of different regions
asia <- c("BGD", "CHN", "HKG", "IDN", "IND", "JPN", "KHM", "KOR", "MYS", "PHL", "SGP", "THA", "TWN", "VNM")

europe <- c("ALB", "HRV", "CYP", "FRA", "GRC", "ITA", "MLT", "NOR", "PRT", "ESP", "TUR")

other <- c("AUS", "CRI", "EGY", "IRN", "ISR", "MUS", "MEX", "PAN", "SAU", "TUN", "ARE", "USA")

## Now we need to get the proportion of catch that falls into each species category (i.e. pelagic or non-pelagic) 
prop_df <- fao_finfish_taxa_prod %>%
  dplyr::mutate(species_cat = ifelse(str_detect(species, "obia|mberjack|Marine fishes nei"), "pelagic", "non_pelagic")) %>%
  group_by(iso3c, species_cat) %>%
  summarise(tonnes = sum(fao_tonnes_production)) %>%
  ungroup() %>%
  group_by(iso3c) %>%
  mutate(total_tonnes = sum(tonnes)) %>%
  ungroup()

prop_pel_df <- prop_df %>%
  dplyr::filter(species_cat == "pelagic") %>%
  dplyr::mutate(prop_pelagic = tonnes/total_tonnes) %>%
  dplyr::mutate(prop_non_pelagic = 1-prop_pelagic) %>%
  dplyr::select(iso3c, prop_pelagic, prop_non_pelagic, total_tonnes) %>%
  dplyr::filter(prop_pelagic == 1)

prop_non_pel_df <- prop_df %>%
  dplyr::filter(species_cat == "non_pelagic") %>%
  dplyr::mutate(prop_non_pelagic = tonnes/total_tonnes) %>%
  dplyr::mutate(prop_pelagic = 1-prop_non_pelagic) %>%
  dplyr::select(iso3c, prop_pelagic, prop_non_pelagic, total_tonnes)

prop_total_iso3c <- rbind(prop_pel_df, prop_non_pel_df) ## rbind them

sum(prop_total_iso3c$total_tonnes) # 3901214
sum(fao_finfish_taxa_prod$fao_tonnes_production) # 3901214

## Read in production data 
production_sf <- st_read(file.path("/home/sgclawson/github/food_systems/aquaculture/marine/marine_fish_general/finfish_nutrient/prod_data_by_cell_iso3c.shp")) %>%
  dplyr::rename("tonnes_per_farm" = "tnns_p_")

## join with our proportion data so that we can get tonnes of pelagic and tonnes of non-pelagic per farm 
production_spp_props <- production_sf %>%
  left_join(prop_total_iso3c, by = "iso3c") %>%
  dplyr::mutate(tonnes_pelagic = prop_pelagic*tonnes_per_farm, 
                tonnes_non_pelagic = prop_non_pelagic*tonnes_per_farm)

## Now we can do the actual analysis outlined above

nutrient_data <- production_spp_props %>%
  dplyr::mutate(tonnes_pelagic_growth = tonnes_pelagic*0.95,
                tonnes_non_pelagic_growth = tonnes_non_pelagic*0.95) %>%   ## multiply by 0.95 to correct for growth outside seapen

  dplyr::mutate(N_poll_pel = tonnes_pelagic_growth*152,
                P_poll_pel = tonnes_pelagic_growth*25,
                N_poll_non_pel = case_when(
                  iso3c %in% asia ~ tonnes_non_pelagic_growth*112,
                  iso3c %in% europe ~ tonnes_non_pelagic_growth*80,
                  iso3c %in% other ~ tonnes_non_pelagic_growth*96
                ),
                P_poll_non_pel = case_when(
                  iso3c %in% asia ~ tonnes_non_pelagic_growth*18,
                  iso3c %in% europe ~ tonnes_non_pelagic_growth*16,
                  iso3c %in% other ~ tonnes_non_pelagic_growth*17
                )) %>%  ## get all of out pollution values by species group and region

  dplyr::mutate(N_poll = N_poll_pel + N_poll_non_pel,
                P_poll = P_poll_pel + P_poll_non_pel) %>% ## get total pollution values for each cell in kg

  # dplyr::mutate(po4eq_kg = N_poll*0.42 + P_poll*3.07) %>% ## get po4eq: cut this (01/14/2021)

  dplyr::mutate(total_N_P_t = (N_poll + P_poll)/1000,
                N_poll_t = N_poll/1000, 
                P_poll_t = P_poll/1000) %>%
  dplyr::select(total_N_P_t, N_poll_t, P_poll_t, geometry)

sum(nutrient_data$total_N_P_t) # 507965.9

## Save final N+P layer
finfish_raster_check <- rasterize(nutrient_data, food_raster, field = "total_N_P_t", fun=sum)
plot(log(finfish_raster_check+1))
cellStats(finfish_raster_check, stat = "sum", na.rm = TRUE) # 507965.9

finfish_raster <- rasterize(nutrient_data, food_raster, field = "total_N_P_t", fun=sum, background=0)
plot(log(finfish_raster+1))

finfish_raster_df <- as.data.frame(finfish_raster, xy=TRUE) %>%
  dplyr::select(x,y, N_P_tonnes=layer)

cellStats(finfish_raster, stat = "sum") # 507965.9
sum(finfish_raster_df$N_P_tonnes) # 507965.9
sum(nutrient_data$total_N_P_t, na.rm = TRUE) # 507965.9

## save to dataprep folder with the correct naming conventions: 
write.csv(finfish_raster_df, file.path(path_prep, "marine_fish_general_nutrient.csv"))

final_rast <- rasterFromXYZ(finfish_raster_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 507965.9

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 507965.9
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_final, "marine_marine-fish-general_aquaculture_meat_nutrient.tif"), overwrite = T)

## Save final N layer
finfish_raster_check <- rasterize(nutrient_data, food_raster, field = "N_poll_t", fun=sum)
plot(log(finfish_raster_check+1))
cellStats(finfish_raster_check, stat = "sum", na.rm = TRUE) # 435852.5

finfish_raster <- rasterize(nutrient_data, food_raster, field = "N_poll_t", fun=sum, background=0)
plot(log(finfish_raster+1))

finfish_raster_df <- as.data.frame(finfish_raster, xy=TRUE) %>%
  dplyr::select(x,y, N_tonnes=layer)

cellStats(finfish_raster, stat = "sum") # 435852.5
sum(finfish_raster_df$N_tonnes) # 435852.5
sum(nutrient_data$N_poll_t, na.rm = TRUE) # 435852.5

## save to dataprep folder with the correct naming conventions: 
write.csv(finfish_raster_df, file.path(path_prep, "marine_fish_general_nutrient_N.csv"))

final_rast <- rasterFromXYZ(finfish_raster_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 435852.5

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 435852.5
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_marine-fish-general_aquaculture_meat_nutrientN.tif"), overwrite = T)

## Save final P layer
finfish_raster_check <- rasterize(nutrient_data, food_raster, field = "P_poll_t", fun=sum)
plot(log(finfish_raster_check+1))
cellStats(finfish_raster_check, stat = "sum", na.rm = TRUE) # 72113.4

finfish_raster <- rasterize(nutrient_data, food_raster, field = "P_poll_t", fun=sum, background=0)
plot(log(finfish_raster+1))

finfish_raster_df <- as.data.frame(finfish_raster, xy=TRUE) %>%
  dplyr::select(x,y, P_tonnes=layer)

cellStats(finfish_raster, stat = "sum") # 72113.4
sum(finfish_raster_df$P_tonnes) # 72113.4
sum(nutrient_data$P_poll_t, na.rm = TRUE) # 72113.4

## save to dataprep folder with the correct naming conventions: 
write.csv(finfish_raster_df, file.path(path_prep, "marine_fish_general_nutrient_P.csv"))

final_rast <- rasterFromXYZ(finfish_raster_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 72113.4

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 72113.4
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_marine-fish-general_aquaculture_meat_nutrientP.tif"), overwrite = T)
```