### 2021-02-03 marine/salmon/salmon_disturbance/STEP2_occupancy_distribution.Rmd

---
title: "occupancy_distribution"
author: "Juliette Verstaen"
date: "1/24/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in libaries and paths}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(raster)
library(data.table)
library(raster)
library(rgdal)
library(here)

path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/salmon/"

food_crs <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

```

In this markdown we are calculating:

1. How many live fish are in the water at a given time.
2. The total volume (and area) used by salmon farming

##Methods

##Conversion Table
Read in and save the necessary values from the salmon conversion table
```{r parameters from conversion table}
salmon_conversion <- read_csv(here("aquaculture/marine/salmon/salmon_disturbance/data/salmon_conversion_factors.csv"))

stocking_density <- as.numeric(salmon_conversion[salmon_conversion$name == "stocking_density", "value"])
live_harvested_conversion <- as.numeric(salmon_conversion[salmon_conversion$name == "live_harvested", "value"])
harvest_weight <- as.numeric(salmon_conversion[salmon_conversion$name == "mean_harvest_weight", "value"])
cage_volume <- as.numeric(salmon_conversion[salmon_conversion$name == "cage_volume", "value"])
```

###Calculate the total number of fish in the water (occupancy)
Read in production map

```{r}
production_raster <- rasterFromXYZ(read_csv(paste0(path_prep, "salmon_farm.csv")), crs = food_crs)

#plot(production_raster)
cellStats(production_raster, stat = "sum") # 2746086

```

```{r calculate salmon occupancy}
#as total number of salmon
salmon_occupancy <- (production_raster*1000)/harvest_weight * live_harvested_conversion
writeRaster(salmon_occupancy, file.path(path_prep, "salmon_occupancy/salmon_occupancy.tif"), format = "GTiff", overwrite = TRUE)

# as live weight of salmon, in kg
salmon_live_kg  <- salmon_occupancy*harvest_weight
salmon_live_tonnes  <- salmon_occupancy*harvest_weight/1000

writeRaster(salmon_live_tonnes, file.path(path_prep, "salmon_occupancy/salmon_live_tonnes.tif"), format = "GTiff", overwrite = TRUE)
```

### Calculate the total area (from an aerial view) used for salmon aquaculture

First we determine how many cages are required given cage dimensions, stocking densities, and kg living fish.

We used Gentry 2017 data to estimate 180,000 kg salmon (harvest weight) per cage.

This is based on 9000 m3 cage that we assume to be circular with a depth of 10 m.  We assume a stocking density of 20kg/m3.

```{r}

kg_per_cage <- cage_volume*stocking_density

```

Given the kg of salmon per cell we determine the number of cages needed as:

kg_salmon/180,000

We round this value up to the nearest whole number and multiply by 1.3 to include fallowing cages.

```{r}

cages <- ceiling(salmon_live_kg/kg_per_cage)

cages_pls_fallow <- ceiling(cages * 1.3)

```

2. We the determine how much surface area the cages disturb including the inner cage area, a 5 meter buffer around the cage, and area for infrastructure.

```{r}
cage_depth_m <- 10
buffer <- 5
infrastructure <- 1.5 #50% of area to infrastructure

cage_area <- cage_volume/cage_depth_m
cage_radius <- sqrt(cage_area/pi)
cage_radius_buffer <- cage_radius + buffer
cage_area_with_buffer <- pi*cage_radius_buffer^2
cage_disturbance_m2 <- cage_area_with_buffer*infrastructure

total_disturbance_km2 <- cages_pls_fallow * cage_disturbance_m2/1000000
total_disturbance_prop <- total_disturbance_km2/raster::area(total_disturbance_km2)
plot(total_disturbance_prop)

salmon_aquaculture_area_df <- as.data.frame(total_disturbance_prop, xy=TRUE) %>% 
  dplyr::rename(salmon_disturbance_prop = layer )

write_csv(salmon_aquaculture_area_df, file.path(path_prep, "salmon_disturbance_prop.csv"))

salmon_aquaculture_area_df_km2 <- as.data.frame(total_disturbance_km2, xy=TRUE) %>% 
  dplyr::rename(salmon_disturbance_km2 = layer)

write_csv(salmon_aquaculture_area_df_km2, file.path(path_prep, "salmon_disturbance_km2.csv"))

## save to datalayers folder with the correct naming conventions: 
final_layer_dist <- read_csv(file.path(path_prep, "salmon_disturbance_km2.csv"))

final_rast <- rasterFromXYZ(final_layer_dist, crs = food_crs)
crs(final_rast)
plot(final_rast)
cellStats(final_rast, "sum") # 70.82137
final_rast

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 70.82137
final_rast

writeRaster(final_rast, file.path(path_final, "marine_salmon_aquaculture_meat_disturbance.tif"), overwrite = TRUE)
```

Look at the pieces more closely
```{r}
norway <- raster::crop(total_disturbance_km2, extent(-25, 50, 50, 75))
plot(norway)
writeRaster(norway, file.path(path_prep, "salmon_occupancy/norway.tif"), overwrite=TRUE)

chile <- raster::crop(total_disturbance_km2, extent(-80, -60, -60, -35))
plot(chile)
writeRaster(chile, file.path(path_prep, "salmon_occupancy/chile.tiff"), overwrite=TRUE)

```