### 2021-02-03 marine/bivalve/food_pressures/bivalve_ghg/STEP1_ghg_emissions.Rmd
organized the files more

---
title: "Bivalve Aquaculture GHG"
author: "Gage Clawson (NCEAS, OHI, UCSB)"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description

This markdown calculates the GHG emissions from bivalve aquaculture.

# Data information
We used [Seafood Carbon Emissions Tool data](http://seafoodco2.dal.ca/) to extract ghg emissions associated with carcass weight of bivalve from bivalve farming.

One factor that we added was aquatic N<sub>2</sub>O emissions as discussed in [Hu *et al*., 2012](https://pubs.acs.org/doi/full/10.1021/es300110x) and [MacLeod *et al*., 2019](http://www.fao.org/3/ca7130en/ca7130en.pdf).

# Methods 
 - Take midpoint of ghg low and ghg high from [seafood carbon emissions tool data](http://seafoodco2.dal.ca/) for all species
 - Multiply by % energy (to cut out the feed emissions)
 - Add the N20 value from the MacLeod paper to the emissions intensity estimate we got for bivalve
 - Multiply emissions intensity estimate by tonnes of production in each cell

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(tidyverse)
library(here)
library(raster)
library(sf)
library(janitor)

# Raster templates
source(here("_spatial/template_raster.R"))

# File paths
raw   <- "/home/shares/food-systems/Food_footprint/_raw_data/"
prep_bivalve <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/bivalve"
prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/" 
path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

# Import master_rgns xy df
master_rgns <- read_csv(here("_spatial/_output/food_rgns.csv"))
```

## Import
```{r}
## Read in farmed emissions data

farmed_emissions <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/seafood_carbon_emmisions_tool/d2020/Specie_List_-_11_24_2020,_11_39_21_AM_Farmed.csv")) %>%
  clean_names()

bivalve_production <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/bivalve/bivalve_farm.csv"))

fao_bivalve_spp <- read_csv(here("aquaculture/marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve_taxa_production.csv"))

```

# Wrangle

```{r}
unique(fao_bivalve_spp$species)

unique(farmed_emissions$common_name)
## we only want bivalves that are in fao
# Blue mussel
# Mediterranean mussel
# American cupped oyster
# Pacific cupped oyster 
# New Zealand mussel

bivalve_emissions_spp <- farmed_emissions %>%
  dplyr::filter(common_name %in% c("Blue mussel", "Mediterranean mussel", "American cupped oyster", "Pacific cupped oyster", "New Zealand mussel"))

## grab the mean of ghg low and high (the 25th and 75th quantiles)
bivalve_emissions <- bivalve_emissions_spp %>%
  rowwise() %>%
  dplyr::mutate(ghg_kgCO2_kg_fish_mean = mean(c(ghg_low, ghg_high)))

ghg_kgCO2_kg_fish <- mean(bivalve_emissions$ghg_kgCO2_kg_fish_mean)

## now multiply by the percent of that which is energy emissions (100%)
ghg_energy_kgCO2_kg_fish <- ghg_kgCO2_kg_fish*1 ##  0.2471429 kg CO2eq/ kg fish (live/carcass weight, ratioed)

# Adding Aquatic N2O
# From MacLeod et al. 
# "The amount of N2O per species-group was determined by multiplying the
# production by the N2O emission factor per kg of production (Hu et al., 2012), i.e.
# 1.69 gN2O-N per kg of production, or 0.791 kgCO2e/kgLW production. This equates
# to a conversion rate of N to N2O-N of 1.8 percent, which is higher that the 0.71
# percent used in Henriksson et al. (2014a). "

# 0.791 kgCO2e/kgLW production 
ghg_total_kgCO2_kg_fish <- 0.791 + ghg_energy_kgCO2_kg_fish

## convert to kgco2/tfish
GHG_kgCO2e_tfish <- ghg_total_kgCO2_kg_fish/0.001

## convert to MTco2e/tfish
GHG_MTCO2e_tfish <- GHG_kgCO2e_tfish/1000

## Now multiply this by our production raster
bivalve_production_raster <- rasterFromXYZ(bivalve_production)
plot(log(bivalve_production_raster + 1))
cellStats(bivalve_production_raster, "sum") # 17435623

ghg_em <- bivalve_production_raster*GHG_MTCO2e_tfish
plot(log(ghg_em + 1))
cellStats(ghg_em, "sum") # 18100668
ghg_em

#check
17435623*GHG_MTCO2e_tfish #18100667 - close enough
# it worked

## Transform CRS to be appropriate
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

crs(ghg_em) <- crs(food_raster)
ghg_em

ghg_em_final <- projectRaster(ghg_em, food_raster, method = "ngb")
plot(log(ghg_em_final+1))
cellStats(ghg_em_final, "sum") #18100668
ghg_em_final

## Save the raster to datalayers folder
writeRaster(ghg_em_final, file.path(path_final, "marine_bivalve_aquaculture_meat_ghg.tif"), overwrite = TRUE)

```