### 2021-02-03 marine/bivalve/food_pressures/bivalve_nutrient/bivalve_nutrient.Rmd

---
title: "Creating a file for bivalve farming nutrient pollution"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/23/2020"
output: html_document
---

This markdown creates a nutrient pollution impact layer for bivalve aquaculture. This impact is zero for bivalves, but there is an impact from gastropods. This markdown uses models and species specific parameters from the paper *Aquaculture Production is a Large, Spatially Concentrated Source of Nutrients in Chinese Freshwater and Coastal Seas* (Wang et al. 2020; table 1).

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/bivalve"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

Because we have multiple values and species for nutrient pollution for bivalves (bivalves and gastropods), we will have to do this analysis in a series of steps. Workflow will follow these rules:

1. Apply the gastropod value to places with gastropod production (and in the right proportion)
2. Apply 0 for other bivalves

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P value
3. We report in summed N and P (total excess nutrients), so sum the N and P rasters (and save each separately as a prep file).
4. We report in tonnes, so divide by 1000

**Numbers for each category:** 
Gastropods: 
N = 110 kg/tonne production, P = 15 kg/tonne production

Bivalves:
N = 0, P = 0

```{r}
fao <- read.csv(here("aquaculture/data/fao_mariculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

## Read in fao data and get a species list for salmon
species_list <- read.csv(here("aquaculture/marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

taxon_list <- read.csv(here("aquaculture/marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, Taxon_code, aq_group)

gastropod_list <- taxon_list %>%
  dplyr::filter(Taxon_code == "GAST") %>%
  dplyr::filter(species != "Marine molluscs nei") %>%
  distinct(species)

gastropod_list <- gastropod_list$species

# the list of species we are including
bivalves_list <- species_list$species[species_list$aq_group=="bivalves"]
bivalves_list

## write out shrimp taxa and production for Juliette
fao_bivalves_taxa_prod <- fao %>%
  filter(species %in% bivalves_list) %>%
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

## Now we need to get the proportion of catch that falls into each species category (i.e. pelagic or non-pelagic) 
prop_df <- fao_bivalves_taxa_prod %>%
  dplyr::mutate(species_cat = ifelse(species %in% gastropod_list, "gastropod", "bivalves")) %>%
  group_by(iso3c, species_cat) %>%
  summarise(tonnes = sum(fao_tonnes_production)) %>%
  ungroup() %>%
  group_by(iso3c) %>%
  mutate(total_tonnes = sum(tonnes)) %>%
  ungroup()

prop_bivalve_df <- prop_df %>%
  dplyr::filter(species_cat == "bivalves") %>%
  dplyr::mutate(prop_bivalve = tonnes/total_tonnes) %>%
  dplyr::mutate(prop_gastropod = 1-prop_bivalve) %>%
  dplyr::select(iso3c, prop_gastropod, prop_bivalve, total_tonnes)

sum(prop_bivalve_df$total_tonnes) # 17435623
sum(fao_bivalves_taxa_prod$fao_tonnes_production) # 17435623

## Read in production data 
production_sf <- st_read(file.path("/home/sgclawson/github/food_systems/aquaculture/marine/bivalve/bivalve_nutrient/prod_data_by_cell_iso3c.shp")) %>%
  dplyr::rename("tonnes_per_farm" = "tnns_p_")

## join with our proportion data so that we can get tonnes of pelagic and tonnes of non-pelagic per farm 
production_spp_props <- production_sf %>%
  left_join(prop_bivalve_df, by = "iso3c") %>%
  dplyr::mutate(tonnes_gastropod = prop_gastropod*tonnes_per_farm, 
                tonnes_bivalve = prop_bivalve*tonnes_per_farm)

sum(production_spp_props$tonnes_bivalve) + sum(production_spp_props$tonnes_gastropod) # 17435623

## Now we can do the actual analysis outlined above
nutrient_data <- production_spp_props %>%
  dplyr::mutate(tonnes_gastropod_growth = tonnes_gastropod*0.95,
                tonnes_bivalve_growth = tonnes_bivalve*0.95) %>%   ## multiply by 0.95 to correct for growth outside seapen

  dplyr::mutate(N_poll_gast = tonnes_gastropod_growth*110,
                P_poll_gast = tonnes_gastropod_growth*15,
                N_poll_bivalve = 0*tonnes_bivalve_growth,
                P_poll_bivalve = 0*tonnes_bivalve_growth) %>%  ## get pollution values by species group

  dplyr::mutate(N_poll_kg = N_poll_gast + N_poll_bivalve,
                P_poll_kg = P_poll_gast + P_poll_bivalve) %>% ## get total pollution values for each cell in kg

  # dplyr::mutate(po4eq_kg = N_poll*0.42 + P_poll*3.07) %>% ## get po4eq; not doing this anymore (01/14/2021)

   dplyr::mutate(N_poll_t = N_poll_kg/1000,
                P_poll_t = P_poll_kg/1000) %>% ## get total pollution values for each cell in tonnes

  dplyr::mutate(total_N_P_t = N_poll_t + P_poll_t) %>%
  dplyr::select(N_poll_t, P_poll_t, total_N_P_t)

sum(nutrient_data$total_N_P_t) #  50167.83
sum(production_spp_props$tonnes_gastropod) # 422466

## Save total N_P layer
bivalves_raster_total <- rasterize(nutrient_data, food_raster, field = "total_N_P_t", fun=sum, background=0)
plot(log(bivalves_raster_total+1))

bivalves_raster_df <- as.data.frame(bivalves_raster_total, xy=TRUE) %>%
  dplyr::select(x,y, N_P_tonnes=layer)

cellStats(bivalves_raster_total, stat = "sum") # 50167.83
sum(bivalves_raster_df$N_P_tonnes) # 50167.83
sum(nutrient_data$total_N_P_t, na.rm = TRUE) # 50167.83

## save to dataprep folder with the correct naming conventions: 
write.csv(bivalves_raster_df, file.path(path_prep, "bivalves_nutrient.csv"))

final_rast <- rasterFromXYZ(bivalves_raster_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 50167.83

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 50167.83
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path(path_final, "marine_bivalve_aquaculture_meat_nutrient.tif"), overwrite = T)

## Save N nutrient layer
bivalves_raster_N <- rasterize(nutrient_data, food_raster, field = "N_poll_t", fun=sum, background=0)
plot(log(bivalves_raster_N+1))

bivalves_raster_df <- as.data.frame(bivalves_raster_N, xy=TRUE) %>%
  dplyr::select(x,y, N_tonnes=layer)

cellStats(bivalves_raster_N, stat = "sum") # 44147.69
sum(bivalves_raster_df$N_tonnes) # 44147.69
sum(nutrient_data$N_poll_t, na.rm = TRUE) # 44147.69

## save to dataprep folder with the correct naming conventions: 
write.csv(bivalves_raster_df, file.path(path_prep, "bivalves_nutrient_N.csv"))

final_rast <- rasterFromXYZ(bivalves_raster_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 44147.69

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 44147.69
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_bivalve_aquaculture_meat_nutrientN.tif"), overwrite = T)

## Save P nutrient layer
bivalves_raster_P <- rasterize(nutrient_data, food_raster, field = "P_poll_t", fun=sum, background=0)
plot(log(bivalves_raster_P+1))

bivalves_raster_df <- as.data.frame(bivalves_raster_P, xy=TRUE) %>%
  dplyr::select(x,y, P_tonnes=layer)

cellStats(bivalves_raster_P, stat = "sum") # 6020.14
sum(bivalves_raster_df$P_tonnes) # 6020.14
sum(nutrient_data$P_poll_t, na.rm = TRUE) # 6020.14

## save to dataprep folder with the correct naming conventions: 
write.csv(bivalves_raster_df, file.path(path_prep, "bivalves_nutrient_P.csv"))

final_rast <- rasterFromXYZ(bivalves_raster_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 6020.14

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 6020.14
final_rast

## Save to datalayers 
writeRaster(final_rast, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers/marine_bivalve_aquaculture_meat_nutrientP.tif"), overwrite = T)
```