### 2021-02-03 marine/marine_fish_general/food_pressures/finfish_ghg/STEP1_ghg_emissions.Rmd

---
title: "Marine Fish - General: Aquaculture GHG"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description

This markdown calculates the GHG emissions from general finfish aquaculture.

# Data information
We used [Seafood Carbon Emissions Tool data](http://seafoodco2.dal.ca/) to extract ghg emissions associated with carcass weight of finfish from finfish farming.

One factor that we added was aquatic N<sub>2</sub>O emissions as discussed in [Hu *et al*., 2012](https://pubs.acs.org/doi/full/10.1021/es300110x) and [MacLoid *et al*., 2019](http://www.fao.org/3/ca7130en/ca7130en.pdf).

# Methods 
 - Take midpoint of ghg low and ghg high from [seafood carbon emissions tool data](http://seafoodco2.dal.ca/) for all finfish species
 - Multiply by % energy (to cut out the feed emissions)
 - Add the N20 value from the MacLeod paper to the emissions intensity estimate we got for finfish
 - Multiply emissions intensity estimate by tonnes of production in each cell

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(tidyverse)
library(here)
library(raster)
library(sf)
library(janitor)

# Raster templates
source(here("_spatial/template_raster.R"))

# File paths
raw   <- "/home/shares/food-systems/Food_footprint/_raw_data/"
prep_finfish <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/marine_fish_general"
prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/" 
path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

# Import master_rgns xy df
master_rgns <- read_csv(here("_spatial/_output/food_rgns.csv"))
```

## Import
```{r}
## Read in farmed emissions data

farmed_emissions <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/seafood_carbon_emmisions_tool/d2020/Specie_List_-_11_24_2020,_11_39_21_AM_Farmed.csv")) %>%
  clean_names()

finfish_production <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/marine_fish_general/marine_fish_general_farm.csv"))

fao_finfish_spp <- read_csv(here("aquaculture/marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish_taxa_production.csv"))

```

# Wrangle

```{r}
unique(sort(fao_finfish_spp$species))

unique(sort(farmed_emissions$common_name))
## we only want finfishs that are in fao
 #  Emissions spp   -   FAO spp
 ###############################################
 # "Almaco jack" - Falls under "Amberjacks nei"                               
 # "Barramundi" -  "Barramundi(=Giant seaperch)"                    
 # "California yellowtail" - "Amberjacks nei"
 # "Cobia" - "Cobia"          
 # "European bass" - "European seabass             
 # "Gilt-head seabream" - "Gilthead seabream"
 # "Japanese amberjack" - "Japanese amberjack"  
 # "Turbot" - "Turbot"

## Filter for the appropriate species... also filter for "Marine net pen" and "Cages".. since we don't want any inland aquaculture
finfish_emissions_spp <- farmed_emissions %>%
  dplyr::filter(common_name %in% c("Almaco jack", "Barramundi", "California yellowtail", "Cobia", "European bass", "Gilt-head seabream", "Japanese amberjack", "Turbot")) %>%
  dplyr::filter(system %in% c("Marine net pen", "Cages"))

## after the filter we lost Turbot... there wasn't that much production anyways.

## This looks pretty good though... all of the ghg low and ghg high are generally between 2 and 4... with the exception of cobia. I feel comfortable just taking the midpoint.

## multiply the ghg low and ghg high by the percent energy
## take the mean of new ghg low and high (the 25th and 75th quantiles)
finfish_emissions <- finfish_emissions_spp %>%
  dplyr::mutate(ghg_low_energy = ghg_low*(percent_energy/100),
                ghg_high_energy = ghg_high*(percent_energy/100)) %>%
  rowwise() %>%
  dplyr::mutate(ghg_kgCO2_kg_fish_mean = mean(c(ghg_low_energy, ghg_high_energy)))

ghg_kgCO2_kg_fish <- mean(finfish_emissions$ghg_kgCO2_kg_fish_mean)

ghg_energy_kgCO2_kg_fish <- ghg_kgCO2_kg_fish # 0.381375 kg CO2eq/ kg fish (live/carcass weight, ratioed)

# Adding Aquatic N2O
# From MacLeod et al. 
# "The amount of N2O per species-group was determined by multiplying the
# production by the N2O emission factor per kg of production (Hu et al., 2012), i.e.
# 1.69 gN2O-N per kg of production, or 0.791 kgCO2e/kgLW production. This equates
# to a conversion rate of N to N2O-N of 1.8 percent, which is higher that the 0.71
# percent used in Henriksson et al. (2014a). "

# 0.791 kgCO2e/kgLW production 
ghg_total_kgCO2_kg_fish <- 0.791 + ghg_energy_kgCO2_kg_fish ## 1.172375

## convert to kgco2/tfish
GHG_kgCO2e_tfish <- ghg_total_kgCO2_kg_fish/0.001

## convert to MTco2e/tfish
GHG_MTCO2e_tfish <- GHG_kgCO2e_tfish/1000

finfish_ghg <- finfish_production %>%
  dplyr::mutate(finfish_ghg_tonnes = GHG_MTCO2e_tfish*tonnes_production) %>%
  dplyr::select(x,y, finfish_ghg_tonnes)

sum(finfish_ghg$finfish_ghg_tonnes)

write.csv(finfish_ghg, file.path(prep_finfish, "marine_fish_general_ghg.csv"))

## Now multiply this by our production raster
finfish_production_raster <- rasterFromXYZ(finfish_production)
plot(log(finfish_production_raster + 1))
cellStats(finfish_production_raster, "sum") # 3901214

ghg_em <- finfish_production_raster*GHG_MTCO2e_tfish
plot(log(ghg_em + 1))
cellStats(ghg_em, "sum") # 4573686
ghg_em

#check
3901214*GHG_MTCO2e_tfish #4573686
# it worked

## Transform CRS to be appropriate
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

crs(ghg_em) <- crs(food_raster)
ghg_em

ghg_em_final <- projectRaster(ghg_em, food_raster, method = "ngb")
plot(log(ghg_em_final+1))
cellStats(ghg_em_final, "sum") # 4573686
ghg_em_final

## Save the raster to datalayers folder
writeRaster(ghg_em_final, file.path(path_final, "marine_marine-fish-general_aquaculture_meat_ghg.tif"), overwrite = TRUE)

```