### 2021-02-03 marine/salmon/food_pressures/salmon_ghg/STEP1_ghg_emissions.Rmd

---
title: "Salmon: Aquaculture GHG"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description

This markdown calculates the GHG emissions from salmon aquaculture.

# Data information
We used [Seafood Carbon Emissions Tool data](http://seafoodco2.dal.ca/) to extract ghg emissions associated with carcass weight of salmon from salmon farming.

One factor that we added was aquatic N<sub>2</sub>O emissions as discussed in [Hu *et al*., 2012](https://pubs.acs.org/doi/full/10.1021/es300110x) and [MacLeod *et al*., 2019](http://www.fao.org/3/ca7130en/ca7130en.pdf).

# Methods 
 - Take midpoint of ghg low and ghg high from [seafood carbon emissions tool data](http://seafoodco2.dal.ca/) for salmon, trout, and char species
 - Multiply by % energy (to cut out the feed emissions)
 - Add the N20 value from the MacLeod paper to the emissions intensity estimate we got for salmon
 - Multiply emissions intensity estimate by tonnes of production in each cell

# Preamble
```{r setup, include = FALSE}
# getting packages we want
library(tidyverse)
library(here)
library(raster)
library(sf)
library(janitor)

# Raster templates
source(here("_spatial/template_raster.R"))

# File paths
raw   <- "/home/shares/food-systems/Food_footprint/_raw_data/"
prep_salmon <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/salmon"
prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/" 
path_final <- "/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"

# Import master_rgns xy df
master_rgns <- read_csv(here("_spatial/_output/food_rgns.csv"))
```

## Import
```{r}
## Read in farmed emissions data

farmed_emissions <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/seafood_carbon_emmisions_tool/d2020/Specie_List_-_11_24_2020,_11_39_21_AM_Farmed.csv")) %>%
  clean_names()

salmon_production <- read_csv(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/salmon/salmon_farm.csv"))

fao <- read.csv(here("aquaculture/data/fao_mariculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

## Read in fao data and get a species list for salmon 
species_list <- read.csv(here("aquaculture/marine/species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

# the list of species we are including
salmon_list <- species_list$species[species_list$aq_group=="salmonids"]
salmon_list

# include only salmon
fao_salmon <- fao %>%
  filter(species %in% salmon_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, species, iso3c, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, species, iso3c, year) %>%
  dplyr::filter(year == 2017)

```

# Wrangle

```{r}
unique(sort(fao_salmon$species))
# [1] "Arctic char"                  "Atlantic salmon"              "Chinook(=Spring=King) salmon" "Coho(=Silver) salmon"        
# [5] "European whitefish"           "Rainbow trout"                "Salmonids nei"                "Trouts nei"

unique(sort(farmed_emissions$common_name))

## we only want salmons that are in fao
 #  Emissions spp   -   FAO spp
 ###############################################
salmon_list <- c("Arctic char", "Atlantic salmon", "Chinook salmon", "Rainbow trout", "Sea trout")

## Filter for the appropriate species... also filter for "Marine net pen".. since we don't want any inland aquaculture
salmon_emissions_spp <- farmed_emissions %>%
  dplyr::filter(common_name %in% salmon_list) %>%
  dplyr::filter(system %in% c("Marine net pen"))

## This looks pretty good though... all of the ghg low and ghg high are between 1 and 3. I feel comfortable just taking the midpoint.

## multiply the ghg low and ghg high by the percent energy
## take the mean of new ghg low and high (the 25th and 75th quantiles)
salmon_emissions <- salmon_emissions_spp %>%
  dplyr::mutate(ghg_low_energy = ghg_low*(percent_energy/100),
                ghg_high_energy = ghg_high*(percent_energy/100)) %>%
  rowwise() %>%
  dplyr::mutate(ghg_kgCO2_kg_fish_mean = mean(c(ghg_low_energy, ghg_high_energy)))

ghg_energy_kgCO2_kg_fish <- mean(salmon_emissions$ghg_kgCO2_kg_fish_mean) # 0.159 kg CO2eq/ kg fish (live/carcass weight, ratioed)

# Adding Aquatic N2O
# From MacLeod et al. 
# "The amount of N2O per species-group was determined by multiplying the
# production by the N2O emission factor per kg of production (Hu et al., 2012), i.e.
# 1.69 gN2O-N per kg of production, or 0.791 kgCO2e/kgLW production. This equates
# to a conversion rate of N to N2O-N of 1.8 percent, which is higher that the 0.71
# percent used in Henriksson et al. (2014a). "

# 0.791 kgCO2e/kgLW production 
ghg_total_kgCO2_kg_fish <- 0.791 + ghg_energy_kgCO2_kg_fish ## 0.95

## convert to kgco2/tfish
GHG_kgCO2e_tfish <- ghg_total_kgCO2_kg_fish/0.001

## convert to MTco2e/tfish
GHG_MTCO2e_tfish <- GHG_kgCO2e_tfish/1000

salmon_production_ghg <- salmon_production %>% 
  dplyr::mutate(ghg_tonnes = tonnes_production*GHG_MTCO2e_tfish) %>%
  dplyr::select(x, y, ghg_tonnes)

sum(salmon_production_ghg$ghg_tonnes) # 2608782

write.csv(salmon_production_ghg, file.path(prep_salmon, "salmon_ghg.csv"))

## Now multiply this by our production raster
salmon_production_raster <- rasterFromXYZ(salmon_production)
plot(log(salmon_production_raster + 1))
cellStats(salmon_production_raster, "sum") # 2746086

ghg_em <- salmon_production_raster*GHG_MTCO2e_tfish
plot(log(ghg_em + 1))
cellStats(ghg_em, "sum") # 2608782
ghg_em

norway <- raster::crop(ghg_em, extent(-25, 50, 50, 75))
par(mar=c(1,1,1,1))
plot(norway)
chile <- raster::crop(ghg_em, extent(-80, -60, -60, -35))
par(mar=c(1,1,1,1))
plot(chile)

#check
2746086*GHG_MTCO2e_tfish #2608782
# it worked

## Transform CRS to be appropriate
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

crs(ghg_em) <- crs(food_raster)
ghg_em

ghg_em_final <- projectRaster(ghg_em, food_raster, method = "ngb")
plot(log(ghg_em_final+1))
cellStats(ghg_em_final, "sum") # 2608782
ghg_em_final

## Save the raster to datalayers folder
writeRaster(ghg_em_final, file.path(path_final, "marine_salmon_aquaculture_meat_ghg.tif"), overwrite = TRUE)

```