### 2021-02-04 marine/marine_fish_general/finfish_farms/STEP2_finfish_farm_locations.Rmd

---
title: "Location of Marine fish - general Farms"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/08/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script standardizes location of marine fish - general (finfish) farms to be lat/long point data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

fao_finfish_spp <- read_csv("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish_taxa_production.csv") %>%
  dplyr::filter(year == 2017)

fao_finfish <- read_csv("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv") %>%
  dplyr::filter(year == 2017)

unique(fao_finfish$country)

#  [1] "Albania"                  "Australia"                "Bangladesh"               "Cambodia"                
#  [5] "China"                    "China, Hong Kong SAR"     "Costa Rica"               "Croatia"                 
#  [9] "Cyprus"                   "Egypt"                    "France"                   "Greece"                  
# [13] "India"                    "Indonesia"                "Iran (Islamic Rep. of)"   "Israel"                  
# [17] "Italy"                    "Japan"                    "Korea, Republic of"       "Malaysia"                
# [21] "Malta"                    "Mauritius"                "Mexico"                   "Norway"                  
# [25] "Panama"                   "Philippines"              "Portugal"                 "Saudi Arabia"            
# [29] "Singapore"                "Spain"                    "Taiwan Province of China" "Thailand"                
# [33] "Tunisia"                  "Turkey"                   "United Arab Emirates"     "United States of America"
# [37] "Viet Nam"     
```

## explore
https://gis.asc-aqua.org/arcgis_app/
Maps from ASC

```{r}
ASC_data <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

ASC_finfish <- ASC_data %>%
  dplyr::filter(aq_grop == "marine_fish_general")

unique(ASC_finfish$crt_cnt)

#  [1] "Turkey"                   "Croatia"                  "Greece"                   "Albania"                 
#  [5] "Spain"                    "Japan"                    "Netherlands"              "Denmark"                 
#  [9] "Australia"                "Mexico"                   "Panama"                   "United States of America"
# [13] "Germany"                  "Viet Nam"                 "Thailand"

## what is in ASC and not in FAO
setdiff(ASC_finfish$crt_cnt, fao_finfish$country)
# [1] "Netherlands" "Denmark"     "Germany"

## what is in FAO and not ASC -- this is what we need to find alternative data sources for (25 countries)
setdiff(fao_finfish$country, ASC_finfish$crt_cnt)
#  [1] "Bangladesh"               "Cambodia"                 "China"                    "Costa Rica"               "Cyprus"                  
#  [6] "Egypt"                    "France"                   "India"                    "Indonesia"                "Iran (Islamic Rep. of)"  
# [11] "Israel"                   "Italy"                    "Korea, Republic of"       "Malaysia"                 "Malta"                   
# [16] "Mauritius"                "Norway"                   "Philippines"              "Portugal"                 "Saudi Arabia"            
# [21] "Singapore"                "Taiwan Province of China" "Tunisia"                  "United Arab Emirates"

```

# Raster

Standard raster used for this analysis

```{r}

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

# Map data from ASC

Data comes from https://gis.asc-aqua.org/arcgis_app/

Source:
Aquaculture Stewardship Council Data Mapping Study, n.d. . Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}

# get iso3c codes
iso <-   data.frame(iso3c = countrycode(ASC_finfish$crt_cnt, origin = 'country.name', destination = 'iso3c'))

# read in shp
farms_all <- ASC_finfish %>%
  cbind(iso) %>%
  dplyr::filter(!(crt_cnt %in% c("Denmark", "Germany", "Netherlands")))

mapview(farms_all$geometry)

unique(farms_all$iso3c)

all_pts <- st_centroid(farms_all)

all_pts_latlon <- st_transform(all_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry, iso3c)

farms_all_fin <- data.frame(do.call(rbind, st_geometry(all_pts_latlon))) 
names(farms_all_fin) <- c("lon", "lat")
farms_all_fin <- farms_all_fin %>%
  cbind(iso3c = all_pts_latlon$iso3c)

## save farm location
write.csv(farms_all_fin, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/ASC_finfish_farm_lat_lon.csv"), row.names = FALSE)

```

Now we will map the remaining countries with marine finfish farms that are not included in the ASC dataset. Workflow will be as follows: 
 - Try to obtain exact locations of farms from published source or dataset (such as FAO or a research paper)
 - If there are no exact locations, but the source lists a city or state, place a point in that particular city or state
 -  If there is enough information to know there are a specific number of finfish farms, equally allocate the farms across the coastline. If there are estimates of farms per state/province, equally allocate them across the state/province, to be even more precise.
 - If there are not exact numbers of farms in the country, but you know there are a lot, randomly place points along the coast line (a non-exact number, like we did for Taiwan bivalve/shrimp). 
 - Lastly, if there is no evidence of marine finfish farm locations, place a point on the coast of the country (preferably with some intuition/knowledge of where they might be, i.e. what we've done for Russia for bivalves).

##  Europe - EMODNET: Malta, Greece, Norway, Cyprus, Spain 
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Finfish+Production
https://www.emodnet-humanactivities.eu/view-data.php

Source:
EMODnet Human Activities: Finfish production, 2019. [Accessed December 8, 2020] 
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Finfish+Production

```{r}

## read in shapefile
eu <- st_read(file.path(path,"EMODnet_HA_Aquaculture_Marine_Finfish_20190104/EMODnet_HA_Aquaculture_Marine_Finfish_20190104.shp"))
mapview(eu)

unique(eu$COUNTRY)
# [1] "Spain"          "United Kingdom" "Greece"         "Finland"        "Norway"         "Ireland"        "Malta"          "Denmark"       
# [9] "Cyprus"

unique(eu$FARM_TYPE)

unique(eu$SPECIES_DE)

unique(eu$STATUS)

eu_spp <- eu %>%
  dplyr::filter(str_detect(SPECIES_DE, "Mediterranean marine fish|alibut|eabass|eabream|urbot|Sole|sole|sea bream|sea bass|Sea bream|Sea bass"))

unique(eu_spp$COUNTRY)

## Need to filter for the ones in FAO dataset
# [1] "Spain" - yes          "United Kingdom" - no "Greece" - yes        "Norway" - yes        "Denmark" - no       "Cyprus" - yes  "Malta" - yes

eu_na <- eu_spp %>%
  dplyr::filter(STATUS == "n.a.")
unique(eu_na$COUNTRY)

eu_active <- eu_spp %>%
  dplyr::filter(STATUS == "Active")

unique(eu_active$COUNTRY)

## we will use both "Active" and "n.a." as the status of the farms. I choose this because all of the denmark points are n.a., and they correspond with the HELCOM data that we used for other spp groups. This indicates to me that the farms denoted as "n.a." are probably actually active.

eu_active_na <- eu_spp %>%
  dplyr::filter(STATUS %in% c("Active", "n.a."))

unique(eu_active_na$COUNTRY)

# [1] "Spain"          "United Kingdom" "Greece"         "Malta"          "Norway"         "Denmark"        "Cyprus"

## Now filter for countries that we want: Spain, Greece, Norway, Cyprus, Malta... the others are not included in FAO 2017
eu_fao_countries <- eu_active_na %>%
  dplyr::filter(COUNTRY %in% c("Spain", "Greece", "Norway", "Cyprus", "Malta"))

unique(eu_fao_countries$COUNTRY)
mapview(eu_fao_countries)

eu_pts <- st_centroid(eu_fao_countries)

eu_pts_latlon <- st_transform(eu_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(COUNTRY, geometry)

eu_countries <- eu_pts_latlon %>%
  dplyr::select(COUNTRY)

farms_eu <- data.frame(do.call(rbind, st_geometry(eu_pts_latlon)))
names(farms_eu) <- c("lon", "lat")

farms_eu_final <- cbind(farms_eu, eu_countries) %>%
  dplyr::select(-geometry) %>%
  mutate(iso3c = case_when(
    COUNTRY == "Spain" ~ "ESP",
    COUNTRY == "Cyprus" ~ "CYP", 
    COUNTRY == "Norway" ~ "NOR", 
    COUNTRY == "Greece" ~ "GRC",
    COUNTRY == "Malta" ~ "MLT"
  )) %>%
  dplyr::select(-COUNTRY)

 write_csv(farms_eu_final, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_ESP_CYP_NOR_GRC_MLT.csv"))

```

## Italy
From http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/italy/en/

Source:
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/thailand/en/

```{r}
## read in thailand shp file
italy_file <- file.path(path, "FAO_NASO/NASO/Point/Italy.kml")

italy_aqua <- read_sf(italy_file)

unique(italy_aqua$description)

mapview(italy_aqua)

## We need to filter for just the appropriate species of finfish aquaculture

italy_finfish <- italy_aqua %>%
  dplyr::filter(!(description %in% c("Floating Cages","Oysters<br>Mussels", "Freshwater farm<br><br>Strugeons<br>Stripped bass"))) %>%
  dplyr::filter(!is.na(description))

mapview(italy_finfish)

ita_bivalve_mol <- st_transform(italy_finfish, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
ita_pts <- st_centroid(ita_bivalve_mol)

ita_pts_latlon <- st_transform(ita_pts, crs = "+init=epsg:4326")

mapview(ita_pts_latlon)

farms_ita <- data.frame(do.call(rbind, st_geometry(ita_pts_latlon))) 
names(farms_ita) <- c("lon", "lat")
farms_ita <- farms_ita %>%
  mutate(iso3c="ITA")

write_csv(farms_ita, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_ita.csv"))

farms_ita <-  st_as_sf(farms_ita, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_ita)
```

## France, Israel
http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/mediterranean-basin/en/

 - It says both Tuna and Finfish, but we'll assume france and israel are just finfish...

Source:
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/thailand/en/

```{r}
## read in thailand shp file
france_file <- file.path(path, "FAO_NASO/NASO/Point/Mediterranean.kml")

france_aqua <- read_sf(france_file)

med_countries <- france_aqua %>%
  dplyr::mutate(country = sub("_.*", "", Name)) %>%
  dplyr::mutate(country = sub(" fishpen A", "", country))

unique(med_countries$country)

## We need to filter for just the appropriate species of finfish aquaculture, France, Israel

med_finfish <- med_countries %>%
  dplyr::filter(country %in% c("France", "Israel"))

mapview(med_finfish)

fra_finfish_mol <- st_transform(med_finfish, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
fra_pts <- st_centroid(fra_finfish_mol)

fra_pts_latlon <- st_transform(fra_pts, crs = "+init=epsg:4326")

mapview(fra_pts_latlon)

countries <- fra_pts_latlon %>%
  dplyr::select(country)

farms_fra <- data.frame(do.call(rbind, st_geometry(fra_pts_latlon))) 
names(farms_fra) <- c("lon", "lat")
farms_fra <- farms_fra %>%
  cbind(countries) %>%
  mutate(iso3c = ifelse(country == "France", "FRA", "ISR")) %>%
  dplyr::select(-country, -geometry)

write_csv(farms_fra, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_fra_isr.csv"))

farms_fra <-  st_as_sf(farms_fra, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_fra)
```

## Indonesia
NASO maps 
 - Per the FAO aquaculture sector report, Indonesia basically only grows grouper/milkfish, so I will assume this is all grouper/milkfish.

Source: 
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/thailand/en/

```{r}
indo_file <- file.path(path, "FAO_NASO/NASO/Point/Indonesia.kml")

indo_aqua <- read_sf(indo_file)

mapview(indo_aqua)

idn_finfish_mol <- st_transform(indo_aqua, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
idn_pts <- st_centroid(idn_finfish_mol)

idn_pts_latlon <- st_transform(idn_pts, crs = "+init=epsg:4326")

mapview(idn_pts_latlon)

farms_idn <- data.frame(do.call(rbind, st_geometry(idn_pts_latlon))) 
names(farms_idn) <- c("lon", "lat")
farms_idn <- farms_idn %>%
  mutate(iso3c = "IDN")

write_csv(farms_idn, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_idn.csv"))

farms_idn <-  st_as_sf(farms_idn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_idn)
```

## Portugal
 - https://www.ine.pt/xportal/xmain?xpid=INE&xpgid=ine_publicacoes&PUBLICACOESpub_boui=435690295&PUBLICACOESmodo=2&fbclid=IwAR0-_SHYtw2XGUAXdx71RGer284Vk_u40HYKXyJzeFbpdnsHydKgAfomTvU
 - National Institute of Statistics for Portugal 
 - There is a csv table, "q5_1.csv", which has the number of farms listed by type of system. I will assume that tanks are for finfish, since [this website](https://rea.apambiente.pt/content/aquaculture-production?language=en) says that the tanks are for **fish production**, and floating structures are for **bivalve molluscs**. Located here "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Portugal/EPesca_2019.xlsx"
 - According to the table, by these criteria, there are 112 marine tank farm systems in Portugal. 
 - I will equally allocate 112 points along the coast of Portugal.

Source: 
Estatísticas da Pesca, 2020. . Instituto Nacional de Estatística. [Accessed December 21, 2020]

```{r}
world <- rnaturalearth::ne_coastline(scale = 110) # world map with 10m resolution
plot(world)

portugal <- ne_countries(scale = 10, country = "Portugal", returnclass = "sf")

#info of our spatial vector object
portugal
mapview(portugal)
plot(portugal)

world <- st_as_sf(world)

portugal_world_crs <- st_transform(portugal, crs(world))

portugal_world_crop <- st_crop(world, portugal_world_crs) #%>%
 # dplyr::filter(scalerank == 5) ## filter for the appropriate island
plot(portugal_world_crop)
mapview(portugal_world_crop)

portugal_crop <- as(portugal_world_crop, "Spatial")

portugal_points_lines <- sample.line(portugal_crop, n = 380) ## equally place the number of farms along the coastline 
class(portugal_points_lines)
plot(portugal_points_lines)

portugal_points <- st_as_sf(portugal_points_lines)
mapview(portugal_points)

portugal_points_final <- portugal_points[253:364, ] ## grab the appropriate number of points
mapview(portugal_points_final)

## now prep dataframe for lat lon points
prt_pts <- st_centroid(portugal_points_final)

prt_pts_latlon <- st_transform(prt_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_prt <- data.frame(do.call(rbind, st_geometry(prt_pts_latlon)))
names(farms_prt) <- c("lon", "lat")

farms_prt_final <- farms_prt %>%
  mutate(iso3c = "PRT")

## visualize it 
farms_prt <-  st_as_sf(farms_prt, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_prt)

## write to farms folder 
write_csv(farms_prt_final, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_prt.csv"))
```

## Iran

The only species the FAO reports for Iran is "Marine fishes nei", and I can't find any information on number of marine farms there might be in Iran. Therefore I will place one farm on the Caspian Sea coast.

Source:
NA

```{r}

farms_irn <- data.frame(lat = c(37.392419247868425),
                        lon = c(50.1602963496454)) %>%
  mutate(iso3c = "IRN")

write_csv(farms_irn, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_irn.csv"))

## visualize it 
farms_irn <-  st_as_sf(farms_irn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_irn)
```

## China 
Can't find any datasets of marine finfish farm locations in China. Therefore I will place one farm on the coast of China.

Source: 
NA

```{r}

farms_chn <- data.frame(lat = c(26.649767732301385),
                        lon = c(119.94149300994019)) %>%
  mutate(iso3c = "CHN")

write_csv(farms_chn, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_chn.csv"))

## visualize it 
farms_chn <-  st_as_sf(farms_chn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_chn)
```

## Saudi Arabia
 - There is really only one marine fish aquaculture producer in Saudi Arabia, and it happens to be the same company that produces almost all of its shrimp. It is one of the largest facilities in the world. http://www.naqua.com.sa/barramundi.php
 - I'll place a point there: 20.287253, 40.030864

Source:
National Aquaculture Group, n.d. [Accessed November 12, 2020]. http://www.naqua.com.sa/about-us.php

```{r}

farms_sau <- data.frame(lat = c(20.287253),
                        lon = c(40.030864)) %>%
  mutate(iso3c = "SAU")

write_csv(farms_sau, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_sau.csv"))

## visualize it 
farms_sau <-  st_as_sf(farms_sau, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_sau)
```

## Mauritius

Can't find any datasets of marine finfish farm locations. Therefore I will place one farm on the coast. It probably doesn't matter, since Mauritius is so small anyways.

Source:
NA

```{r}

farms_mus <- data.frame(lat = c(-20.005039247819017),
                        lon = c(57.67022708374523)) %>%
  mutate(iso3c = "MUS")

write_csv(farms_mus, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_mus.csv"))

## visualize it 
farms_mus <-  st_as_sf(farms_mus, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_mus)
```

## United Arab Emirates

http://www.fishfarm.ae/facilities.html
 - One here in Dibba Al Fujairah. 
 - They do sea bream and sea bass, which matches what the FAO reports.

The above is also corroborated by the FAO here: http://www.fao.org/fishery/countrysector/naso_uae/en

I will place a point off of the coast of Dibba Al Fujairah: 25.604715925918068, 56.340214330776234

Source: 
National Aquaculture Sector Overview. United Arab Emirates, 2016. , National Aquaculture Sector Overview Fact Sheets.

```{r}

farms_are <- data.frame(lat = c(25.604715925918068),
                        lon = c(56.340214330776234)) %>%
  mutate(iso3c = "ARE")

write_csv(farms_are, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_are.csv"))

## visualize it 
farms_are <-  st_as_sf(farms_are, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_are)
```

## Tunisia

https://apps.fas.usda.gov/newgainapi/api/report/downloadreportbyfilename?filename=Aquaculture%20Industry%20Small%20but%20Growing_Tunis_Tunisia_5-31-2017.pdf
"At present, there are roughly 50 companies involved in aquaculture production:
 - Shellfish farming: 12 private companies
 - Bluefin tuna fattening: 5 private companies
 - Other marine fish farming: 20 private companies
 - Inland fish farming: 15 private companies"

 "The primary marine aquaculture production area is in the Monastir governorate, along Tunisia’s eastern
shore. Marine aquaculture is practiced using modern and innovative techniques, whether in concrete
facilities with a density of 60 kg/m3
or in floating cages on the high seas. Alevin are exclusively
imported for sea bass and sea bream production while bluefin tuna are imported for fattening."

 - I'll place 20 points along the coast of Monastir

Source: 
Aquaculture Industry Small but Growing: Tunisia, 2017. Fishery and Agriculture Society, Tunis.

```{r}
## load municipality shapefile, downloaded from here: https://gadm.org/download_country_v3.html
tunisia <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/Tunisia/gadm36_TUN_shp/gadm36_TUN_1.shp")) %>%
  dplyr::filter(NAME_1 == "Monastir")

mapview(tunisia)

## OR if you don't municipality and only country info you could use this: 
# tunisia <- rnaturalearth::ne_countries(country = "Tunisia")

world <- rnaturalearth::ne_coastline(scale = 10) # world map with 10m resolution
plot(world)
class(world)

#info of our spatial vector object
tunisia
mapview(tunisia)

world <- st_as_sf(world)

tunisia_world_crs <- st_transform(tunisia, crs(world))

tunisia_world_crop <- st_crop(world, tunisia_world_crs) 
plot(tunisia_world_crop)
mapview(tunisia_world_crop)

tunisia_crop <- as(tunisia_world_crop, "Spatial")
class(tunisia_crop)
mapview(tunisia_crop)

tunisia_points_lines <- sample.line(tunisia_crop, n = 20) ## equally place the number of farms along the coastline 
class(tunisia_points_lines)
plot(tunisia_points_lines)

tunisia_points <- st_as_sf(tunisia_points_lines)
mapview(tunisia_points)

## now prep dataframe for lat lon points
tun_pts <- st_centroid(tunisia_points)

tun_pts_latlon <- st_transform(tun_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_tun <- data.frame(do.call(rbind, st_geometry(tun_pts_latlon)))
names(farms_tun) <- c("lon", "lat")

farms_tun_final <- farms_tun %>%
  mutate(iso3c = "TUN")

## visualize it 
farms_tun <-  st_as_sf(farms_tun, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_tun)

## write to farms folder 
write_csv(farms_tun_final, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_tun.csv"))
```

## Taiwan
The two main species farmed in Taiwan are Milkfish and Grouper. I will place farms for these two species

Milkfish:
https://digitalarchive.worldfishcenter.org/bitstream/handle/20.500.12348/3610/Pub%20TR4%206.pdf?sequence=1&isAllowed=y
 - I will use the respondents from the survey, and place the appropriate number of points in each location (table located on page 6, I will only use "gatherers"). This is a really old paper, but it is the best I can find. 42 total points

 - Putai (Budai): 4 points: 
 23.396775208505524, 120.1486974182384
 23.391435122712778, 120.14508600806049
 23.379465182872867, 120.13344924193156
 23.37173018517046, 120.12923593005728

 - Peimen (Beimen): 4 points
 23.328815202320154, 120.11007291797651
 23.30989854415313, 120.10560972235575
 23.286248936869796, 120.09462339467383
 23.24808538030484, 120.08741361713255

 - Chiku (Cigu): 2 points
 23.098351835751856, 120.0340739167051
 23.08684757757559, 120.0323838672442

 - Yungan: 4 points
 22.84129712135721, 120.19480335124462
 22.83607641104833, 120.19703494905501
 22.823893974670515, 120.19737827179506
 22.808703955682557, 120.19960986960544

 - Lingpien (Linbian) : 6 points
 22.427007004796724, 120.48594464702815
 22.42449445979184, 120.48934227493775
 22.419155150689896, 120.4981761075027
 22.413920334638384, 120.50893526254974
 22.41119815232611, 120.51527750131434
 22.40868532131159, 120.51890163775126

 - Taitung City: 6 points
 22.729748162448605, 121.13759861706552
 22.73478926413642, 121.14388413948092
 22.73995620049436, 121.14975973652137
 22.74096436042943, 121.15167272160434
 22.746257078096264, 121.15631854252005
 22.74965943121689, 121.16246742314381

 - Chungkung (Chengkung): 6 points
 23.094448352409827, 121.3724851599623
 23.100673804589537, 121.38612214670725
 23.110708254441576, 121.39450636818748
 23.118419405368904, 121.4097595903985
 23.12715198105829, 121.40864842851558
 23.13662711164484, 121.40026420703533

 - Erhchieh (River delta below Yilan City): 6 points
 24.718398599156366, 121.83377270959099
 24.72276233035146, 121.83292494893261
 24.7123233460693, 121.83961283857079
 24.734911520797553, 121.82878034126954
 24.707959248912765, 121.8381057085115
 24.70256809420227, 121.83829409976889

 - Suao: 4 points
 24.59561929860139, 121.87509035189372
 24.591466187548917, 121.87405621071602
 24.466470314555245, 121.84309342644401
 24.43396678054247, 121.80292466585698

Source:
Lee, C.-S., 1983. Production and marketing of milkfish in Taiwan: an economic analysis.

```{r}
milkfish_points <- data.frame(lat = c(23.396775208505524, 23.391435122712778, 23.379465182872867, 23.37173018517046, 23.328815202320154, 23.30989854415313, 23.286248936869796, 23.24808538030484, 23.098351835751856, 23.08684757757559, 22.84129712135721, 22.83607641104833, 22.823893974670515, 22.808703955682557, 22.427007004796724, 22.42449445979184, 22.419155150689896, 22.413920334638384, 22.41119815232611, 22.40868532131159, 22.729748162448605, 22.73478926413642, 22.73995620049436, 22.74096436042943, 22.746257078096264, 22.74965943121689,  23.094448352409827, 23.100673804589537, 23.110708254441576, 23.118419405368904, 23.12715198105829, 23.13662711164484, 24.718398599156366, 24.72276233035146, 24.7123233460693, 24.734911520797553, 24.707959248912765, 24.70256809420227, 24.59561929860139, 24.591466187548917, 24.466470314555245, 24.43396678054247),

                              lon = c(120.1486974182384, 120.14508600806049, 120.13344924193156, 120.12923593005728, 120.11007291797651, 120.10560972235575, 120.09462339467383, 120.08741361713255, 120.0340739167051, 120.0323838672442, 120.19480335124462, 120.19703494905501, 120.19737827179506, 120.19960986960544, 120.48594464702815, 120.48934227493775, 120.4981761075027, 120.50893526254974, 120.51527750131434, 120.51890163775126, 121.13759861706552, 121.14388413948092, 121.14975973652137, 121.15167272160434, 121.15631854252005, 121.16246742314381, 121.3724851599623, 121.38612214670725, 121.39450636818748, 121.4097595903985, 121.40864842851558, 121.40026420703533, 121.83377270959099, 121.83292494893261, 121.83961283857079, 121.82878034126954, 121.8381057085115, 121.83829409976889, 121.87509035189372, 121.87405621071602, 121.84309342644401, 121.80292466585698))
```

Grouper:
https://www.researchgate.net/publication/49753615_Real-Time_Quantitative_PCR_Assay_for_Monitoring_of_Nervous_Necrosis_Virus_Infection_in_Grouper_Aquaculture
 - Although this is just a sample of grouper farm locations, I will use this and place the points accordingly. Table and map on the second page. 24 total farms.

 - Kunshen: 5 points
 22.963578173481825, 120.16265849315296
 22.958915389973615, 120.16450385288077
 22.958796842534618, 120.16420344548321
 22.95887587417216, 120.16630629726609
 22.962708853156307, 120.16480426027832

 - Budai: 2 points
23.366020730714546, 120.12803212666465
23.344101510137286, 120.11599409273818

 - Linyuan: 4 points
 22.5060215233288, 120.361518692894
 22.495712950055186, 120.37199003646582
 22.489210224135196, 120.38503630058813
 22.478900398348003, 120.40220243759113

 - Jiading: 5 points
 22.910490254668936, 120.17469982002679
 22.900053951559105, 120.17658809509712
 22.88771910346936, 120.18208125893808
 22.87475049020418, 120.18637279318882
 22.85735162493687, 120.18688777729894

 - Cigu: 3 points
23.11902593130365, 120.05874863883409
23.124310527454163, 120.07531112355083
23.130061175003874, 120.08122629666394

 - An-Nan: 2 points
 23.033784880994748, 120.07405573439489
 23.01988231566828, 120.09946161715932

 - Anping: 2 points
 22.990894265701648, 120.14268135694375
 22.969242998130003, 120.15847420298654

 - Linbian: 1 points
 22.412837508670247, 120.5141089682018

Source:
Kuo, H.-C., Wang, T.-Y., Chen, P.-P., Chen, Y.-M., Chuang, H.-C., Chen, T.-Y., 2011. Real-Time Quantitative PCR Assay for Monitoring of Nervous Necrosis Virus Infection in Grouper Aquaculture. Journal of Clinical Microbiology 49, 1090–1096. https://doi.org/10.1128/JCM.01016-10

```{r}
grouper_points <- data.frame(lat = c(23.366020730714546, 23.344101510137286, 23.11902593130365, 23.124310527454163, 23.130061175003874, 22.412837508670247, 22.5060215233288, 22.495712950055186, 22.489210224135196, 22.478900398348003, 22.963578173481825, 22.958915389973615, 22.958796842534618, 22.95887587417216, 22.962708853156307, 22.910490254668936, 22.900053951559105, 22.88771910346936, 22.87475049020418, 22.85735162493687, 23.033784880994748, 23.01988231566828, 22.990894265701648, 22.969242998130003),
                             lon = c(120.12803212666465, 120.11599409273818, 120.05874863883409, 120.07531112355083, 120.08122629666394, 120.5141089682018, 120.361518692894, 120.37199003646582, 120.38503630058813, 120.40220243759113, 120.16265849315296, 120.16450385288077, 120.16420344548321,  120.16630629726609, 120.16480426027832, 120.17469982002679, 120.17658809509712, 120.18208125893808, 120.18637279318882, 120.18688777729894, 120.07405573439489, 120.09946161715932, 120.14268135694375, 120.15847420298654))

```

Combine milkfish and grouper points

```{r}
all_points_twn <- rbind(milkfish_points, grouper_points) %>%
  dplyr::mutate(iso3c = "TWN")

farms_twn <-  st_as_sf(all_points_twn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_twn)

## write to farms folder 
write_csv(all_points_twn, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_twn.csv"))

```

## Costa Rica 
There is only one species farmed in Costa Rice, Spotted Rose Snapper.

https://www.researchgate.net/profile/Angel_Herrera-Ulloa/publication/235638071_Spotted_rose_snapper_Lutjanus_guttatus_aquaculture_research_and_development_as_socio-economic_alternative_for_Costa_Rican_fishing_communities/links/5acd4d66aca2723a333f53f8/Spotted-rose-snapper-Lutjanus-guttatus-aquaculture-research-and-development-as-socio-economic-alternative-for-Costa-Rican-fishing-communities.pdf
 - States that there are 3 pilot farms along the Gulf of Nicoya.

This says that Martec is the only company in the world which farms spotted rose snapper: https://www.seafoodsource.com/news/aquaculture/industrias-martec-plots-expansion-of-its-spotted-rose-snapper-offshore-aquaculture-operation

https://www.martec.co.cr/#aquaculture
 - Also states that they are the only company in the world to produce Spotted Rose Snapper commercially, and their farms are 3 miles off shore.... based on this I will place one point 3 miles offshore of where "Martec Industries" is located on google maps.

9.423478373794573, -84.20910189917947

Source:
Martec Industries Aquaculture , n.d. . Martec.  https://www.martec.co.cr/#aquaculture

```{r}
farms_cri <- data.frame(lat = c(9.423478373794573),
                        lon = c(-84.20910189917947)) %>%
  mutate(iso3c = "CRI")

write_csv(farms_cri, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_cri.csv"))

## visualize it 
farms_cri <-  st_as_sf(farms_cri, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_cri)
```

## Egypt

There is a map provided by the FAO (2010) here: http://www.fao.org/fishery/countrysector/naso_egypt/en
 - It has both production and locations.... I think I will need to hand place them, but we can use this. 
 - I will only use the ones that have Mullet, Seabass, Seabream, or Meagre, and that are along the coast, using ponds or cages.

 - Mersa Matruh: 31.360400495764903, 27.236078999801336: Production = 5
 - Alexandria: 31.15023941171644, 29.81109569981684: Production = 1521
 - Behira: 31.319036258284083, 30.225935332660324 : Production = 20376
 - Dakahlia: 31.491306679824632, 31.4278324391214 : Production = 5460
 - Damietta: 31.52619128473819, 31.849352783373792 : Production = 55551
 - Port Said: 31.275726202441984, 32.30190584291966 : Production = 16233 
 - Suez: 29.944394112057978, 32.54627145839493 : Production = 750
 - South Sinai: 27.80632012228518, 34.29387671355847 : Production = 30

Source:
FAO Fisheries & Aquaculture - National Aquaculture Sector Overview - Egypt, 2010. , National Aquaculture Sector Overview Fact Sheets.

```{r}
farms_egy <- data.frame(lat = c(31.360400495764903, 31.15023941171644, 31.319036258284083, 31.491306679824632, 31.52619128473819, 31.275726202441984, 29.944394112057978, 27.80632012228518),
                        lon = c(27.236078999801336, 29.81109569981684, 30.225935332660324, 31.4278324391214, 31.849352783373792, 32.30190584291966, 32.54627145839493,  34.29387671355847),
                        rgn_production = c(5, 1521, 20376, 5460, 55551, 16233, 750, 30),
                        subrgn_name = c("Mersa Matruh", "Alexandria", "Behira", "Dakahlia", "Damietta", "Port Said", "Suez", "South Sinai")) %>%
  mutate(iso3c = "EGY")

write_csv(farms_egy, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_egy.csv"))

## visualize it 
farms_egy <-  st_as_sf(farms_egy, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_egy)
```

## India

There is not much information on locations of aquaculture facilities. However, this shows "Mandapam Hatchery" 80000 tonnes of marine finfish in 2015... The total for 2017 is 96000. I think it is reasonable to place a point on Mandapam, at the CMFRI farm, and leave it at that. 
http://www.cmfri.org.in/division/mariculture

9.28696221512509, 79.13007737814245

Source:
Indian Council of Agriculture Research - Central Marine Fisheries Research Institute, 2015.  http://www.cmfri.org.in/division/mariculture

```{r}
farms_ind <- data.frame(lat = c(9.28696221512509),
                        lon = c(79.13007737814245)) %>%
  mutate(iso3c = "IND")

write_csv(farms_ind, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_ind.csv"))

## visualize it 
farms_ind <-  st_as_sf(farms_ind, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_ind)
```

## Singapore
Barramundi and milkfish

https://www.sfa.gov.sg/food-farming/food-farms/farming-in-singapore
 - This says 109 sea based fish farms.
 - Seems like a pretty reliable source.

I'll place 109 farms equally along the coast of Singapore.

Source: 
Food Farms in Singapore, 2019. . Singapore Food Agency. [Accessed December 17, 2020] https://www.sfa.gov.sg/food-farming/food-farms/farming-in-singapore

```{r}
world <- rnaturalearth::ne_coastline(scale = 10) # world map with 10m resolution
plot(world)

singapore <- ne_countries(scale = 50, country = "Singapore", returnclass = "sf")

#info of our spatial vector object
singapore
mapview(singapore)
plot(singapore)

world <- st_as_sf(world)

singapore_world_crs <- st_transform(singapore, crs(world))

singapore_world_crop <- st_crop(world, singapore_world_crs) %>%
  dplyr::filter(scalerank == 3) ## filter for the appropriate area
plot(singapore_world_crop)
mapview(singapore_world_crop)

singapore_crop <- as(singapore_world_crop, "Spatial")

singapore_points_lines <- sample.line(singapore_crop, n = 225) ## equally place the number of farms along the coastline 
class(singapore_points_lines)
plot(singapore_points_lines)

singapore_points <- st_as_sf(singapore_points_lines)
mapview(singapore_points)

singapore_points_final <- singapore_points[106:214, ] ## grab the appropriate number of points
mapview(singapore_points_final)

## now prep dataframe for lat lon points
sgp_pts <- st_centroid(singapore_points_final)

sgp_pts_latlon <- st_transform(sgp_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_sgp <- data.frame(do.call(rbind, st_geometry(sgp_pts_latlon)))
names(farms_sgp) <- c("lon", "lat")

farms_sgp_final <- farms_sgp %>%
  mutate(iso3c = "SGP")

## visualize it 
farms_sgp <-  st_as_sf(farms_sgp, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_sgp)

## write to farms folder 
write_csv(farms_sgp_final, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_sgp.csv"))
```

## Malaysia

FAO Sector overview:
"The culture of marine fish in floating net-cages in
lagoons and sheltered coastal waters occupies an area of 102 ha, and is concentrated mainly in the western
coast of Peninsular Malaysia."

 - I can only find the broad suitable areas: https://www.dof.gov.my/ 
 - Broad suitable areas https://www.dof.gov.my/index.php/pages/view/78. I'll use these and use the area as a production estimate. Saved onto aurora /home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Malaysia/malaysia_finfish_farms.csv.

Source:
Aquaculture Industrial Zone Area for High Impact Projects (HIP), n.d. . Department of Fisheries Malaysia. [Accessed November 12, 2020] https://www.dof.gov.my/index.php/pages/view/78

```{r}
malaysia_finfish_loc <- read_csv("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Malaysia/malaysia_finfish_farms.csv") %>%
  clean_names()

```

Farms: 
 - SEMPANG (PT Lot 389, 390), MELAKA : 2.131733588116292, 102.40118069215845
 - KG. AYAU, ROMPIN, PAHANG : 2.8071936098459522, 103.49758024234029
 - PERMATANG PAUH ISLAND, PAHANG : 5.384539285193906, 100.35040003558889
 - KG. SABAK FERI, ROMPIN, PAHANG : 2.8056383383493353, 103.49881667894354
 - SEMPANG (PT Lot 387), MELAKA : 2.1305137929032876, 102.40627304737939

Now lets map the farms 
```{r}
farms_mys <- data.frame(lat=c(2.131733588116292, 2.8071936098459522, 5.384539285193906, 2.8056383383493353, 2.1305137929032876),
                        lon=c(102.40118069215845, 103.49758024234029, 100.35040003558889, 103.49881667894354, 102.40627304737939)) %>%
  mutate(iso3c = "MYS")

write_csv(farms_mys, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_mys.csv"))

## visualize it 
farms_mys <-  st_as_sf(farms_mys, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_mys)
```

## Bangladesh

There isn't much information. I will place a point randomly on the coast of Bangladesh.

Source:
NA

```{r}
farms_bgd <- data.frame(lat=c(21.802433819366538),
                        lon=c(90.40418040967)) %>%
  mutate(iso3c = "BGD")

write_csv(farms_bgd, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_bgd.csv"))

## visualize it 
farms_bgd <-  st_as_sf(farms_bgd, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_bgd)
```

## Cambodia

The only species is Barramundi.

https://www.researchgate.net/publication/340828970_Asian_seabass_Lates_calcarifer_farming_practices_in_Cambodia
 - Farms are definitely located in Prey Nob, Stoeng Hav, and Thomnob Rolork. 
 - This is the only information I can find on Cambodian Barramundi fishing, so I will place 3 points in these 3 places. It is really small production anyways.

Prey Nob: 10.53427341558665, 103.57357940185497
Stoeng Hav: 10.748668380457524, 103.6089106992865
Thomnob Rolork: 10.647924480438327, 103.5006922685572

Source:
Chankakada, M., Sros, K., Ramana, O., Phearum, H., Laining, A., Rimmer, M.A., Somony, T., Viseth, H., 2020. Asian Seabass Farming Practices in Cambodia 5.

```{r}
farms_khm <- data.frame(lat=c(10.53427341558665, 10.748668380457524, 10.647924480438327),
                        lon=c(103.57357940185497, 103.6089106992865, 103.5006922685572)) %>%
  mutate(iso3c = "KHM")

write_csv(farms_khm, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_khm.csv"))

## visualize it 
farms_khm <-  st_as_sf(farms_khm, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_khm)
```

## Korea, Republic of (South Korea)

http://www.fao.org/fishery/countrysector/naso_korea/en
 - Lists 612 finfish farms. I will place 612 farms along the coast of South Korea.

Source:
National Aquaculture Sector Overview. Korea, 2005. , National Aquaculture Sector Overview Fact Sheets.

```{r}
world <- rnaturalearth::ne_coastline(scale = 110) # world map with 10m resolution
plot(world)

south_korea <- ne_countries(scale = 50, country = "South Korea", returnclass = "sf")

#info of our spatial vector object
south_korea
mapview(south_korea)
plot(south_korea)

world <- st_as_sf(world)

south_korea_world_crs <- st_transform(south_korea, crs(world))

south_korea_world_crop <- st_crop(world, south_korea_world_crs) #%>%
 # dplyr::filter(scalerank == 3) ## filter for the appropriate area

south_korea_world_crop <- south_korea_world_crop[2, ]
plot(south_korea_world_crop)
mapview(south_korea_world_crop)

south_korea_crop <- as(south_korea_world_crop, "Spatial")

south_korea_points_lines <- sample.line(south_korea_crop, n = 612) ## equally place the number of farms along the coastline 
class(south_korea_points_lines)
plot(south_korea_points_lines)

south_korea_points <- st_as_sf(south_korea_points_lines)
mapview(south_korea_points)

## now prep dataframe for lat lon points
kor_pts <- st_centroid(south_korea_points)

kor_pts_latlon <- st_transform(kor_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_kor <- data.frame(do.call(rbind, st_geometry(kor_pts_latlon)))
names(farms_kor) <- c("lon", "lat")

farms_kor_final <- farms_kor %>%
  mutate(iso3c = "KOR")

## visualize it 
farms_kor <-  st_as_sf(farms_kor, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_kor)

## write to farms folder 
write_csv(farms_kor_final, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_kor.csv"))
```

## Philippines
The only  species listed is Milkfish. 
There is no exact number listed, however I know that there are a lot of milkfish farms in the Philippines, as it is their oldest aquaculture method. There are definitely over 300. I will randomly place over 300 farms along the coast of the philippines.

Source:
NA

```{r}
world <- rnaturalearth::ne_coastline(scale = 110) # world map with 10m resolution
plot(world)

philippines <- ne_countries(scale = 50, country = "Philippines", returnclass = "sf")

#info of our spatial vector object
philippines
mapview(philippines)
plot(philippines)

world <- st_as_sf(world)

philippines_world_crs <- st_transform(philippines, crs(world))

philippines_world_crop <- st_crop(world, philippines_world_crs) #%>%
 # dplyr::filter(scalerank == 3) ## filter for the appropriate area

philippines_world_crop <- philippines_world_crop[-1, ]
plot(philippines_world_crop)
mapview(philippines_world_crop)

philippines_crop <- as(philippines_world_crop, "Spatial")

philippines_points_lines <- sample.line(philippines_crop, n = 43) ## equally place the number of farms along the coastline 
class(philippines_points_lines)
plot(philippines_points_lines)

philippines_points <- st_as_sf(philippines_points_lines)
mapview(philippines_points)

## now prep dataframe for lat lon points
phl_pts <- st_centroid(philippines_points)

phl_pts_latlon <- st_transform(phl_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_phl <- data.frame(do.call(rbind, st_geometry(phl_pts_latlon)))
names(farms_phl) <- c("lon", "lat")

farms_phl_final <- farms_phl %>%
  mutate(iso3c = "PHL")

## visualize it 
farms_phl <-  st_as_sf(farms_phl, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_phl)

## write to farms folder 
write_csv(farms_phl_final, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_phl.csv"))
```

## Hong Kong

https://www.afcd.gov.hk/english/fisheries/fish_aqu/fish_aqu_mpo/fish_aqu_mpo.html
From HK government.

This shows a map fish culture zones, fish ponds, and oyster production areas. I will place a point in all of the "Fish culture zones" shown on the map. Most of the fish cages are visible on google maps when you search these locations.

Sha Tau Kok: 22.537974350807968, 114.22465750899093
Ap Chau: 22.549021238266082, 114.27266261841545
Kat O: 22.549243173819686, 114.29054140885648
0 Pui Tong: 22.545901930448537, 114.29912254835972
Sai Lau Kong: 22.524524976629923, 114.28385013462695
Wong Wan: 22.507936457313377, 114.30641749235996
Tap Mun: 22.47796934657302, 114.35412559932351
Kau Lau Wan: 22.459641511472412, 114.360272945696
Sham Wan: 22.4435502213444, 114.35400029960742
Lo Fu Wat: 22.46853714662761, 114.2720207303674
Yung Shue Au: 22.42337637093222, 114.27819733339493
Leung Shuen Wan: 22.349528341786087, 114.35178911904846
Tiu Cham Wan: 22.367924650829952, 114.31997002657292
Tai Tau Chau: 22.368469602475702, 114.3230438982345 
Kai Lung Wan: 22.3691151827537, 114.30143307907038
Kau Sai: 22.341819373881275, 114.31744391160476
Ma Nam Wat: 22.35464650188533, 114.26990023341041
Po Toi O: 22.274992746117093, 114.29389242171153
Po Toi: 22.164051938726015, 114.25252757025574
Sok Kwu Wan: 22.20591578796148, 114.12923505969019
Lo Tik Wan: 22.221779736133797, 114.12958553008323
Ma Wan: 22.352165476146805, 114.05569941361628
Yim Tin Tsai: 22.45358281100093, 114.2079128367368
Cheung Sha Wan: 22.235400816169882, 114.00599898096631
Yim Tin Tsai (East): 22.454708931476908, 114.22148845386273
Tung Lung Chau: 22.25579039550771, 114.28554310187128

Source:
Marine fish culture, pond fish culture and oyster culture, n.d. . Agriculture, Fisheries, and Conservation Department; The Government of the Hong Kong Special Administrative Region.  https://www.afcd.gov.hk/english/fisheries/fish_aqu/fish_aqu_mpo/fish_aqu_mpo.html

```{r}
farms_hkg <- data.frame(lat_lon = c("22.537974350807968, 114.22465750899093", "22.549021238266082, 114.27266261841545", "22.549243173819686, 114.29054140885648", "22.545901930448537, 114.29912254835972", "22.524524976629923, 114.28385013462695", "22.507936457313377, 114.30641749235996", "22.47796934657302, 114.35412559932351", "22.459641511472412, 114.360272945696", "22.4435502213444, 114.35400029960742", "22.46853714662761, 114.2720207303674", "22.42337637093222, 114.27819733339493", "22.349528341786087, 114.35178911904846", "22.367924650829952, 114.31997002657292", "22.368469602475702, 114.3230438982345", "22.3691151827537, 114.30143307907038", "22.341819373881275, 114.31744391160476", "22.35464650188533, 114.26990023341041", "22.274992746117093, 114.29389242171153", "22.164051938726015, 114.25252757025574", "22.20591578796148, 114.12923505969019", "22.221779736133797, 114.12958553008323", "22.352165476146805, 114.05569941361628", "22.45358281100093, 114.2079128367368", "22.235400816169882, 114.00599898096631", "22.454708931476908, 114.22148845386273", "22.25579039550771, 114.28554310187128")) %>%
  separate(lat_lon, into = c("lat", "lon"), sep = ",") %>%
  mutate(iso3c = "HKG")

write_csv(farms_hkg, here("marine/marine_fish_general/finfish_farms/data/finfish_sites/farms_hkg.csv"))

## visualize it 
farms_hkg <-  st_as_sf(farms_hkg, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_hkg)
```

# Join cell count dataframes

```{r}
farm_files <- list.files(here::here("marine/marine_fish_general/finfish_farms/data/finfish_sites"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

write.csv(all_farms, here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon.csv"), row.names = FALSE)

```