### 2021-02-04 marine/shrimp/shrimp_farms/STEP1_fao_shrimp_production.Rmd

---
title: "FAO shrimp production"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "10/09/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

All data is ultimately adjusted to FAO production data.  This script cleans the fao mariculture data and subsets it to to the relevant years of data and species (shrimp).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(data.table)
library(countrycode)
library(here)
library(stringr)

path <- "/home/shares/food-systems/Food_footprint/_raw_data"
source(here('marine/salmon/salmon_farm/R/fao_mar_clean.R')) # functions specific to mariculture dealing with compound countries
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/fao_fxn.R') # function for cleaning FAO files
source(here('workflow/common.R'))

```

Subset the data to include only the species (shrimp) and countries we are interested in. Then summarize by total tonnes of shrimp per country/year. (There are multiple shrimp species)

```{r}

fao <- read.csv(here("data/fao_aquaculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

## Read in fao data and get a species list for salmon 
species_list <- read.csv(here("marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

# the list of species we are including
shrimp_list <- species_list$species[species_list$aq_group=="shrimps_prawns"]
shrimp_list

## write out shrimp taxa and production for Juliette
fao_shrimp_taxa_prod <- fao %>%
  filter(species %in% shrimp_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

write.csv(fao_shrimp_taxa_prod, here("marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp_taxa_production.csv"), row.names=FALSE)

# include only shrimp
fao_shrimp <- fao %>%
  filter(species %in% shrimp_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, year)

write.csv(fao_shrimp, here("marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp.csv"), row.names=FALSE)

## Do some exploration:

 explore <- read_csv(here("marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp.csv")) %>%
   filter(year==2017) %>%
   arrange(-fao_tonnes_production) %>%
   mutate(percent = round((fao_tonnes_production/sum(fao_tonnes_production)*100),2))

sum(explore$fao_tonnes_production)

asc_countries <- c("Australia", "Bangladesh", "Belize", "China", "Colombia", "Ecuador", "Guatemala", "Honduras", "India", "Indonesia", "Madagascar", "Mexico", "New Caledonia", "Nicaragua", "Panama", "Peru", "Thailand", "Venezuela, Boliv Rep of", "Viet Nam")

explore_test <- explore %>%
  dplyr::filter(!(country %in% asc_countries))

sum(explore_test$percent) 
 100 - 6.38 # 93.62

asc_explore <- explore %>%
  dplyr::filter(country %in% asc_countries)

## missing venezuela

sum(asc_explore$fao_tonnes_production)
sum(asc_explore$percent) + sum(explore_test$percent)
sum(asc_explore$fao_tonnes_production) + sum(explore_test$fao_tonnes_production)

sum(explore$fao_tonnes_production)
```