### 2021-02-10 marine/bivalve/bivalve_farms/STEP2_bivalve_farm_locations.Rmd
finished up making sure no bivalve points fall on land

---
title: "Location of Bivalve Farms"
author: "Gage Clawson (USCB, NCEAS, OHI)"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script standardizes location of bivalve farms to be lat/long point data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

fao_bivalve_spp <- read_csv("marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve_taxa_production.csv") %>%
  dplyr::filter(year == 2017)

fao_bivalve <- read_csv("marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve.csv") %>%
  dplyr::filter(year == 2017)

unique(fao_bivalve$country)
#  [1] "Australia"                "Brazil"                   "Bulgaria"                 "Cambodia"                
#  [5] "Canada"                   "Chile"                    "China"                    "China, Hong Kong SAR"    
#  [9] "Croatia"                  "Cuba"                     "Denmark"                  "France"                  
# [13] "French Polynesia"         "Germany"                  "Greece"                   "Guernsey"                
# [17] "India"                    "Indonesia"                "Ireland"                  "Italy"                   
# [21] "Japan"                    "Jersey"                   "Korea, Dem. People's Rep" "Korea, Republic of"      
# [25] "Malaysia"                 "Mexico"                   "Netherlands"              "New Zealand"             
# [29] "Norway"                   "Peru"                     "Philippines"              "Portugal"                
# [33] "Russian Federation"       "Slovenia"                 "South Africa"             "Spain"                   
# [37] "Sweden"                   "Taiwan Province of China" "Thailand"                 "United Kingdom"          
# [41] "United States of America" "Viet Nam"

coast_raw <- st_read("/home/shares/food-systems/Aquaculture_mapping/Data/Jared/Marine_Brackish/raw_data/Coastline/lines.shp")
# mapview(coast_raw)
```

## explore
https://gis.asc-aqua.org/arcgis_app/
Maps from ASC

```{r}
ASC_data <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

ASC_bivalve <- ASC_data %>%
  dplyr::filter(aq_grop == "bivalves")

unique(ASC_bivalve$crt_cnt)
#  [1] "China"                                               
#  [2] "New Zealand"                                         
#  [3] "South Africa"                                        
#  [4] "Australia"                                           
#  [5] "Korea, Republic of"                                  
#  [6] "Japan"                                               
#  [7] "Peru"                                                
#  [8] "Mexico"                                              
#  [9] "France"                                              
# [10] "Chile"                                               
# [11] "United Kingdom of Great Britain and Northern Ireland"
# [12] "Viet Nam"                                            
# [13] "Portugal"

## what is in ASC and not in FAO
setdiff(ASC_bivalve$crt_cnt, fao_bivalve$country)
#[1] "United Kingdom of Great Britain and Northern Ireland"

## what is in FAO and not ASC -- this is what we need to find alternative data sources for (29 countries)
setdiff(fao_bivalve$country, ASC_bivalve$crt_cnt)
#  [1] "Brazil"                   "Bulgaria"                 "Cambodia"                 "Canada"                  
#  [5] "China, Hong Kong SAR"     "Croatia"                  "Cuba"                     "Denmark"                 
#  [9] "French Polynesia"         "Germany"                  "Greece"                   "Guernsey"                
# [13] "India"                    "Indonesia"                "Ireland"                  "Italy"                   
# [17] "Jersey"                   "Korea, Dem. People's Rep" "Malaysia"                 "Netherlands"             
# [21] "Norway"                   "Philippines"              "Russian Federation"       "Slovenia"                
# [25] "Spain"                    "Sweden"                   "Taiwan Province of China" "Thailand"                
# "United States of America"

mapview(ASC_bivalve)
```

# Raster

Standard raster used for this analysis

```{r}

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

For each point, we will add a column for "data quality". It will be scored as very good (1.0), good (0.667), fair (0.333) or poor (0.0). 
 - Very Good qualifications: Locations taken from leases, or from a map with specific locations (i.e. EMOD, ASC, HELCOM etc.).
 - Good qualifications: Locations are projected or it is a only sample of total locations (NASO maps). Google maps estimates. 
 - Fair qualifications: A specific number of farms are placed on the coast, without knowing exact locations. Or Per province/state locations (i.e. placing a point randomly in each province) or sub regional location. 
 - Bad qualifications: No data at all (randomly placed point(s) along the coast, a non-specific number).

# Map data from ASC

Data comes from https://gis.asc-aqua.org/arcgis_app/

Source: 
Aquaculture Stewardship Council Data Mapping Study, n.d. . Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}

# get iso3c codes
iso <-   data.frame(iso3c = countrycode(ASC_bivalve$crt_cnt, origin = 'country.name', destination = 'iso3c'))

# read in shp
farms_all <- ASC_bivalve %>%
  cbind(iso)

mapview(farms_all$geometry)

unique(farms_all$iso3c)

all_pts <- st_centroid(farms_all)

all_pts_latlon <- st_transform(all_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry, iso3c)

farms_all_fin <- data.frame(do.call(rbind, st_geometry(all_pts_latlon))) 
names(farms_all_fin) <- c("lon", "lat")
farms_all_fin <- farms_all_fin %>%
  cbind(iso3c = all_pts_latlon$iso3c) %>%
  mutate(data_quality = 1.0)

## save farm location
write.csv(farms_all_fin, here("marine/bivalve/bivalve_farms/data/bivalve_sites/ASC_bivalve_farm_lat_lon.csv"), row.names = FALSE)

```

Now we will map the remaining countries with bivalve farms that are not included in the ASC dataset. Workflow will be as follows: 
 - Try to obtain exact locations of farms from published source or dataset (such as FAO or a research paper)
 - If there are no exact locations, but the source lists a city or state, place a point in that particular city or state
 -  If there is enough information to know there are a specific number of bivalve farms, equally allocate the farms across the coastline. If there are estimates of farms per state/province, equally allocate them across the state/province, to be even more precise. 
 - If there are not exact numbers of farms in the country, but you know there are a lot, randomly place points along the coast line (a non-exact number, like we did for Taiwan). 
 - Lastly, if there is no evidence of marine bivalve farm locations, place a point on the coast of the country (preferably with some intuition/knowledge of where they might be, i.e. what we've done for Russia).

## USA 
This has a lot of maps, but almost all don't allow downloads: https://coastalscience.noaa.gov/research/marine-spatial-ecology/coastal-aquaculture-planning-portal-capp/?tx_category=shellfish-algae-planning-and-siting

## US Northeast - NY, CT, ME, NH
 - Regional Ocean Council 2018

Source: 
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 2020] https://www.northeastoceandata.org/data-download/

```{r}

us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))

unique(us_ne$allSpecies)

unique(fao_bivalve_spp$species)

## need to filter allSpecies for anything with "Oyster", "Mussel", "Shellfish", "Clam", "Quahog",

# id bivalves and active sites
bivalve_species <- grep("Oyster|Mussel|Shellfish|Clam|Quahog", unique(us_ne$allSpecies), value=TRUE)

us_ne_bivalve <- us_ne %>%
  filter(allSpecies %in% bivalve_species) %>%
  filter(status == "Active")
unique(us_ne_bivalve$state)

us_ne_bivalve_mol <- st_transform(us_ne_bivalve, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
us_ne_bivalve_pts <- st_centroid(us_ne_bivalve_mol)
# check
plot(st_geometry(us_ne_bivalve_mol))
plot(st_geometry(st_centroid(us_ne_bivalve_mol)), add=TRUE)

us_ne_bivalve_pts_latlon <- st_transform(us_ne_bivalve_pts, crs = "+init=epsg:4326")

farms_us_ne <- data.frame(do.call(rbind, st_geometry(us_ne_bivalve_pts_latlon))) # get lat lon
names(farms_us_ne) <- c("lon", "lat") 
farms_us_ne <- farms_us_ne %>%
  mutate(iso3c = "USA") %>%
  mutate(data_quality = 1.0)

# all_farms_sf <- st_as_sf(farms_us_ne, coords = c("lon", "lat"), 
#                          crs = crs(gadm_latlon), agr = "constant")
# 
# mapview(all_farms_sf)

write_csv(farms_us_ne, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_ne_usa.csv"))

```

## USA - WA, RI 
 - Aquaculture, Office for Coastal Management (OCM), 2018

Source: 
Aquaculture, Office for Coastal Management (OCM), 2018. [Accessed January 17, 2020] https://www.fisheries.noaa.gov/inport/item/53129/

```{r}
us_aqua <- st_read(file.path(path,"USA/US/us_aqua/us_aqua.shp"))

mapview(us_aqua$geometry)

unique(us_aqua$state)

## filter for species !=NA, any that are not bivalves, those that are active or conditionally active (seasonal), and filter out those we already have (NY, ME, CT, NH)

us_aqua_species <- us_aqua %>%
   dplyr::filter(!is.na(species)) %>%
  dplyr::filter(!(species %in% c("finfish", "Atlantic salmon", "Atlantic cod", "Seaweed", "Shrimp", "Coral", "clam seed, oyster seed", "Gelidium spp., Macrocystis pyrifera, M. angustifolia, M. intergrifolia, Pelagophycus spp. Euchema uncinatum (male plants only)", "oyster seed", "In progress", "Not Available", "Hatchery", "softshell", "softshell, butter", "softshell, varnish", "butter"))) %>%
  dplyr::filter(state != "Federal waters") %>%
  dplyr::filter(!(status %in% c(NA, "Pending", "Restricted-Relay", "Not Available", "Conditional Closure", "Prohibited"))) %>%
dplyr::filter(!(state %in% c("Maine", "Connecticut", "New York"))) ## filter out the states we already have

unique(us_aqua_species$state)

unique(us_aqua_species$species)

unique(us_aqua_species$status)

mapview(us_aqua_species$geometry)

us_aqua_bivalve_mol <- st_transform(us_aqua_species, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
us_aqua_bivalve_pts <- st_centroid(us_aqua_bivalve_mol)
# check
plot(st_geometry(us_aqua_bivalve_mol))
plot(st_geometry(st_centroid(us_aqua_bivalve_mol)), add=TRUE)

us_aqua_bivalve_pts_latlon <- st_transform(us_aqua_bivalve_pts, crs = "+init=epsg:4326")

farms_us_aqua <- data.frame(do.call(rbind, st_geometry(us_aqua_bivalve_pts_latlon))) 
names(farms_us_aqua) <- c("lon", "lat")
farms_us_aqua <- farms_us_aqua %>%
  mutate(iso3c = "USA") %>%
  mutate(data_quality = 1.0)

all_farms_sf <- st_as_sf(farms_us_aqua, coords = c("lon", "lat"),
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

write_csv(farms_us_ne, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_wa_ri_usa.csv"))

```

## HELCOM shellfish mariculture points - Sweden, Germany

Source: 
HELCOM Map and Data Service, n.d. . The Baltic Marine Environment Protection Commission. [Accessed October 12, 2020]

```{r}

## Shellfish points
enc <- readr::guess_encoding(file.path(path, "HELCOM_Baltic/Shellfish_mariculture_points/Shellfish_mariculture_points.shp"), n_max = 1000)[,"encoding"] %>% as.character()

# Read in shapefile
farms_baltic <- st_read(file.path(path, "HELCOM_Baltic/Shellfish_mariculture_points/Shellfish_mariculture_points.shp"), options = "ENCODING=WINDOWS-1252")

mapview(farms_baltic$geometry)

baltic_pts <- st_centroid(farms_baltic)

bal_pts_latlon <- st_transform(baltic_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry, COUNTRY, NAME)

bal_subregions <- bal_pts_latlon %>%
  dplyr::select(COUNTRY, NAME)

farms_bal <- data.frame(do.call(rbind, st_geometry(bal_pts_latlon)))
names(farms_bal) <- c("lon", "lat")

farms_bal <- cbind(farms_bal, bal_subregions) %>%
  mutate(iso3c = ifelse(COUNTRY == "DE", "DEU", "SWE")) %>%
  dplyr::select(lon, lat, iso3c, subregion = NAME) %>%
  mutate(data_quality = 1.0)

write.csv(farms_bal, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_swe_deu.csv"), row.names = FALSE)

```

## HELCOM shellfish mariculture areas - Denmark

Source:
HELCOM Map and Data Service, n.d. . The Baltic Marine Environment Protection Commission. [Accessed October 12, 2020]

```{r}
## SHELLFISH polygons
# shape file
farms_baltic_area <- st_read(file.path(path, "HELCOM_Baltic/Shellfish_mariculture_areas/Shellfish_mariculture_areas.shp"), options = "ENCODING=WINDOWS-1252")

mapview(farms_baltic_area$geometry)

baltic_pts_area <- st_centroid(farms_baltic_area)

bal_pts_latlon_area <- st_transform(baltic_pts_area, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_bal_area <- data.frame(do.call(rbind, st_geometry(bal_pts_latlon_area)))
names(farms_bal_area) <- c("lon", "lat")

farms_bal_area <- farms_bal_area %>%
  mutate(iso3c = "DNK") %>%
  dplyr::select(lon, lat, iso3c) %>%
  mutate(data_quality = 1.0)

all_farms_sf <- st_as_sf(farms_bal_area, coords = c("lon", "lat"),
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

write.csv(farms_bal_area, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_dnk.csv"), row.names = FALSE)
```

## British colombia
Shellfish hatcheries: 
https://catalogue.data.gov.bc.ca/dataset/shellfish-hatcheries-coastal-resource-information-management-system-crims/resource/f161583a-1d73-4c92-bf8d-ea67727a89d3

Source: 
Shellfish Hatcheries - Coastal Resource Information Management System (CRIMS) , 2019. Ministry of Forests, Lands, Natural Resource Operations and Rural Development. [Accessed November 30, 2020] https://catalogue.data.gov.bc.ca/dataset/shellfish-hatcheries-coastal-resource-information-management-system-crims

```{r}

## read shapefile
bc <- st_read(file.path(path, "Canada/CRIMS_SHELLFISH_HATCHERIES/SHLFISH_HT_point.shp"))

mapview(bc)
bc_pts <- st_centroid(bc)

bc_pts_latlon <- st_transform(bc_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_bc <- data.frame(do.call(rbind, st_geometry(bc_pts_latlon)))
names(farms_bc) <- c("lon", "lat")
farms_bc <- farms_bc %>%
  mutate(iso3c="CAN") %>%
  mutate(data_quality = 1.0)

write_csv(farms_bc, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_bc_can.csv"))

```

## Nova Scotia
Obtained data from here: https://data.novascotia.ca/Fishing-and-Aquaculture/Aquaculture-License-and-Lease-GIS-Database/h57h-p9mm

Source: 
Nova Scotia Marine Aquaculture Leases , 2020. . Nova Scotia Fisheries and Aquaculture. [Accessed February 18, 2020] https://data.novascotia.ca/Fishing-and-Aquaculture/Nova-Scotia-Marine-Aquaculture-Leases/h57h-p9mm

```{r}
## read shapefile
ns <- st_read(file.path(path,"Canada/Nova_Scotia_Marine_Aquaculture_Sites/geo_export_fcc6f2ef-82bb-4445-98f0-a82a56cd0d00.shp"))
mapview(ns)

unique(ns$aqua_speci)

# id bivalves: filter out anything that is not a bivalve
not_bivalve <- c("Atlantic Salmon; Arctic Char; Rainbow Trout", "Rainbow Trout; Atlantic Salmon; Arctic Char; Brook Trout", "Rainbow Trout; Arctic Char; Brook Trout; Atlantic Salmon","Rainbow Trout", "Atlantic Salmon; Rainbow Trout; Atlantic Halibut; Atlantic Cod; Haddock", "Atlantic Salmon", "Atlantic Salmon; Rainbow Trout", "Rainbow Trout; Atlantic Salmon"  )

ns_bivalve <- ns %>%
  filter(!(aqua_speci %in% not_bivalve))

ns_bivalve_mol <- st_transform(ns_bivalve, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
ns_pts <- st_centroid(ns_bivalve_mol)

ns_pts_latlon <- st_transform(ns_pts, crs = "+init=epsg:4326")

mapview(ns_pts_latlon)

farms_ns <- data.frame(do.call(rbind, st_geometry(ns_pts_latlon))) 
names(farms_ns) <- c("lon", "lat")
farms_ns <- farms_ns %>%
  mutate(iso3c="CAN") %>%
  mutate(data_quality = 1.0)

write_csv(farms_ns, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_ns_can.csv"))

```

##  Europe - EMODNET: Ireland, Bulgaria, Italy, Norway, Netherlands, Greece
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Shellfish+Production
https://www.emodnet-humanactivities.eu/view-data.php

Source: 
EMODnet Human Activities: Shellfish aquaculture, 2019. [Accessed November 30, 2020] https://www.emodnet-humanactivities.eu/search-results.php?dataname=Shellfish+Production

```{r}

## read in shapefile
eu <- st_read(file.path(path,"EMODnet_HA_Aquaculture_Shellfish_20191126/EMODnet_HA_Aquaculture_Shellfish_20191126.shp"))
mapview(eu)

unique(eu$COUNTRY)
# [1] "France"         "Ireland"        "Bulgaria"       "United Kingdom" "Italy"          "Norway"        
# [7] "Denmark"        "Netherlands"    "Greece"

unique(eu$FARM_TYPE)

eu_na <- eu %>%
  dplyr::filter(STATUS == "n.a.")
unique(eu_na$COUNTRY)

eu_active <- eu %>%
  dplyr::filter(STATUS == "Active")

unique(eu_active$COUNTRY)

## we will use both "Active" and "n.a." as the status of the farms. I choose this because all of the denmark points are n.a., and they correspond with the HELCOM data that we used. This indicates to me that the farms denoted as "n.a." are probably actually active.

eu_active_na <- eu %>%
  dplyr::filter(STATUS %in% c("Active", "n.a."))

unique(eu_active_na$COUNTRY)

# [1] "France"         "Ireland"        "Bulgaria"       "United Kingdom" "Italy"          "Norway"        
# [7] "Denmark"        "Netherlands"    "Greece"

## Now filter for countries that we want: Ireland, Bulgaria, Italy, Norway, Netherlands, and Greece. UK and DNK are already accounted for in our other datasets. 
eu_fao_countries <- eu_active_na %>%
  dplyr::filter(!(COUNTRY %in% c("France", "United Kingdom", "Denmark")))

unique(eu_fao_countries$COUNTRY)
mapview(eu_fao_countries)

eu_pts <- st_centroid(eu_fao_countries)

eu_pts_latlon <- st_transform(eu_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(COUNTRY, geometry)

eu_countries <- eu_pts_latlon %>%
  dplyr::select(COUNTRY)

farms_eu <- data.frame(do.call(rbind, st_geometry(eu_pts_latlon)))
names(farms_eu) <- c("lon", "lat")

farms_eu_final <- cbind(farms_eu, eu_countries) %>%
  dplyr::select(-geometry) %>%
  mutate(iso3c = case_when(
    COUNTRY == "Ireland" ~ "IRL",
    COUNTRY == "Bulgaria" ~ "BGR", 
    COUNTRY == "Norway" ~ "NOR", 
    COUNTRY == "Netherlands" ~ "NLD",
    COUNTRY == "Italy" ~ "ITA", 
    COUNTRY == "Greece" ~ "GRC"
  )) %>%
  dplyr::select(-COUNTRY) %>%
  mutate(data_quality = 1.0)

 write_csv(farms_eu_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_IRL_BGR_NOR_NLD_ITA_GRC.csv"))

```

## Spain
TODO: Figure out how to download exact locations

https://servicio.pesca.mapama.es/acuivisor/
 - This is a map of all aquaculture facilities in Spain. It will not allow for download of the locations of the facilities. However, I have downloaded all of the names of each shellfish installation, with the company name, and what general location they are located in. I will filter for unique cities, and place points in each place.

How to download: 
 - Make sure you only have the basemap and the "Instalaciones" layers selected.
 - Click on the filter tab and filter for "Marina" aquaculture, and for these shellfish species: 
 Abalón japonés, Almeja babosa, Almeja fina, Almeja japonesa, Almejas nep (Ruditapes spp), Almejas nep (Veneridae), Berberecho, Berberecho verde, Bivalvos, Bígaro, Coquina, Escupiña grabada, Lapa, Navaja recta, Navaja, Oreja de mar, Ostión u ostra japonesa, Ostra u ostra plana, Ostras u ostras planas, Vieira, Volandeira, Zamburiña
 - Download the data and save as a csv.

General location to place points:
 - Alicante: 38.345509445679276, -0.48793746088812184
 - Almeria: 36.83403538312481, -2.46312530015449
 - Asturias: 43.6586646127094, -5.838989914402157
 - Baleares: 39.578993240236755, 2.6312097506887446
 - Cadiz: 36.55518316813145, -6.2759664948571094
 - Cantabria: 43.488143305644364, -3.808940707432179
 - Castellon: 39.98227981322492, 0.025845987383098443
 - Gerona: 41.893397069777485, 3.1931418366760647
 - Granada: 36.69547948740147, -3.403621978170133
 - La Coruna: 43.36941893593161, -8.412121809979775
 - Las Palmas: 28.17841973419213, -15.411897095500123
 - Lugo: Inland, Don't include
 - Malaga: 36.705816371679525, -4.422146790963811
 - Murcia: Inland, don't include
 - Pontevedra: 42.42511968211785, -8.658739084602496
 - Sevilla: Inland, don't include
 - Santa Cruz de Tenerife: 28.464013540472575, -16.25186419732133
 - Tarragona: 41.109958666992036, 1.2538616546744752
 - Valencia: 39.47564987283762, -0.3208672158378773
 - Vizcaya: 43.436972390962005, -2.9200783381913014

Source: 
Acuivisor, 2019. . Ministerio de Agricultura, Pesca Y Alimentacion, Gobierno de Espana. [Accessed December 2, 2020] https://servicio.pesca.mapama.es/acuivisor/

```{r}
## Read in dataset of company names:
spain_shellfish <- read_csv(here("marine/bivalve/bivalve_farms/data/spain_shellfish_aquaculture.csv"))

head(unique(spain_shellfish$Empresa))

communitiy <- data.frame(companies = unique(spain_shellfish$Provincia))

## there are 21 different cities where shellfish is farmed.

farms_esp <- data.frame(lat = c(38.345509445679276, 36.83403538312481, 43.6586646127094, 39.578993240236755, 36.55518316813145, 43.488143305644364, 39.98227981322492, 41.893397069777485, 36.69547948740147, 43.36941893593161, 28.17841973419213, 36.705816371679525, 42.42511968211785, 28.464013540472575, 41.109958666992036, 39.47564987283762, 43.436972390962005),
                        lon = c(-0.48793746088812184, -2.46312530015449, -5.838989914402157, 2.6312097506887446, -6.2759664948571094, -3.808940707432179, 0.025845987383098443, 3.1931418366760647,-3.403621978170133,  -8.412121809979775, -15.411897095500123, -4.422146790963811, -8.658739084602496, -16.25186419732133, 1.2538616546744752, -0.3208672158378773, -2.9200783381913014)) %>%
  mutate(iso3c = "ESP") %>%
  mutate(data_quality = 0.333)

write_csv(farms_esp, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_esp.csv"))

## visualize it 
farms_esp <-  st_as_sf(farms_esp, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_esp)

```

## Jersey 
 - Government of Jersey: https://www.gov.je/SiteCollectionDocuments/Industry%20and%20finance/ID%20Mollusc%20farms%20registered%20businesses%20page%201.pdf
  - https://www.gov.je/SiteCollectionDocuments/Industry%20and%20finance/ID%20Molluscs%20farms%20registered%20businesses%20page%202%2020181030%20DM.pdf

 - 5 licensed farms (only including "mollusc farm open")
 - Jersey Oyster Company Limited: 49.17300525845569, -2.0325974703629277
 - Seymour Oyster Company Ltd: 49.22644208989364, -2.0327040476242284
 - La Rocque Fisheries Limited: 49.17293132589725, -2.0312731919583795
 - Jersey Sea Farms: 49.231523271584344, -2.057638560247779
 - Island Dived Scallops: 49.20231914535721, -2.03533532472839

Source: 
Fish and shellfish farming, 2020. . Government of Jersey: Information and public services for the Island of Jersey. [Accessed December 4, 2020]  https://www.gov.je/Industry/FarmingFishing/Fishing/Pages/FishFarming.aspx

```{r, evalute = F}

farms_jey <- data.frame(lat = c(49.17300525845569, 49.22644208989364, 49.17293132589725, 49.231523271584344, 49.20231914535721),
                        lon = c(-2.0325974703629277, -2.0327040476242284, -2.0312731919583795, -2.057638560247779, -2.03533532472839)) %>%
  mutate(iso3c = "JEY") %>%
  mutate(data_quality = 1.0)

write_csv(farms_jey, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_jey.csv"))

## visualize it 
farms_jey <-  st_as_sf(farms_jey, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_jey)

```

## Guernsey 
 - https://www.gov.gg/fishing#:~:text=Sea%20Fisheries%20is%20responsible%20for,one%20licensed%20site%20on%20Herm.
 - Sea Fisheries is responsible for the licensing of all shellfish farms within the Bailiwick. There are a total of seven licensed aquaculture sites distributed around Guernsey's west and north coasts, as well as one licensed site on Herm. There is also an oyster farm located in a disused quarry in the north of the island.
 - Allocate 7 farms equally across coastline

Source: 
Fishing and Aquaculture, 2020. . Government of the States of Guernsey. [Accessed December 3, 2020] https://www.gov.gg/fishing#:~:text=Sea%20Fisheries%20is%20responsible%20for,one%20licensed%20site%20on%20Herm

```{r}
## read in country shapefile
Guernsey <- ne_countries(scale = 10, country = "Guernsey", returnclass = "sf")

#info of our spatial vector object
mapview(Guernsey)
plot(Guernsey)

## transform to same crs at the coast file so we can crop
Guernsey_world_crs <- st_transform(Guernsey, crs(coast_raw))

# crop the to the coast file, and select the main island 
Guernsey_world_crop <- st_crop(coast_raw, Guernsey_world_crs) %>%
  dplyr::filter(FID %in% c(220920, 220921, 220922)) %>% ## filter for the appropriate island
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>% ## merge all the geometries into one
  ungroup()

plot(Guernsey_world_crop)
mapview(Guernsey_world_crop)
class(Guernsey_world_crop)
crs(Guernsey_world_crop)

g_meters <- st_transform(Guernsey_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(g_meters)

buff <- st_buffer(g_meters, dist = 1000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)
plot(buff)

linestring <- st_cast(buff, "LINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

buffer_outline <- linestring[1,]  ## grab the outline row 
mapview(buffer_outline)

## now reproject back to original 
buffer_outline_fixed <- st_transform(buffer_outline, crs = crs(Guernsey_world_crop))

## make a spatiallinesdataframe so we can use sample.line
Guernsey_crop <- as(buffer_outline_fixed, "Spatial")
class(Guernsey_crop)
mapview(Guernsey_crop)

Guernsey_points_lines <- sample.line(Guernsey_crop, n = 7) ## equally place the number of farms along the coastline 
class(Guernsey_points_lines)
plot(Guernsey_points_lines)

Guernsey_points <- st_as_sf(Guernsey_points_lines)
mapview(Guernsey_points, color = "red") + mapview(Guernsey_world_crop)

st_distance(Guernsey_points, Guernsey_world_crop)
# Units: [m]
#          [,1]
# [1,] 992.6726
# [2,] 992.6273
# [3,] 992.6395
# [4,] 992.3816
# [5,] 992.3755
# [6,] 992.7368
# [7,] 992.6771
## close enough to 1 km

## now prep dataframe for lat lon points
ggy_pts <- st_centroid(Guernsey_points)

ggy_pts_latlon <- st_transform(ggy_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_ggy <- data.frame(do.call(rbind, st_geometry(ggy_pts_latlon)))
names(farms_ggy) <- c("lon", "lat")

farms_ggy_final <- farms_ggy %>%
  mutate(iso3c = "GGY") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_ggy <-  st_as_sf(farms_ggy, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_ggy)

## write to farms folder 
write_csv(farms_ggy_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_ggy.csv"))
```

## Slovenia
 - http://www.fao.org/fishery/countrysector/naso_slovenia/en#:~:text=From%20the%20345%20aquaculture%20facilities,facilities%20are%20for%20shellfish%20farming.
 - From the 345 aquaculture facilities in Slovenia, there are 26 marine aquaculture facilities, 1 of which is for fish farming, while the remaining **20 facilities are for shellfish farming**.

Source: 
National Aquaculture Sector Overview. Slovenia, 2005. , National Aquaculture Sector Overview Fact Sheets. Food and Agriculture Organization of the United Nations.

```{r}
# world <- rnaturalearth::ne_coastline(scale = 10) #world map with 10m resolution
# plot(world)

slovenia <- ne_countries(scale = 10, country = "Slovenia", returnclass = "sf")

#info of our spatial vector object
mapview(slovenia)
plot(slovenia)

world <- st_as_sf(world)

slovenia_world_crs <- st_transform(slovenia, crs(coast_raw))

slovenia_world_crop <- st_crop(coast_raw, slovenia_world_crs) %>% ## crop to slovenia extent
  filter(FID %in% c(46261, 46262)) %>%
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>% ## merge all the geometries into one
  ungroup()

plot(slovenia_world_crop)
mapview(slovenia_world_crop)

s_meters <- st_transform(slovenia_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(s_meters)

buff <- st_buffer(s_meters, dist = 1000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)
plot(buff)

linestring <- st_cast(buff, "LINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(slovenia_world_crop))

linestring_fix <- st_difference(buffer_outline_fixed, slovenia) ## now grab only the outline 
mapview(linestring_fix)

## make a spatiallinesdataframe so we can use sample.line
slovenia_crop <- as(linestring_fix, "Spatial")
class(slovenia_crop)
mapview(slovenia_crop)

slovenia_points_lines <- sample.line(slovenia_crop, n = 41) ## equally place the number of farms along the coastline 
class(slovenia_points_lines)
plot(slovenia_points_lines)

slovenia_points <- st_as_sf(slovenia_points_lines) %>%
  mutate(new_id = 1:41)
mapview(slovenia_points)

## Now filter out the points that fall outside of Slovenia
slovenia_points_final <- slovenia_points[2:21,]
mapview(slovenia_points_final)

mean(st_distance(slovenia_points_final, slovenia_world_crop))
# 998.6394 [m]
## close enough to 1 km

## prep to get lat lon 
svn_pts <- st_centroid(slovenia_points_final)

svn_pts_latlon <- st_transform(svn_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_svn <- data.frame(do.call(rbind, st_geometry(svn_pts_latlon)))
names(farms_svn) <- c("lon", "lat")

farms_svn_final <- farms_svn %>%
  mutate(iso3c = "SVN") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_svn <-  st_as_sf(farms_svn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_svn)

## write to farms folder 
write_csv(farms_svn_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_svn.csv"))
```

## Croatia
 - This states there are 266 shellfish farm locations in Croatia: https://ec.europa.eu/fisheries/sites/fisheries/files/docs/pages/spatial-planning-for-aquaculture-hr.pdf
 - However, I can't find much else information. Due to this, I will allocate the 266 farms equally along the coast of Croatia.

Source: 
Lukin, A., 2014. Spatial Planning For Marine Aquaculture, Croatia, Zadar County Case. Croatian Ministry of Agriculture, Directorate of Fisheries.

```{r}
world <- ne_coastline(scale = 110) #world map with 110m resolution
plot(world)

croatia <- ne_countries(scale = 10, country = "Croatia", returnclass = "sf")

#info of our spatial vector object
mapview(croatia)
plot(croatia)

world <- st_as_sf(world)

croatia_world_crs <- st_transform(croatia, crs(world))

croatia_world_crop <- st_crop(world, croatia_world_crs)

croatia_world_crop <- st_intersection(croatia_world_crop, croatia)
mapview(croatia_world_crop)
# %>% ## crop to croatia extent
#   filter(!(FID %in% c(46278:46281, 643461:643463, 156618, 46260:46269, 46178:46182, 46177, 206670:206671))) %>%
#   mutate(new_id = 1) %>%
#   group_by(new_id) %>%
#   summarise(geometry = st_union(geometry)) %>% ## merge all the geometries into one
#   ungroup()
# mapview(croatia_world_crop)

c_meters <- st_transform(croatia_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(c_meters)

buff <- st_buffer(c_meters, dist = 40000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)
plot(buff)

linestring <- st_cast(buff, "LINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(croatia_world_crop))

linestring_fix <- st_difference(buffer_outline_fixed, croatia) ## now grab only the outline 
mapview(linestring_fix)

croatia_crop <- as(linestring_fix, "Spatial")
plot(croatia_crop)
mapview(croatia_crop)

croatia_points_lines <- sample.line(croatia_crop, n = 640) ## equally place the number of farms along the coastline 
class(croatia_points_lines)
plot(croatia_points_lines)

croatia_points <- st_as_sf(croatia_points_lines) %>%
  mutate(new_id = 1:640)
mapview(croatia_points)

## Now filter out the points that fall outside of croatia: rows 57-70
croatia_points_final <- croatia_points %>%
  dplyr::filter(new_id %in% c(2:24, 395:640)) %>%
  dplyr::filter(!(new_id %in% c(419, 533, 561)))
mapview(croatia_points_final)

mean(st_distance(croatia_points_final, croatia_world_crop))

## get lat lon
hrv_pts <- st_centroid(croatia_points_final)

hrv_pts_latlon <- st_transform(hrv_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_hrv <- data.frame(do.call(rbind, st_geometry(hrv_pts_latlon)))
names(farms_hrv) <- c("lon", "lat")

farms_hrv_final <- farms_hrv %>%
  mutate(iso3c = "HRV") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_hrv <-  st_as_sf(farms_hrv, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_hrv)

## write to farms folder 
write_csv(farms_hrv_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_hrv.csv"))
```

## Cuba
 - Four sites taken from this paper: https://scielo.conicyt.cl/scielo.php?script=sci_arttext&pid=S0718-560X2017000100024

Cayo Esquivel del Sur: 23.034188, -80.082047
Cayo de la Cruz: 22.994283, -79.994291
Cayo Hachuela: 22.952677, -79.963228
Cayos de la Enfermeria: 22.910246, -79.958924

Source: 
Mazon Suastegui, J.M., Rivero Suarez, S.E., Betanzos Vega, A., Saucedo, P.E., Rodriguez Jaramillo, C., Acosta Salmon, H., 2017. Potential of sites in northern Cuba for developing an industry of the native mangrove oyster (Crassostrea rhizophorae). lajar 45, 218–222. https://doi.org/10.3856/vol45-issue1-fulltext-24

```{r}

farms_cub <- data.frame(lat = c(23.0341889, 22.994283, 22.952677, 22.910246),
                        lon = c(-80.082047, -79.994291, -79.963228, -79.958924)) %>%
  mutate(iso3c = "CUB") %>%
  mutate(data_quality = 0.667)

write_csv(farms_cub, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_cub.csv"))

## visualize it 
farms_cub <-  st_as_sf(farms_cub, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_cub)
```

## Taiwan
- I can't find any information on the locations of shellfish farms in Taiwan, however, they have a lot of production. I will randomly place a point in the Qigu lagoon (see article). 
 - https://www.newsecuritybeat.org/2019/10/taiwan-beach-snow/

Source:
NA

```{r}

farms_twn <- data.frame(lat = c(23.12185817551473),
                        lon = c(120.05482898155049)) %>%
  mutate(iso3c = "TWN") %>%
  mutate(data_quality = 0.0)

write_csv(farms_twn, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_twn.csv"))

## visualize it 
farms_twn <-  st_as_sf(farms_twn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_twn)

```

OLD WAY: 
```{r}

# gadm <- st_read("/home/shares/food-systems/Aquaculture_mapping/Data/Jared/Marine_Brackish/manipulated_data/GADM36/gadm36molcountries_simp.shp")
# plot(gadm)
# 
#  TWN_shp <- gadm %>%
#   filter(GID_0 == "TWN") %>%
#   st_transform(., food_crs) %>%
#   mutate(Region = 1)
#  mapview(TWN_shp)
# 
# st_write(TWN_shp, "/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/taiwan/taiwan.shp")
# 
# twn_reg_ras <- fasterize::fasterize(TWN_shp, fake_raster, fun = "sum")
# 
# 
# plot(fake_raster)
# plot(twn_reg_ras)
# 
# writeRaster(twn_reg_ras, "/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/taiwan/taiwan.tif")
# 
# 
# twn_reg_ras <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/taiwan/taiwan.tif")
# 
# twn_spatial <- as(TWN_shp, "Spatial")
# 
# twn_coast <- rasterize(twn_spatial, twn_reg_ras, getCover = TRUE)
# 
# plot(twn_coast)
# cellStats(twn_coast, "sum") #464.63
# 
# # filter to include >0 <1, o.w. NA
# 
# twn_coast_df <- as.data.frame(twn_coast, xy=TRUE) %>%
#   dplyr::filter(layer <1 & layer >0) %>% ## filter out anything not on taiwan coast
#   mutate(iso3c = "TWN") %>%
#   dplyr::select(lon=x, lat=y, iso3c)
# 
# farms_twn <-  st_as_sf(twn_coast_df, coords = c("lon", "lat"), 
#                  crs = crs(gadm_latlon), agr = "constant")
# 
# mapview(farms_twn)
# 
# 
# write_csv(twn_coast_df, here::here("marine/bivalve/bivalve_farms/data/farm_sites/farms_twn.csv"))
# 
# 
# twn <- read_csv(here::here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_twn.csv"))
# farms_twn <-  st_as_sf(twn, coords = c("lon", "lat"), 
#                  crs = crs(gadm_latlon), agr = "constant")
# 
# mapview(farms_twn)
```

## French Polynesia
 - http://www.fao.org/3/t5816e/t5816e04.htm

 - http://www.ressources-marines.gov.pf/aquaculteurs/laquaculture-en-polynesie-francaise/les-filieres-de-laquaculture-en-polynesie-francaise/
 - Aquaculture is: for molluscs: the giant clam (Tridacna maxima).
 - To date (2019), two lagoons are authorized for the collection of giant clam spat, the lagoon of Reao and that of Tatakoto.

Backed up by this: https://aquaculture.spc.int/index.php?option=com_countries&view=country&id=6
Giant clams (Tridacna mamoibroodies_bxima) spat collection and stock enhancement techniques are mastered in Eastern Tuamotu atolls. Small scale collectors should be born in 2007.

I will place the two points at these two lagoons (the production is really small, 1200 tonnes):
 - Tatakoto: -17.349601342130857, -138.40992718639146
 - Reao: -18.50754847683744, -136.38467411084235

Source: 
The aquaculture sectors in French Polynesia , n.d. . Department of Marine Resources Pu Fa’Ahotu Moana. http://www.ressources-marines.gov.pf/aquaculteurs/laquaculture-en-polynesie-francaise/les-filieres-de-laquaculture-en-polynesie-francaise/

```{r}
farms_pyf <- data.frame(lat = c(-17.349601342130857, -18.50754847683744),
                        lon = c(-138.40992718639146, -136.38467411084235)) %>%
  mutate(iso3c = "PYF") %>%
  mutate(data_quality = 1.0)

write_csv(farms_pyf, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_pyf.csv"))

## visualize it 
farms_pyf <-  st_as_sf(farms_pyf, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_pyf)
```

## India 
https://www.researchgate.net/publication/285488252_Mussel_Farming_and_Its_Potential_in_India
 - Has a potential map.

http://www.fao.org/fishery/countrysector/naso_india/en
 - Among maritime states, Kerala was the first to recognize the advantages of utilizing mussel farming technology in rural development, from a meagre production in 1997 where cultured mussel production rose to 1 250 tonnes in 2002 with over 250 mussel farms being established in the estuaries of Kerala.

Based on this, I will place 250 mussel farms along the coast in the Kerala state in India.

Source:
FAO Fisheries & Aquaculture - National Aquaculture Sector Overview - India, 2014. National Aquaculture Sector Overview Fact Sheets.

```{r}
# world <- ne_coastline(scale = 10) #world map with 10m resolution
# plot(world)

## read in state shapefile
kerala_state <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/India/Kerala/state.shp"))

#info of our spatial vector object
mapview(kerala_state)
plot(kerala_state)

# world <- st_as_sf(world)

kerala_state_world_crs <- st_transform(kerala_state, crs(coast_raw))

kerala_state_world_crop <- st_crop(coast_raw, kerala_state_world_crs) %>% ## crop to kerala extent
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>% ## merge all the geometries into one
  ungroup()
mapview(kerala_state_world_crop)

k_meters <- st_transform(kerala_state_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(k_meters)

buff <- st_buffer(k_meters, dist = 7000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)
plot(buff)

linestring <- st_cast(buff, "LINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(kerala_state_world_crop))

india <- ne_countries(scale = 10, country = "India", returnclass = "sf")
mapview(india)

linestring_fix <- st_difference(buffer_outline_fixed, india) ## now grab only the outline 
mapview(linestring_fix)
mapview(kerala_state_world_crop)
mapview(kerala_state)

kerala_state_crop <- as(linestring_fix, "Spatial")
plot(kerala_state_crop)
mapview(kerala_state_crop)

kerala_state_points_lines <- sample.line(kerala_state_crop, n = 280) ## equally place the number of farms along the coastline 
class(kerala_state_points_lines)
plot(kerala_state_points_lines)

kerala_state_points <- st_as_sf(kerala_state_points_lines) %>%
  mutate(new_id = 1:280) 
mapview(kerala_state_points)

# Now filter out the points that fall outside of kerala_state: rows 57-70
kerala_state_points_final <- kerala_state_points %>%
  filter(new_id %in% c(31:280))
mapview(kerala_state_points_final)

mean(st_distance(kerala_state_points_final, kerala_state_world_crop))

## get lat lon
ind_pts <- st_centroid(kerala_state_points_final)

ind_pts_latlon <- st_transform(ind_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_ind <- data.frame(do.call(rbind, st_geometry(ind_pts_latlon)))
names(farms_ind) <- c("lon", "lat")

farms_ind_final <- farms_ind %>%
  mutate(iso3c = "IND") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_ind <-  st_as_sf(farms_ind, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_ind)

## write to farms folder 
write_csv(farms_ind_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_ind.csv"))
```

## Thailand
NASO Maps by FAO: 
http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/thailand/en/
 - This shows both Shrimp and Bivalve aquaculture sites in Thailand. Although it is not comprehensive, it is the best information I could find. I'll filter for the bivalve sites.

Source:
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/thailand/en/

```{r}
## read in thailand shp file
thai_file <- file.path(path, "Thailand/Thailand.kmz")

thai_aqua <- read_sf(thai_file)

mapview(thai_aqua)

## We need to filter for just bivalve aquaculture. The best way that I found to do this was to look at the .kmz file in google earth pro, and determine with rows are bivalves..

## NASO 4,5,6,7 are the bivalves:
unique(thai_aqua$Name)

thai_bivalve <- thai_aqua %>%
  dplyr::filter(Name %in% c("[NASO / 4]", "[NASO / 5]", "[NASO / 6]", "[NASO / 7]"))

mapview(thai_bivalve)

tha_bivalve_mol <- st_transform(thai_bivalve, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
tha_pts <- st_centroid(tha_bivalve_mol)

tha_pts_latlon <- st_transform(tha_pts, crs = "+init=epsg:4326")

mapview(tha_pts_latlon)

farms_tha <- data.frame(do.call(rbind, st_geometry(tha_pts_latlon))) 
names(farms_tha) <- c("lon", "lat")
farms_tha <- farms_tha %>%
  mutate(iso3c = "THA") %>%
  mutate(data_quality = 0.667)

write_csv(farms_tha, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_tha.csv"))

farms_tha <-  st_as_sf(farms_tha, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_tha)
```

## Indonesia 
http://www.fao.org/3/AB716E/AB716E08.htm#:~:text=The%20commercially%20important%20bivalve%20species,is%20largely%20experimental%20at%20present.&text=Total%20shellfish%20production%20for%201984,for%20about%202.2%20%25%20of%20this.

 - This has a map of major oyster capture and culture areas in Indonesia... This will have to do. Approximate locations listed below.

 From left to right on the map: 
 3.2182751224662542, 99.69952320753342
 -5.964150035506321, 106.35924782378112
 -6.2646012824936275, 108.0659620387556
 -6.75680397048712, 108.64496221771543
 -6.975573928794459, 112.61871244712762
 -4.984732967006041, 120.31326531641196
 0.8890896205268102, 123.52557614041675
 -3.436748178542682, 129.1920950239657
 -5.142358892765042, 137.4727448207172

Source:
NETWORK OF AQUACULTURE CENTRES IN ASIA: Status of oyster culture in selected Asian countries, 1988.

```{r}
farms_idn <- data.frame(lat = c(3.2182751224662542, -5.964150035506321, -6.2646012824936275, -6.75680397048712, -6.975573928794459, -4.984732967006041, 0.889089620526810, -3.436748178542682, -5.142358892765042),
                        lon = c(99.69952320753342, 106.35924782378112, 108.0659620387556, 108.64496221771543, 112.61871244712762, 120.31326531641196, 123.52557614041675, 129.1920950239657, 137.4727448207172)) %>%
  mutate(iso3c = "IDN") %>%
  mutate(data_quality = 0.667)

write_csv(farms_idn, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_idn.csv"))

## visualize it 
farms_idn <-  st_as_sf(farms_idn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_idn)

```

## Russia
 - This probably doesn't really matter, because the production is so low, and the coastline is so big...
 - I can't find any good information on shellfish farming in Russia. I will place a point along the coast.
 - It does seem like scallop farming primarily occurs on the east coast, so I will place it there.

Source:
NA

```{r}

farms_rus <- data.frame(lat = c(43.053493816917545),
                        lon = c(131.95698485517275)) %>%
  mutate(iso3c = "RUS") %>%
  mutate(data_quality = 0.0)

write_csv(farms_rus, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_rus.csv"))

## visualize it 
farms_rus <-  st_as_sf(farms_rus, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_rus)

```

## Korea, Dem. People's Rep
**Revisiting Changing Patterns of North Koreaʹs Fisheries Production: 1990s‐2000s**
"By the 1980s, there has been 60 or more fish farming companies in North
Korea's national distribution"

 - I'll place 60 farms along the coast of north korea. Although this specifies "fish farms", it is better than nothing. And there are probably more than 60 farms (shell fish or not), nowadays.

Source:
Park, S., Hong, S., 2012. Revisiting Changing Patterns of North Koreaʹs Fisheries Production: 1990s‐2000s 19.

```{r}
# world <- ne_coastline(scale = 110) #world map with 110m resolution
# plot(world)

north_korea <- ne_countries(scale = 110, country = "North Korea", returnclass = "sf")

#info of our spatial vector object
mapview(north_korea)
plot(north_korea)

# world <- st_as_sf(world)

north_korea_world_crs <- st_transform(north_korea, crs(coast_raw))

north_korea_world_crop <- st_crop(coast_raw, north_korea_world_crs) %>% ## crop to north korea extent
  dplyr::filter(!(FID %in% c(44790, 135067, 44788:44789, 135092, 135103, 87219, 44955))) %>%
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>%
  ungroup()

mapview(north_korea_world_crop)

nk_meters <- st_transform(north_korea_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(nk_meters)

buff <- st_buffer(nk_meters, dist = 7000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)
plot(buff)

linestring <- st_cast(buff, "MULTILINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(north_korea_world_crop))

linestring_fix <- st_difference(buffer_outline_fixed, north_korea) ## now grab only the outline 
mapview(linestring_fix)

north_korea_crop <- as(linestring_fix, "Spatial")
plot(north_korea_crop)
mapview(north_korea_crop)

north_korea_points_lines <- sample.line(north_korea_crop, n = 85) ## equally place the number of farms along the coastline 
class(north_korea_points_lines)
plot(north_korea_points_lines)

north_korea_points <- st_as_sf(north_korea_points_lines) %>%
  mutate(new_id = 1:85)
mapview(north_korea_points)

## Now filter out the points that fall outside of north_korea: rows 57-70
north_korea_points_final <- north_korea_points %>%
  filter(new_id %in% c(1:32, 46:47, 55:80 ))
mapview(north_korea_points_final)

# extract lat lon
prk_pts <- st_centroid(north_korea_points_final)

prk_pts_latlon <- st_transform(prk_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_prk <- data.frame(do.call(rbind, st_geometry(prk_pts_latlon)))
names(farms_prk) <- c("lon", "lat")

farms_prk_final <- farms_prk %>%
  mutate(iso3c = "PRK") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_prk <-  st_as_sf(farms_prk, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_prk)

## write to farms folder 
write_csv(farms_prk_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_prk.csv"))
```

## Brazil 
http://www.fao.org/fishery/countrysector/naso_brazil/en
"Mollusc farming was introduced into Brazil in the 1960s but only since 1989 has this activity developed as an important economic alternative for small-scale fishermen. Besides the generation of income and employment, bivalve culture has had a noticeable effect in contributing to the stabilisation of traditional communities on their original lands. Today, there are no large companies in this sector and small farmers are organized into associations and cooperatives. Santa Catarina State in the South region is responsible for 93 percent of Brazilian bivalve mollusc production."

https://onlinelibrary.wiley.com/doi/full/10.1111/raq.12107
 "Currently, the total bivalve production reaches 19 082 tons, with **729 marine farmers** scattered among the 12 municipalities mentioned above, and marine aquaculture has now reached an important position in Santa Catarina State as a significant contributor to food production, providing employment and livelihood opportunities and representing a legitimate use of the marine environment for sustainable food production"

Of the 837 aquaculture areas grouped inside the **27 aquaculture parks** originally proposed by the PLDM, 798 areas and 24 parks are already licensed and attended by **618 shellfish farmers** who represent 74% of all shellfish farmers in Santa Catarina State

 - Based on this paper, I will assign 27 points along the coastline of Santa Catarina state. I'm assuming that the 3 remaining unlicensed aquaculture parks are licensed now.

Source:
Suplicy, F.M., Vianna, L.F. de N., Rupp, G.S., Novaes, A.L.T., Garbossa, L.H.P., de Souza, R.V., Guzenski, J., da Costa, S.W., Silva, F.M., dos Santos, A.A., 2017. Planning and management for sustainable coastal aquaculture development in Santa Catarina State, south Brazil. Rev Aquacult 9, 107–124. https://doi.org/10.1111/raq.12107

```{r}

## read in state shape 
sta_catarina <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/Brazil/Santa_Catarina/dc172qt8462.shp"))
mapview(sta_catarina)

# world <- ne_coastline(scale = 110) #world map with 110m resolution
# plot(world)

# world <- st_as_sf(world)

sta_catarina_world_crs <- st_transform(sta_catarina, crs(coast_raw))

sta_catarina_world_crop <- st_crop(coast_raw, sta_catarina_world_crs) %>% ## crop to santa catarina extent
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>%
  ungroup()

mapview(sta_catarina_world_crop)

sc_meters <- st_transform(sta_catarina_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(sc_meters)

buff <- st_buffer(sc_meters, dist = 7000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)
plot(buff)

linestring <- st_cast(buff, "MULTILINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

linestring <- st_cast(linestring, "LINESTRING")

brazil <- ne_countries(scale = 10, country = "Brazil", returnclass = "sf")
mapview(brazil)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(brazil))

linestring_fix <- st_difference(buffer_outline_fixed, brazil) ## now grab only the outline 
mapview(linestring_fix)
plot(sta_catarina_world_crop)
mapview(sta_catarina_world_crop)

sta_catarina_crop <- as(linestring_fix, "Spatial")
plot(sta_catarina_crop)
mapview(sta_catarina_crop)

sta_catarina_points_lines <- sample.line(sta_catarina_crop, n = 4) ## equally place the number of farms along the coastline 
class(sta_catarina_points_lines)
plot(sta_catarina_points_lines)

sta_catarina_points <- st_as_sf(sta_catarina_points_lines) %>%
  mutate(new_id = 1:28) %>%
  filter(new_id != 12)
mapview(sta_catarina_points)

## get lat lon
bra_pts <- st_centroid(sta_catarina_points)

bra_pts_latlon <- st_transform(bra_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_bra <- data.frame(do.call(rbind, st_geometry(bra_pts_latlon)))
names(farms_bra) <- c("lon", "lat")

farms_bra_final <- farms_bra %>%
  mutate(iso3c = "BRA") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_bra <-  st_as_sf(farms_bra, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_bra)

## write to farms folder 
write_csv(farms_bra_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_bra.csv"))
```

## Philippines

http://www.fao.org/3/AB710E/AB710E18.htm#ann12
 - This has the number of farms per province. I will allocate those number equally along the coastline of each province that has farms

LOCATION	Number of farms	
Zambales	 	 	NA
Bulacan	 	 	NA
Mindoro Oriental	 NA
Palawan	 	 	NA
Quezon	 	  NA
Capiz	**300**		NA
Iloilo	 	 	NA
Aklan	 	 	NA
Negros Occidental	**6** 
Cebu	 	 	NA
Bohol	 	 	NA
Leyte	 	 	NA
Samar	**250**	
Total	**556**

Source:
REPORT OF THE WORKSHOP AND STUDY TOUR ON MOLLUSC SANITATION AND MARKETING, 1989. . Food and Agriculture Organization of the United Nations.

```{r}
## read in province shape
provinces <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/philippines/provinces/Provinces.shp"))

mapview(provinces)

## filter provinces for the ones we want
provinces_farms <- provinces %>%
  dplyr::filter(NAME_1 %in% c("Capiz", "Negros Occidental", "Samar"))

mapview(provinces_farms)

# world <- ne_coastline(scale = 50) #world map with 50m resolution
# plot(world)

# start with Capiz
capiz <- provinces_farms %>%
  dplyr::filter(NAME_1 == "Capiz")
mapview(capiz)
plot(capiz)

# world <- st_as_sf(world)

capiz_world_crs <- st_transform(capiz, crs(coast_raw))

capiz_world_crop <- st_crop(coast_raw, capiz_world_crs) %>% ## crop to extent
  dplyr::filter(!(FID %in% c(443449, 439815, 443448, 439801))) %>%
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>%
  ungroup()

mapview(capiz_world_crop)

cp_meters <- st_transform(capiz_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(cp_meters)

buff <- st_buffer(cp_meters, dist = 7000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)

linestring <- st_cast(buff, "MULTILINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

linestring <- st_cast(linestring, "LINESTRING")

philippines <- ne_countries(scale = 10, country = "Philippines", returnclass = "sf")
mapview(philippines)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(philippines))

linestring_fix <- st_difference(buffer_outline_fixed, philippines) ## now grab only the outline 
mapview(linestring_fix)

capiz_crop <- as(linestring_fix, "Spatial")
plot(capiz_crop)
mapview(capiz_crop)

capiz_points_lines <- sample.line(capiz_crop, n = 151) ## equally place the number of farms along the coastline 
class(capiz_points_lines)
plot(capiz_points_lines)

capiz_points <- st_as_sf(capiz_points_lines) %>%
  mutate(new_id = 1:302) %>%
  filter(!(new_id %in% c(150, 146))) %>%
  dplyr::select(-new_id)
mapview(capiz_points)

# now do negros
negros <- provinces_farms %>%
  dplyr::filter(NAME_1 == "Negros Occidental")
mapview(negros)

negros_world_crs <- st_transform(negros, crs(coast_raw))

negros_world_crop <- st_crop(coast_raw, negros_world_crs) %>%
  filter(FID %in% c(224916:224917, 224908:224912)) %>%
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>%
  ungroup()

mapview(negros_world_crop)

n_meters <- st_transform(negros_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(n_meters)

buff <- st_buffer(n_meters, dist = 5000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)

linestring <- st_cast(buff, "MULTILINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

linestring <- st_cast(linestring, "LINESTRING")

philippines <- ne_countries(scale = 10, country = "Philippines", returnclass = "sf")
mapview(philippines)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(philippines))

linestring_fix <- st_difference(buffer_outline_fixed, philippines) ## now grab only the outline 
mapview(linestring_fix)

negros_crop <- as(linestring_fix, "Spatial")
plot(negros_crop)
mapview(negros_crop)

negros_points_lines <- sample.line(negros_crop, n = 6) ## equally place the number of farms along the coastline 
class(negros_points_lines)
plot(negros_points_lines)

negros_points <- st_as_sf(negros_points_lines)
mapview(negros_points)

# now do samar 
samar <- provinces_farms %>%
  dplyr::filter(NAME_1 == "Samar")
mapview(samar)

samar_world_crs <- st_transform(samar, crs(coast_raw))

# marabut to veriato

samar_world_crop <- st_crop(coast_raw, samar_world_crs) %>%
  filter(FID %in% c(222688:222691, 222671:222673)) %>%
  mutate(new_id = 1) %>%
  group_by(new_id) %>%
  summarise(geometry = st_union(geometry)) %>%
  ungroup()

mapview(samar_world_crop)

sm_meters <- st_transform(samar_world_crop, crs = 7801) ## transform to a crs that can buffer in meters 
mapview(sm_meters)

buff <- st_buffer(sm_meters, dist = 3000) # buffer out to 1 km so that we can place points along a line 1 km from the coastline
mapview(buff)

linestring <- st_cast(buff, "MULTILINESTRING") ## make a line string so we can filter for the outline 
mapview(linestring)

linestring <- st_cast(linestring, "LINESTRING")

philippines <- ne_countries(scale = 10, country = "Philippines", returnclass = "sf")
mapview(philippines)

## now reproject back to original 
buffer_outline_fixed <- st_transform(linestring, crs = crs(philippines))

linestring_fix <- st_difference(buffer_outline_fixed, philippines) ## now grab only the outline 
mapview(linestring_fix)

samar_crop <- as(linestring_fix, "Spatial")
plot(samar_crop)
mapview(samar_crop)

samar_points_lines <- sample.line(samar_crop, n = 283) ## equally place the number of farms along the coastline 
class(samar_points_lines)
plot(samar_points_lines)

samar_points <- st_as_sf(samar_points_lines) %>%
  mutate(new_id = 1:283) %>%
  filter(!(new_id %in% c(2:13, 108:109, 98:100, 79, 74:76, 62:66, 52:53, 14:16, 95:96))) %>%
  dplyr::select(-new_id)
mapview(samar_points)

## bind all three states together
all_points <- rbind(samar_points, negros_points, capiz_points)
mapview(all_points)

# get lat lon points
phl_pts <- st_centroid(all_points)

phl_pts_latlon <- st_transform(phl_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(geometry)

farms_phl <- data.frame(do.call(rbind, st_geometry(phl_pts_latlon)))
names(farms_phl) <- c("lon", "lat")

farms_phl_final <- farms_phl %>%
  mutate(iso3c = "PHL") %>%
  mutate(data_quality = 0.333)

## visualize it 
farms_phl <-  st_as_sf(farms_phl, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_phl)

## write to farms folder 
write_csv(farms_phl_final, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_phl.csv"))
```

## Malaysia
http://www.fao.org/fishery/countrysector/naso_malaysia/en
"Brackish water aquaculture is characterized by the extensive culture of bivalve molluscs, occupying an area of 7 659 ha, mostly in the western coastal waters where there is an abundance of mud flats suitable for the culture of blood cockles"

 - I'll place a point on the western coast:
 - 4.074359958576369, 100.73343376478226

Source: 
NA

```{r}

farms_mys <- data.frame(lat = c(4.074359958576369),
                        lon = c(100.73343376478226)) %>%
  mutate(iso3c = "MYS") %>%
  mutate(data_quality = 0.0)

write_csv(farms_mys, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_mys.csv"))

## visualize it 
farms_mys <-  st_as_sf(farms_mys, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_mys)

```

## Cambodia 
http://www.fao.org/fishery/countrysector/naso_cambodia/en
 - There is no good information on the location of farms, I will just place a point on the coast. 
 - 11.069927194335508, 103.64594001534627

Source:
NA

```{r}

farms_khm <- data.frame(lat = c(11.069927194335508),
                        lon = c(103.64594001534627)) %>%
  mutate(iso3c = "KHM") %>%
  mutate(data_quality = 0.0)

write_csv(farms_khm, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_khm.csv"))

## visualize it 
farms_khm <-  st_as_sf(farms_khm, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_khm)

```

## Hong Kong
https://www.afcd.gov.hk/english/fisheries/fish_aqu/fish_aqu_mpo/fish_aqu_mpo.html

All farms are located in Deep Bay, specifically shown in the map.

 - I will place a point in Deep Bay

Source: 
Marine fish culture, pond fish culture and oyster culture, n.d. . Agriculture, Fisheries, and Conservation Department; The Government of the Hong Kong Special Administrative Region.  https://www.afcd.gov.hk/english/fisheries/fish_aqu/fish_aqu_mpo/fish_aqu_mpo.html

```{r}

farms_hkg <- data.frame(lat = c(22.48384043911231),
                        lon = c(113.98143449612208)) %>%
  mutate(iso3c = "HKG") %>%
  mutate(data_quality = 0.667)

write_csv(farms_hkg, here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_hkg.csv"))

## visualize it 
farms_hkg <-  st_as_sf(farms_hkg, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_hkg)

```

# Join cell count dataframes

```{r}
farm_files <- list.files(here::here("marine/bivalve/bivalve_farms/data/bivalve_sites"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

all_farms_1 <- all_farms %>%
  dplyr::select(-data_quality)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

write.csv(all_farms_1, here::here("marine/bivalve/bivalve_farms/data/global_bivalve_farm_lat_lon.csv"), row.names = FALSE)

write.csv(all_farms, here::here("marine/bivalve/bivalve_farms/data/global_bivalve_farm_lat_lon_data_quality.csv"))
```