### 2021-02-12 marine/crustaceans/crustacean_farms/STEP2_crustacean_farm_locations.Rmd

---
title: "Location of crustacean Farms"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script standardizes location of crustacean farms to be lat/long point data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

fao_crustacean_spp <- read_csv("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans_taxa_production.csv") %>%
  dplyr::filter(year == 2017)
fao_crustacean <- read_csv("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans.csv") %>%
  dplyr::filter(year == 2017)

unique(fao_crustacean$country)

```

# Raster

Standard raster used for this analysis

```{r}

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

Now we will map the remaining countries with shrimp farms that are not included in the ASC dataset. Workflow will be as follows: 
 - Try to obtain exact locations of farms from published source or dataset (such as FAO or a research paper)
 - If there are no exact locations, but the source lists a city or state, place a point in that particular city or state
 -  If there is enough information to know there are a specific number of shrimp farms, equally allocate the farms across the coastline. If there are estimates of farms per state/province, equally allocate them across the state/province, to be even more precise. 
 - If there are not exact numbers of farms in the country, but you know there are a lot, randomly place points along the coast line (a non-exact number, like we did for Taiwan). 
 - Lastly, if there is no evidence of marine shrimp farm locations, place a point on the coast of the country (preferably with some intuition/knowledge of where they might be).

For each point, we will add a column for "data quality". It will be scored as very good (5), good (4), fair (3), bad (2), and very bad (1). 
 5: Very Good qualifications - Locations taken from leases, or from a map with specific locations
   - Must not intersect with land mask
   - "Land mask"

4: Good qualifications - Locations are projected or it is a only sample of total locations
   - Must not intersect with land mask
   - "Land mask"

3: Fair qualifications - A specific number of farms are placed on the coast, without knowing exact locations. 
   - Must be allocated spatially
   - Must not intersect with land mask
   - Flagged as "Allocate spatially"

2: Bad qualifications -  We know that aquaculture is limited to a specific subregion/state, but do not know how many or where the farms are. 
   - Must be allocated using suitability analysis.
   - Flagged as 2 and "state/province suitability"

1: Very Bad qualifications - No data at all.
   - Must be allocated using suitability analysis.
   - Flagged as 1 and "country suitability"

## Indonesia
Can't find any infomration on crab farm locations in Indonesia. I will place one farm, randomly on the coast of Indonesia.

Source: 
NA

```{r}

farms_idn <- data.frame(lat = c(0.3359943014669836),
                        lon = c(117.55582991782983)) %>%
  mutate(iso3c = "IDN") %>%
  mutate(data_quality = 1) %>%
  mutate(action_needed = "country based suitability")

write_csv(farms_idn, here("marine/crustaceans/crustacean_farms/data/crustacean_sites/farms_idn.csv"))

## visualize it 
farms_idn <-  st_as_sf(farms_idn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_idn)
```

## China

Can't find any datasets of marine crustacean farm locations in China. I will place one farm randomly on the coast of China.

Source: 
NA

```{r}

farms_chn <- data.frame(lat = c(29.378009267567716),
                        lon = c(121.95043909100868)) %>%
  mutate(iso3c = "CHN") %>%
  mutate(data_quality = 1) %>%
  mutate(action_needed = "country based suitability")

write_csv(farms_chn, here("marine/crustaceans/crustacean_farms/data/crustacean_sites/farms_chn.csv"))

## visualize it 
farms_chn <-  st_as_sf(farms_chn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_chn)
```

## Bangladesh

https://www.researchgate.net/publication/318250461_Mud_crab_aquaculture_and_fisheries_in_coastal_Bangladesh

"Mud crabs have already gained popularity among coastal communities in greater Khulna and Chittagong regions (Azam et al. 1998) and are being recognized as a candidate species for culture in brackishwater environments (Ballao et al. 1999). There are about 300,000 people directly or indirectly connected with mud crab farming activities."

 - I'll place points in the Khulna and Chittagong regions, since I can't find any other information on mud crab aquaculture locations in Bangladesh.

 - Khulna: 21.876081791221054, 89.49499697244659
 - Chittagong: 22.325919306644522, 91.76441856678511

Source: 
Rahman, M., Islam, M.A., Haque, S.M., Wahab, A., 2017. Mud Crab Aquaculture and Fisheries in Coastal Bangladesh 7.

```{r}

farms_bgd <- data.frame(lat = c(21.876081791221054, 22.325670145257305),
                        lon = c(89.49499697244659, 91.76091935028879)) %>%
  mutate(iso3c = "BGD") %>%
  mutate(data_quality = 2) %>%
  mutate(action_needed = "state/province based suitability")

write_csv(farms_bgd, here("marine/crustaceans/crustacean_farms/data/crustacean_sites/farms_bgd.csv"))

## visualize it 
farms_bgd <-  st_as_sf(farms_bgd, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_bgd)
```

## Myanmar

http://www.fao.org/3/ad497e/ad497e05.htm
"Mud crab farming is also important in Myanmar. Mud crab (Scylla serrata, and related species) culture in mangroves or tidal flats is practiced mainly in the Ayeyarwaddy delta area, in Rakhine and southern parts of Myanmar."

 - I'll place points in Ayeyarwaddy region and Rakhine state.
 - Ayeyarwaddy: 15.930407995930784, 94.76647475697742
 - Rakhine: 18.615093163569437, 94.24290607527016

Source:
Myanmar aquaculture and inland fisheries, 2003. Food and Agriculture Organization of the United Nations Regional Office For Asia and the Pacific.

```{r}

farms_mmr <- data.frame(lat = c(15.930407995930784, 18.615093163569437),
                        lon = c(94.76647475697742, 94.24290607527016)) %>%
  mutate(iso3c = "MMR") %>%
  mutate(data_quality = 2) %>%
  mutate(action_needed = "state/province based suitability")

write_csv(farms_mmr, here("marine/crustaceans/crustacean_farms/data/crustacean_sites/farms_mmr.csv"))

## visualize it 
farms_mmr <-  st_as_sf(farms_mmr, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_mmr)
```

## Viet nam

https://aciar.gov.au/sites/default/files/legacy/node/693/ias36_pdf_18085.pdf
"In addition to harvesting wild stock, mud crabs are now being farmed in
many coastal provinces in Vietnam, including Quang Ninh, Hai Phong,
Thanh Hoa, Ninh Binh, Nghe An, Thua Thien–Hue, Ba Ria–Vung Tau,
Ho Chi Minh City, Ben Tre, Tra Vinh, Soc Trang, Minh Hai, Kien Giang
and Bac Lieu."

"It is difficult to know how many people participate in this industry.
Apparently, no formal surveys have been conducted, there are no known
official statistics, and crab farming is typically undertaken together with
other activities, such as shrimp and/or fish aquaculture, other primary
production such as rice growing, or even as a sideline to paid employment."

"Table 4 gives the sole set of official statistics on crab aquaculture area and
production in Vietnam in 2004. The statistics cover both mud crabs and
blue swimmer crabs, and were collated by the Vietnamese Ministry of
Fisheries from reports of provincial fishery departments. "

 - Table on page 20: This is a table from 2004, but it will have to do. It also has production estimates, which is nice. I will randomly place a point along the coast in each province, and allocate the production accordingly in step 3.. 
  - NOTE: for the production, we are assuming that not reported within a province == 0 production, due to the lack of data.

  - Quang Ninh: prod = 0: 21.186050529545593, 107.60858937642442
  - Hai Phong: prod = 1650: 20.840947248646557, 106.77685003845498
  - Thai Binh: prod = 0: 20.454148639095266, 106.59728856581118
  - Nam Dinh: prod = 0: 20.146554460540862, 106.3298019452741
  - Ninh Binh: prod = 1005: 19.955496263213185, 106.03097468374126
  - Thanh Hoa: prod = 970: 19.726817666482322, 105.87661730467457
  - Nghe An: prod = 0: 18.982876968029743, 105.6165097440173
  - Ha Tinh: prod = 200: 18.456817613512555, 105.92726656633255
  - Quang Binh: prod = 110: 17.56001328746647, 106.58039999954293
  - Quang Tri: prod = 20: 16.905826522213328, 107.19194145855845
  - TT Hue: prod = 82: 16.495115996149718, 107.74891583338649
  - Da Nang: prod = 0: 16.167986126005765, 108.15075180169012
  - Quang Nam: prod = 135: 15.699026568754403, 108.48016289264334
  - Quang Ngai: prod = 0: 15.010595152983287, 108.92174225624275
  - Binh Dinh: prod = 130: 14.17668770026659, 109.19617116389627
  - Phu Yen: prod = 0: 13.335760529041393, 109.29183511586322
  - Khanh Hoa: prod = 0: 12.299111935360338, 109.23699322130719
  - Ninh Thuân: prod = 5: 11.62183927318579, 109.1546891359726
  - Binh Thuân: prod = 0: 10.830102342204183, 108.03702110643216
  - Long An: prod = 250: 10.510731358280973, 106.7310999868549
  - Tien Giang: prod = 0: 10.30006571666483, 106.78122716883196
  - Ben Tre: prod = 637: 10.037866832581582, 106.69637036338817
  - Tra Vinh: prod = 1196: 9.592223563185232, 106.54126727713718
  - Soc Trang: prod = 0: 9.389580381208823, 106.1857446280396
  - Bac Lieu: prod = 3466: 9.147921315425062, 105.59613926925006
  - Ca Mau: prod = 0: 8.674166818318303, 105.12379673906345
  - Kien Giang: prod = 170: 10.087872204898957, 104.88743812284027
  - Can Tho: prod = 0: 10.055231, 105.789267

Viet Nam: 
Lindner, B., 2005. IMPACTS OF MUD CRAB HATCHERY TECHNOLOGY IN VIETNAM 68.

```{r}

farms_vnm <- data.frame(lat = c(21.186050529545593, 20.840947248646557, 20.454148639095266, 20.146554460540862, 19.955496263213185, 19.726817666482322, 18.982876968029743, 18.456817613512555, 17.56001328746647, 16.905826522213328, 16.495115996149718, 16.167986126005765, 15.699026568754403, 15.010595152983287, 14.17668770026659, 13.335760529041393, 12.299111935360338, 11.62183927318579, 10.830102342204183, 10.510731358280973, 10.30006571666483, 10.037866832581582, 9.592223563185232, 9.389580381208823, 9.147921315425062, 8.674166818318303, 10.087872204898957, 10.055819944690162),
                        lon = c(107.60858937642442, 106.77685003845498, 106.59728856581118, 106.3298019452741, 106.03097468374126, 105.87661730467457, 105.6165097440173, 105.92726656633255, 106.58039999954293, 107.19194145855845, 107.74891583338649, 108.15075180169012, 108.48016289264334, 108.92174225624275, 109.19617116389627, 109.29183511586322, 109.23699322130719, 109.1546891359726, 108.03702110643216, 106.7310999868549, 106.78122716883196, 106.69637036338817, 106.54126727713718, 106.1857446280396, 105.59613926925006, 105.12379673906345,  104.88743812284027, 105.79073953273914),
                        prod = c(0, 1650, 0, 0, 1005, 970, 0, 200, 110, 20, 82, 0, 135, 0, 130, 0, 0, 5, 0, 250, 0, 637, 1196, 0, 3466, 0, 170, 0), sub_rgn = c("Quang Ninh", "Hai Phong", "Thai Binh", "Nam Dinh", "Ninh Binh", "Thanh Hoa", "Nghe An", "Ha Tinh", "Quang Binh", "Quang Tri", "TT Hue", "Da Nang", "Quang Nam", "Quang Ngai", "Binh Dinh", "Phu Yen", "Khanh Hoa", "Ninh Thuân", "Binh Thuân", "Long An", "Tien Giang", "Ben Tre", "Tra Vinh", "Soc Trang", "Bac Lieu", "Ca Mau", "Kien Giang", "Can Tho")) %>%
  mutate(iso3c = "VNM") %>%
  mutate(data_quality = 2) %>%
  mutate(action_needed = "state/province based suitability")

write_csv(farms_vnm, here("marine/crustaceans/crustacean_farms/data/crustacean_sites/farms_vnm.csv"))

## visualize it 
farms_vnm_map <-  st_as_sf(farms_vnm, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_vnm_map)
```

## Philippines

All former reports can be found here: 
https://psa.gov.ph/content/fisheries-statistics-philippines

Fisheries statistics of the Philippines: https://psa.gov.ph/sites/default/files/Fisheries%20Statistics%20of%20the%20Philippines%2C%202017-2019.pdf

Table 125: TABLE 125 Volume of Mud Crab Production in Aquaculture by Region and Province,
 Philippines: 2017 - 2019

 - I've saved the table in Aurora as a .csv.

 - I'll place a point in each location listed in the table, and use the production estimates to allocate the FAO production equally (although the number is almost the same as reported in the FAO, off by 3 tonnes).

Consider using this data for future iterations... 
https://openstat.psa.gov.ph/PXWeb/pxweb/en/DB/search/?searchquery=aquaculture&rxid=bdf9d8da-96f1-4100-ae09-18cb3eaeb313

Source: 
Fisheries Statistics of the Philippines, 2017-2019, 2020. . Philippine Statistics Authority. [Accessed December 18, 2020]

```{r}
phl_mudcrab_prod <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Philippines/phl_mudcrab_production.csv")) %>%
  dplyr::select(-X6) %>%
  mutate(`2017` = as.numeric(`2017`), 
         `2018` = as.numeric(`2018`),
         `2019` = as.numeric(`2019`)) %>%
  pivot_longer(cols = c(`2017`, `2018`, `2019`), names_to = "year", values_to = "prod") %>%
  dplyr::filter(!is.na(prod), 
                year == 2017)

## now we need to find coordinates for each province. I will randomly select a lat lon point along the coast of each province:

provinces <- data.frame(Province = unique(sort(phl_mudcrab_prod$Province))) %>%
  mutate(lat_lon = c("9.190611500208238, 125.52994229122065", "11.737731148020183, 122.36341263788152", "13.21836252920181, 123.77011641931539", "11.096612509189843, 122.04466274447995", "14.588280500720717, 120.39232212075018", "10.062398891617871, 124.08857333922121", "14.759345796261423, 120.79195178365572", "14.278379190276917, 122.80695157663077", "13.728018225223346, 123.21602051808212", "11.544441602769396, 122.87251489793287", "13.600373850277089, 124.08856214435265", "7.226273453773761, 125.8615542691359", "7.29064081849023, 125.70838297490391", "6.8067269257546394, 125.38044959114181", "7.021850825639651, 126.44668668618641", "10.249781995409927, 125.64231490812028", "11.457019560196713, 125.50124809292508", "18.35920136671901, 121.62457647441146", "14.294681717859676, 120.71661934172276", "10.565159685797113, 122.51279683720878", "16.5330902343553, 120.31511351999148", "8.187575265346036, 124.12440346213432", "10.86506672011331, 124.72028684317488", "17.08773382905305, 120.43638606220428", "17.205031284221167, 120.4139227279512", "10.891325726593882, 122.7831164266059", "6.994603962157642, 123.99059538388416", "13.344206870917525, 121.82133665840661", "12.576951933159716, 123.28286940887746", "14.510536378070825, 121.08748285435844", "8.483931963597566, 123.80942490173003", "8.846965587470466, 125.16333794567215", "9.943820309588942, 122.44630440874657", "12.509837523018708, 124.65010656730024", "13.415437014491157, 121.0922973705213", "10.038412865363988, 118.7206892394855", "14.855971440318308, 120.58766325017285", "16.124387999329027, 120.09336724359946", "14.664908885542868, 121.61549543341468", "12.585717059110207, 122.27016916837972", "12.007625357698121, 124.69126813617964", "12.818661572681938, 123.83950186846029", "6.373081834977807, 124.08215994073517", "9.555721530351239, 125.78630860627861", "9.013346905513, 126.23742298508941", "6.9053988560744335, 122.0691745144182", "8.18224915422801, 122.95491554623727", "7.81172076996685, 123.42354992829671", "7.703880319874889, 122.52615049321432", "15.272079732639275, 120.00987788280406")) %>%
  separate(lat_lon, into = c("lat", "lon"), sep = ",")

## Now join together with our larger dataset

phl_lat_lon <- phl_mudcrab_prod %>%
  left_join(provinces, by = "Province") %>%
  dplyr::select(lat, lon, prod, "sub_rgn" = "Province") %>%
  mutate(iso3c = "PHL") %>%
  mutate(data_quality = 2) %>%
  mutate(action_needed = "state/province based suitability")

write_csv(phl_lat_lon, here("marine/crustaceans/crustacean_farms/data/crustacean_sites/farms_phl.csv"))

## visualize it 
farms_phl_map <-  st_as_sf(phl_lat_lon, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_phl_map)
```

# Join cell count dataframes

```{r}
farm_files <- list.files(here::here("marine/crustaceans/crustacean_farms/data/crustacean_sites"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

write.csv(all_farms, here::here("marine/crustaceans/crustacean_farms/data/global_crustacean_farm_lat_lon_data_quality.csv"), row.names = FALSE)

all_farms_1 <- all_farms %>%
  dplyr::select(-data_quality, -action_needed)

write.csv(all_farms_1, here::here("marine/crustaceans/crustacean_farms/data/global_crustacean_farm_lat_lon.csv"), row.names = FALSE)
```