### 2021-02-12 marine/salmon/salmon_farm/STEP2_salmon_farm_locations.Rmd

---
title: "Location of Salmon Farms"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "10/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script standarizes location of salmon farms to be lat/long point data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"
fao_salmon <- read_csv("marine/salmon/salmon_farm/data/fao_mariculture_salmon.csv") %>%
     dplyr::filter(year == 2017)

sort(unique(fao_salmon$country))

#  [1] "Australia"                "Canada"                   "Chile"                    "Denmark"                 
#  [5] "Faroe Islands"            "Finland"                  "Iceland"                  "Ireland"                 
#  [9] "Japan"                    "New Zealand"              "Norway"                   "Russian Federation"      
# [13] "Sweden"                   "Turkey"                   "United Kingdom"           "United States of America"
```

# Raster

Standard raster used for this analysis

```{r}

source(here("_spatial/template_raster.R"))

```

```{r}

coast_raw <- st_read("/home/shares/food-systems/Aquaculture_mapping/Data/Jared/Marine_Brackish/raw_data/Coastline/lines.shp")

mapview(head(coast_raw))

```

# Global regions

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```
Now we will map the remaining countries with shrimp farms that are not included in the ASC dataset. Workflow will be as follows: 
 - Try to obtain exact locations of farms from published source or dataset (such as FAO or a research paper)
 - If there are no exact locations, but the source lists a city or state, place a point in that particular city or state
 -  If there is enough information to know there are a specific number of shrimp farms, equally allocate the farms across the coastline. If there are estimates of farms per state/province, equally allocate them across the state/province, to be even more precise. 
 - If there are not exact numbers of farms in the country, but you know there are a lot, randomly place points along the coast line (a non-exact number, like we did for Taiwan). 
 - Lastly, if there is no evidence of marine shrimp farm locations, place a point on the coast of the country (preferably with some intuition/knowledge of where they might be).

For each point, we will add a column for "data quality". It will be scored as very good (5), good (4), fair (3), bad (2), and very bad (1). 
 5: Very Good qualifications - Locations taken from leases, or from a map with specific locations
   - Must not intersect with land mask
   - "Land mask"

4: Good qualifications - Locations are projected or it is a only sample of total locations
   - Must not intersect with land mask
   - "Land mask"

3: Fair qualifications - A specific number of farms are placed on the coast, without knowing exact locations. 
   - Must be allocated spatially
   - Must not intersect with land mask
   - Flagged as "Allocate spatially"

2: Bad qualifications -  We know that aquaculture is limited to a specific subregion/state, but do not know how many or where the farms are. 
   - Must be allocated using suitability analysis.
   - Flagged as 2 and "state/province suitability"

1: Very Bad qualifications - No data at all.
   - Must be allocated using suitability analysis.
   - Flagged as 1 and "country suitability"

# Chile

Data for Chile comes from https://mapas.subpesca.cl/ideviewer/?locale=en-gb.

There is also a list of concessions here: http://www.subpesca.cl/portal/619/w3-article-92935.html

I was in contact with Valeria Tamayo at UCSB who helped me with some spanish translations concerning this data.

Source:
Undersecretariat for Fisheries and Aquaculture, 2020. [Accessed Feb 2020] https://mapas.subpesca.cl/ideviewer/?locale=en-gb

```{r}

# Column meaning and names
#0 LINK_DOC - URL of the SUBPESCA resolution approving the aquaculture or resolution technical project that cartographically regulates the aquaculture concession
#1 CARTA - Name of the geographic reference letter
#2 DATUM - Reference datum name
#3 HUSO - Spindle Name
#4 COMUNA - Name of the commune
#5 N_PERT - Pert number, code of the application for admission to process
#6 N_CODIGOCE - Farming center code according to the national aquaculture registry
#7 N_RESOLSSP - Resolution number of the undersecretary of fisheries and aquaculture that apprehends the aquaculture technical project
#8 F_RESOLSSP - Date of issuance of the resolution of the undersecretary of fisheries and aquaculture approving the aquaculture technical project
#9 N_RESOLSSF - Resolution number of the undersecretary for the armed forces granted by the aquaculture concession
#10 F_RESOLSSF - Date of issuance of the resolution of the undersecretary for the armed forces granted by the aquaculture concession
#11 TOPONIMIO - Geographical location of the aquaculture concession
#12 ESPECIES - Name of crop species
#13 AGRUPCONCE - Identifier of the salmonid concession group or sanitary neighborhood
#14 SUPERFICIE - Surface in hectares
#15 TITULAR - Name of the holder of the aquaculture concession
#16 COORDENADS - List of geographical coordinates of the aquaculture concession corresponding to each of the vertices of the polygon
#17 REGION - Region name
#18 GRUPO_ESP - Group of crop species
#19 TIPO_PORCI - Type of space where the aquaculture concession is located
#20 ESTADO - Processing status of the aquaculture concession
#21 ESTADO_SOL - Status of the aquaculture application process

chile_dat <- st_read(file.path(path, "Chile"), layer = "Concesiones_de_Acuicultura")

```

We filter to only those that produce salmon, but note that this is ALL leases, not neccesarily only active sites. 
```{r}

#Filter to only salmon farming
chile <- chile_dat %>% 
  filter(grepl("SALMON", REP_SUB_13))

# convert polygons to points
chile_pts <- st_centroid(chile)
chile_pts <- st_transform(chile_pts, food_crs)

# save as a dataframe
farms_chile <- data.frame(do.call(rbind, st_geometry(chile_pts))) 
names(farms_chile) <- c("lon", "lat")
farms_chile <- farms_chile %>%
  mutate(iso3c="CHL") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_chile, here("marine/salmon/salmon_farm/data/farm_sites/farms_chile.csv"))

```

# Norway

We use the number of salmon licenses to get a proxy for the relative distribution of production along the Norwegian coast.

## Load csv with lat long
License data.

Source: 
Fiskeridirektoratet, 2019. 2019 aquaculture license data. [accessed May 2019] https://fiskeridir.no/Akvakultur/Registre-og-skjema/Akvakulturregisteret

```{r}

enc <- readr::guess_encoding(file.path(path, "Norway/Akvakulturregisteret.csv"), n_max = 1000)[,"encoding"] %>% as.character()

norway_csv <- read.csv(file = file.path(path, "/Norway/Akvakulturregisteret.csv"), skip = 1, header = TRUE, sep = ";", fileEncoding = enc)

#Variable translations
#Akvakulturregisteret – variables:
#TILL_NR - License number
#ORG.NR/PERS.NR - Company official number
#NAVN - company name
#ADRESSE - address
#POSTNR - post number
#POSTSTED - regional center
#TILDELINGSTIDSPUNKT - date that the license got accepted
#TIDSBEGRENSET - time bound of license
#TILL_KOMNR - license municipality number
#TILL_KOM - License municipality
#FORMÅL - purpose of use - Purpose (you probaly want "kommersiell" =commercial)
#PRODUKSJONSFORM - Production form (Matfisk = grow out)
#ART - Species (Laks=salmon, ørret=trout, regnbueørret = rainbow trout)
#TILL_KAP - Maximum allowed biomass for the license - some kind of capacity
#TILL_ENHET - units of measure (TN = ton, KG = kilogram)
#LOK_NR - Location number (identifies actual group of pens)
#LOK_NAVN - Location name
#LOK_KOMNR - Location municipality number
#LOK_KOM - Location municipality
#LOK_PLASS - Location place (land or sea) - sea 
#VANNMILJØ - Water environment (freshwater, saltwater or both) - saltvann and freskvann/saltvann
#LOK_KAP - Location max allowed biomass
#LOK_ENHET - 
#N_GEOWGS84 - GPS North
#Ø_GEOWGS84 - GPS East

```

We filter the data to only include commercial marine salmon production. We also only include farms that are growing salmon for food.

The dataset also has rows with the same lat/lon coordinates, so we only use unique lat/lon after filtering to get single points for each location.

http://www.fao.org/fishery/countrysector/naso_norway/en
"By the end of 2003, 832 licences for ongrowing in sea water".  And 1,018 here (includes salmon and rainbow trout): https://www.ssb.no/en/statbank/table/08967/tableViewLayout1/. 
I get a similar value of 893 by including all years to 2013 and removing duplicate regions.  So, I am assuming that most of these licenses in these data are active.  Regardless, this should provide a reasonable estimate of production distribution.

For 2018 (https://www.ssb.no/en/statbank/table/08967/tableViewLayout1/) including salmon and rainbow trout: 1,160
These data have 920 licenses for salmon only. These values seem reasonably aligned to me.  I am using this information to make sure these data do not include inactive licenses, how to best account for multiple license listings at the same location (conclusion: count only once), and whether we need to subset by year enacted (no).

```{r}

#FORMAL = Kommersiell, Slaktemerd, 
unique(norway_csv$FORMÅL)

#Kommersiell = Commerical
#Forskning = Research
#Slaktemerd  = Slaughter  cage             
#Fangstbasert akvakultur = Catch-based aquaculture, i.e., wild caught fish are stored in pens to hold them until prices increase during non-fishing season, https://www.sciencedirect.com/science/article/pii/S0308597X18302252
#Undervisning    =  Teaching
#Manntall =   Census            
#Visning =  Display
# Fiskepark = Fishpark
# Utvikling = Development

unique(norway_csv$PRODUKSJONSFORM)

#MATFISK = food fish                             
#SETTEFISK  = hatcheries                        
#STAMFISK = broodstock                         
#STAMDYR = broodstock                            
#AKVAKULTURDYR I TIDLIGE LIVSSTADIER =AQUACULTURE ANIMALS IN EARLY LIFE STAGES
#AKVAKULTURDYR TIL KONSUM = AQUACULTURE ANIMALS FOR CONSUME          
#MATFISK-GRØNN KONVERTERT = FOOD FISH GREEN CONVERTED         
#MATFISK-GRØNN A  = Food fish green A              
#MATFISK - (5% MTB ØKNING)  = FOOD - (5% MTB INCREASE)         
#ALGER TIL  fôr/konsum    = ALGER FOR FOOD / CONSUMPTION
#SKALLDYR   = SHELLFISH
#ANDRE FORMÅL   = Other objective                     
#MATFISK-GRØNN C = Food fish green C                    
#ALGER TIDLIG FASE  = ALGER EARLY PHASE
#HAVBEITE = ranching                           
#MATFISK-GRØNN B   = food fish green B                 
#ØKOLOGISK MATFISK   =  organic food fish                
#MATFISK-GRØNN A (5% MTB ØKNING) = FOOD FISH GREEN A (5% MTB INCREASE

unique(norway_csv$ART)
species <- "Laks"

unique(norway_csv$VANNMILJØ)
# FERSKVANN = freshwater
# BRAKKVANN = brackishwater
# SALTVANN = saltwater

# Only include Commercial production
purpose <- c("Kommersiell")

#Only include salmon for food
objective <- c("MATFISK", "AKVAKULTURDYR TIL KONSUM", "MATFISK-GRØNN KONVERTERT", "MATFISK - (5% MTB ØKNING)", "MATFISK-GRØNN C", "HAVBEITE", "MATFISK-GRØNN B", "ØKOLOGISK MATFISK", "MATFISK-GRØNN A (5% MTB ØKNING)")

# Subset
sub_norway <- norway_csv %>% 
  filter(ART == species,
         FORMÅL %in% purpose,
         PRODUKSJONSFORM %in% objective,
         VANNMILJØ %in% c("SALTVANN")) #only 1 FERSKVANN/SALTVANN, not going to have a large effect on results if we cut/include

#There are multiple liscenses for the same place.  This may be for multiple pens or across time.  
# We will count them as one location given that the data then match well with other sources
sub_norway %>%  
arrange(N_GEOWGS84, Ø_GEOWGS84)

farms_nor <- sub_norway %>% 
  distinct(lon=Ø_GEOWGS84, lat=N_GEOWGS84, iso3c="NOR") %>% # Counting duplicate liscense locations once
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_nor, here("marine/salmon/salmon_farm/data/farm_sites/farms_nor.csv"))

```

# Scotland

Based on these data for the UK region, it appears that salmon aquaculture is occurring only in Scotland.

```{r}

## another dataset for entire UK region
## want to make sure salmon aquaculture is occurring only in Scottland in UK
test <- st_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/UK/EMOD/EMODnet_HA_Aquaculture_Marine_Finfish_20190104.shp") %>%
  filter(COUNTRY == "United Kingdom",
         SALMON == "Yes", 
         STATUS == "Active")
## based on plot, this is all in Scottland.

mapview(test)
```

Source: 
http://aquaculture.scotland.gov.uk/
```{r}
## get scotland data: http://aquaculture.scotland.gov.uk/
# Load in data and select sites that have been producing in the last 3 years

#### TODO #### This could probably be improved by setting up a way to webscrape these data... not sure how this was compiled previously.

scotland_csv <- as.data.frame(read_csv(file.path(path, "Scotland/Scotland_site_details.csv")))

# Subset to salmon farms
sub_scotland <- scotland_csv %>% 
  dplyr::select(ID = `Marine Scotland Site ID`, Easting, Northing, `Producing in Last 3 Years`, `Aquaculture Type`, `Water Type`, Species, subregion = `Local Authority` ) %>% 
  filter(`Producing in Last 3 Years` == c("Yes", "New"),
         `Water Type` %in% c("Seawater", "Freshwater and Seawater"), 
         grepl("Salmon", Species))

new_names <- data.frame(subregion = c("Argyll and Bute", "North Ayrshire", "Western Isles", "Shetland", "Orkney", "Highland"),
                         subregion_new = c("Argyll & Clyde", "Argyll & Clyde", "Outer Hebrides", "Shetland Isles", "Orkney Islands", 
                                       "North Coast & West Highlands"))

sub_scotland <- sub_scotland %>%
  left_join(new_names, by="subregion") %>%
  dplyr::select(Easting, Northing, subregion = subregion_new)

coordinates(sub_scotland) <- cbind(sub_scotland$Easting, sub_scotland$Northing)
crs(sub_scotland) <- "+init=epsg:7405"
scotland <- spTransform(sub_scotland, CRS('+init=epsg:4326'))

scotland <- st_as_sf(scotland) %>% st_zm()

farms_scotland <- data.frame(do.call(rbind, st_geometry(scotland))) 
names(farms_scotland) <- c("lon", "lat")

farms_scotland$subregion <- sub_scotland$subregion
farms_scotland <- farms_scotland %>%
  mutate(iso3c = "GBR") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_scotland, here("marine/salmon/salmon_farm/data/farm_sites/farms_scotland.csv"))

```

# Faroe Islands

We consider Faroe Islands as a single region. We have data from ASC to use for our farms.

Source: 
Aquaculture Stewardship Council Data Mapping Study, n.d. . Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}
ASC_salmon_sf <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

fro_salmon_sf <- ASC_salmon_sf %>%
  dplyr::filter(crt_cnt == "Faroe Islands")

mapview(fro_salmon_sf)

fro_pts <- st_centroid(fro_salmon_sf)

fro_pts_latlon <- st_transform(fro_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry) %>%
  mutate(iso3c = "FRO")

farms_fro <- data.frame(do.call(rbind, st_geometry(fro_pts_latlon))) 
names(farms_fro) <- c("lon", "lat")
farms_all_fro <- farms_fro %>%
  cbind(iso3c = fro_pts_latlon$iso3c) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_all_fro, here("marine/salmon/salmon_farm/data/farm_sites/farms_fro.csv"))

```

# Australia

Data is from the national map (https://nationalmap.gov.au/#share=s-9CWx9q1BlcDp56T59yiCGgZDU3Q). Instructions on how to access raw data are here: https://nationalmap.gov.au/help/help.html.

Julia Blanchard's RA helped extract the data by manually going through the map and selecting individual farm leases to record their locations, then reading the DPIPWE documentation to identify the species farmed at each location.

According to this source: https://data.gov.au/dataset/ds-dga-c12a0597-b535-48b3-a602-1f8382d8dae6/details
In 2015-2016, >98% of salmonid aquaculture was in Tasmania.  So limiting to Tasmania is probably fine.

TODO: 
There is probably a way to automate this by web scraping, like we did for Australia tuna aquaculture.

Source: 
Tasmanian Government Department of Primary Industries, Parks, Water and Environment, 2019. Sea fishing & aquaculture. [Accessed Jun 2019] https://dpipwe.tas.gov.au/sea-fishing-aquaculture/marine-farming-aquaculture

Australian Government National Map, 2019. Marine framing zones. [accessed June 2019]
https://nationalmap.gov.au/#share=s-9CWx9q1BlcDp56T59yiCGgZDU3Q

```{r}

#Tasmania
tas_csv <- read.csv(file = file.path(path, "/Australia/NationalMap_AquacultureFarms.csv"), header = T)

farms_tas <- tas_csv %>% 
  filter(species_farmed1 == "salmon" | species_farmed2 == "salmon" | species_farmed3 == "salmon") %>%
  dplyr::select(lon, lat) %>%
  mutate(iso3c="AUS") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_tas, here("marine/salmon/salmon_farm/data/farm_sites/farms_tas.csv"))

```

# Canada
TODO: This could probably updated with more current data.

An annual average of 122,300 tonnes of salmon was produced in the last five years (2011-2015). Salmon has represented an average of 70% of total Canadian aquaculture volume. British Columbia’s salmon production represents over 60% of this total. (https://www.dfo-mpo.gc.ca/aquaculture/sector-secteur/species-especes/salmon-saumon-eng.htm).

These data suggest production occurs in: Nova Scotia, New Brunswick, and BC.

British Colombia data with species information: https://catalogue.data.gov.bc.ca/dataset/saltwater-finfish-tenures-coastal-resource-information-management-system

PEI data is available, but no salmon: http://www.arcgis.com/home/webmap/viewer.html?webmap=16aa8830c7084a8a92ce066b525978b4

Newfoundland and Labrador data was extracted from figures found on the web: https://www.fishaq.gov.nl.ca/aquaculture/index.html
(sites are for 2015)

## British colombia

Source: 
Saltwater Finfish Tenures - Coastal Resource Information Management System (CRIMS) , 2019. Ministry of Forests, Lands, Natural Resource Operations and Rural Development. [Accessed January 17, 2020] https://catalogue.data.gov.bc.ca/dataset/saltwater-finfish-tenures-coastal-resource-information-management-system

```{r}

bc <- st_read(file.path(path, "Canada/CRIMS_S_WATER_FINFISH_TENURES/WAT_FNF_TN_polygon.shp"))

bc_pts <- st_centroid(bc)

bc_pts_latlon <- st_transform(bc_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry)

farms_bc <- data.frame(do.call(rbind, st_geometry(bc_pts_latlon))) 
names(farms_bc) <- c("lon", "lat")
farms_bc <- farms_bc %>%
  mutate(iso3c="CAN") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_bc, here("marine/salmon/salmon_farm/data/farm_sites/farms_bc_can.csv"))

```

## New Brunswick
Source: 
NB Agriculture, Aquaculture and Fisheries, 2019. Finfish. [Accessed Oct 2019]
https://hub.arcgis.com/datasets/DAAF-MAAP::masmps-data

```{r}

nb <- st_read(file.path(path,"/Canada/New_brunswick_Finfish_mol.shp"))
mapview(nb)
nb_pts <- st_centroid(nb)

nb_pts_latlon<-st_transform(nb_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry)

farms_nb <- data.frame(do.call(rbind, st_geometry(nb_pts_latlon))) 
names(farms_nb) <- c("lon", "lat")
farms_nb <- farms_nb %>%
  mutate(iso3c="CAN") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_nb, here("marine/salmon/salmon_farm/data/farm_sites/farms_nb_can.csv"))

```

## Nova Scotia
Originally tried using points along the entire coastline.  But this was far too many points.  Here is a good site:
https://www.nsgc.gov.ns.ca/fishaqua/provinceSalmon.htm (other aquaculture species also available).

Obtained data from here: https://data.novascotia.ca/Fishing-and-Aquaculture/Aquaculture-License-and-Lease-GIS-Database/h57h-p9mm
(Feb 18 2020)

Source: 
Nova Scotia Marine Aquaculture Leases , 2020. . Nova Scotia Fisheries and Aquaculture. [Accessed February 18, 2020] https://data.novascotia.ca/Fishing-and-Aquaculture/Nova-Scotia-Marine-Aquaculture-Leases/h57h-p9mm

```{r}

ns <- st_read(file.path(path,"Canada/Nova_Scotia_Marine_Aquaculture_Sites/geo_export_fcc6f2ef-82bb-4445-98f0-a82a56cd0d00.shp"))
mapview(ns)
# id salmon
salmon <- grep("Salmon|salmon", unique(ns$aqua_speci), value=TRUE)

ns_salmon <- ns %>%
  filter(aqua_speci %in% salmon)

ns_salmon_mol <- st_transform(ns_salmon, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
ns_pts <- st_centroid(ns_salmon_mol)

ns_pts_latlon <- st_transform(ns_pts, crs = "+init=epsg:4326")

mapview(ns_pts_latlon)

farms_ns <- data.frame(do.call(rbind, st_geometry(ns_pts_latlon))) 
names(farms_ns) <- c("lon", "lat")
farms_ns <- farms_ns %>%
  mutate(iso3c="CAN") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_ns, here("marine/salmon/salmon_farm/data/farm_sites/farms_ns_can.csv"))

```

## Newfoundland and Labrador

Source: 
Newfoundland Labrador Fisheries, Forestry and Agriculture, 2015. [Accessed February 18, 2020] https://www.gov.nl.ca/flr/lands/ownership/commercial/aquaculture/

```{r}

new_lab <- read.csv(file.path(path,"Canada/NewFoundland_Labrador/NewFoundland_Labrador_salmon.csv"))

farms_new_lab <- new_lab %>%
  mutate(iso3c = "CAN") %>%
  dplyr::select(lon, lat, iso3c) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_new_lab, here("marine/salmon/salmon_farm/data/farm_sites/farms_new_lab_can.csv"))

```

# United States
Salmon aquaculture occurs along the Atlantic coast and in Puget sound.

Puget Sound data (2018 data): http://www.moldychum.com/violations-prompt-washington-state-cancel-atlantic-salmon-farm-lease/

Source:

## Puget Sound

Source: 
Our Sound, Our Salmon, 2017. [Accessed January 31, 2020] http://www.moldychum.com/violations-prompt-washington-state-cancel-atlantic-salmon-farm-lease/

 - this will receive a 0.667 (good quality), since I do not believe it is comprehensive, and is only a sample of points.

```{r}

farms_ps <- read.csv(file.path(path, "USA/PugetSound/USA_PugetSound_AtSalmon.csv")) %>%
  mutate(iso3c = "USA") %>%
  dplyr::select(lon, lat, iso3c) %>%
  mutate(data_quality = 4) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_ps, here("marine/salmon/salmon_farm/data/farm_sites/farms_ps_usa.csv"))

```

## US Northeast
Source:
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 17, 2020] https://www.northeastoceandata.org/data-download/

```{r}

us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))

# id salmon and active sites
salmon_species <- grep("Salmon|salmon", unique(us_ne$allSpecies), value=TRUE)

us_ne_salmon <- us_ne %>%
  filter(allSpecies %in% salmon_species) %>%
  filter(status == "Active")

us_ne_salmon_mol <- st_transform(us_ne_salmon, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
us_ne_salmon_pts <- st_centroid(us_ne_salmon_mol)
# check
plot(st_geometry(us_ne_salmon_mol))
plot(st_geometry(st_centroid(us_ne_salmon_mol)), add=TRUE)

us_ne_salmon_pts_latlon <- st_transform(us_ne_salmon_pts, crs = "+init=epsg:4326")

farms_us_ne <- data.frame(do.call(rbind, st_geometry(us_ne_salmon_pts_latlon))) 
names(farms_us_ne) <- c("lon", "lat")
farms_us_ne <- farms_us_ne %>%
  mutate(iso3c = "USA") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_us_ne, here("marine/salmon/salmon_farm/data/farm_sites/farms_ne_usa.csv"))

```

# New Zealand

We use the data from ASC for our New Zealand farms.

Source: 
Aquaculture Stewardship Council Data Mapping Study, n.d. . Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}
ASC_salmon_sf <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

nzl_salmon_sf <- ASC_salmon_sf %>%
  dplyr::filter(crt_cnt == "New Zealand") %>%
  dplyr::filter(spcs__c == "Salmon")

mapview(nzl_salmon_sf)

nzl_pts <- st_centroid(nzl_salmon_sf)

nzl_pts_latlon <- st_transform(nzl_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry) %>%
  mutate(iso3c = "NZL")

farms_nzl <- data.frame(do.call(rbind, st_geometry(nzl_pts_latlon))) 
names(farms_nzl) <- c("lon", "lat")
farms_all_nzl <- farms_nzl %>%
  cbind(iso3c = nzl_pts_latlon$iso3c) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_all_nzl, here("marine/salmon/salmon_farm/data/farm_sites/farms_nz.csv"))

```

# Ireland
Downloaded from here (Jan 31 2020): https://www.agriculture.gov.ie/seafood/engineering/publications/gisdata/

Source:
Ireland Department of Agriculture, Food and the Marine., 2019. Aquaculture licence GIS data. [accessed January 31, 2020]  https://www.agriculture.gov.ie/seafood/engineering/publications/gisdata/

```{r}

ireland <- st_read(file.path(path,"Ireland/AquacultureLicenceGISData23122019/Aqua_Sites_Redacted_23-12-2019.shp"))

ire_salmon <- ireland %>%
  dplyr::filter(Species_Na %in% "Salmon")

ire_pts <- st_centroid(ire_salmon)

ire_pts_latlon <- st_transform(ire_pts, crs = "+init=epsg:4326")

farms_ire <- data.frame(do.call(rbind, st_geometry(ire_pts_latlon))) 
names(farms_ire) <- c("lon", "lat")
farms_ire <- farms_ire %>%
  mutate(iso3c="IRL") %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_ire, here("marine/salmon/salmon_farm/data/farm_sites/farms_ire.csv"))

```

# Russia

https://salmonbusiness.com/russian-aquaculture-harvested-over-over-17000-tonnes-of-fish-last-year/

https://www.intrafish.com/finance/russian-aquaculture-sets-huge-salmon-production-target-as-earnings-spike/2-1-662980

Russian Aquaculture is the only in Russia company farming Atlantic salmon. The company has 49 sites for salmon and trout farming in the Barents and the White Seas and in Karelia (Trout), with annual potential production capacity of around 50,000 metric tons of salmonids.

Appears that most of Russia's salmon aquaculture occurs in the Barents Sea in the Murmansk region.  Trout are in the White Seas and lakes of the Republic of Karelia (not sure if this is freshwater).  There are multiple salmon sites within the region, but we do not have specific locations.

```{r}

# lat long estimates from google earth
farms_russia <- data.frame(lon=c(33.7149),
                    lat=c(69.4953)) %>%
  mutate(iso3c = "RUS") %>%
  mutate(data_quality = 1) %>%
  mutate(action_needed = "country based suitability")

write_csv(farms_russia, here("marine/salmon/salmon_farm/data/farm_sites/farms_russia.csv"))

```

# Japan
We are using the ASC map data for Japan farms.

Source: 
Aquaculture Stewardship Council Data Mapping Study, n.d. Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}
ASC_salmon_sf <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

JPN_salmon_sf <- ASC_salmon_sf %>%
  dplyr::filter(crt_cnt == "Japan") %>%
  dplyr::filter(aq_grop == "salmonids")

mapview(JPN_salmon_sf)

JPN_pts <- st_centroid(JPN_salmon_sf)

JPN_pts_latlon <- st_transform(JPN_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry) %>%
  mutate(iso3c = "JPN")

farms_JPN <- data.frame(do.call(rbind, st_geometry(JPN_pts_latlon))) 
names(farms_JPN) <- c("lon", "lat")
farms_all_JPN <- farms_JPN %>%
  cbind(iso3c = JPN_pts_latlon$iso3c) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_all_JPN, here("marine/salmon/salmon_farm/data/farm_sites/farms_japan.csv"))

```

# Denmark
I've found some good information on farms in Denmark. There is a HELCOM dataset with finfish locations, and there are 20 farms. Other sources indicate there are 19 salmon specific farms in Denmark, which seems close enough to me.

Columns of interest: 
 - geometry to get lat/lon
 - PR_AV_TON : Average tonnes of production per year

Source: 
HELCOM Map and Data Service, n.d. The Baltic Marine Environment Protection Commission. [Accessed October 12, 2020]

```{r}

enc <- readr::guess_encoding(file.path(path, "HELCOM_Baltic/Finfish_mariculture.shp"), n_max = 1000)[,"encoding"] %>% as.character()

# lat long estimates from google earth
farms_denmark <- st_read(file.path(path, "HELCOM_Baltic/Finfish_mariculture.shp"), options = "ENCODING=WINDOWS-1252") %>%
  filter(COUNTRY == "DK")

mapview(farms_denmark$geometry)

dk_pts <- st_centroid(farms_denmark)

dk_pts_latlon <- st_transform(dk_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry, NAME)

dk_subregions <- dk_pts_latlon %>%
  dplyr::select(NAME)

farms_dk <- data.frame(do.call(rbind, st_geometry(dk_pts_latlon))) 
names(farms_dk) <- c("lon", "lat")
farms_dk <- farms_dk %>%
  mutate(iso3c="DNK")

farms_dk <- cbind(farms_dk, dk_subregions ) %>%
  dplyr::select(lon, lat, iso3c, subregion = NAME) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write.csv(farms_dk, here("marine/salmon/salmon_farm/data/farm_sites/farms_denmark.csv"), row.names = FALSE)

#farms_dk <- read_csv(here("aquaculture/marine/salmon/salmon_farm/data/farm_sites/farms_denmark.csv"))
```

# Turkey
There is salmon and trout aquaculture in Turkey.

Page 33 of this indicates there are 12 marine rainbow trout farms in turkey: http://tudav.org/wp-content/uploads/2018/04/marine_aquaculture_in_turkey.pdf

It says that half of those farms (6 farms) are located in Persembe bay. I will allocate the remaining 6 equally across Sinop bay, Sana bay, and Rize bay (2 farms each).

Source: 
Candan, A., Karatas, S., Kücüktas, H., Okumus, I. (Eds.), 2007. Marine aquaculture in Turkey. Türk Denzi Araştırmaları Vakfı, İstanbul.

```{r}

# Sinop: 2 farms, randomly picked locations
# lat 42.011686, lon 35.193134
# 42.015462, 35.168353

# Persembe: 6 farms
# 41.073031, 37.776490
# 41.071171, 37.776501
# 41.103902, 37.788303
# 41.106756, 37.788893
# 41.111485, 37.788643
# 41.114387, 37.788943

# Sana: 2 farms
# 40.991640, 39.812112
# 40.994117, 39.803345

# Rize: 2 farms
# 41.038866, 40.562498,
# 41.036462, 40.558582

farms_turkey <- data.frame(lon=c(35.193, 35.168, 37.776,  37.776, 37.788, 37.789, 37.789, 37.789, 39.812, 39.803, 40.562, 40.558),
                    lat=c(42.012, 42.015, 41.073, 41.071, 41.103, 41.106, 41.111, 41.114, 40.991, 40.994, 41.038, 41.036)) %>%
  mutate(iso3c = "TUR") %>%
  mutate(data_quality = 2) %>%
  mutate(action_needed = "state/province based suitability")

write_csv(farms_turkey, here("marine/salmon/salmon_farm/data/farm_sites/farms_tur.csv"))

```

# Finland

Source: 
HELCOM Map and Data Service, n.d. The Baltic Marine Environment Protection Commission. [Accessed October 12, 2020]

```{r}

enc <- readr::guess_encoding(file.path(path, "HELCOM_Baltic/Finfish_mariculture.shp"), n_max = 1000)[,"encoding"] %>% as.character()

# lat long estimates from google earth
farms_fin <- st_read(file.path(path, "HELCOM_Baltic/Finfish_mariculture.shp"), options = "ENCODING=WINDOWS-1252") %>%
  filter(COUNTRY == "FI")

mapview(farms_fin$geometry)

fi_pts <- st_centroid(farms_fin)

fi_pts_latlon <- st_transform(fi_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry, NAME)

fi_subregions <- fi_pts_latlon %>%
  dplyr::select(NAME)

farms_fi <- data.frame(do.call(rbind, st_geometry(fi_pts_latlon))) 
names(farms_fi) <- c("lon", "lat")
farms_fi <- farms_fi %>%
  mutate(iso3c="FIN")

farms_fi <- cbind(farms_fi, fi_subregions ) %>%
  dplyr::select(lon, lat, iso3c, subregion = NAME) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write.csv(farms_fi, here("marine/salmon/salmon_farm/data/farm_sites/farms_finland.csv"), row.names = FALSE)

farms_fi <- read_csv(here("marine/salmon/salmon_farm/data/farm_sites/farms_finland.csv"))
```

# Sweden 
Source:
HELCOM Map and Data Service, n.d. The Baltic Marine Environment Protection Commission. [Accessed October 12, 2020]

```{r}

enc <- readr::guess_encoding(file.path(path, "HELCOM_Baltic/Finfish_mariculture.shp"), n_max = 1000)[,"encoding"] %>% as.character()

# lat long estimates from google earth
farms_swe <- st_read(file.path(path, "HELCOM_Baltic/Finfish_mariculture.shp"), options = "ENCODING=WINDOWS-1252") %>%
  filter(COUNTRY == "SE") %>%
  dplyr::filter(PRODUCTION %in% c("Rainbow trout (Oncorhynchus mykiss)", "Salmon, Rainbow trout (Oncorhynchus mykiss), Char (salvelinus)", "Salmon, Trout (Salmo trutta)", "Salmon", "Salmon, Trout (Salmo trutta), Rainbow trout (Oncorhynchus mykiss), Char (salvelinus)", "Perch (perca fluviatilis), Pikeperch (Sander lucioperca), Rainbow trout (Oncorhynchus mykiss)", "Trout (Salmo trutta)", "Salmon, Trout (Salmo trutta) - breeding", "Perch (perca fluviatilis), Grayling (thymallus thymallus), Rainbow trout (Oncorhynchus mykiss), Char (salvelinus), tropical fish species"))

mapview(farms_swe$geometry)

unique(farms_swe$PRODUCTION)

sw_pts <- st_centroid(farms_swe)

sw_pts_latlon <- st_transform(sw_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry, NAME)

sw_subregions <- sw_pts_latlon %>%
  dplyr::select(NAME)

farms_sw <- data.frame(do.call(rbind, st_geometry(sw_pts_latlon))) 
names(farms_sw) <- c("lon", "lat")
farms_sw <- farms_sw %>%
  mutate(iso3c="SWE")

farms_sw <- cbind(farms_sw, sw_subregions ) %>%
  dplyr::select(lon, lat, iso3c, subregion = NAME) %>%
  filter(!is.na(lat)) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write.csv(farms_sw, here("marine/salmon/salmon_farm/data/farm_sites/farms_sweden.csv"), row.names = FALSE)

#farms_sw <- read_csv(here("marine/salmon/salmon_farm/data/farm_sites/farms_sweden.csv"))
```

## Iceland

We are using the ASC data for this:

Source: 
Aquaculture Stewardship Council Data Mapping Study, n.d. . Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}
ASC_salmon_sf <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

ISL_salmon_sf <- ASC_salmon_sf %>%
  dplyr::filter(crt_cnt == "Iceland")

mapview(ISL_salmon_sf)

ISL_pts <- st_centroid(ISL_salmon_sf)

ISL_pts_latlon <- st_transform(ISL_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry) %>%
  mutate(iso3c = "ISL")

farms_ISL <- data.frame(do.call(rbind, st_geometry(ISL_pts_latlon))) 
names(farms_ISL) <- c("lon", "lat")
farms_all_ISL <- farms_ISL %>%
  cbind(iso3c = ISL_pts_latlon$iso3c) %>%
  mutate(data_quality = 5) %>%
  mutate(action_needed = "Land mask")

write_csv(farms_all_ISL, here("marine/salmon/salmon_farm/data/farm_sites/farms_isl.csv"))

```

# Join cell count dataframes

```{r}

farm_files <- list.files(here("marine/salmon/salmon_farm/data/farm_sites"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

all_farms_1 <- all_farms %>%
  dplyr::select(-data_quality, -action_needed)

write.csv(all_farms_1, here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon.csv"), row.names = FALSE)

write.csv(all_farms, here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon_quality.csv"), row.names = FALSE)

# all_farms <- read_csv(here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon_quality.csv"), col_types = "ddccd")

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

unique(all_farms$iso3c)
```