### 2021-02-22 marine/explore_spatial_allocation.Rmd
finished calculating distance to shore for each mariculture practice

---
title: "Explore spatial allocation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script explores the distribution/data quality of aquaculture within each country.

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

source(here::here("_spatial/template_raster.R"))

gordon_coast <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/outline_st.shp"))
```

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

Read in and summarise all data quality farm data 
```{r}
bivalves <- read_csv(here::here("marine/bivalve/bivalve_farms/data/global_bivalve_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "bivalve") %>%
  dplyr::select(-X1)

shrimp <- read_csv(here::here("marine/shrimp/shrimp_farms/data/global_shrimp_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "shrimp") %>%
  dplyr::select(-area_ha, -prod)

marine_fish <- read_csv(here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "marine_fish_general") %>%
  dplyr::select(-rgn_production)

salmon <- read_csv(here::here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon_quality.csv")) %>%
  mutate(group = "salmon")

crustaceans <- read_csv(here::here("marine/crustaceans/crustacean_farms/data/global_crustacean_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "crustaceans") %>%
  dplyr::select(-prod)

tuna <- read_csv(here::here("marine/tuna/tuna_farms/data/global_tuna_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "tuna") %>%
  mutate(subregion = NA)

all_farms <- rbind(shrimp, bivalves, marine_fish, salmon, crustaceans, tuna)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

## find average distance to shore per each spp group

## do bivalves 
datalist_biv = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "bivalve")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows) %>%
    filter(row_number > 2322) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_biv[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_biv = do.call(rbind, datalist_biv)

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_biv, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/bivalve_dist.shp"))

## do tunas 
datalist_tuna = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "tuna")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows) #%>%
    #filter(row_number > 2322) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_tuna[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_tuna = do.call(rbind, datalist_tuna)

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_tuna, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/tuna_dist.shp"))

## do marine fish  
datalist_mf = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "marine_fish_general")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows) #%>%
    #filter(row_number > 2322) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_mf[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_mf = do.call(rbind, datalist_mf)

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_mf, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/marine_fish_general_dist.shp"))

## do salmon 
datalist_salmon = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "salmon")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows)#  %>%
# filter(row_number < 2345, row_number > 2235) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 2235 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_salmon[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_salmon = do.call(rbind, datalist_salmon)
## ~100 per hour = there are 3400 rows... = 34 hours to run....

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_salmon, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/salmon_dist.shp"))

## do shrimp
datalist_shrimp = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "shrimp")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
     mutate(row_number = 1:num_rows)
    # filter(row_number > 3464) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_shrimp[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

  dist_data_shrimp = do.call(rbind, datalist_shrimp)

st_write(dist_data_shrimp, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/shrimp_dist.shp"))

## there are no crustaceans that fall in that category
```

```{r}
n_farms <- all_farms %>%
  group_by(iso3c, data_quality, group) %>%
  summarise(n_farms = n()) %>%
  ungroup() %>%
  group_by(iso3c) %>%
  mutate(avg_quality = mean(data_quality)) %>%
  ungroup()

## n_farms shows us the number of points that will need spatial allocation within each country according to our new rules (any of the rows that score 3. Those that score 1 or 2, will need some sort of suitability analysis.)

n_countries <- n_farms %>%
  group_by(data_quality, group) %>%
  summarise(n_countries = n()) %>%
  ungroup() %>%
  group_by(group) %>%
  mutate(total_countries = sum(n_countries)) %>%
  ungroup() %>%
  mutate(percent = n_countries/total_countries)

## n_countries shows us the number of countries that will need spatial allocation within each species group according to our new rules (any of the rows that score 3 in the data quality column. Those that score 1 or 2, will need some sort of suitability analysis, based on country or provice/state.)

n_countries_allocated <- n_countries %>%
  dplyr::filter(data_quality == 3)
sum(n_countries_allocated$n_countries)
## 15 countries/species group combos that will need spatial allocation according to our rules (7-10km to port, depth requirements)

n_countries_test <- all_farms_sf %>%
  filter(data_quality == 3) %>%
  group_by(iso3c, action_needed, data_quality, group, subregion) %>%
  summarise(count = n())

n_countries_no_data <- n_countries %>%
  dplyr::filter(data_quality == 1)
sum(n_countries_no_data$n_countries) # 15
## 15 countries/species group combos that will need country level spatial allocation according to our rules (7-10km to port, depth requirements)

n_states_no_data <- n_countries %>%
  dplyr::filter(data_quality == 2)
sum(n_states_no_data$n_countries) # 10
## 14 countries/species group combos that will need subregional spatial allocation according to our rules (7-10km to port, depth requirements)

## these were the rules for the data quality column: 
# For each point, we will add a column for "data quality". It will be scored as very good (5), good (4), fair (3), bad (2), and very bad (1). 
#  5: Very Good qualifications - Locations taken from leases, or from a map with specific locations
#    - Must not intersect with land mask
#    - "Land mask"
# 
# 4: Good qualifications - Locations are projected or it is a only sample of total locations
#    - Must not intersect with land mask
#    - "Land mask"
# 
# 3: Fair qualifications - A specific number of farms are placed on the coast, without knowing exact locations. 
#    - Must be allocated spatially
#    - Must not intersect with land mask
#    - Flagged as "allocate spatially"
# 
# 2: Bad qualifications -  We know that aquaculture is limited to a specific subregion/state, but do not know how many or where the farms are. 
#    - Must be allocated using suitability analysis.
#    - Flagged as 2 and "state/province suitability"
# 
# 1: Very Bad qualifications - No data at all.
#    - Must be allocated using suitability analysis.
#    - Flagged as 1 and "country suitability"  
```