### 2021-03-09 marine/explore_spatial_allocation.Rmd
made a dataset for number of farms modeling

---
title: "Explore spatial allocation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script explores the distribution/data quality of aquaculture within each country.

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)
library(countrycode)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

source(here::here("_spatial/template_raster.R"))

gordon_coast <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/outline_st.shp"))
```

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

Read in and combine all farm data 
```{r}
bivalves <- read_csv(here::here("marine/bivalve/bivalve_farms/data/global_bivalve_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "bivalve") %>%
  dplyr::select(-X1)

shrimp <- read_csv(here::here("marine/shrimp/shrimp_farms/data/global_shrimp_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "shrimp") %>%
  dplyr::select(-area_ha, -prod)

marine_fish <- read_csv(here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "marine_fish_general") %>%
  dplyr::select(-rgn_production)

salmon <- read_csv(here::here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon_quality.csv")) %>%
  mutate(group = "salmon")

crustaceans <- read_csv(here::here("marine/crustaceans/crustacean_farms/data/global_crustacean_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "crustaceans") %>%
  dplyr::select(-prod)

tuna <- read_csv(here::here("marine/tuna/tuna_farms/data/global_tuna_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "tuna") %>%
  mutate(subregion = NA)

all_farms <- rbind(shrimp, bivalves, marine_fish, salmon, crustaceans, tuna)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

```

Read in and combine all FAO tonnage data per country
```{r}
bivalves_tonnes <- read_csv(here::here("marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve.csv")) %>%
  mutate(group = "bivalve")

shrimp_tonnes <- read_csv(here::here("marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp.csv")) %>%
  mutate(group = "shrimp") %>%
  filter(year == 2017)

marine_fish_tonnes <- read_csv(here::here("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv")) %>%
  mutate(group = "marine_fish_general") %>%
  filter(year == 2017)

salmon_tonnes <- read_csv(here::here("marine/salmon/salmon_farm/data/fao_mariculture_salmon.csv")) %>%
  mutate(group = "salmon") %>%
  filter(year == 2017)

crustaceans_tonnes <- read_csv(here::here("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans.csv")) %>%
  mutate(group = "crustaceans") %>%
  filter(year == 2017)

tuna_tonnes <- read_csv(here::here("marine/tuna/tuna_farms/data/fao_mariculture_tuna.csv")) %>%
  mutate(group = "tuna") %>%
  filter(year == 2017)

all_species_country_tonnes <- rbind(bivalves_tonnes, shrimp_tonnes, marine_fish_tonnes, salmon_tonnes, crustaceans_tonnes, tuna_tonnes)
```

```{r}
n_farms <- all_farms %>%
  group_by(iso3c, data_quality, group) %>%
  summarise(n_farms = n()) %>%
  ungroup()

## n_farms shows us the number of points that will need spatial allocation within each country according to our new rules (any of the rows that score 3. Those that score 1 or 2, will need some sort of suitability analysis.)

## make a dataset with the country, species type (farm type), tonnage, and # of farms, for those that score data quality %in% c(3,4,5)
sum(all_species_country_tonnes$fao_tonnes_production) # 29623515

n_farms_final <- n_farms %>%
  filter(data_quality %in% c(3,4,5)) %>%
  group_by(iso3c, group) %>%
  summarise(n_farms = sum(n_farms)) %>%
  left_join(all_species_country_tonnes, by = c("iso3c", "group")) %>%
  dplyr::select(iso3c, country, year, group, n_farms, fao_tonnes_production)
sum(n_farms_final$fao_tonnes_production) # 26812832

n_farms_tonnes_no_info <- n_farms %>%
  filter(data_quality %in% c(1, 2)) %>%
  group_by(iso3c, group) %>%
  summarise(n_farms = sum(n_farms)) %>%
  left_join(all_species_country_tonnes, by = c("iso3c", "group")) %>%
  mutate(n_farms = NA)
sum(n_farms_tonnes_no_info$fao_tonnes_production) # 2810682

26812832 + 2810682 # 29623514 # perfect

n_farms_final_final <- rbind(n_farms_final, n_farms_tonnes_no_info)
sum(n_farms_final_final$fao_tonnes_production) # 29623515

## save n_farms_final to use to make a model of farms and production
write.csv(n_farms_final_final, "data/n_farms_production_data.csv", row.names = FALSE)

```

Now let's do some exploration:

```{r}
n_countries <- n_farms %>%
  group_by(data_quality, group) %>%
  summarise(n_countries = n()) %>%
  ungroup() %>%
  group_by(group) %>%
  mutate(total_countries = sum(n_countries)) %>%
  ungroup() %>%
  mutate(percent = n_countries/total_countries)

test2 <- n_farms %>%
  group_by(data_quality) %>%
  summarise(n_countries = n())

countries <- data.frame(iso3c = unique(all_farms_sf$iso3c)) %>%
  mutate(country = countrycode(iso3c, origin = 'iso3c', destination = 'country.name'))

## n_countries shows us the number of countries that will need spatial allocation within each species group according to our new rules (any of the rows that score 3 in the data quality column. Those that score 1 or 2, will need some sort of suitability analysis, based on country or provice/state.)

n_countries_allocated <- n_countries %>%
  dplyr::filter(data_quality == 3)
sum(n_countries_allocated$n_countries)
## 15 countries/species group combos that will need spatial allocation according to our rules (7-10km to port, depth requirements)

n_countries_test <- all_farms_sf %>%
  filter(data_quality == 3) %>%
  group_by(iso3c, action_needed, data_quality, group, subregion) %>%
  summarise(count = n())

n_countries_no_data <- n_countries %>%
  dplyr::filter(data_quality == 1)
sum(n_countries_no_data$n_countries) # 15
## 15 countries/species group combos that will need country level spatial allocation according to our rules (7-10km to port, depth requirements)

n_states_no_data <- n_countries %>%
  dplyr::filter(data_quality == 2)
sum(n_states_no_data$n_countries) # 10
## 14 countries/species group combos that will need subregional spatial allocation according to our rules (7-10km to port, depth requirements)

## these were the rules for the data quality column: 
# For each point, we will add a column for "data quality". It will be scored as very good (5), good (4), fair (3), bad (2), and very bad (1). 
#  5: Very Good qualifications - Locations taken from leases, or from a map with specific locations
#    - Must not intersect with land mask
#    - "Land mask"
# 
# 4: Good qualifications - Locations are projected or it is a only sample of total locations
#    - Must not intersect with land mask
#    - "Land mask"
# 
# 3: Fair qualifications - A specific number of farms are placed on the coast, without knowing exact locations. 
#    - Must be allocated spatially
#    - Must not intersect with land mask
#    - Flagged as "allocate spatially"
# 
# 2: Bad qualifications -  We know that aquaculture is limited to a specific subregion/state, but do not know how many or where the farms are. 
#    - Must be allocated using suitability analysis.
#    - Flagged as 2 and "state/province suitability"
# 
# 1: Very Bad qualifications - No data at all.
#    - Must be allocated using suitability analysis.
#    - Flagged as 1 and "country suitability"  
```