### 2021-03-23 marine/archive/LAST_STEP_distance_to_shore_calcs.Rmd

---
title: "Distance to shore calcs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script calculates the distance to shore for each aquaculture point that falls under data quality = 4 or data quality = 5.

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

source(here::here("_spatial/template_raster.R"))

gordon_coast <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/outline_st.shp"))
```

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

Read in and summarise all data quality farm data 
```{r}
bivalves <- read_csv(here::here("marine/bivalve/bivalve_farms/data/global_bivalve_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "bivalve") %>%
  dplyr::select(-X1)

shrimp <- read_csv(here::here("marine/shrimp/shrimp_farms/data/global_shrimp_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "shrimp") %>%
  dplyr::select(-area_ha, -prod)

marine_fish <- read_csv(here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "marine_fish_general") %>%
  dplyr::select(-rgn_production)

salmon <- read_csv(here::here("marine/salmon/salmon_farm/data/global_salmon_farm_lat_lon_quality.csv")) %>%
  mutate(group = "salmon")

crustaceans <- read_csv(here::here("marine/crustaceans/crustacean_farms/data/global_crustacean_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "crustaceans") %>%
  dplyr::select(-prod)

tuna <- read_csv(here::here("marine/tuna/tuna_farms/data/global_tuna_farm_lat_lon_data_quality.csv")) %>%
  mutate(group = "tuna") %>%
  mutate(subregion = NA)

all_farms <- rbind(shrimp, bivalves, marine_fish, salmon, crustaceans, tuna)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

## find average distance to shore per each spp group

## do bivalves 
datalist_biv = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "bivalve")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows) %>%
    filter(row_number > 2322) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_biv[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_biv = do.call(rbind, datalist_biv)

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_biv, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/bivalve_dist.shp"))

## do tunas 
datalist_tuna = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "tuna")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows) #%>%
    #filter(row_number > 2322) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_tuna[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_tuna = do.call(rbind, datalist_tuna)

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_tuna, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/tuna_dist.shp"))

## do marine fish  
datalist_mf = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "marine_fish_general")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows) #%>%
    #filter(row_number > 2322) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_mf[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_mf = do.call(rbind, datalist_mf)

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_mf, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/marine_fish_general_dist.shp"))

## do salmon 
datalist_salmon = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "salmon")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
    mutate(row_number = 1:num_rows)#  %>%
# filter(row_number < 2345, row_number > 2235) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 2235 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_salmon[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

dist_data_salmon = do.call(rbind, datalist_salmon)
## ~100 per hour = there are 3400 rows... = 34 hours to run....

## save the first round just in case... it only made it through 1404 rows. Delete the rows already done and continue
st_write(dist_data_salmon, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/salmon_dist.shp"))

## do shrimp
datalist_shrimp = list()

  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "shrimp")

  num_rows <- nrow(group_df)

  group_df_final <- group_df %>%
     mutate(row_number = 1:num_rows)
    # filter(row_number > 3464) ## now delete the rows that have already been completed

num_rows_final <- group_df_final$row_number

for(i in num_rows_final){

    # i = 1411 # testing
  final_df <- group_df_final %>%
    filter(row_number == i)

  datalist_shrimp[[i]] <- cbind(final_df, data.frame(distance = st_distance(final_df, gordon_coast)))

}

  dist_data_shrimp = do.call(rbind, datalist_shrimp)

st_write(dist_data_shrimp, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/shrimp_dist.shp"))

## there are no crustaceans that fall in that category.. we will probably just need to use the shrimp distance for crustaceans
  group_df <- all_farms_sf %>%
    filter(action_needed == "Land mask") %>%
    filter(group == "crustaceans")

## Now compile them together and write as a shp. And calculate some summary stats.

all_farms_bad <- all_farms_sf %>%
  dplyr::filter(action_needed != "Land mask")

biv_dist <- st_read(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/bivalve_dist.shp"))

shrimp_dist <- st_read(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/shrimp_dist.shp"))

tuna_dist <- st_read(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/tuna_dist.shp"))

mf_dist <- st_read(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/marine_fish_general_dist.shp"))

salmon_dist <- st_read(file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/salmon_dist.shp"))

crust_dist <- all_farms_sf %>%
  dplyr::filter(group == "crustaceans") %>%
  dplyr::filter(action_needed == "Land mask") %>%
  dplyr::filter(distance = NA) ## there are none...

all_farms_distance <- rbind(biv_dist, shrimp_dist, tuna_dist, mf_dist, salmon_dist)

## save this file
st_write(all_farms_distance, file.path("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/point_placement/all_farms_distance.shp"))

## Now calculate some summary stats
stats <- all_farms_distance %>%
  mutate(distanc_km = distanc/1000) %>%
  group_by(group) %>%
  summarise(mean_dist = mean(distanc_km), 
            sd_dist = sd(distanc_km),
            max_dist = max(distanc_km), 
            min_dist = min(distanc_km), 
            quant_90 = quantile(distanc_km, 0.9),
            quant_10 = quantile(distanc_km, 0.1), 
            points = n()) %>%
  ungroup()

data.table(stats)

## lets explore some outliers...

shrimp_outliers <- shrimp_dist %>%
  mutate(distanc_km = distanc/1000) %>%
  filter(distanc_km > 10)

mapview(shrimp_outliers) ## it appears all of the shrimp outliers are on land... which makes sense, because marine shrimp ponds are usually on land. I think they are ok. But I wonder if we should only grab those points that dont fall on land for our spatial allocation. Or maybe just place the shrimp ones very close to the shore.

biv_outliers <- biv_dist %>%
  mutate(distanc_km = distanc/1000) %>%
  filter(distanc_km > 10)

mapview(biv_outliers) ## similarly, the largest outliers are on land. Aside from one in New Zealand, but it comes from the ASC dataset, and I think is probably a mistake.. removing it doesn't make much of a difference anyways.

biv_fix <- biv_dist %>%
  filter(distanc < 200000)
mean(biv_fix$distanc)/1000 # 2.502882

## tuna seems fine

mf_outliers <- mf_dist %>%
  mutate(distanc_km = distanc/1000) %>%
  filter(distanc_km > 10)

mapview(mf_outliers) ## similarly, the largest outliers are on land.

salmon_outliers <- salmon_dist %>%
  mutate(distanc_km = distanc/1000) %>%
  filter(distanc_km > 10)

mapview(salmon_outliers) ## similarly, the largest outliers are on land/along waterways/lakes.

mapview(biv_dist)
mapview(shrimp_dist)
mapview(salmon_dist)
mapview(mf_dist)
mapview(tuna_dist)
```