### 2021-03-23 marine/marine_fish_general/finfish_farms/STEP2_finfish_farm_locations.Rmd

---
title: "Location of Marine fish - general Farms"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/08/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script standardizes location of marine fish - general (finfish) farms to be lat/long point data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

fao_finfish_spp <- read_csv("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish_taxa_production.csv") %>%
  dplyr::filter(year == 2017)

fao_finfish <- read_csv("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv") %>%
  dplyr::filter(year == 2017)

unique(fao_finfish$country)

#  [1] "Albania"                  "Australia"                "Bangladesh"               "Cambodia"                
#  [5] "China"                    "China, Hong Kong SAR"     "Costa Rica"               "Croatia"                 
#  [9] "Cyprus"                   "Egypt"                    "France"                   "Greece"                  
# [13] "India"                    "Indonesia"                "Iran (Islamic Rep. of)"   "Israel"                  
# [17] "Italy"                    "Japan"                    "Korea, Republic of"       "Malaysia"                
# [21] "Malta"                    "Mauritius"                "Mexico"                   "Norway"                  
# [25] "Panama"                   "Philippines"              "Portugal"                 "Saudi Arabia"            
# [29] "Singapore"                "Spain"                    "Taiwan Province of China" "Thailand"                
# [33] "Tunisia"                  "Turkey"                   "United Arab Emirates"     "United States of America"
# [37] "Viet Nam"

```

# Raster

Standard raster used for this analysis

```{r}

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

Now we will map the remaining countries with shrimp farms. Workflow will be as follows: 
 - Try to obtain exact locations of farms from published source or dataset (such as FAO or a research paper)
 - If there are not exact locations or comprehensive locations, find the number of farms. 
 - If there is no information at all, note it. We will gapfill later.

For each point, we will add a column for "data quality". It will be scored as very good (5), good (4), fair (3), bad (2), and very bad (1).

5: Comprehensive number of exact farms

4: Subset of exact locations, need to add a specific number of farms to existing locations. 
  - country|species|farm locations
  - country|species|additional points

3: A specific number of farms are known, but no locations 
 - country|species|additional points

2.5: A specific number of farms that are not comprehensive, so we will have to gapfill on top of those, if necessary. 
 - country, species, n_existing_points

2: Exact locations are a subset of total farms. But the number of total farms are unknown. Need to be gapfilled by tonnage.

1: No information on number of farms at all. Need to be gapfilled by tonnage. 
 - country|species|# additional farms (this will be NA)

## Read in ASC data
https://gis.asc-aqua.org/arcgis_app/
Maps from ASC

```{r}
ASC_data <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp"))

ASC_finfish <- ASC_data %>%
  dplyr::filter(aq_grop == "marine_fish_general")

unique(ASC_finfish$crt_cnt)

#  [1] "Turkey"                   "Croatia"                  "Greece"                   "Albania"                 
#  [5] "Spain"                    "Japan"                    "Netherlands"              "Denmark"                 
#  [9] "Australia"                "Mexico"                   "Panama"                   "United States of America"
# [13] "Germany"                  "Viet Nam"                 "Thailand"

## what is in ASC and not in FAO
setdiff(ASC_finfish$crt_cnt, fao_finfish$country)
# [1] "Netherlands" "Denmark"     "Germany"

## what is in FAO and not ASC -- this is what we need to find alternative data sources for (25 countries)
setdiff(fao_finfish$country, ASC_finfish$crt_cnt)
#  [1] "Bangladesh"               "Cambodia"                 "China"                    "Costa Rica"               "Cyprus"                  
#  [6] "Egypt"                    "France"                   "India"                    "Indonesia"                "Iran (Islamic Rep. of)"  
# [11] "Israel"                   "Italy"                    "Korea, Republic of"       "Malaysia"                 "Malta"                   
# [16] "Mauritius"                "Norway"                   "Philippines"              "Portugal"                 "Saudi Arabia"            
# [21] "Singapore"                "Taiwan Province of China" "Tunisia"                  "United Arab Emirates"

```

# Map data from ASC

Data comes from https://gis.asc-aqua.org/arcgis_app/

Source:
Aquaculture Stewardship Council Data Mapping Study, n.d. . Aquaculture Stewardship Council. [Accessed November 11, 2020] 
https://gis.asc-aqua.org/arcgis_app/

```{r}

# get iso3c codes
iso <-   data.frame(iso3c = countrycode(ASC_finfish$crt_cnt, origin = 'country.name', destination = 'iso3c'))

# read in shp
farms_all <- ASC_finfish %>%
  cbind(iso) %>%
  dplyr::filter(!(crt_cnt %in% c("Denmark", "Germany", "Netherlands", "Greece", "Spain"))) ## don't need Spain or greece since it is already represented in EMODnet data

mapview(farms_all$geometry)

unique(farms_all$iso3c)

all_pts <- st_centroid(farms_all)

all_pts_latlon <- st_transform(all_pts, crs = "+init=epsg:4326") %>% 
  dplyr::select(geometry, iso3c)

farms_all_fin <- data.frame(do.call(rbind, st_geometry(all_pts_latlon))) 
names(farms_all_fin) <- c("lon", "lat")
farms_all_fin <- farms_all_fin %>%
  cbind(iso3c = all_pts_latlon$iso3c)

## save farm location
write.csv(farms_all_fin, here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv"), row.names = FALSE)

```

Now we need to map for each country, and assign a flag for gapfilling which will be completed later. I will start with all of the countries list in the ASC data.

## Turkey

Need to add 301 points to existing 22 (a final total of 323 points)

Data is from the ASC so it is a subset.

However, the FAO gives us specific numbers: http://www.fao.org/fishery/countrysector/naso_turkey/en
 - 309 Seabass/seabream farms 
 - 1 seabass farm, 13 seabass and bream and new species

 Data quality = 4

```{r}
farms_tur <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "TUR") %>%
  mutate(data_quality = 4)

write_csv(farms_tur, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_tur.csv"))

## Now we will make a file that tells us the number of additional farms that need to be added to be comprehensive
tur_additional <- data.frame(iso3c = "TUR", type = "marine_fish_general", num_additional_farms = 301, data_quality = 4, total_farms = 323)
write.csv(tur_additional, here("marine/marine_fish_general/finfish_farms/data/subset_known_add_points_4/add_farms_tur.csv"), row.names = FALSE)
```

## Croatia

Need to add 48 points to existing 11 (a final total of 59 points)

Current data is 11 points from ASC

This claims there are 59 total farms in Croatia, so we will add more to the current points there: 
Lukin, A., 2014. Spatial Planning For Marine Aquaculture, Croatia, Zadar County Case. Croatian Ministry of Agriculture, Directorate of Fisheries.

```{r}
farms_hrv <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "HRV") %>%
  mutate(data_quality = 4)

write_csv(farms_hrv, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_hrv.csv"))

## Now we will make a file that tells us the number of additional farms that need to be added to be comprehensive
hrv_additional <- data.frame(iso3c = "HRV", type = "marine_fish_general", num_additional_farms = 48, data_quality = 4, total_farms = 59)
write.csv(hrv_additional, here("marine/marine_fish_general/finfish_farms/data/subset_known_add_points_4/add_farms_hrv.csv"), row.names = FALSE)
```

## Albania

Need to add 22 points to existing 1 (a final total of 23 points)

Current data is from the ASC, which is a subset of farms.

http://www.fao.org/fishery/countrysector/naso_albania/en
“There are 23 marine cage farms, all of them situated in the Ionic Sea. They produce sea bream and sea bass”

```{r}
farms_alb <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "ALB") %>%
  mutate(data_quality = 4)

write_csv(farms_alb, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_alb.csv"))

## Now we will make a file that tells us the number of additional farms that need to be added to be comprehensive
alb_additional <- data.frame(iso3c = "ALB", type = "marine_fish_general", num_additional_farms = 22, data_quality = 4, total_farms = 23)
write.csv(alb_additional, here("marine/marine_fish_general/finfish_farms/data/subset_known_add_points_4/add_farms_alb.csv"), row.names = FALSE)
```

## Japan
Current data is from ASC, only 8 farms, so definitely not comprehensive, just a drop in the bucket compared to the production of Japan. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_jpn <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "JPN") %>%
  mutate(data_quality = 2)

write_csv(farms_jpn, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_jpn.csv"))

```

## Australia 
Current data is from ASC, only 5 farms, so probably not comprehensive. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_aus <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "AUS") %>%
  mutate(data_quality = 2)

write_csv(farms_aus, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_aus.csv"))

```

## Mexico

Current data is from ASC, only 2 farms, so maybe not comprehensive. There are likely more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_mex <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "MEX") %>%
  mutate(data_quality = 2)

write_csv(farms_mex, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_mex.csv"))

```

## Panama
Current data is from ASC, only 1 farms, so probably not comprehensive. There are likely more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_pan <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "PAN") %>%
  mutate(data_quality = 2)

write_csv(farms_pan, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_pan.csv"))

```

## USA
Current data is from ASC, only 1 farms, so definitely not comprehensive. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_usa <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "USA") %>%
  mutate(data_quality = 2)

write_csv(farms_usa, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_usa.csv"))

```

## Viet Nam
Current data is from ASC, only 5 farms, so definitely not comprehensive, just a drop in the bucket compared to the production of Viet Nam. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_vnm <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "VNM") %>%
  mutate(data_quality = 2)

write_csv(farms_vnm, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_vnm.csv"))

```

## Thailand
Current data is from ASC, only 1 farms, so definitely not comprehensive, just a drop in the bucket compared to the production of Thailand. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
farms_tha <- read.csv(here("marine/marine_fish_general/finfish_farms/data/ASC_finfish_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "THA") %>%
  mutate(data_quality = 2)

write_csv(farms_tha, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_tha.csv"))

```

##  Europe - EMODNET: Malta, Greece, Norway, Cyprus, Spain 
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Finfish+Production
https://www.emodnet-humanactivities.eu/view-data.php

Source:
EMODnet Human Activities: Finfish production, 2019. [Accessed December 8, 2020] 
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Finfish+Production

```{r}

## read in shapefile
eu <- st_read(file.path(path,"EMODnet_HA_Aquaculture_Marine_Finfish_20190104/EMODnet_HA_Aquaculture_Marine_Finfish_20190104.shp"))
mapview(eu)

unique(eu$COUNTRY)
# [1] "Spain"          "United Kingdom" "Greece"         "Finland"        "Norway"         "Ireland"        "Malta"          "Denmark"       
# [9] "Cyprus"

unique(eu$FARM_TYPE)

unique(eu$SPECIES_DE)

unique(eu$STATUS)

eu_spp <- eu %>%
  dplyr::filter(str_detect(SPECIES_DE, "Mediterranean marine fish|alibut|eabass|eabream|urbot|Sole|sole|sea bream|sea bass|Sea bream|Sea bass"))

unique(eu_spp$COUNTRY)

## Need to filter for the ones in FAO dataset
# [1] "Spain" - yes          "United Kingdom" - no "Greece" - yes        "Norway" - yes        "Denmark" - no       "Cyprus" - yes  "Malta" - yes

eu_na <- eu_spp %>%
  dplyr::filter(STATUS == "n.a.")
unique(eu_na$COUNTRY)

eu_active <- eu_spp %>%
  dplyr::filter(STATUS == "Active")

unique(eu_active$COUNTRY)

## we will use both "Active" and "n.a." as the status of the farms. I choose this because all of the denmark points are n.a., and they correspond with the HELCOM data that we used for other spp groups. This indicates to me that the farms denoted as "n.a." are probably actually active.

eu_active_na <- eu_spp %>%
  dplyr::filter(STATUS %in% c("Active", "n.a."))

unique(eu_active_na$COUNTRY)

# [1] "Spain"          "United Kingdom" "Greece"         "Malta"          "Norway"         "Denmark"        "Cyprus"

## Now filter for countries that we want: Spain, Greece, Norway, Cyprus, Malta... the others are not included in FAO 2017
eu_fao_countries <- eu_active_na %>%
  dplyr::filter(COUNTRY %in% c("Spain", "Greece", "Norway", "Cyprus", "Malta"))

unique(eu_fao_countries$COUNTRY)
mapview(eu_fao_countries)

eu_pts <- st_centroid(eu_fao_countries)

eu_pts_latlon <- st_transform(eu_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(COUNTRY, geometry)

eu_countries <- eu_pts_latlon %>%
  dplyr::select(COUNTRY)

farms_eu <- data.frame(do.call(rbind, st_geometry(eu_pts_latlon)))
names(farms_eu) <- c("lon", "lat")

farms_eu_final <- cbind(farms_eu, eu_countries) %>%
  dplyr::select(-geometry) %>%
  mutate(iso3c = case_when(
    COUNTRY == "Spain" ~ "ESP",
    COUNTRY == "Cyprus" ~ "CYP", 
    COUNTRY == "Norway" ~ "NOR", 
    COUNTRY == "Greece" ~ "GRC",
    COUNTRY == "Malta" ~ "MLT"
  )) %>%
  dplyr::select(-COUNTRY) %>%
  mutate(data_quality = 5)

write_csv(farms_eu_final, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_ESP_CYP_NOR_GRC_MLT.csv"))

```

## Italy
From http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/italy/en/
 - We will use these as existing farm points, but need to add the remaining farms that are known about.

Source:
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/

Need to add 81 points to existing 9 (a final total of 90 points)

Only a subset of points based on NASO satellite maps

However FAO reports there were 90 finfish farms in 2013...

"Ninety farms (11 percent of the total number of aquaculture farms) operated in 2013. Seabass and seabream as well as other species such as grey mullets and European eel are also cultured under extensive and semi-extensive conditions in coastal areas (e.g. “valli”, ponds and lagoons) of the northern Adriatic regions and in Sardinia, Sicily, Apulia and Lazio."
http://www.fao.org/fishery/countrysector/naso_italy/en

```{r}
## read in thailand shp file
italy_file <- file.path(path, "FAO_NASO/NASO/Point/Italy.kml")

italy_aqua <- read_sf(italy_file)

unique(italy_aqua$description)

mapview(italy_aqua)

## We need to filter for just the appropriate species of finfish aquaculture

italy_finfish <- italy_aqua %>%
  dplyr::filter(!(description %in% c("Floating Cages","Oysters<br>Mussels", "Freshwater farm<br><br>Strugeons<br>Stripped bass"))) %>%
  dplyr::filter(!is.na(description))

mapview(italy_finfish)

ita_bivalve_mol <- st_transform(italy_finfish, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
ita_pts <- st_centroid(ita_bivalve_mol)

ita_pts_latlon <- st_transform(ita_pts, crs = "+init=epsg:4326")

mapview(ita_pts_latlon)

farms_ita <- data.frame(do.call(rbind, st_geometry(ita_pts_latlon))) 
names(farms_ita) <- c("lon", "lat")
farms_ita <- farms_ita %>%
  mutate(iso3c="ITA") %>%
  mutate(data_quality = 4)

write_csv(farms_ita, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_ita.csv"))

farms_ita <-  st_as_sf(farms_ita, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_ita)

## Now we will make a file that tells us the number of additional farms that need to be added to be comprehensive
ita_additional <- data.frame(iso3c = "ITA", type = "marine_fish_general", num_additional_farms = 81, data_quality = 4, total_farms = 90)
write.csv(ita_additional, here("marine/marine_fish_general/finfish_farms/data/subset_known_add_points_4/add_farms_ita.csv"), row.names = FALSE)
```

## France, Israel
http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/mediterranean-basin/en/

 - It says both Tuna and Finfish, but we'll assume france and israel are just finfish...

Source:
NASO aquaculture maps collection , 2010. Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites

 - These are both not comprehensive, and I can't find any information on specific numbers of farms. So both will be labeled as data quality = 2 and need to be gapfilled based on the remaining tonnes.

```{r}
## read in thailand shp file
france_file <- file.path(path, "FAO_NASO/NASO/Point/Mediterranean.kml")

france_aqua <- read_sf(france_file)

med_countries <- france_aqua %>%
  dplyr::mutate(country = sub("_.*", "", Name)) %>%
  dplyr::mutate(country = sub(" fishpen A", "", country))

unique(med_countries$country)

## We need to filter for just the appropriate species of finfish aquaculture, France, Israel

med_finfish <- med_countries %>%
  dplyr::filter(country %in% c("France", "Israel"))

mapview(med_finfish)

fra_finfish_mol <- st_transform(med_finfish, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
fra_pts <- st_centroid(fra_finfish_mol)

fra_pts_latlon <- st_transform(fra_pts, crs = "+init=epsg:4326")

mapview(fra_pts_latlon)

countries <- fra_pts_latlon %>%
  dplyr::select(country)

farms_fra <- data.frame(do.call(rbind, st_geometry(fra_pts_latlon))) 
names(farms_fra) <- c("lon", "lat")
farms_fra <- farms_fra %>%
  cbind(countries) %>%
  mutate(iso3c = ifelse(country == "France", "FRA", "ISR")) %>%
  dplyr::select(-country, -geometry) %>%
  mutate(data_quality = 2)

write_csv(farms_fra, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_fra_isr.csv"))

```

## Indonesia
NASO maps 
 - Per the FAO aquaculture sector report, Indonesia basically only grows grouper/milkfish, so I will assume this is all grouper/milkfish.

Source: 
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/

Current data is a subset, only 6 farms, so definitely not comprehensive, just a drop in the bucket compared to the production of Indonesia. There are likely many more farms than this. Because we can't find any information on more specific locations or specific number of farms, we will gapfill later based on global tonnes/farm. For now we need to mark the data quality as 2.

```{r}
indo_file <- file.path(path, "FAO_NASO/NASO/Point/Indonesia.kml")

indo_aqua <- read_sf(indo_file)

idn_finfish_mol <- st_transform(indo_aqua, crs = "+init=epsg:3035") # convert to equal area projection to find centroids. 
idn_pts <- st_centroid(idn_finfish_mol)

idn_pts_latlon <- st_transform(idn_pts, crs = "+init=epsg:4326")

mapview(idn_pts_latlon)

farms_idn <- data.frame(do.call(rbind, st_geometry(idn_pts_latlon))) 
names(farms_idn) <- c("lon", "lat")
farms_idn <- farms_idn %>%
  mutate(iso3c = "IDN") %>%
  mutate(data_quality = 2)

write_csv(farms_idn, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_idn.csv"))

```

## Portugal
 - https://www.ine.pt/xportal/xmain?xpid=INE&xpgid=ine_publicacoes&PUBLICACOESpub_boui=435690295&PUBLICACOESmodo=2&fbclid=IwAR0-_SHYtw2XGUAXdx71RGer284Vk_u40HYKXyJzeFbpdnsHydKgAfomTvU
 - National Institute of Statistics for Portugal 
 - There is a csv table, "q5_1.csv", which has the number of farms listed by type of system. I will assume that tanks are for finfish, since [this website](https://rea.apambiente.pt/content/aquaculture-production?language=en) says that the tanks are for **fish production**, and floating structures are for **bivalve molluscs**. Located here "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Portugal/EPesca_2019.xlsx"
 - According to the table, by these criteria, there are 112 marine tank farm systems in Portugal. 
 - I will equally allocate 112 points along the coast of Portugal in suitable locations.

Source: 
Estatísticas da Pesca, 2020. Instituto Nacional de Estatística. [Accessed December 21, 2020]

```{r}
farms_prt <- data.frame(iso3c = "PRT", type = "marine_fish_general", num_additional_farms = 112, data_quality = 3)

write_csv(farms_prt, here("marine/marine_fish_general/finfish_farms/data/known_number_farms_3/farms_prt.csv"))

```

## Iran

The only species the FAO reports for Iran is "Marine fishes nei", and I can't find any information on number of marine farms there might be in Iran. This will have to be fully gapfilled.

Source:
NA

```{r}
irn_gf <- data.frame(iso3c = "IRN", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(irn_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_irn.csv"))

```

## China 
Can't find any datasets of marine finfish farm locations in China. This will need to be fully gapfilled.

Source: 
NA

```{r}
chn_gf <- data.frame(iso3c = "CHN", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(chn_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_chn.csv"))

```

## Saudi Arabia
 - There is really only one marine fish aquaculture producer in Saudi Arabia, and it happens to be the same company that produces almost all of its shrimp. It is one of the largest facilities in the world. http://www.naqua.com.sa/barramundi.php
 - I'll place a point there: 20.287253, 40.030864

 - I beleive this is comprehensive.

Source:
National Aquaculture Group, n.d. [Accessed November 12, 2020]. http://www.naqua.com.sa/about-us.php

```{r}

farms_sau <- data.frame(lat = c(20.27293672234808),
                        lon = c(40.03614937279944)) %>%
  mutate(iso3c = "SAU") %>%
  mutate(data_quality = 5)

write_csv(farms_sau, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_sau.csv"))

## visualize it 
farms_sau <-  st_as_sf(farms_sau, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_sau)
```

## Mauritius

Can't find any datasets of marine finfish farm locations. This will need to be fully gapfilled.

Source:
NA

```{r}
mus_gf <- data.frame(iso3c = "MUS", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(mus_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_mus.csv"))

```

## United Arab Emirates

http://www.fishfarm.ae/facilities.html
 - One here in Dibba Al Fujairah. 
 - They do sea bream and sea bass, which matches what the FAO reports.

The above is also corroborated by the FAO here: http://www.fao.org/fishery/countrysector/naso_uae/en

I will place a point off of the coast of Dibba Al Fujairah: 25.604715925918068, 56.340214330776234

Source: 
National Aquaculture Sector Overview. United Arab Emirates, 2016. , National Aquaculture Sector Overview Fact Sheets.

```{r}

farms_are <- data.frame(lat = c(25.604715925918068),
                        lon = c(56.340214330776234)) %>%
  mutate(iso3c = "ARE") %>%
  mutate(data_quality = 5)

write_csv(farms_are, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_are.csv"))

## visualize it 
farms_are <-  st_as_sf(farms_are, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_are)
```

## Tunisia

https://apps.fas.usda.gov/newgainapi/api/report/downloadreportbyfilename?filename=Aquaculture%20Industry%20Small%20but%20Growing_Tunis_Tunisia_5-31-2017.pdf
"At present, there are roughly 50 companies involved in aquaculture production:
 - Shellfish farming: 12 private companies
 - Bluefin tuna fattening: 5 private companies
 - Other marine fish farming: 20 private companies
 - Inland fish farming: 15 private companies"

 "The primary marine aquaculture production area is in the Monastir governorate, along Tunisia’s eastern
shore. Marine aquaculture is practiced using modern and innovative techniques, whether in concrete
facilities with a density of 60 kg/m3
or in floating cages on the high seas. Alevin are exclusively
imported for sea bass and sea bream production while bluefin tuna are imported for fattening."

 - I'll place 20 points along the coast of Monastir in suitable locations.

Source: 
Aquaculture Industry Small but Growing: Tunisia, 2017. Fishery and Agriculture Society, Tunis.

```{r}
farms_tun <- data.frame(iso3c = "TUN", type = "marine_fish_general", num_additional_farms = 20, data_quality = 3)

write_csv(farms_tun, here("marine/marine_fish_general/finfish_farms/data/known_number_farms_3/farms_tun.csv"))
```

## Taiwan
The two main species farmed in Taiwan are Milkfish and Grouper. I will place farms for these two species

Milkfish:
https://digitalarchive.worldfishcenter.org/bitstream/handle/20.500.12348/3610/Pub%20TR4%206.pdf?sequence=1&isAllowed=y
 - I will use the respondents from the survey, and place the appropriate number of points in each location (table located on page 6, I will only use "gatherers"). This is a really old paper, but it is the best I can find. 42 total points.

 - Putai (Budai): 4 points: 
 23.396775208505524, 120.1486974182384
 23.391435122712778, 120.14508600806049
 23.379465182872867, 120.13344924193156
 23.37173018517046, 120.12923593005728

 - Peimen (Beimen): 4 points
 23.328815202320154, 120.11007291797651
 23.30989854415313, 120.10560972235575
 23.286248936869796, 120.09462339467383
 23.24808538030484, 120.08741361713255

 - Chiku (Cigu): 2 points
 23.098351835751856, 120.0340739167051
 23.08684757757559, 120.0323838672442

 - Yungan: 4 points
 22.84129712135721, 120.19480335124462
 22.83607641104833, 120.19703494905501
 22.823893974670515, 120.19737827179506
 22.808703955682557, 120.19960986960544

 - Lingpien (Linbian) : 6 points
 22.427007004796724, 120.48594464702815
 22.42449445979184, 120.48934227493775
 22.419155150689896, 120.4981761075027
 22.413920334638384, 120.50893526254974
 22.41119815232611, 120.51527750131434
 22.40868532131159, 120.51890163775126

 - Taitung City: 6 points
 22.729748162448605, 121.13759861706552
 22.73478926413642, 121.14388413948092
 22.73995620049436, 121.14975973652137
 22.74096436042943, 121.15167272160434
 22.746257078096264, 121.15631854252005
 22.74965943121689, 121.16246742314381

 - Chungkung (Chengkung): 6 points
 23.094448352409827, 121.3724851599623
 23.100673804589537, 121.38612214670725
 23.110708254441576, 121.39450636818748
 23.118419405368904, 121.4097595903985
 23.12715198105829, 121.40864842851558
 23.13662711164484, 121.40026420703533

 - Erhchieh (River delta below Yilan City): 6 points
 24.718398599156366, 121.83377270959099
 24.72276233035146, 121.83292494893261
 24.7123233460693, 121.83961283857079
 24.734911520797553, 121.82878034126954
 24.707959248912765, 121.8381057085115
 24.70256809420227, 121.83829409976889

 - Suao: 4 points
 24.59561929860139, 121.87509035189372
 24.591466187548917, 121.87405621071602
 24.466470314555245, 121.84309342644401
 24.43396678054247, 121.80292466585698

Source:
Lee, C.-S., 1983. Production and marketing of milkfish in Taiwan: an economic analysis.

```{r}
milkfish_points <- data.frame(lat = c(23.396775208505524, 23.391435122712778, 23.379465182872867, 23.37173018517046, 23.328815202320154, 23.30989854415313, 23.286248936869796, 23.24808538030484, 23.098351835751856, 23.08684757757559, 22.84129712135721, 22.83607641104833, 22.823893974670515, 22.808703955682557, 22.427007004796724, 22.42449445979184, 22.419155150689896, 22.413920334638384, 22.41119815232611, 22.40868532131159, 22.729748162448605, 22.73478926413642, 22.73995620049436, 22.74096436042943, 22.746257078096264, 22.74965943121689,  23.094448352409827, 23.100673804589537, 23.110708254441576, 23.118419405368904, 23.12715198105829, 23.13662711164484, 24.718398599156366, 24.72276233035146, 24.7123233460693, 24.734911520797553, 24.707959248912765, 24.70256809420227, 24.59561929860139, 24.591466187548917, 24.466470314555245, 24.43396678054247),

                              lon = c(120.1486974182384, 120.14508600806049, 120.13344924193156, 120.12923593005728, 120.11007291797651, 120.10560972235575, 120.09462339467383, 120.08741361713255, 120.0340739167051, 120.0323838672442, 120.19480335124462, 120.19703494905501, 120.19737827179506, 120.19960986960544, 120.48594464702815, 120.48934227493775, 120.4981761075027, 120.50893526254974, 120.51527750131434, 120.51890163775126, 121.13759861706552, 121.14388413948092, 121.14975973652137, 121.15167272160434, 121.15631854252005, 121.16246742314381, 121.3724851599623, 121.38612214670725, 121.39450636818748, 121.4097595903985, 121.40864842851558, 121.40026420703533, 121.83377270959099, 121.83292494893261, 121.83961283857079, 121.82878034126954, 121.8381057085115, 121.83829409976889, 121.87509035189372, 121.87405621071602, 121.84309342644401, 121.80292466585698))
```

Grouper:
https://www.researchgate.net/publication/49753615_Real-Time_Quantitative_PCR_Assay_for_Monitoring_of_Nervous_Necrosis_Virus_Infection_in_Grouper_Aquaculture
 - Although this is just a sample of grouper farm locations, I will use this and place the points accordingly. Table and map on the second page. 24 total farms.

 - Kunshen: 5 points
 22.963578173481825, 120.16265849315296
 22.958915389973615, 120.16450385288077
 22.958796842534618, 120.16420344548321
 22.95887587417216, 120.16630629726609
 22.962708853156307, 120.16480426027832

 - Budai: 2 points
23.366020730714546, 120.12803212666465
23.344101510137286, 120.11599409273818

 - Linyuan: 4 points
 22.5060215233288, 120.361518692894
 22.495712950055186, 120.37199003646582
 22.489210224135196, 120.38503630058813
 22.478900398348003, 120.40220243759113

 - Jiading: 5 points
 22.910490254668936, 120.17469982002679
 22.900053951559105, 120.17658809509712
 22.88771910346936, 120.18208125893808
 22.87475049020418, 120.18637279318882
 22.85735162493687, 120.18688777729894

 - Cigu: 3 points
23.11902593130365, 120.05874863883409
23.124310527454163, 120.07531112355083
23.130061175003874, 120.08122629666394

 - An-Nan: 2 points
 23.033784880994748, 120.07405573439489
 23.01988231566828, 120.09946161715932

 - Anping: 2 points
 22.990894265701648, 120.14268135694375
 22.969242998130003, 120.15847420298654

 - Linbian: 1 points
 22.412837508670247, 120.5141089682018

Source:
Kuo, H.-C., Wang, T.-Y., Chen, P.-P., Chen, Y.-M., Chuang, H.-C., Chen, T.-Y., 2011. Real-Time Quantitative PCR Assay for Monitoring of Nervous Necrosis Virus Infection in Grouper Aquaculture. Journal of Clinical Microbiology 49, 1090–1096. https://doi.org/10.1128/JCM.01016-10

```{r}
grouper_points <- data.frame(lat = c(23.366020730714546, 23.344101510137286, 23.11902593130365, 23.124310527454163, 23.130061175003874, 22.412837508670247, 22.5060215233288, 22.495712950055186, 22.489210224135196, 22.478900398348003, 22.963578173481825, 22.958915389973615, 22.958796842534618, 22.95887587417216, 22.962708853156307, 22.910490254668936, 22.900053951559105, 22.88771910346936, 22.87475049020418, 22.85735162493687, 23.033784880994748, 23.01988231566828, 22.990894265701648, 22.969242998130003),
                             lon = c(120.12803212666465, 120.11599409273818, 120.05874863883409, 120.07531112355083, 120.08122629666394, 120.5141089682018, 120.361518692894, 120.37199003646582, 120.38503630058813, 120.40220243759113, 120.16265849315296, 120.16450385288077, 120.16420344548321,  120.16630629726609, 120.16480426027832, 120.17469982002679, 120.17658809509712, 120.18208125893808, 120.18637279318882, 120.18688777729894, 120.07405573439489, 120.09946161715932, 120.14268135694375, 120.15847420298654))

```

Combine milkfish and grouper points. These are not comprehensive points, so the rest of the production is going to need to be gapfilled.

```{r}
all_points_twn <- rbind(milkfish_points, grouper_points) %>%
  dplyr::mutate(iso3c = "TWN") %>%
  mutate(data_quality = 2)

farms_twn <-  st_as_sf(all_points_twn, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_twn)

## write to farms folder 
write_csv(all_points_twn, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_twn.csv"))

```

## Costa Rica 
There is only one species farmed in Costa Rice, Spotted Rose Snapper.

https://www.researchgate.net/profile/Angel_Herrera-Ulloa/publication/235638071_Spotted_rose_snapper_Lutjanus_guttatus_aquaculture_research_and_development_as_socio-economic_alternative_for_Costa_Rican_fishing_communities/links/5acd4d66aca2723a333f53f8/Spotted-rose-snapper-Lutjanus-guttatus-aquaculture-research-and-development-as-socio-economic-alternative-for-Costa-Rican-fishing-communities.pdf
 - States that there are 3 pilot farms along the Gulf of Nicoya.

This says that Martec is the only company in the world which farms spotted rose snapper: https://www.seafoodsource.com/news/aquaculture/industrias-martec-plots-expansion-of-its-spotted-rose-snapper-offshore-aquaculture-operation

https://www.martec.co.cr/#aquaculture
 - Also states that they are the only company in the world to produce Spotted Rose Snapper commercially, and their farms are 3 miles off shore.... based on this I will place one point 3 miles offshore of where "Martec Industries" is located on google maps.

9.423478373794573, -

 - This is likely just a subset of total farms, so I will gapfill the remaining production. Data quality = 2

Source:
Martec Industries Aquaculture , n.d. . Martec.  https://www.martec.co.cr/#aquaculture

```{r}
farms_cri <- data.frame(lat = c(9.423478373794573),
                        lon = c(-84.20910189917947)) %>%
  mutate(iso3c = "CRI") %>%
  mutate(data_quality = 2)

write_csv(farms_cri, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_cri.csv"))

## visualize it 
farms_cri <-  st_as_sf(farms_cri, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_cri)
```

## Egypt

There is a map provided by the FAO (2010) here: http://www.fao.org/fishery/countrysector/naso_egypt/en
 - It has both production and locations.... I think I will need to hand place them, but we can use this. 
 - I will only use the ones that have Mullet, Seabass, Seabream, or Meagre, and that are along the coast, using ponds or cages.

 - Mersa Matruh: 31.360400495764903, 27.236078999801336: Production = 5
 - Alexandria: 31.15023941171644, 29.81109569981684: Production = 1521
 - Behira: 31.319036258284083, 30.225935332660324 : Production = 20376
 - Dakahlia: 31.491306679824632, 31.4278324391214 : Production = 5460
 - Damietta: 31.52619128473819, 31.849352783373792 : Production = 55551
 - Port Said: 31.275726202441984, 32.30190584291966 : Production = 16233 
 - Suez: 29.944394112057978, 32.54627145839493 : Production = 750
 - South Sinai: 27.80632012228518, 34.29387671355847 : Production = 30

Source:
FAO Fisheries & Aquaculture - National Aquaculture Sector Overview - Egypt, 2010. , National Aquaculture Sector Overview Fact Sheets.

 - The map shown here is the MAIN aquaculture farms, meaning that there could be more. I will gapfill the rest.

```{r}
farms_egy <- data.frame(lat = c(31.360400495764903, 31.15023941171644, 31.319036258284083, 31.491306679824632, 31.52619128473819, 31.275726202441984, 29.944394112057978, 27.80632012228518),
                        lon = c(27.236078999801336, 29.81109569981684, 30.225935332660324, 31.4278324391214, 31.849352783373792, 32.30190584291966, 32.54627145839493,  34.29387671355847),
                        rgn_production = c(5, 1521, 20376, 5460, 55551, 16233, 750, 30),
                        subregion = c("Mersa Matruh", "Alexandria", "Behira", "Dakahlia", "Damietta", "Port Said", "Suez", "South Sinai")) %>%
  mutate(iso3c = "EGY") %>%
  mutate(data_quality = 2)

write_csv(farms_egy, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_egy.csv"))

## visualize it 
farms_egy <-  st_as_sf(farms_egy, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_egy)
```

## India

There is not much information on locations of aquaculture facilities. This will be fully gapfilled.

http://www.cmfri.org.in/division/mariculture

```{r}
ind_gf <- data.frame(iso3c = "IND", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(ind_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_ind.csv")) 
```

## Singapore
Barramundi and milkfish

https://www.sfa.gov.sg/food-farming/food-farms/farming-in-singapore
 - This says 109 sea based fish farms.
 - Seems like a pretty reliable source.

I'll place 109 farms equally along the coast of Singapore.

Source: 
Food Farms in Singapore, 2019. Singapore Food Agency. [Accessed December 17, 2020] https://www.sfa.gov.sg/food-farming/food-farms/farming-in-singapore

```{r}
farms_sgp <- data.frame(iso3c = "SGP", type = "marine_fish_general", num_additional_farms = 109, data_quality = 3)

write_csv(farms_sgp, here("marine/marine_fish_general/finfish_farms/data/known_number_farms_3/farms_sgp.csv"))
```

## Malaysia

FAO Sector overview:
"The culture of marine fish in floating net-cages in
lagoons and sheltered coastal waters occupies an area of 102 ha, and is concentrated mainly in the western
coast of Peninsular Malaysia."

 - I can only find the broad suitable areas: https://www.dof.gov.my/ 
 - Broad suitable areas https://www.dof.gov.my/index.php/pages/view/78. I'll use these and use the area as a production estimate. Saved onto aurora /home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Malaysia/malaysia_finfish_farms.csv.

Source:
Aquaculture Industrial Zone Area for High Impact Projects (HIP), n.d. . Department of Fisheries Malaysia. [Accessed November 12, 2020] https://www.dof.gov.my/index.php/pages/view/78

 - This is just a subset of farms, since this only represents industrial practices. We will need to gapfill some.

```{r}
malaysia_finfish_loc <- read_csv("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Malaysia/malaysia_finfish_farms.csv") %>%
  clean_names()

```

Farms: 
 - SEMPANG (PT Lot 389, 390), MELAKA : 2.131733588116292, 102.40118069215845
 - KG. AYAU, ROMPIN, PAHANG : 2.8071936098459522, 103.49758024234029
 - PERMATANG PAUH ISLAND, PAHANG : 5.384539285193906, 100.35040003558889
 - KG. SABAK FERI, ROMPIN, PAHANG : 2.8056383383493353, 103.49881667894354
 - SEMPANG (PT Lot 387), MELAKA : 2.1305137929032876, 102.40627304737939

Now lets map the farms 
```{r}
farms_mys <- data.frame(lat=c(2.131733588116292, 2.8071936098459522, 5.384539285193906, 2.8056383383493353, 2.1305137929032876),
                        lon=c(102.40118069215845, 103.49758024234029, 100.35040003558889, 103.49881667894354, 102.40627304737939)) %>%
  mutate(iso3c = "MYS") %>%
  mutate(data_quality = 2) %>%
  mutate(subregion = c("Sempang, Malacca", "Rompin, Pahang", "Permatang Pauh, Pahang", "Rompin, Pahang", "Sempang, Malacca"))

write_csv(farms_mys, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_mys.csv"))

## visualize it 
farms_mys <-  st_as_sf(farms_mys, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_mys)
```

## Bangladesh

There isn't much information. I will place a point randomly on the coast of Bangladesh.

Source:
NA

```{r}
bgd_gf <- data.frame(iso3c = "BGD", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(bgd_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_bgd.csv")) 
```

## Cambodia

The only species is Barramundi.

https://www.researchgate.net/publication/340828970_Asian_seabass_Lates_calcarifer_farming_practices_in_Cambodia
 - Farms are definitely located in Prey Nob, Stoeng Hav, and Thomnob Rolork. 
 - This is the only information I can find on Cambodian Barramundi fishing. 
Source:
Chankakada, M., Sros, K., Ramana, O., Phearum, H., Laining, A., Rimmer, M.A., Somony, T., Viseth, H., 2020. Asian Seabass Farming Practices in Cambodia 5.

 - Although this source is really good, there are no specific locations or numbers. This will need to be fully gapfilled.

```{r}
khm_gf <- data.frame(iso3c = "KHM", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(khm_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_khm.csv")) 
```

## Korea, Republic of (South Korea)

http://www.fao.org/fishery/countrysector/naso_korea/en
 - Lists 612 finfish farms. I will place 612 farms along the coast of South Korea.

Source:
National Aquaculture Sector Overview. Korea, 2005. , National Aquaculture Sector Overview Fact Sheets.

```{r}
farms_kor <- data.frame(iso3c = "KOR", type = "marine_fish_general", num_additional_farms = 612, data_quality = 3)

write_csv(farms_kor, here("marine/marine_fish_general/finfish_farms/data/known_number_farms_3/farms_kor.csv"))
```

## Philippines
The only  species listed is Milkfish. 
There is no exact number listed, however I know that there are a lot of milkfish farms in the Philippines, as it is their oldest aquaculture method. I will just place 1 point randomly along the coast.

Source:
NA

```{r}
phl_gf <- data.frame(iso3c = "PHL", type = "marine_fish_general", num_additional_farms = NA, data_quality = 1)
write_csv(phl_gf, here("marine/marine_fish_general/finfish_farms/data/no_info_1/gf_phl.csv"))

```

## Hong Kong

https://www.afcd.gov.hk/english/fisheries/fish_aqu/fish_aqu_mpo/fish_aqu_mpo.html
From HK government.

This shows a map fish culture zones, fish ponds, and oyster production areas. I will place a point in all of the "Fish culture zones" shown on the map. Most of the fish cages are visible on google maps when you search these locations.

Sha Tau Kok: 22.537974350807968, 114.22465750899093
Ap Chau: 22.549021238266082, 114.27266261841545
Kat O: 22.549243173819686, 114.29054140885648
0 Pui Tong: 22.545901930448537, 114.29912254835972
Sai Lau Kong: 22.524524976629923, 114.28385013462695
Wong Wan: 22.507936457313377, 114.30641749235996
Tap Mun: 22.47796934657302, 114.35412559932351
Kau Lau Wan: 22.459641511472412, 114.360272945696
Sham Wan: 22.4435502213444, 114.35400029960742
Lo Fu Wat: 22.46853714662761, 114.2720207303674
Yung Shue Au: 22.42337637093222, 114.27819733339493
Leung Shuen Wan: 22.349528341786087, 114.35178911904846
Tiu Cham Wan: 22.367924650829952, 114.31997002657292
Tai Tau Chau: 22.368469602475702, 114.3230438982345 
Kai Lung Wan: 22.3691151827537, 114.30143307907038
Kau Sai: 22.341819373881275, 114.31744391160476
Ma Nam Wat: 22.35464650188533, 114.26990023341041
Po Toi O: 22.274992746117093, 114.29389242171153
Po Toi: 22.164051938726015, 114.25252757025574
Sok Kwu Wan: 22.20591578796148, 114.12923505969019
Lo Tik Wan: 22.221779736133797, 114.12958553008323
Ma Wan: 22.352165476146805, 114.05569941361628
Yim Tin Tsai: 22.45358281100093, 114.2079128367368
Cheung Sha Wan: 22.235400816169882, 114.00599898096631
Yim Tin Tsai (East): 22.454708931476908, 114.22148845386273
Tung Lung Chau: 22.25579039550771, 114.28554310187128

Source:
Marine fish culture, pond fish culture and oyster culture, n.d. . Agriculture, Fisheries, and Conservation Department; The Government of the Hong Kong Special Administrative Region.  https://www.afcd.gov.hk/english/fisheries/fish_aqu/fish_aqu_mpo/fish_aqu_mpo.html

```{r}
farms_hkg <- data.frame(lat_lon = c("22.537974350807968, 114.22465750899093", "22.549021238266082, 114.27266261841545", "22.549243173819686, 114.29054140885648", "22.545901930448537, 114.29912254835972", "22.524524976629923, 114.28385013462695", "22.507936457313377, 114.30641749235996", "22.47796934657302, 114.35412559932351", "22.459641511472412, 114.360272945696", "22.4435502213444, 114.35400029960742", "22.46853714662761, 114.2720207303674", "22.42337637093222, 114.27819733339493", "22.349528341786087, 114.35178911904846", "22.367924650829952, 114.31997002657292", "22.368469602475702, 114.3230438982345", "22.3691151827537, 114.30143307907038", "22.341819373881275, 114.31744391160476", "22.35464650188533, 114.26990023341041", "22.274992746117093, 114.29389242171153", "22.164051938726015, 114.25252757025574", "22.20591578796148, 114.12923505969019", "22.221779736133797, 114.12958553008323", "22.352165476146805, 114.05569941361628", "22.45358281100093, 114.2079128367368", "22.235400816169882, 114.00599898096631", "22.454708931476908, 114.22148845386273", "22.25579039550771, 114.28554310187128")) %>%
  separate(lat_lon, into = c("lat", "lon"), sep = ",") %>%
  mutate(iso3c = "HKG") %>%
  mutate(data_quality = 5)

write_csv(farms_hkg, here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5/farms_hkg.csv"))

## visualize it 
farms_hkg <-  st_as_sf(farms_hkg, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_hkg)
```

# Join cell count dataframes

```{r}
farm_files <- list.files(here::here("marine/marine_fish_general/finfish_farms/data/known_farm_sites_2_4_5"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

all_farms_1 <- all_farms %>%
  dplyr::select(-data_quality)

write.csv(all_farms_1, here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon.csv"), row.names = FALSE)

write.csv(all_farms, here::here("marine/marine_fish_general/finfish_farms/data/global_finfish_farm_lat_lon_data_quality.csv"), row.names = FALSE)

```