### 2021-03-23 marine/tuna/tuna_farms/STEP2_tuna_farm_locations.Rmd

---
title: "Location of Tuna farms"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "12/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script standardizes location of tuna farms to be lat/long point data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(stringr)
library(rgdal)
library(PBSmapping)
library(maptools)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(janitor)
library(rnaturalearth)
library(spatialEco)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

fao_tuna_spp <- read_csv("marine/tuna/tuna_farms/data/fao_mariculture_tuna_taxa_production.csv") %>%
  dplyr::filter(year == 2017)
fao_tuna <- read_csv("marine/tuna/tuna_farms/data/fao_mariculture_tuna.csv") %>%
  dplyr::filter(year == 2017)

unique(fao_tuna$country)

# Need to map these 7 countries
# [1] "Australia" "Croatia"   "Japan"     "Malta"     "Mexico"    "Spain"     "Turkey"

```

# Raster

Standard raster used for this analysis

```{r}

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- projectRaster(gadm, food_raster)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

```{r}

gadm_df <- tabularaster::as_tibble(gadm_latlon, value = TRUE, cells = TRUE) %>% 
  mutate(cellvalue = round(cellvalue, 0))

gadm_reg <- dplyr::left_join(gadm_df, gadm_codes, by = c("cellvalue" = "reg_id"))

```

Now we will map the remaining countries with shrimp farms. Workflow will be as follows: 
 - Try to obtain exact locations of farms from published source or dataset (such as FAO or a research paper)
 - If there are not exact locations or comprehensive locations, find the number of farms. 
 - If there is no information at all, note it. We will gapfill later.

For each point, we will add a column for "data quality". It will be scored as very good (5), good (4), fair (3), bad (2), and very bad (1).

5: Comprehensive number of exact farms

4: Subset of exact locations, need to add a specific number of farms to existing locations. 
  - country|species|farm locations
  - country|species|additional points

3: A specific number of farms are known, but no locations 
 - country|species|additional points

2.5: A specific number of farms that are not comprehensive, so we will have to gapfill on top of those, if necessary. 
 - country, species, n_existing_points

2: Exact locations are a subset of total farms. But the number of total farms are unknown. Need to be gapfilled by tonnage.

1: No information on number of farms at all. Need to be gapfilled by tonnage. 
 - country|species|# additional farms (this will be NA)

##  Europe - EMODNET: Spain, Malta
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Finfish+Production
https://www.emodnet-humanactivities.eu/view-data.php

Source:
EMODnet Human Activities: Marine Finfish Aquaculture, 2019. [Accessed December 8, 2020] 
https://www.emodnet-humanactivities.eu/search-results.php?dataname=Finfish+Production

 - Comprehensive

```{r}

## read in shapefile
eu <- st_read(file.path(path,"EMODnet_HA_Aquaculture_Marine_Finfish_20190104/EMODnet_HA_Aquaculture_Marine_Finfish_20190104.shp"))
mapview(eu)

unique(eu$COUNTRY)
# [1] "Spain"          "United Kingdom" "Greece"         "Finland"        "Norway"         "Ireland"        "Malta"          "Denmark"       
# [9] "Cyprus"

unique(eu$FARM_TYPE)

unique(eu$SPECIES_DE)

unique(eu$STATUS)

eu_spp <- eu %>%
  dplyr::filter(str_detect(SPECIES_DE, "Tuna|tuna"))

unique(eu_spp$COUNTRY)
# [1] "Spain" "Malta"

eu_na <- eu_spp %>%
  dplyr::filter(STATUS == "n.a.")
unique(eu_na$COUNTRY)

eu_active <- eu_spp %>%
  dplyr::filter(STATUS == "Active")

unique(eu_active$COUNTRY)

## we will use both "Active" and "n.a." as the status of the farms. I choose this because all of the denmark points are n.a. for salmon and bivalve, and they correspond with the HELCOM data that we used for other spp groups. This indicates to me that the farms denoted as "n.a." are probably actually active.

eu_active_na <- eu_spp %>%
  dplyr::filter(STATUS %in% c("Active", "n.a."))

unique(eu_active_na$COUNTRY)

# [1] "Spain"     "Malta"

## Now filter for countries that we want: Spain, Greece, Norway, Cyprus, Malta... the others are not included in FAO 2017
eu_fao_countries <- eu_active_na %>%
  dplyr::filter(COUNTRY %in% c("Spain", "Malta"))

unique(eu_fao_countries$COUNTRY)
mapview(eu_fao_countries)

eu_pts <- st_centroid(eu_fao_countries)

eu_pts_latlon <- st_transform(eu_pts, crs = "+init=epsg:4326") %>%
  dplyr::select(COUNTRY, geometry)

eu_countries <- eu_pts_latlon %>%
  dplyr::select(COUNTRY)

farms_eu <- data.frame(do.call(rbind, st_geometry(eu_pts_latlon)))
names(farms_eu) <- c("lon", "lat")

farms_eu_final <- cbind(farms_eu, eu_countries) %>%
  dplyr::select(-geometry) %>%
  mutate(iso3c = case_when(
    COUNTRY == "Spain" ~ "ESP",
    COUNTRY == "Malta" ~ "MLT"
  )) %>%
  dplyr::select(-COUNTRY) %>%
  mutate(data_quality = 5)

 write_csv(farms_eu_final, here("marine/tuna/tuna_farms/data/known_farm_sites_2_4_5/farms_ESP_MLT.csv"))

```

## Croatia, Turkey
http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/mediterranean-basin/en/

Source:
NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/

Croatia: 
Need to add 14 points to existing 1 (a final total of 15 points)

Current data is from NASO maps collection and is only a subset of farms

NASO aquaculture maps collection , 2010. . Food and Agriculture Organization of the United Nations. [Accessed January 17, 2020] http://www.fao.org/fishery/naso-maps/selected-aquaculture-sites/

We need to add 14 farms to match the total listed by this government source: 
Lukin, A., 2014. Spatial Planning For Marine Aquaculture, Croatia, Zadar County Case. Croatian Ministry of Agriculture, Directorate of Fisheries.

Turkey:
Need to add 5 points to existing 3 (a final total of 8 points)

Data is from NASO maps collection, so only a subset

However the FAO gives us specific numbers 
 - http://www.fao.org/fishery/countrysector/naso_turkey/en
 - 8 farms

```{r}
## read in thailand shp file
med_file <- file.path(path, "FAO_NASO/NASO/Point/Mediterranean.kml")

med_aqua <- read_sf(med_file)

med_countries <- med_aqua %>%
  dplyr::mutate(country = sub("_.*", "", Name)) %>%
  dplyr::mutate(country = sub(" fishpen A", "", country))

unique(med_countries$country)

## We need to filter for just the appropriate countries (that we don't already have): Croatia, Turkey

med_tuna <- med_countries %>%
  dplyr::filter(country %in% c("Croatia", "Turkey"))

mapview(med_tuna)
mapview(med_aqua)

med_tuna_mol <- st_transform(med_tuna, crs = "+init=epsg:3035") # convert to equal area projection to find centroids.
med_pts <- st_centroid(med_tuna_mol)

med_pts_latlon <- st_transform(med_pts, crs = "+init=epsg:4326")

mapview(med_pts_latlon)

countries <- med_pts_latlon %>%
  dplyr::select(country)

farms_med <- data.frame(do.call(rbind, st_geometry(med_pts_latlon)))
names(farms_med) <- c("lon", "lat")
farms_med <- farms_med %>%
  cbind(countries) %>%
  mutate(iso3c = ifelse(country == "Croatia", "HRV", "TUR")) %>%
  dplyr::select(-country, -geometry) %>%
  mutate(data_quality = 4)

write_csv(farms_med, here("marine/tuna/tuna_farms/data/known_farm_sites_2_4_5/farms_hrv_tur.csv"))

farms_med <-  st_as_sf(farms_med, coords = c("lon", "lat"),
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_med)

#### Croatia
## Now we will make a file that tells us the number of additional farms that need to be added to be comprehensive
hrv_additional <- data.frame(iso3c = "HRV", type = "tuna", num_additional_farms = 14, data_quality = 4, total_farms = 15)
write.csv(hrv_additional, here("marine/tuna/tuna_farms/data/subset_known_add_points_4/add_farms_hrv.csv"), row.names = FALSE)

#### Turkey 
tur_additional <- data.frame(iso3c = "TUR", type = "tuna", num_additional_farms = 5, data_quality = 4, total_farms = 8)
write.csv(tur_additional, here("marine/tuna/tuna_farms/data/subset_known_add_points_4/add_farms_tur.csv"), row.names = FALSE)

```

## Australia
https://www.pir.sa.gov.au/aquaculture/marine_aquaculture/species
"Southern Bluefin Tuna
The farming of Southern Bluefin Tuna (Thunnus maccoyi) generates the highest farm-gate sales in South Australia’s industry. Tuna farming is unique to South Australia. Its development in 1991 drew attention to the potential of South Australia’s aquaculture. All of Australia’s Southern Bluefin Tuna are farmed in waters off Port Lincoln.

Southern Bluefin Tuna are mostly found in the world’s southern oceans."

This shows leases: https://www.aginsight.sa.gov.au/?Bookmark=Ox473rlD

Source: https://www.pir.sa.gov.au/aquaculture/aquaculture_public_register

 - To download the appropriate data, go here: https://www.aginsight.sa.gov.au/?Bookmark=Ox473rlD. Filter for the appropriate layers, i.e. Tuna leases. Click "Select focus", and draw a circle (or square) around your area of interest. You will then be prompted to download an excel or csv file which will contain the lease holder information, and a url to the lease. 
 - We will scrape the coordinates from each lease document url that is shown in the download from the aquaculture public register to obtain the lat lon points needed.

Source:  
Aquaculture public register - Government of South Australia, 2017. [Accessed November 14, 2020] https://www.pir.sa.gov.au/aquaculture/aquaculture_public_register

```{r}
south_au <- read_csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/Australia/tuna/south_au_leases.csv"))

tuna_leases <- south_au %>%
  dplyr::filter(str_detect(`Registration Category`, "Tuna|Miscellaneous"))

## now we need to look at each link and make sure they are tuna (because some of them are listed as "Miscellaneous")

unique(tuna_leases$`Lease Document`)

# row 1 : tuna
# row 2 : tuna
# row 3 : tuna
# row 4 : tuna
# row 5 : tuna
# row 6 : doesn't indicate tuna, filter out 
# row 7 : tuna
# row 8 : tuna
# row 9 : tuna
# row 10: tuna
# row 11: tuna
# row 12: tuna
# row 13: doesn't indicate tuna, filter out 
# row 14: tuna
# row 15: tuna
# row 16: tuna
# row 17: tuna

## Filter out rows 6 and 13
tuna_leases_df <- tuna_leases[c(-6, -13), ]

tuna_links <- c(tuna_leases_df$`Lease Document`)

#install.packages("rvest")
library("rvest")

coords_list <- list()

for(i in 2:seq_along(tuna_links)){

  # i = 1
# Read the webpage into R
webpage <- read_html(tuna_links[i])

# Parse the webpage for coords
coords <- html_nodes(webpage, "dd div")

# Extract the coords
coord_numbers <- html_text(coords)

coords_df <- as.data.frame(coord_numbers) %>%
  dplyr::filter(str_detect(coord_numbers, "°")) ## filter for coordinates

## Format to lat lon: decimal degrees. Right now they are in degrees mins secs 
lat_lon_df <- do.call("cbind", split(coords_df, rep(c(1, 2), length.out = nrow(coords_df)))) %>%
  dplyr::rename("lat" = "coord_numbers") %>%
  dplyr::rename("lon" = "coord_numbers")

## covert to appropriate coordinate system
lat_lon_df_clean <- data.frame(lapply(lat_lon_df, function(x) {
                gsub("[^0-9.]", " ", x)
             })) %>%
  separate(lat, into = c("lat_d", "lat_m", "lat_s"), sep = "  ") %>%
  separate(lon, into = c("lon_d", "lon_m", "lon_s"), sep = "  ") %>%
  mutate_each(funs(as.numeric)) %>%
  transmute(lat = lat_d + lat_m/60 + lat_s/60^2,
            lon = lon_d + lon_m/60 + lon_s/60^2)

coords_list[[i]] <- lat_lon_df_clean

}

## unlist the list to get our coordinates 
lat_lon_final <- coords_list %>%
    bind_rows() %>%
  mutate(lat = -lat) %>% ## change lat to have a negative in front of it (because it S)
  mutate(iso3c = "AUS") %>%
  mutate(data_quality = 5) 
## now that we have coordinates, place them on a map and save

farms_aus_map <-  st_as_sf(lat_lon_final, coords = c("lon", "lat"),
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_aus_map)

write_csv(lat_lon_final, here("marine/tuna/tuna_farms/data/known_farm_sites_2_4_5/farms_AUS.csv"))
```

## Japan
http://www.fao.org/fishery/countrysector/naso_japan/en
Currently 92 firms are operating the Atlantic bluefin tuna aquaculture, and the total number of aquaculture firms and cages are regulated by the government since 2012.

I'll place 92 farms along Japan's coast

Source:
FAO Fisheries & Aquaculture - National Aquaculture Sector Overview - Japan, 2017. National Aquaculture Sector Overview Fact Sheets.

```{r}
farms_jpn <- data.frame(iso3c = "JPN", type = "tuna", num_additional_farms = 92, data_quality = 3)

write_csv(farms_jpn, here("marine/tuna/tuna_farms/data/known_number_farms_3/farms_jpn.csv"))
```

## Mexico 
https://www.gob.mx/inapesca/acciones-y-programas/atun-aleta-azul
Only located in Baja California and Baja California Sur

 - 8 commercial farms in Baja California
 - 3 commercial farms in Baja California Sur

 Seafood Watch:
 https://www.seafoodwatch.org/recommendation/tuna/tuna-pacific-bluefin-mexico-marine-net-pen?species=386

 "In Mexico, the tuna farming industry is primarily located 15 km north of Ensenada, Baja California, in Salsipuedes Bay, with some farms in Todos Los Santos Bay and others off the coast of Ensenada."

 "Mexico’s current National Aquaculture Charter indicates that 6 aquaculture companies were operating 13 net pens with a maximum capacity of 6,800 MT in 2011 (SAGARPA 2013), though personal communication with an industry expert indicates that the active companies have been consolidated into two."

**Based on this, I will place 13 farms in Salsipuedes Bay.**

Source:
Monterey Bay Aquarium Seafood Watch Report on Pacific Bluefin Tuna Farming in Mexico, 2016.

```{r}

farms_mex <- data.frame(lat = c(31.972882899404116, 31.96836856413186, 31.963417102445995, 31.95861101663323, 31.955698114940013, 31.952930772763118, 31.95030900327055, 31.945939221180044, 31.932974311561416, 31.929769220541903, 31.95016334721594, 31.959630510399904, 31.96764042474878),
                          lon = c(-116.81987407899483, -116.8094027351974, -116.80339458711688, -116.7956698252991, -116.79343822744065, -116.79000499996607, -116.78777340210759, -116.78193691540082, -116.77575710594658, -116.77180889435084, -116.77060726473474, -116.78262356089573, -116.79395321156183), 
                          iso3c = "MEX") %>%
  mutate(data_quality = 5)

farms_mex_map <-  st_as_sf(farms_mex, coords = c("lon", "lat"),
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_mex_map)

write_csv(farms_mex, here("marine/tuna/tuna_farms/data/known_farm_sites_2_4_5/farms_MEX.csv"))

```

# Join cell count dataframes

```{r}
farm_files <- list.files(here::here("marine/tuna/tuna_farms/data/known_farm_sites_2_4_5"), full=TRUE)
all_farms = plyr::ldply(farm_files, read_csv)

all_farms_1 <- all_farms %>%
  dplyr::select(-data_quality)

all_farms_sf <- st_as_sf(all_farms, coords = c("lon", "lat"), 
                         crs = crs(gadm_latlon), agr = "constant")

mapview(all_farms_sf)

write.csv(all_farms_1, here::here("marine/tuna/tuna_farms/data/global_tuna_farm_lat_lon.csv"), row.names = FALSE)

write.csv(all_farms, here::here("marine/tuna/tuna_farms/data/global_tuna_farm_lat_lon_data_quality.csv"), row.names = FALSE)

```