### 2021-03-23 marine_mapping_archive_v1/STEP1_species_groups/STEP2_ASC_data_prep.Rmd

---
title: "Prep ASC data"
output: html_document
---

Prep the aquaculture stewardship data and save as a shp to use for the different species groups.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
library(countrycode)
library(here)
library(stringr)
library(sf)
library(mapview)

path <- "/home/shares/food-systems/Food_footprint/_raw_data"
source(here('marine/salmon/salmon_farm/R/fao_mar_clean.R')) # functions specific to mariculture dealing with compound countries
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/fao_fxn.R') # function for cleaning FAO files
source(here('workflow/common.R'))
```

```{r}
fao <- read.csv(here("data/fao_aquaculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

species_list <- read.csv(here("marine/STEP1_species_groups/int/marine_groups.csv"))

fao_all <- species_list %>%
  # filter(species %in% shrimp_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>% 
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  # dplyr::filter(aq_group == "bivalves") %>%
  # dplyr::filter(year == 2017) # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, aq_group, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, year) %>%
  dplyr::filter(year == 2017)
```

```{r}
ASC_data <- read_csv(file.path(here::here("data/ASC_farms.csv"))) ## downloaded from ASC website

unique(fao_all$aq_group)
# [1] "marine_fish_general" "bivalves"            "salmonids"           "shrimps_prawns"     
# [5] "tuna"                "crustaceans"

unique(ASC_data$species__c)

unique(ASC_data$farm_species)

unique(ASC_data$cert_country)

ASC_classified <- ASC_data %>%
  dplyr::filter(species__c %notin% c("Freshwater Trout", "Pangasius", "Seaweed (Algae)", "Tilapia")) %>%
  mutate(aq_group = case_when(
    species__c %in% c("Shrimp") ~ "shrimps_prawns",
    species__c %in% c("Abalone", "Bivalve") ~ "bivalves",
    species__c %in% c("Salmon") ~ "salmonids",
    species__c %in% c("Seabass, Seabream and Meagre", "Seriola and Cobia", "Tropical Marine Finfish") ~ "marine_fish_general"
  ))

## Now we need to make the ASC data into an sf
ASC_all_sf <- ASC_classified %>%
  mutate(geometry = str_replace_all(geom, "POINT|[()]", "")) %>%
  mutate(geometry = trimws(geometry)) %>%
  mutate(x= as.numeric(gsub(" .*$", "", geometry))) %>%
  dplyr::filter(!is.na(geom)) %>%
  mutate(y = as.numeric(sub("^\\S+\\s+", '', geometry))) %>%
  dplyr::select(-geometry, -geom, -X10, -X11, -X12, -X13, -X14)  %>%
  st_as_sf(coords = c("x", "y"), crs = 4326)

mapview(ASC_all_sf$geometry)

## now we need to write as a shapefile 
# st_write(ASC_all_sf, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/ASC_all_sf.shp", overwrite = TRUE)

st_write(ASC_all_sf, "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/ASC_data/ASC_all_sf.shp", overwrite = TRUE)
```

See how many countries from FAO are represented in ASC for marine_fish_general category

```{r}
fao_marine_fish <- fao_all %>%
  dplyr::filter(aq_group == "marine_fish_general")

ASC_marine_fish <- ASC_all_sf %>%
  dplyr::filter(aq_group == "marine_fish_general")

setdiff(fao_marine_fish$country, ASC_marine_fish$cert_country)

#  [1] "Bangladesh"               "Cambodia"                 "China"                   
#  [4] "China, Hong Kong SAR"     "Costa Rica"               "Cyprus"                  
#  [7] "Egypt"                    "France"                   "India"                   
# [10] "Indonesia"                "Iran (Islamic Rep. of)"   "Israel"                  
# [13] "Italy"                    "Korea, Republic of"       "Malaysia"                
# [16] "Malta"                    "Mauritius"                "Norway"                  
# [19] "Philippines"              "Portugal"                 "Saudi Arabia"            
# [22] "Singapore"                "Taiwan Province of China" "Tunisia"                 
# [25] "United Arab Emirates"

setdiff(ASC_marine_fish$cert_country, fao_marine_fish$country)
# [1] "Netherlands" "Denmark"     "Germany"

unique(ASC_marine_fish$cert_country)
```

See how many countries from FAO are represented in ASC for bivalves category

```{r}
fao_bivalve <- fao_all %>%
  dplyr::filter(aq_group == "bivalves")

ASC_bivalve <- ASC_all_sf %>%
  dplyr::filter(aq_group == "bivalves")

setdiff(fao_bivalve$country, ASC_bivalve$cert_country)

#  [1] "Brazil"                   "Bulgaria"                 "Cambodia"                
#  [4] "Canada"                   "China, Hong Kong SAR"     "Croatia"                 
#  [7] "Cuba"                     "Denmark"                  "French Polynesia"        
# [10] "Germany"                  "Greece"                   "Guernsey"                
# [13] "India"                    "Indonesia"                "Ireland"                 
# [16] "Italy"                    "Jersey"                   "Korea, Dem. People's Rep"
# [19] "Malaysia"                 "Netherlands"              "Norway"                  
# [22] "Philippines"              "Russian Federation"       "Slovenia"                
# [25] "Spain"                    "Sweden"                   "Taiwan Province of China"
# [28] "Thailand"                        "United States of America"

setdiff(ASC_bivalve$cert_country, fao_bivalve$country)
# United kingdom spelled differently.. don't worry about it

unique(ASC_bivalve$cert_country)
```

Tuna
```{r}
fao_tuna <- fao_all %>%
  dplyr::filter(aq_group == "tuna")

ASC_tuna <- ASC_all_sf %>%
  dplyr::filter(aq_group == "tuna")

unique(fao_tuna$country)
```

Crustaceans
```{r}
fao_crust <- fao_all %>%
  dplyr::filter(aq_group == "crustaceans")

ASC_crust <- ASC_all_sf %>%
  dplyr::filter(aq_group == "crustaceans")

unique(fao_crust$country)
```