### 2021-03-23 marine_mapping_archive_v1/bivalve/bivalve_farms/STEP3_rasterize_production.Rmd

---
title: "Rasterizing prodution per each region"
author: "Gage Clawson (USCB, NCEAS, OHI)"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

# read in points and converted to spatial dataframe
```{r}

farms <- read.csv(here::here("marine/bivalve/bivalve_farms/data/global_bivalve_farm_lat_lon.csv"))

farms_sf <-  st_as_sf(farms, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_sf)
```

# determine subregions where points are located

```{r}

farms_sf$reg_id <- raster::extract(gadm_latlon, farms_sf)

farms_sf <- farms_sf %>%
  left_join(gadm_codes, by="reg_id") %>%
  dplyr::mutate(NAME_1 = as.character(NAME_1))

farms_sf$check <-  paste(farms_sf$iso3c, farms_sf$GID_0, sep=":")
table(farms_sf$check)

## read in those needed to check 
ireland <- read_csv(here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_IRL_BGR_NOR_NLD_ITA_GRC.csv")) %>%
  dplyr::filter(iso3c == "IRL")

britain <- read_csv(here("marine/bivalve/bivalve_farms/data/bivalve_sites/ASC_bivalve_farm_lat_lon.csv")) %>%
  dplyr::filter(iso3c == "GBR")

taiwan <- read_csv(here("marine/bivalve/bivalve_farms/data/bivalve_sites/farms_twn.csv"))

## Need to be changed to this 
# GBR:JEY - GBR
# HRV:MNE - HRV
# TWN:CHN - TWN
# IRL:GBR - IRL
# PRK:RUS - PRK
# SVN:HRV - SVN
# SVN:ITA - SVN
# HKG:CHN - HKG

# some annoying mismatches due to raster resolution. Fix these

unique(farms_sf$NAME_1)

farms_sf_correct <- farms_sf %>%
  mutate(GID_0 = as.character(GID_0)) %>%
  mutate(NAME_0 = as.character(NAME_0)) %>%
  mutate(NAME_1 = as.character(NAME_1)) %>%
  mutate(GID_0 = ifelse(check %in% "GBR:JEY", "GBR", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "GBR:JEY", "United Kingdom", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "HRV:MNE", "HRV", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HRV:MNE", "Croatia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "TWN:CHN", "TWN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "TWN:CHN", "Taiwan", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "IRL:GBR", "IRL", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "IRL:GBR", "Ireland", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "PRK:RUS", "PRK", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "PRK:RUS", "North Korea", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "SVN:HRV", "SVN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "SVN:HRV", "Slovenia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "SVN:ITA", "SVN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "SVN:ITA", "Slovenia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "HKG:CHN", "HKG", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HKG:CHN", "Hong Kong", NAME_0))

farms_sf_correct$check <-  paste(farms_sf_correct$iso3c, farms_sf_correct$GID_0, sep=":")
table(farms_sf_correct$check)

```

# FAO bivalve production data per country
```{r}

fao <- read.csv(here::here("marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve.csv")) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::select(iso3c, fao_tonnes_production)

setdiff(fao$iso3c, farms_sf$iso3c)
setdiff(farms_sf$iso3c, fao$iso3c)

```

# join country production to farms by iso3c

```{r}

farms_sf_correct <- farms_sf_correct %>%
  left_join(fao, by="iso3c")

summary(farms_sf_correct) # make sure no production NA values

```

# get areas with no regional production data and calculate production per farm

```{r}

## calculate farm production
farms_prod <- farms_sf_correct %>%
  group_by(NAME_0) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = fao_tonnes_production/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry)

```

# combine and check it all matches ok

```{r}

test <- farms_prod %>%
  dplyr::group_by(iso3c) %>%
  dplyr::summarize(total_tonnes = sum(tonnes_per_farm, na.rm=TRUE)) %>%
  left_join(fao, by = "iso3c") %>%
  mutate(difference = fao_tonnes_production - total_tonnes)

farms_sf_prod <- st_as_sf(farms_prod)

## write the production data so that we can use it for the nutrient layer. 
st_write(farms_sf_prod, "marine/bivalve/bivalve_nutrient/prod_data_by_cell_iso3c.shp")

```

## convert to raster, with the value of each cell equalling the sum tonnes per farm for all farms landing in cell.

```{r}

farms_sf <- st_as_sf(farms_prod)
mapview(farms_sf)

sum(farms_sf$tonnes_per_farm) # 17435623

bivalve_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
plot(log(bivalve_raster_check+1))
cellStats(bivalve_raster_check, stat = "sum", na.rm = TRUE) # 17435623

bivalve_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)

plot(log(bivalve_raster+1))

bivalve_raster_df <- as.data.frame(bivalve_raster, xy=TRUE) %>%
  dplyr::select(x,y, tonnes_production=layer)

cellStats(bivalve_raster, stat = "sum") # 17435623
sum(bivalve_raster_df$tonnes_production) # 17435623
sum(farms_sf$tonnes_per_farm, na.rm = TRUE) # 17435623
sum(fao$fao_tonnes_production) # 17435623

write_csv(bivalve_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/bivalve/bivalve_farm.csv")

```