### 2021-04-05 marine/marine_fish_general/finfish_farms/STEP4_rasterize_production.Rmd

---
title: "Rasterizing prodution per each region"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "10/09/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this script we allocated the production per farm (equally if there is no subregional production estimates).

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)

source(here("_spatial/template_raster.R"))

```

# Global regions

```{r}

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

# read in points and converted to spatial dataframe
```{r}

data <- read_csv("marine/output/all_marine_farms.csv") %>% 
  st_as_sf(., 
           coords = c("lon", "lat"),
           crs = crs(food_raster), 
           agr = "constant")

all_farms <- list.files("data/temp_data/", pattern = "rds$", full.names = TRUE) %>% 
  map_df(., readRDS) %>% 
  rename(iso3c = iso3) %>% 
  mutate(source = "modeled") %>% 
  rbind(data %>% mutate(source = "real"))

finfish_farms <- all_farms %>%
  filter(type == "marine_fish_general")

```

# determine subregions where points are located

```{r}

finfish_farms$reg_id <- raster::extract(gadm_latlon, finfish_farms)

farms_sf <- finfish_farms %>%
  left_join(gadm_codes, by="reg_id") %>%
  dplyr::mutate(NAME_1 = as.character(NAME_1))

farms_sf$check <-  paste(farms_sf$iso3c, farms_sf$GID_0, sep=":")
table(farms_sf$check)

mapview(farms_sf)
# some annoying mismatches due to raster resolution. Fix these

# ALB:MNE
# CHN:HKG
# CHN:MAC
# CYP:XAD
# EGY:ISR
# ESP:PRT
# GRC:ALB
# GRC:TUR
# HKG:CHN
# HRV:SVN
# IDN:PNG
# IDN:TLS
# ISR:EGY
# ITA:FRA
# MYS:SGP
# MYS:THA
# PRT:ESP
# SGP:MYS
# TUR:BGR
# TUR:GRC
# TWN:CHN

farms_sf_correct <- farms_sf %>%
  mutate(GID_0 = as.character(GID_0)) %>%
  mutate(NAME_0 = as.character(NAME_0)) %>%
  mutate(NAME_1 = as.character(NAME_1)) %>%
  mutate(GID_0 = ifelse(check %in% "ALB:MNE", "ALB", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "ALB:MNE", "Albania", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:HKG", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:HKG", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CHN:MAC", "CHN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CHN:MAC", "China", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CYP:XAD", "CYP", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CYP:XAD", "Cyprus", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "EGY:ISR", "EGY", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "EGY:ISR", "Egypt", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "ESP:PRT", "ESP", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "ESP:PRT", "Spain", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "GRC:ALB", "GRC", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "GRC:ALB", "Greece", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "GRC:TUR", "GRC", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "GRC:TUR", "Greece", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "HKG:CHN", "HKG", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HKG:CHN", "Hong Kong", NAME_0)) %>%   
  mutate(GID_0 = ifelse(check %in% "HRV:SVN", "HRV", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HRV:SVN", "Croatia", NAME_0)) %>%    
  mutate(GID_0 = ifelse(check %in% "IDN:PNG", "IDN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "IDN:PNG", "Indonesia", NAME_0)) %>% 
  mutate(GID_0 = ifelse(check %in% "IDN:TLS", "IDN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "IDN:TLS", "Indonesia", NAME_0)) %>% 
  mutate(GID_0 = ifelse(check %in% "ISR:EGY", "ISR", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "ISR:EGY", "Israel", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "ITA:FRA", "ITA", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "ITA:FRA", "Italy", NAME_0)) %>%   
  mutate(GID_0 = ifelse(check %in% "MYS:SGP", "MYS", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MYS:SGP", "Malaysia", NAME_0)) %>%   
  mutate(GID_0 = ifelse(check %in% "MYS:THA", "MYS", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "MYS:THA", "Malaysia", NAME_0)) %>% 
  mutate(GID_0 = ifelse(check %in% "PRT:ESP", "PRT", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "PRT:ESP", "Portugal", NAME_0)) %>% 
  mutate(GID_0 = ifelse(check %in% "SGP:MYS", "SGP", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "SGP:MYS", "Singapore", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "TUR:BGR", "TUR", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "TUR:BGR", "Turkey", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "TUR:GRC", "TUR", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "TUR:GRC", "Turkey", NAME_0)) %>%  
  mutate(GID_0 = ifelse(check %in% "TWN:CHN", "TWN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "TWN:CHN", "Taiwan", NAME_0))

farms_sf_correct$check <-  paste(farms_sf_correct$iso3c, farms_sf_correct$GID_0, sep=":")
table(farms_sf_correct$check)

mapview(farms_sf_correct) # this all looks reasonable, no crazy points
```

# FAO salmon production data per country
```{r}

fao <- read.csv(here("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv")) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::select(iso3c, fao_tonnes_production)

```

# join country production to farms by iso3c

```{r}

farms_sf_correct <- farms_sf_correct %>%
  left_join(fao, by="iso3c")

summary(farms_sf_correct) # make sure no production NA values

```

# get areas with no regional production data and calculate production per farm

```{r}

## calculate farm production
farms_prod <- farms_sf_correct %>%
  group_by(NAME_0) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = fao_tonnes_production/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry)

```

# combine and check it all matches ok

```{r}

test <- farms_prod %>%
  group_by(iso3c) %>%
  summarize(total_tonnes = sum(tonnes_per_farm, na.rm=TRUE)) %>%
  left_join(fao) %>%
  mutate(difference = fao_tonnes_production - total_tonnes)

```

## convert to raster, with the value of each cell equalling the sum tonnes per farm for all farms landing in cell.

```{r}

farms_sf <- st_as_sf(farms_prod)
mapview(farms_sf)

finfish_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
mapview(finfish_raster_check)

finfish_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)
plot(log(finfish_raster + 1))

finfish_raster_df <- as.data.frame(finfish_raster, xy=TRUE) %>%
  dplyr::select(x,y, tonnes_production=layer)

cellStats(finfish_raster, stat = "sum") # 3901214
sum(finfish_raster_df$tonnes_production) # 3901214
sum(farms_sf$tonnes_per_farm, na.rm = TRUE) # 3901214
sum(fao$fao_tonnes_production) #3901214

write_csv(finfish_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/marine_fish_general/marine_fish_general_farm_updated.csv")

```