### 2021-04-23 analysis/summary_tables.Rmd
fixed missed nova scotia salmon points.. one more farm needs to be added to russia salmon now

---
title: "Summary information"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "03/06/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```

# Global regions

```{r}

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

# read in points and converted to spatial dataframe
```{r}

data <- read_csv(file.path("marine/output/all_marine_farms.csv")) %>% 
  st_as_sf(., 
           coords = c("lon", "lat"),
           crs = crs(food_raster), 
           agr = "constant")

all_farms_to_place <- read_csv(file.path("marine/output/all_points_to_add.csv")) %>%
  distinct(iso3c, type, estimated)

all_farms <- list.files("data/temp_data/", pattern = "rds$", full.names = TRUE) %>% 
  map_df(., readRDS) %>% 
  rename(iso3c = iso3) %>% 
  left_join(all_farms_to_place) %>%
  mutate(source = "modeled") %>%
  rbind(data %>% mutate(source = "real", estimated = 0)) %>%
  mutate(data_type = case_when(
    source == "real" ~ "known location", 
    source == "modeled" & estimated == 0 ~ "modeled location, known number of farms",
    source == "modeled" & estimated == 1 ~ "modeled location, estimated number of farms"
  ))
summary(all_farms)

by_type <- all_farms %>%
  group_by(type) %>%
  summarise(n()) %>%
  st_drop_geometry()

by_country_type <- all_farms %>%
  st_drop_geometry() %>%
  group_by(iso3c, type) %>%
  summarise(n())

by_data_type_country <- all_farms %>%
  st_drop_geometry() %>%
  group_by(iso3c, type, data_type) %>%
  summarise(n()) %>%
  ungroup()

known_countries <- by_data_type_country %>%
  filter(data_type == "known location") %>%
  distinct(iso3c)

by_data_type <- all_farms %>%
  st_drop_geometry %>%
  group_by(type, data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)

sources <- read_csv("analysis/data/mariculture_mapping_sources.csv") %>%
  clean_names() %>%
  mutate(type = case_when(species == "Salmonids" ~ "salmon",
                          species == "Unfed or algae fed shellfish" ~ "bivalve",
                          species == "Shrimp" ~ "shrimp", 
                          species == "Tuna" ~ "tuna", 
                          species == "General Marine Fish" ~ "marine_fish_general", 
                          species == "Crustaceans" ~ "crustaceans"))

## join all farms with sources data 
all_farms_sources <- all_farms %>%
  left_join(sources, by = c("iso3c", "type")) %>%
  mutate(data_source = case_when(
    data_type == "modeled location, known number of farms" ~ datasource_number_of_farms,
    data_type == "modeled location, estimated number of farms" ~ "NA", 
    data_type == "known location" ~ datasource_farm_locations
  )) %>%
  dplyr::select(iso3c, country, species, data_type, details, data_source, geometry)

test <- all_farms_sources %>%
  filter(iso3c == "IRN",
         species == "Shrimp")
```