### 2021-04-26 marine_mapping_OLD_v1/STEP1_species_groups/summarise_tonnes_for_paper.Rmd

---
title: "Summarise total tonnes for report"
output: html_document
editor_options: 
  chunk_output_type: console
---

Putting this script here to check total numbers of aquaculture.

```{r}
library(tidyverse)
library(here)
library(countrycode)

source(here('aquaculture/R/fao_mar_clean.R')) # functions specific to mariculture dealing with compound countries
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/fao_fxn.R') # function for cleaning FAO files

```

```{r}

fao <- read_csv("/home/shares/food-systems/Food_footprint/_raw_data/FAO_aquaculture/v_2020/FAO_GlobalAquacultureProduction_Quantity_1950_2018.csv") ## read in the foa data

fao <- fao %>%
  rename(country = `Country (Country)`,
         species = `ASFIS species (ASFIS species)`, 
         fao_major_area = `FAO major fishing area (FAO major fishing area)`, 
         environment = `Environment (Environment)`) %>%
  dplyr::select(-Unit)
table(fao$environment)  
## convert to long format and clean FAO codes:
fao <- fao %>%
  gather(key="year", value="value", -country, -species, -fao_major_area, -environment) %>%
    fao_clean_data() %>%
  dplyr::filter(year >= 2010) %>%
  group_by(country, species, fao_major_area, environment) %>%
  dplyr::mutate(total_harvest = sum(value, na.rm=TRUE)) %>%
  filter(total_harvest > 0) %>% # data is cut if country/species/major area hasn't been harvested since 2010
  dplyr::select(-total_harvest) %>%
  ungroup()

## convert to iso3c codes
fao <- fao %>% 
  filter(!country == "Totals") %>% 
  filter(!is.na(species)) %>%
  mutate(country = ifelse(country == "R\xe9union", "Reunion", country)) %>%
  filter(!country %in% c("Yugoslavia SFR", "Un. Sov. Soc. Rep.")) %>%  # these countries no longer exist  
  mutate(country = ifelse(country=='Zanzibar', "Tanzania", country)) %>%  #semiautonomous region of Tanzania
  mutate(country = ifelse(country=="Eswatini", "Swaziland", country)) %>%  # also goes by this name
  mar_split() %>%   # fixes channel islands, breaks into Guernsey and Jersey
  mutate(iso3c = countrycode(country, "country.name", "iso3c"))

```

## explore freshwater
```{r}

fao <- read_csv("aquaculture/data/fao_mariculture_clean.csv")

fw_test <- fao %>%
  dplyr::filter(environment == "Freshwater") %>%
  dplyr::filter(year == 2017)

sum(fw_test$value, na.rm = TRUE) # 48671218

fw <- fao %>%
  dplyr::filter(environment == "Freshwater") %>%
  dplyr::filter(year == 2017) %>%
  group_by(species) %>%
  summarize(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(total)

sum(fw$total, na.rm = TRUE) # 48671218
data.frame(fw)

```

## explore marine
```{r}
## Read in list of species from OHI so that we can add higher taxonomic classifications
mar_sp_list <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2020/gh-pages/globalprep/mar/v2020/raw/species_list.csv")

marine <- fao %>%
  filter(environment %in% c("Marine", "Brackishwater")) %>%
  dplyr::filter(year == 2017) %>%
  group_by(country, species) %>%
  summarize(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(total)

#### get seaweeds totals 
marine_seaweeds <- marine %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
  dplyr::filter(Taxon_code == "AL"|
                species == "Red oyas")

sum(marine_seaweeds$total, na.rm = TRUE) # 32560138 tonnes of seaweeds not included

#### get marine animals totals by filtering out the seaweeds
marine_animals_total <- marine %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
  dplyr::filter(Taxon_code != "AL",
                  species != "Red oyas") ## filter out seaweeds
sum(marine_animals_total$total, na.rm = TRUE) # 30991774 - total marine animals production (including those spp that we will not include in the analysis)

#### Now get what we included for marine animals:
## filter out all seaweeds 
## filter out all faux freshwater species listed as marine or brackishwater: "Freshwater|catfish|Cyprinids|ilapia|iver"
## filter out all jellyfish
## filter out and species/country producing <500 tonnes
marine_included_analysis <- marine %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
   dplyr::filter(Taxon_code != "AL",
                species != "Red oyas") %>%
  dplyr::filter(!str_detect(species, "Freshwater|catfish|Cyprinids|ilapia|iver|ellyfish")) %>%
  dplyr::filter(total >= 500)
sum(marine_included_analysis$total, na.rm = TRUE) # 29630029 tonnes were included in our analysis

#### Now lets get a list of marine animals not included in our analysys (just do the opposite of above)
marine_animal_spp_not_included_analysis <- marine %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
   dplyr::filter(str_detect(species, "Freshwater|catfish|Cyprinids|ilapia|iver|ellyfish")) %>%
  dplyr::filter(total > 0)
unique(marine_animal_spp_not_included_analysis$species)

marine_other_not_included_analysis <- marine %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
  dplyr::filter(Taxon_code != "AL",
                species != "Red oyas") %>%
  dplyr::filter(!str_detect(species, "Freshwater|catfish|Cyprinids|ilapia|iver|ellyfish")) %>%
  dplyr::filter(total < 500) %>%
  dplyr::filter(total > 0)

marine_all_not_included_analysis <- rbind(marine_animal_spp_not_included_analysis, marine_other_not_included_analysis)
sum(marine_all_not_included_analysis$total, na.rm = TRUE) # 1361746 - matches the difference... perfect

unique(marine_all_not_included_analysis$species)

```

This is how we filtered for our production analysis: 
 - we filtered out these species :
[1] "Tilapias nei"          "River Plata mussel"    "River eels nei"        "Giant river prawn"     "Bagrid catfish"       
 [6] "Freshwater fishes nei" "Mozambique tilapia"    "Nile tilapia"          "Blackchin tilapia"     "Sabaki tilapia"       
[11] "Blue tilapia"          "North African catfish" "Cyprinids nei"         "Jellyfishes nei"  
 - we filtered out species within countries that have production < 500