### 2021-04-26 marine_mapping_OLD_v1/shrimp/shrimp_farms/STEP3_rasterize_production.Rmd

---
title: "Rasterizing prodution per each region"
author: "Gage Clawson (USCB, NCEAS, OHI)"
date: "10/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)

source(here::here("_spatial/template_raster.R"))

```

# Global regions

```{r}

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

# read in points and converted to spatial dataframe
```{r}
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)

farms <- read.csv(here::here("marine/shrimp/shrimp_farms/data/global_shrimp_farm_lat_lon.csv"))

farms_sf <-  st_as_sf(farms, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

mapview(farms_sf)

test <- rasterize(farms_sf, food_raster, fun = "count")
plot(log(test + 1))

```

# determine subregions where points are located

```{r}

farms_sf$reg_id <- raster::extract(gadm_latlon, farms_sf)

farms_sf <- farms_sf %>%
  left_join(gadm_codes, by="reg_id") %>%
  dplyr::mutate(NAME_1 = as.character(NAME_1))

farms_sf$check <-  paste(farms_sf$iso3c, farms_sf$GID_0, sep=":")
table(farms_sf$check)

# CRI:PAN - CRI
# HND:NIC - HND
# KHM:THA - KHM
# TWN:CHN - TWN

mapview(filter(farms_sf, check=="HND:HND"))
mapview(filter(farms_sf, check=="HND:NIC"))
mapview(filter(farms_sf, check =="KHM:KHM"))
mapview(filter(farms_sf, check == "KHM:THA"))
mapview(filter(farms_sf, check == "TWN:CHN"))
mapview(filter(farms_sf, check == "CHN:CHN"))
mapview(filter(farms_sf, check == "CRI:PAN"))
mapview(farms_sf)

# some annoying mismatches due to raster resolution. Fix these

unique(farms_sf$NAME_1)

farms_sf_correct <- farms_sf %>%
  mutate(GID_0 = as.character(GID_0)) %>%
  mutate(NAME_0 = as.character(NAME_0)) %>%
  mutate(NAME_1 = as.character(NAME_1)) %>%
  mutate(GID_0 = ifelse(check %in% "HND:NIC", "HND", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "HND:NIC", "Honduras", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "KHM:THA", "KHM", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "KHM:THA", "Cambodia", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "TWN:CHN", "TWN", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "TWN:CHN", "Taiwan", NAME_0)) %>%
  mutate(GID_0 = ifelse(check %in% "CRI:PAN", "CRI", GID_0)) %>%
  mutate(NAME_0 = ifelse(check %in% "CRI:PAN", "Costa Rica", NAME_0))

farms_sf_correct$check <-  paste(farms_sf_correct$iso3c, farms_sf_correct$GID_0, sep=":")
table(farms_sf_correct$check)

```

# FAO shrimp production data per country
```{r}

fao <- read.csv(here::here("marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp.csv")) %>%
  dplyr::filter(year == 2017) %>%
  dplyr::select(iso3c, fao_tonnes_production)

setdiff(fao$iso3c, farms_sf$iso3c)
setdiff(farms_sf$iso3c, fao$iso3c)

```

# join country production to farms by iso3c

```{r}

farms_sf_correct <- farms_sf_correct %>%
  left_join(fao, by="iso3c")

summary(farms_sf_correct) # make sure no production NA values

```

# get areas with regional production data and calculate production per farm
```{r}

## Saudi Arabia is the only one with regional production
farms_rgn_prod <- farms_sf_correct %>%
  dplyr::filter(!is.na(prod))

## calculate farm production
farms_rgn_prod <- farms_rgn_prod %>%
  mutate(prop_production = prod/sum(prod)) %>%
  mutate(regional_prod = fao_tonnes_production*prop_production) %>%
  group_by(NAME_1) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = regional_prod/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry)

```

# get areas with no regional production data and calculate production per farm

```{r}

## calculate farm production
farms_prod <- farms_sf_correct %>%
  group_by(NAME_0) %>%
  add_tally() %>%
  ungroup() %>%
  mutate(tonnes_per_farm = fao_tonnes_production/n) %>%
  dplyr::select(iso3c, tonnes_per_farm, geometry) %>%
  dplyr::filter(iso3c != "SAU") ## filter out SAU since we have regional production estimates

```

# combine and check it all matches ok

```{r}

farm_production <- rbind(farms_prod, farms_rgn_prod)

test <- farm_production %>%
  dplyr::group_by(iso3c) %>%
  dplyr::summarize(total_tonnes = sum(tonnes_per_farm, na.rm=TRUE)) %>%
  left_join(fao, by = "iso3c") %>%
  mutate(difference = fao_tonnes_production - total_tonnes)

```

## convert to raster, with the value of each cell equalling the sum tonnes per farm for all farms landing in cell.

```{r}

farms_sf <- st_as_sf(farm_production)
mapview(farms_sf)

sum(farms_sf$tonnes_per_farm) # 5094133

shrimp_raster_check <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum)
plot(log(shrimp_raster_check+1))
cellStats(shrimp_raster_check, stat = "sum", na.rm = TRUE) # 5094133

mapview(shrimp_raster_check)

shrimp_raster <- rasterize(farms_sf, food_raster, field = "tonnes_per_farm", fun=sum, background=0)
mapview(log(shrimp_raster+1))

plot(log(shrimp_raster+1))

shrimp_raster_df <- as.data.frame(shrimp_raster, xy=TRUE) %>%
  dplyr::select(x,y, tonnes_production=layer)

cellStats(shrimp_raster, stat = "sum") # 5094133
sum(shrimp_raster_df$tonnes_production) # 5094133
sum(farms_sf$tonnes_per_farm, na.rm = TRUE) # 5094133
sum(fao$fao_tonnes_production) # 5094133

write_csv(shrimp_raster_df, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/shrimp/shrimp_farm.csv")

```