### 2021-04-27 analysis/paper_stats.Rmd

---
title: "Summary stats for paper"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "04/26/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script takes our final farms dataset and calculates summary statistics that will be used in the paper.

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```

# Global regions

```{r}

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))
```

# Read in farms data and convert to sf

```{r}
all_farms_sf <- read_csv("all_marine_aquaculture_farms_sources.csv") %>%
    st_as_sf(., 
           coords = c("X", "Y"),
           crs = crs(food_raster), 
           agr = "constant")

all_farms <- all_farms_sf %>%
  st_drop_geometry()

```

```{r}
sources <- read_csv("analysis/data/mariculture_mapping_sources.csv") %>%
  clean_names() %>%
  mutate(type = case_when(species == "Salmonids" ~ "salmon",
                          species == "Unfed or algae fed shellfish" ~ "bivalve",
                          species == "Shrimp" ~ "shrimp", 
                          species == "Tuna" ~ "tuna", 
                          species == "General Marine Fish" ~ "marine_fish_general", 
                          species == "Crustaceans" ~ "crustaceans"))

## Numbers by data type
by_data_type <- all_farms %>%
  group_by(data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)

by_data_type
# A tibble: 3 x 4
#   data_type                                   n_obs total_obs   pct
#   <chr>                                       <int>     <int> <dbl>
# 1 known location                               9750     95341  10.2
# 2 modeled location, estimated number of farms 75072     95341  78.7
# 3 modeled location, known number of farms     10519     95341  11.0

# How many countries have spatial locations information?
known_countries <- all_farms %>%
  filter(data_type == "known location") %>%
  distinct(iso3c)

# Some of the world’s largest aquaculture producers had a noticeable paucity of aquaculture location data, particularly China, whose represented 77% of the total estimated points, and only had 15 known location points

chn_data_estimated <- all_farms %>%
  filter(iso3c == "CHN") %>%
  filter(data_type == "modeled location, estimated number of farms")
# 58151 points
58151/75073 # 0.7745927 = 77%

chn_data_known <- all_farms %>%
  filter(iso3c == "CHN") %>%
  filter(data_type == "known location") # 15 locations

# Chile, Norway, Peru, the United States, Canada and Ireland provided a substantial number of known locations (>1000).
top_known <- all_farms %>%
  filter(data_type == "known location") %>%
  group_by(iso3c) %>%
    summarise(n_obs = n())

# Only 5 countries (Cambodia, Mauritius, Myanmar, North Korea, and Russia) had no form of farm data at all, however, these countries only account for 0.5% of marine aquaculture production included.

no_info_at_all <- sources %>%
  dplyr::filter(known_farms_location == 0, known_farms_no_location == 0) ## only 5 with no information at all

mar_production <- read_csv("marine/STEP1_species_groups/int/tonnes_per_country_group.csv")

no_info_production <- mar_production %>%
  filter(country %in% c("Cambodia", "Mauritius", "Myanmar", "Korea, Dem. People's Rep", "Russian Federation"))
sum(no_info_production$total_tonnes) # 152800.1

sum(mar_production$total_tonnes) # 29623515
152800.1/29623515

chn_production <- mar_production %>%
  filter(country == "China")
sum(chn_production$total_tonnes)
17696822/29623515

# Per species group, salmon was the most data rich (98% of points had location information, 1.13% of points were known with no location information, 0.87% estimated), followed by Tuna (34% of points had location information, 66% of points were known with no location information, 0% estimated).

by_species_data_type <- all_farms %>%
  group_by(species, data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)
```

Check some of the data
```{r}
by_type <- all_farms %>%
  group_by(species) %>%
  summarise(n_obs = n())

by_country_type <- all_farms %>%
  group_by(iso3c, species) %>%
  summarise(n_obs = n()) %>%
  left_join(sources) %>%
  dplyr::select(iso3c, species, n_obs, number_of_total_farms) %>%
  mutate(difference = n_obs - number_of_total_farms)

by_data_type_country <- all_farms %>%
  group_by(iso3c, species, data_type) %>%
  summarise(n_obs = n()) %>%
  ungroup() %>%
  left_join(sources) %>%
  dplyr::select(iso3c, species, data_type, n_obs, number_of_total_farms, known_farms_location, known_farms_no_location, estimated_farms)

known <- by_data_type_country %>%
  filter(data_type == "known location") %>%
  dplyr::select(iso3c, species, data_type, n_obs, known_farms_location) %>%
  mutate(difference = n_obs - known_farms_location)

number_known <- by_data_type_country %>% 
  filter(data_type == "modeled location, known number of farms") %>%
  dplyr::select(iso3c, species, data_type, n_obs, known_farms_no_location) %>%
  mutate(difference = n_obs - known_farms_no_location)

estimated <- by_data_type_country %>% 
  filter(data_type == "modeled location, estimated number of farms") %>%
  dplyr::select(iso3c, species, data_type, n_obs, estimated_farms) %>%
  mutate(difference = n_obs - estimated_farms)

```