### 2021-05-11 analysis/prod_farm_rel_explore.Rmd
made the figures!

---
title: "Explore production and farm trends"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "04/28/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("analysis/R/comp_farms_production.R")

## Bivalve
bivalve <- comp_farms_production("bivalve")

plot_farms_prod(bivalve)

lm_farms_prod(bivalve)

## cambodia has 8300 tonnes of production, plugging into model we get
8300*8.740e-03 - 4.778e+01 # ~25 farms

8300/mean(bivalve$tonnes_per_farm) # ~ 30 farms rounding up

## Salmon
salmon <- comp_farms_production("salmon")

plot_farms_prod(salmon)

lm_farms_prod(salmon)

## Russia has 13016 tonnes of production, plugging into model we get
13016*1.060e-03 + 7.318e+01 # ~87 farms

13016/mean(salmon$tonnes_per_farm) # ~ 15 farms rounding up

## Shrimp
shrimp <- comp_farms_production("shrimp")

plot_farms_prod(shrimp)

lm_farms_prod(shrimp)

## Myanmar has 55309.51 tonnes of production

55309.51*3.888e-03 + 2.024e+01 # ~236 farms

55309.51/mean(shrimp$tonnes_per_farm) # ~ 165 farms rounding up

## mfgeneral
finfish <- comp_farms_production("finfish")

plot_farms_prod(finfish)

lm_farms_prod(finfish)

## Myanmar has 55309.51 tonnes of production

1419359*2.669e-03 + 4.052e+01 # ~3828.789 farms

1419359/mean(finfish$tonnes_per_farm) # ~ 1639.569 farms
```

All of the models prove to be significant. However, they estimate a lot more farms than just using the mean production per farm values... so we are likely being conservative with our current number-of-farm estimations.