### 2021-05-20 marine/bivalve/bivalve_farms/STEP3_gapfill_biv.Rmd
checked the median global tonnes/farm

---
title: "Gapfilling the number of farms to add for bivalve countries based on tonnes of production"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in all the relevant farm datasets

```{r}

farm_files <- list.files(here("marine/bivalve/bivalve_farms/data/known_farm_sites_2_4_5"), full=TRUE)
all_known_farms = plyr::ldply(farm_files, read_csv) %>%
  mutate(type = "bivalve")

number_files <- list.files(here("marine/bivalve/bivalve_farms/data/known_number_farms_3"), full=TRUE)
all_number_3 = plyr::ldply(number_files, read_csv)

number_files_2.5 <- list.files(here("marine/bivalve/bivalve_farms/data/known_subset_2.5"), full=TRUE)
all_number_2.5 = plyr::ldply(number_files_2.5, read_csv)

subset_farms_files <- list.files(here("marine/bivalve/bivalve_farms/data/subset_known_add_points_4"), full=TRUE)
subset_add_4 = plyr::ldply(subset_farms_files, read_csv)

no_info_files <- list.files(here("marine/bivalve/bivalve_farms/data/no_info_1"), full=TRUE)
no_info_1 = plyr::ldply(no_info_files, read_csv)

subset_gapfill_2 <- all_known_farms %>%
  filter(data_quality == 2)
```

Read in relevant production data
```{r}
bivalve_tonnes <- read_csv(here::here("marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve.csv")) %>%
  mutate(type = "bivalve") %>%
  filter(year == 2017)
```

Calculate a global tonnes per farm for bivalve

```{r}
## Filter for data_quality = 4 or 5 and group by country to get number of farms
comprehensive_farms <- all_known_farms %>%
  filter(data_quality %in% c(4,5)) %>% 
  group_by(iso3c, type, data_quality) %>%
  summarise(n_farms = n()) %>% ## now we can see the places that has additional farms to add
  ungroup()

## Join this with the additional farms we need to add to some places, so we can calculate a comprehensive tonnes/farm for bivalve

data_quality_3 <- all_number_3 %>%
    rename("n_farms" = "num_additional_farms")

data_quality_4 <- subset_add_4 %>%
    rename("n_farms" = "num_additional_farms")

comprehensive_farms_total <- comprehensive_farms %>%
  full_join(data_quality_3) %>%
  full_join(data_quality_4) %>%
  group_by(iso3c, type) %>%
  summarise(n_farms = sum(n_farms))

tonnes_per_farm <- comprehensive_farms_total %>%
  left_join(bivalve_tonnes) %>%
  mutate(tonnes_per_farm = fao_tonnes_production/n_farms)

avg_global_bivalve_tonnes_per_farm = mean(tonnes_per_farm$tonnes_per_farm)
avg_global_bivalve_tonnes_per_farm # 283.4135

med_global_bivalve_tonnes_per_farm <- median(tonnes_per_farm$tonnes_per_farm)
med_global_bivalve_tonnes_per_farm # 134.3808
```

Now let's gapfill those that are data quality = 2 and data quality = 1, and data quality = 2.5

```{r}
## start with the subset of locations that need gapfilling based on tonnes/farm
gapfill_2 <- subset_gapfill_2 %>%
  group_by(iso3c, type) %>%
  summarise(n_existing_farms = n()) %>%
  ungroup() %>%
  left_join(bivalve_tonnes) %>%
  mutate(avg_tonnes_per_farm = avg_global_bivalve_tonnes_per_farm) %>%
  mutate(tonnes_accounted_for = avg_tonnes_per_farm*n_existing_farms) %>%
  mutate(tonnes_remaining = fao_tonnes_production - tonnes_accounted_for) %>%
  mutate(tonnes_remaining = ifelse(tonnes_remaining <= 0, 0, tonnes_remaining)) %>%
  mutate(num_additional_farms = ceiling(tonnes_remaining/avg_tonnes_per_farm)) %>%
  dplyr::select(iso3c, type, num_additional_farms, avg_tonnes_per_farm) %>%
  mutate(estimated = 1) %>%
  filter(num_additional_farms != 0)

## now lets do the locations that have no info and need gapfilling based on tonnes/farm
gapfill_1 <- no_info_1 %>%
  left_join(bivalve_tonnes) %>%
  mutate(avg_tonnes_per_farm = avg_global_bivalve_tonnes_per_farm) %>%
  mutate(num_additional_farms = ceiling(fao_tonnes_production/avg_tonnes_per_farm)) %>%
  dplyr::select(iso3c, type, num_additional_farms, avg_tonnes_per_farm) %>%
  mutate(estimated = 1) %>%
  filter(num_additional_farms != 0)

## now lets do the locations that have some information regarding number of farms 
gapfill_2.5 <- all_number_2.5 %>%
  left_join(bivalve_tonnes) %>%
  mutate(avg_tonnes_per_farm = avg_global_bivalve_tonnes_per_farm) %>%
  mutate(tonnes_accounted_for = avg_tonnes_per_farm*n_existing_farms) %>%
  mutate(tonnes_remaining = fao_tonnes_production - tonnes_accounted_for) %>%
  mutate(tonnes_remaining = ifelse(tonnes_remaining <= 0, 0, tonnes_remaining)) %>%
  mutate(num_additional_farms = ceiling(tonnes_remaining/avg_tonnes_per_farm)) %>%
  mutate(num_additional_farms = num_additional_farms + n_existing_farms) %>% 
  dplyr::select(iso3c, type, num_additional_farms, avg_tonnes_per_farm) %>%
  mutate(estimated = ifelse(iso3c == "IND", 0, 1)) %>%
  filter(num_additional_farms != 0)

## Wrangle those that just need a specific number of points added (data_quality = 4)

number_add_points_4 <- subset_add_4 %>%
  dplyr::select(-data_quality, -total_farms) %>%
  mutate(estimated = 0, 
         avg_tonnes_per_farm = avg_global_bivalve_tonnes_per_farm)

```

Now join the gapfilled number of farms with data quality = 3 and write the final file of points that need to allocated

```{r}
number_additional_points <- all_number_3 %>%
  mutate(estimated = 0,
         avg_tonnes_per_farm = avg_global_bivalve_tonnes_per_farm) %>%
  dplyr::select(-data_quality) %>%
  rbind(gapfill_1, gapfill_2, gapfill_2.5, number_add_points_4)

## write this out to our data folder

write.csv(number_additional_points, here("marine/bivalve/bivalve_farms/output/bivalve_points_to_add_final.csv"), row.names = FALSE)
```

Now lets make the file with all of the known farms

```{r}
all_known_farms_final <- all_known_farms %>%
  dplyr::select(-data_quality, -subregion)

write.csv(all_known_farms_final, here("marine/bivalve/bivalve_farms/output/all_known_bivalve_farms_final.csv"), row.names = FALSE)

```

Data check

```{r}
test <- all_known_farms %>% filter(data_quality == 5| iso3c == "CUB") %>%
  group_by(iso3c) %>%
  summarise(n_farms = n()) %>%
  dplyr::select(iso3c)

test_2 <- number_additional_points %>%
  dplyr::select(iso3c)

test_all <- rbind(test, test_2)
length(unique(test_all$iso3c)) ## 42 countries

length(unique(bivalve_tonnes$iso3c)) ## 42 countries

## perfect

```