### 2021-06-11 archive/marine_mapping_OLD_v1/tuna/food_pressures/tuna_nutrient/step1_nutrient.Rmd

---
title: "Tuna nutrient pollution"
author: "Gage Clawson (NCEAS, UCSB, OHI)"
date: "01/06/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    includes: 
toc: true
---

This markdown uses models and species specific parameters from **Nitrogen and phosphorus budget in coastal and marine cage aquaculture and impacts of effluent loading on ecosystem: review and analysis towards model development**

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(raster)
library(readr)
library(data.table)

source(here("_spatial/template_raster.R"))

path_prep <- "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/tuna"
path_final <-"/home/shares/food-systems/Food_footprint/all_food_systems/datalayers"
path_pre <- "/home/shares/food-systems/Food_footprint/all_food_systems/predatalayers"

production <- read_csv(file.path(path_prep, "tuna_farm.csv"))
food_raster <- raster(nrows=2160, ncols=4320, xmn=-180, xmx=180, ymn=-90, ymx=90)
```

What I've found (we ended up only using global values):

**Preliminary model of nitrogen loads from southern bluefin tuna aquaculture** 
https://pubmed.ncbi.nlm.nih.gov/17669437/ 
 - Only lists N pollution values for Southern Bluefin Tuna in Australia. 
 - A range from 260 - 502 kg N / tonne growth 
 - Would use the mean: 281 kg N / tonnes growth

**Marine cage culture and the environment: effects on water quality and primary production** 
https://www.int-res.com/articles/aei2014/6/q006p151.pdf
 - "Some research studies have found significant increases in dissolved nitrogen. For example, AguadoGiménez et al. (2006) calculated the estimated dissolved nitrogen outputs of a bluefin tuna fattening operation in the Mediterranean could be 2−5.6 times greater than a comparable sea bream farm, with the potential for significant environmental impacts during peak production periods."
 - From Lex: seabream is 52 kg N / tonne = [104, 291.2] kg N / tonne : This would be for Europe

Aguado-Giminez: https://onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2109.2006.01541.x

"The phosphorus output from a tuna fattening operation was calculated to be about 3−5 times higher than from sea
bream or sea bass farms because of differences in digestibility and feed formulation (Aguado-Giménez et al. 2006)"
 -  From Lex: Seabass is 17 kg P / tonne = [51, 85] kg P / tonne : for Europe
 - From Lex: Asian Seabass is [11, 16] kg P / tonne = [(33, 48), (55, 80)] kg P / tonne

**Nitrogen and phosphorus budget in coastal and marine cage aquaculture and impacts of effluent loading on ecosystem: review and analysis towards model development**
https://reader.elsevier.com/reader/sd/pii/S0025326X04003042?token=FD7F20D6183C00EC61F7CEED7BA11815BFA6A54438409343DAA77A2D5DEDDAD0BF5B8B2A55EED17F347FDE6E64F4FD72

 - General Cage Aquaculture
   - 132.5 kg N / tonne
   - 25 kg P / tonne

Steps: 
1. Multiply tonnes production by 0.95 (this corrects for the growth that occurs outside the seapen on land), this is the value Lex uses
2. Then multiply growth by the N value and the P values
3. We report in PO4eq: Multiply P by by 3.07 to get PO4eq and N by 0.42 to get PO4eq. Sum the N and P PO4eq rasters.
4. We report in tonnes, so divide by 1000

 USE global values: 
  - General Cage Aquaculture
   - 132.5 kg N / tonne
   - 25 kg P / tonne

```{r}

tuna_pollution <- production %>%
  dplyr::mutate(prod_growth = tonnes_production*0.95) %>%
  dplyr::mutate(prod_growth_N_poll = prod_growth*132.5,
                prod_growth_P_poll = prod_growth*25) %>% ## in kg pollution now
  # dplyr::mutate(PO4eq_kg = prod_growth_N_poll*0.42 + prod_growth_P_poll*3.07) %>% cut (01/14/2021)
  dplyr::mutate(N_P_tonnes = (prod_growth_N_poll + prod_growth_P_poll)/1000, 
                P_poll_t = prod_growth_P_poll/1000, 
                N_poll_t = prod_growth_N_poll/1000) %>%
  dplyr::select(x, y, N_P_tonnes, P_poll_t, N_poll_t)

sum(tuna_pollution$N_P_tonnes) # 5546.982
sum(production$tonnes_production) # 37072.56

## save to dataprep folder with the correct naming conventions:
write.csv(tuna_pollution, file.path(path_prep, "tuna_nutrient.csv"))

## Save N + P layer
N_P_df <- tuna_pollution %>%
  dplyr::select(x, y, N_P_tonnes)

final_rast <- rasterFromXYZ(N_P_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 5546.982

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 5546.982
final_rast

## Save to datalayers
writeRaster(final_rast, file.path(path_final, "marine_tuna_aquaculture_meat_nutrient.tif"), overwrite = T)

## Save N layer
N_df <- tuna_pollution %>%
  dplyr::select(x, y, N_poll_t)

final_rast <- rasterFromXYZ(N_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 4666.508

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 4666.508
final_rast

## Save to datalayers
writeRaster(final_rast, file.path(path_pre, "marine_tuna_aquaculture_meat_nutrientN.tif"), overwrite = T)

## Save P layer
P_df <- tuna_pollution %>%
  dplyr::select(x, y, P_poll_t)

final_rast <- rasterFromXYZ(P_df, crs = food_crs)
crs(final_rast)
plot(final_rast)
final_rast
cellStats(final_rast, "sum") # 880.4733

final_area_fix <- final_rast/raster::area(final_rast)

final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")

final_rast <- final_rast_proj*raster::area(final_rast_proj)
cellStats(final_rast, "sum") # 880.4733
final_rast

## Save to datalayers
writeRaster(final_rast, file.path(path_pre, "marine_tuna_aquaculture_meat_nutrientP.tif"), overwrite = T)

```

ARCHIVE: 
```{r}
# fao <- read.csv(here("aquaculture/data/fao_mariculture_clean.csv")) %>%
#   filter(environment %in% c("Marine", "Brackishwater"))
# 
# 
# ## Read in fao data and get a species list for salmon
# species_list <- read.csv(here("aquaculture/marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
#   distinct(species, aq_group)
# 
# # the list of species we are including
# tuna_list <- species_list$species[species_list$aq_group=="tuna"]
# tuna_list
# 
# ## write out shrimp taxa and production for Juliette
# fao_tuna_taxa_prod <- fao %>%
#   filter(species %in% tuna_list) %>%
#   filter(!is.na(value)) %>%
#   arrange(-value, country) %>%
#   filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
#   group_by(country, iso3c, species, year) %>%
#   dplyr::filter(year == 2017) %>%
#   summarize(fao_tonnes_production = sum(value)) %>%
#   arrange(country, iso3c, species, year)
# 
# ## set up lists of different regions
# asia <- c("JPN")
# 
# europe <- c("HRV", "MLT", "ESP", "TUR") 
# 
# australia <- c("AUS")
# 
# mexico <- c("MEX")
# 
# 
# 
# ## Read in production data 
# production_sf <- st_read(file.path("/home/sgclawson/github/food_systems/aquaculture/marine/tuna/tuna_nutrient/prod_data_by_cell_iso3c.shp")) %>%
#   dplyr::rename("tonnes_per_farm" = "tnns_p_")
# 
# ## Now we can do the actual analysis outlined above
# 
# nutrient_data <- production_sf %>%
#   dplyr::mutate(tonnes_growth = tonnes_per_farm*0.95) %>%   ## multiply by 0.95 to correct for growth outside seapen
#   
#   dplyr::mutate(N_poll = case_when(
#                   iso3c %in% asia ~ tonnes_growth*132.5,
#                   iso3c %in% europe ~ tonnes_growth*197.6,
#                   iso3c %in% mexico ~ tonnes_growth*132.5,
#                   iso3c %in% australia ~ tonnes_growth*281
#                 ),
#                 P_poll = case_when(
#                   iso3c %in% asia ~ tonnes_growth*54,
#                   iso3c %in% europe ~ tonnes_growth*68,
#                   iso3c %in% australia ~ tonnes_growth*25,
#                   iso3c %in% mexico ~ tonnes_growth*25
#                 )) %>%  ## get all of out pollution values by species group and region
#   
#   dplyr::mutate(po4eq_kg = N_poll*0.42 + P_poll*3.07) %>% ## get po4eq
#   
#   dplyr::mutate(po4eq_tonnes = po4eq_kg/1000) %>%
#   dplyr::select(po4eq_tonnes, geometry)
# 
# sum(nutrient_data$po4eq_tonnes) # 7603.354
# sum(fao_tuna_taxa_prod$fao_tonnes_production) # 37072.56
# 
# tuna_raster_check <- rasterize(nutrient_data, food_raster, field = "po4eq_tonnes", fun=sum)
# plot(log(tuna_raster_check+1))
# cellStats(tuna_raster_check, stat = "sum", na.rm = TRUE) # 7603.354
# 
# 
# tuna_raster <- rasterize(nutrient_data, food_raster, field = "po4eq_tonnes", fun=sum, background=0)
# plot(log(tuna_raster+1))
# 
# tuna_raster_df <- as.data.frame(tuna_raster, xy=TRUE) %>%
#   dplyr::select(x,y, po4eq_tonnes=layer)
# 
# cellStats(tuna_raster, stat = "sum") # 7603.354
# sum(tuna_raster_df$po4eq_tonnes) # 7603.354
# sum(nutrient_data$po4eq_tonnes, na.rm = TRUE) # 7603.354
# 
# ## save to dataprep folder with the correct naming conventions: 
# write.csv(tuna_raster_df, file.path(path_prep, "tuna_nutrient.csv"))
# 
# 
# final_rast <- rasterFromXYZ(tuna_raster_df, crs = food_crs)
# crs(final_rast)
# plot(final_rast)
# final_rast
# cellStats(final_rast, "sum") # 7603.354
# 
# final_area_fix <- final_rast/raster::area(final_rast)
# 
# final_rast_proj <- projectRaster(final_area_fix, food_raster, method = "ngb")
# 
# final_rast <- final_rast_proj*raster::area(final_rast_proj)
# cellStats(final_rast, "sum") # 7603.354
# final_rast
# 
# ## Save to datalayers 
# writeRaster(final_rast, file.path(path_final, "marine_tuna_aquaculture_meat_nutrient.tif"), overwrite = T)
```