### 2021-07-09 analysis/SI_table_3.Rmd

---
title: "SI tables"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "04/26/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script takes our final farms dataset and creates the .

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)
library(countrycode)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```