### 2021-07-09 analysis/validation_testing.Rmd

---
title: "Validation Testing"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "07/08/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script tests our validation set of points against our actual known points.

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library("rnaturalearth")
library("rnaturalearthdata")
library(cowplot) 
library(ggpubr)
library(RColorBrewer)
library(OpenStreetMap)
library(ggspatial)
library(ggplot2)
library(patchwork)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```