### 2021-07-09 marine/marine_fish_general/finfish_farms/STEP1_fao_finfish_production.Rmd

---
title: "FAO marine finfish production"
author: "Gage Clawson"
date: "10/09/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

All data is ultimately adjusted to FAO production data.  This script cleans the fao mariculture data and subsets it to to the relevant years of data and species (marine fish - general).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(data.table)
library(countrycode)
library(here)
library(stringr)

path <- "/home/shares/food-systems/Food_footprint/_raw_data"
source(here('marine/salmon/salmon_farm/R/fao_mar_clean.R')) # functions specific to mariculture dealing with compound countries
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/fao_fxn.R') # function for cleaning FAO files
source(here('workflow/common.R'))

```

Subset the data to include only the species (marine fish general) and countries we are interested in. Then summarize by total tonnes of finfish per country/year.

```{r}

fao <- read.csv(here("data/fao_aquaculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

## Read in fao data and get a species list for marine fish
species_list <- read.csv(here("marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

# the list of species we are including
finfish_list <- species_list$species[species_list$aq_group=="marine_fish_general"]
finfish_list

## write out marine fish taxa and production for Juliette
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year)

write.csv(fao_finfish_taxa_prod, here("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish_taxa_production.csv"), row.names=FALSE)

# include only marine finfish
fao_finfish <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, year)

write.csv(fao_finfish, here("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv"), row.names=FALSE)

## look at all groups

all_groups_included <- aquaculture_groups %>%
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value < 500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(year, iso3c, aq_group) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(iso3c, aq_group) %>%
  #filter(year == 2017) %>%
  ungroup()
  #mutate(total_prod = sum(fao_tonnes_production)) %>%
  #mutate(prop_prod = fao_tonnes_production/total_prod)

```

```{r}
## write out bivalve taxa and production for Juliette
fao_finfish_taxa_prod <- fao %>%
  filter(species %in% finfish_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

sum(fao_finfish_taxa_prod$value) # 3916717

unique(fao_finfish_taxa_prod$country) # 37 countries 
```