### 2021-07-09 workflow/calc_dist_to_port.Rmd

---
title: "Calculate average distance to ports from known mariculture locations"
author: "Gage Clawson"
date: "06/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(spatialEco)
library(tidyverse)
library(raster)
library(rgeos)
library(rmapshaper)
library(mapview)

# load in crs, coastlines, eez, masks, country shapes, ports, depth, and rasters
source("../_spatial/template_raster.R")

```

Read in all known farms

```{r}
data <- read_csv("../marine/output/all_marine_farms.csv") %>% 
  st_as_sf(., 
           coords = c("lon", "lat"),
           crs = crs(food_raster), 
           agr = "constant") 
```

Read in ports dataset 
```{r}
the_crs <- crs(fake_raster, asText = TRUE)

ports_raw <- st_read("/home/shares/food-systems/Food_footprint/_raw_data/world_port_index/Commercial_ports.shp") %>%
  st_transform(the_crs) %>%
  filter(HARB_SIZE_ != "L") # remove L ports
```

Calculate the distance from each known location to the closest port

```{r}
# st_nearest_points

# st_distance

test_data <- data[1:3, ]

dist_ports <- as.data.frame(st_distance(ports_raw, test_data))

min(dist_ports$V1)
min(dist_ports$V2)
min(dist_ports$V3)

mapview(test_data)

test_df_final <- dist_ports %>%
  pivot_longer(cols = c(V1, V2, V3)) %>%
  group_by(name) %>%
  summarise(min_dist = min(value))

## ok so that works... i just to calculate the distance from each point in the whole known location dataset, and find the min of each column

dist_ports <- as.data.frame(st_distance(ports_raw, data)) ## calculate distance to each port from each known mariculture location

write.csv(dist_ports, "int/distance_to_ports_raw.csv")

write.csv(dist_ports, "/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/distance_from_known_to_ports_raw.csv", row.names = FALSE)

dist_ports <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/distance_from_known_to_ports_raw.csv")

dist_ports_final <- dist_ports %>%
  pivot_longer(cols = everything()) %>%
  group_by(name) %>%
  summarise(closest_port_km = as.numeric(min(value)*0.001)) %>%
  ungroup()

## add in species group identifier

data_df <- data %>%
  mutate(row_num = row_number()) %>%
  mutate(name = paste("V", row_num, sep =""))

spp_group_dist <- dist_ports_final %>%
  ungroup() %>%
  left_join(data_df) # %>%
  # filter(closest_port_km >= 1)

mean(spp_group_dist$closest_port_km) # 31.30463 km
sd(spp_group_dist$closest_port_km) # 32.333 km
range(spp_group_dist$closest_port_km) #  0.1312871 to 531.2067646
median(spp_group_dist$closest_port_km) # 18.85009 km

summary(spp_group_dist)

spp_group_dist_final <- spp_group_dist %>%
  group_by(type) %>%
  summarise(dist_port = mean(closest_port)) %>%
  ungroup()

```