### 2021-07-19 analysis/match_data_sources.Rmd

---
title: "Match point to sources"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "04/26/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script takes the farms dataset produced from prior steps, and matches to our data source spreadsheet, so that each point in the final database has a source associated with it (or an NA, if it was estimated).

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```

# Global regions

```{r}

# this file includes ocean regions:
gadm <- raster("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm36mol_pyfill_10km.tif")

gadm_latlon <- raster::projectRaster(gadm, food_raster, method="ngb", over=TRUE)
# plot(gadm_latlon)

gadm_codes <- read.csv("/home/shares/food-systems/Food_footprint/_raw_data/Country_regions/gadm_reg_codes.csv") %>% 
  mutate(reg_id = as.numeric(reg_id))

```

# read in points and convert to spatial dataframe
```{r}

data <- read_csv(file.path("marine/output/all_marine_farms.csv")) %>% 
  st_as_sf(., 
           coords = c("lon", "lat"),
           crs = crs(food_raster), 
           agr = "constant")

all_farms_to_place <- read_csv(file.path("marine/output/all_points_to_add.csv")) %>%
  distinct(iso3c, type, estimated)

all_farms <- readRDS("data/all_modeled_marine_farms_31km.rds") %>%
  left_join(all_farms_to_place) %>%
    mutate(data_type = case_when(
    source == "real" ~ "known locations", 
    source == "modeled" & estimated == 0 ~ "modeled locations, known number of farms",
    source == "modeled" & estimated == 1 ~ "modeled locations, estimated number of farms"
  )) %>%
  mutate(estimated = ifelse(is.na(estimated), 0, estimated))

summary(all_farms)

test <- all_farms %>%
  st_drop_geometry() %>%
  group_by(data_type) %>% 
  summarise(n()) ## perfect!!

test <- all_farms %>%
  dplyr::filter(is.na(estimated))

```

# Match to our sources dataset 
```{r}

# read in our sources spreadsheet (manually uploaded by GC)
sources <- read_csv("analysis/data/mariculture_mapping_sources.csv") %>%
  clean_names() %>%
  mutate(type = case_when(species == "Salmonids" ~ "salmon",
                          species == "Unfed or algae fed shellfish" ~ "bivalve",
                          species == "Shrimp" ~ "shrimp", 
                          species == "Tuna" ~ "tuna", 
                          species == "General Marine Fish" ~ "marine_fish_general", 
                          species == "Crustaceans" ~ "crustaceans")) %>%
mutate(species_group = case_when(species == "Salmonids" ~ "Salmonidae fish",
                          species == "Unfed or algae fed shellfish" ~ "Unfed or algae fed bivalve molluscs",
                          species == "Shrimp" ~ "Shrimps and prawns", 
                          species == "Tuna" ~ "Bluefin tuna", 
                          species == "General Marine Fish" ~ "General marine fish", 
                          species == "Crustaceans" ~ "Non-shrimp crustaceans")) %>% 
  mutate(data_year_number_of_farms = as.character(data_year_number_of_farms)) %>%
  mutate(data_type_2 = case_when(
    known_farms_location > 0 & known_farms_no_location == 0 & estimated_farms == 0 ~ "A",
    known_farms_location == 0 & known_farms_no_location > 0 & estimated_farms == 0 ~ "B",
    known_farms_location > 0 & known_farms_no_location > 0 & estimated_farms == 0 ~ "C",
    known_farms_location > 0 & known_farms_no_location == 0 & estimated_farms > 0 ~ "D",
    known_farms_location == 0 & known_farms_no_location == 0 & estimated_farms > 0 ~ "F",
  )) %>%
  mutate(data_type_2 = ifelse(str_detect(details, "Although"), "D", data_type_2)) %>%
  mutate(data_type_2 = ifelse(iso3c == "IND" & species == "Unfed or algae fed shellfish", "E", data_type_2)) ## fix some mismatches

## data types with comprehensive # of farms are A, B, and C. These all recieve their production/# of farms to get prod per farm.

## data types D, and E recieve global prod per farm estimates.

## read in all species production

biv_prod <- read_csv("marine/bivalve/bivalve_farms/data/fao_mariculture_bivalve.csv") %>%
  mutate(type = "bivalve")
crust_prod <- read_csv("marine/crustaceans/crustaceans_farms/data/fao_mariculture_crustaceans.csv") %>%
  mutate(type = "crustaceans")
finfish_prod <- read_csv("marine/marine_fish_general/finfish_farms/data/fao_mariculture_finfish.csv") %>%
  mutate(type = "marine_fish_general")
salmon_prod <- read_csv("marine/salmon/salmon_farms/data/fao_mariculture_salmon.csv") %>%
  mutate(type = "salmon")
shrimp_prod <- read_csv("marine/shrimp/shrimp_farms/data/fao_mariculture_shrimp.csv") %>%
  mutate(type = "shrimp")
tuna_prod <- read_csv("marine/tuna/tuna_farms/data/fao_mariculture_tuna.csv") %>%
  mutate(type = "tuna")

all_prod <- rbind(biv_prod, crust_prod, finfish_prod, salmon_prod, shrimp_prod, tuna_prod) %>%
  filter(year == 2017)

spp_prod_per_farm <- read_csv("analysis/data/prod_per_farm.csv")

sources_prod <- sources %>%
  left_join(all_prod, by = c("iso3c", "type")) %>%
  left_join(spp_prod_per_farm) %>% 
  mutate(tonnes_per_farm = case_when(
        str_detect(details, "Although") ~ fao_tonnes_production/number_of_total_farms,
    data_type_2 %in% c("A", "B", "C") ~ fao_tonnes_production/number_of_total_farms,
    data_type_2 %in% c("D", "F") ~ spp_prod_per_farm,
    data_type_2 %in% c("E") ~ fao_tonnes_production/number_of_total_farms  )) %>%
  dplyr::select(-country.y, -spp_prod_per_farm) %>%
  rename("country" = "country.x")

sources_prod_SI_table <- sources_prod %>%
  dplyr::select(country, species_group, fao_tonnes_production, number_of_total_farms, tonnes_per_farm, data_type_2) %>%
  arrange(country)

write.csv(sources_prod_SI_table, "analysis/tables/SI_table_3.csv", row.names = FALSE)

## join all farms with sources data 
all_farms_sources <- all_farms %>%
  left_join(sources_prod, by = c("iso3c", "type")) %>%
  mutate(data_source = case_when(
    data_type == "modeled locations, known number of farms" ~ datasource_number_of_farms,
    data_type == "modeled locations, estimated number of farms" ~ "NA", 
    data_type == "known locations" ~ datasource_farm_locations), 
  data_year = case_when(
    data_type == "modeled locations, known number of farms" ~ data_year_number_of_farms,
    data_type == "modeled locations, estimated number of farms" ~ "NA", 
    data_type == "known locations" ~ data_year_farm_locations)) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, details, data_source, data_year, tonnes_per_farm, geometry)

unique(all_farms_sources$data_year)

```

# Fix instances where a known location has multiple sources
```{r}
test <- all_farms_sources %>%
  filter(iso3c == "AUS", species == "Unfed or algae fed shellfish")
unique(all_farms_sources$species)

test <- all_farms_sources %>%
  filter(iso3c == "IND", species == "Unfed or algae fed shellfish")

## Have multiple location data sources...  fix these so that each point has only one of the datasources

# Salmon: CAN, USA
# Shrimp: USA 
# Bivalve: AUS, CAN, USA
# General marine fish: TWN

## add in subregional information
all_farms_sources_gadm <- all_farms_sources
all_farms_sources_gadm$reg_id <- raster::extract(gadm_latlon, all_farms_sources)

all_farms_sources_gadm <- all_farms_sources_gadm %>%
  left_join(gadm_codes, by="reg_id") %>%
  dplyr::mutate(NAME_1 = as.character(NAME_1))

## Canada Salmon fix: 
CAN_salmon_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "CAN", 
               species == "Salmonids") %>%
  separate(col = data_source, into = c("ds_1", "ds_2", "ds_3", "ds_4"), sep = ";") %>%
  separate(col = data_year, into = c("dy_1", "dy_2", "dy_3", "dy_4"), sep = ",") %>%
  mutate(data_source = case_when(
    NAME_1 == "British Columbia" ~ ds_1, 
    NAME_1 %in% c("New Brunswick", "Maine") ~ ds_2, 
    NAME_1 %in% c("Newfoundland and Labrador") ~ ds_4,
    NAME_1 == "Nova Scotia" ~ ds_3

  )) %>%
  mutate(data_year = case_when(
    NAME_1 == "British Columbia" ~ dy_1, 
    NAME_1 %in% c("New Brunswick", "Maine") ~ dy_2, 
    NAME_1 %in% c("Newfoundland and Labrador") ~ dy_4,
    NAME_1 == "Nova Scotia" ~ dy_3

  )) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

check <- CAN_salmon_fix %>%
  st_drop_geometry() %>%
  group_by(data_source) %>%
  summarise(n()) ## numbers are correct

## Canada Bivalve fix: 
CAN_bivalve_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "CAN", 
               species == "Unfed or algae fed shellfish") %>%
  separate(col = data_source, into = c("ds_1", "ds_2"), sep = ";") %>%
  mutate(data_source = case_when(
    NAME_1 == "British Columbia" ~ ds_1, 
    NAME_1 == "Nova Scotia" ~ ds_2

  )) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

check <- CAN_bivalve_fix %>%
  st_drop_geometry() %>%
  group_by(data_source) %>%
  summarise(n()) ## numbers are correct

## USA Salmon fix: 
USA_salmon_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "USA", 
               species == "Salmonids") %>%
  separate(col = data_source, into = c("ds_1", "ds_2"), sep = ";") %>%
  separate(col = data_year, into = c("dy_1", "dy_2"), sep = ",") %>%
  mutate(data_source = case_when(
    NAME_1 %in% c("New Brunswick", "Maine") ~ ds_2, 
    NAME_1 == "Washington" ~ ds_1
  )) %>%
    mutate(data_year = case_when(
    NAME_1 %in% c("New Brunswick", "Maine") ~ dy_2, 
    NAME_1 == "Washington" ~ dy_1
  )) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

check <- USA_salmon_fix %>%
  st_drop_geometry() %>%
  group_by(data_source) %>%
  summarise(n()) ## numbers are correct

## USA Shrimp fix: 
USA_shrimp_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "USA", 
               species == "Shrimp") %>%
  separate(col = data_source, into = c("ds_1", "ds_2", "ds_3", "ds_4", "ds_5"), sep = ";") %>%
  separate(col = data_year, into = c("dy_1", "dy_2"), sep = ",") %>%
  mutate(data_source = case_when(
    NAME_1 %in% c("Texas") ~ ds_2, 
    NAME_1 == "Alabama" ~ ds_3, 
    NAME_1 == "Hawaii" ~ ds_4, 
    NAME_1 == "Florida" ~ ds_5
  )) %>%
    mutate(data_year = case_when(
    NAME_1 %in% c("Texas") ~ dy_1, 
    NAME_1 == "Alabama" ~ dy_2, 
    NAME_1 == "Hawaii" ~ dy_2, 
    NAME_1 == "Florida" ~ dy_2
  )) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

## USA Bivalve fix: 
USA_bivalve_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "USA", 
               species == "Unfed or algae fed shellfish", 
               data_type == "known location") %>%
  separate(col = data_source, into = c("ds_1", "ds_2"), sep = ";") %>%
  mutate(data_source = ifelse(NAME_1 == "Washington"| NAME_1 == "Rhode Island", ds_2, ds_1)) %>% 
  mutate(data_year = 2018) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

check <- USA_bivalve_fix %>%
  st_drop_geometry() %>%
  group_by(data_source) %>%
  summarise(n()) ## numbers are correct

## AUS Bivalve fix: 
AUS_bivalve_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "AUS", 
               species == "Unfed or algae fed shellfish", 
               data_type == "known location") %>%
  separate(col = data_source, into = c("ds_1", "ds_2"), sep = ";") %>%
  mutate(data_source = ifelse(NAME_1 == "Victoria", ds_1, ds_2)) %>%
  mutate(data_year = 2021) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

## TWN marine fish fix: 
# read in taiwan sources sheet
twn_sources <- read_csv(here("marine/marine_fish_general/finfish_farms/data/taiwan_farms_sources.csv"))

twn_sources_sf <- st_as_sf(twn_sources, coords = c("lon", "lat"), 
                 crs = crs(gadm_latlon), agr = "constant")

## rows 1-40 are lee, rows 41-66 are kuo
TWN_finfish_fix <- all_farms_sources_gadm %>%
  dplyr::filter(iso3c == "TWN", 
               species == "General Marine Fish", 
               data_type == "known location") %>%
  separate(col = data_source, into = c("ds_1", "ds_2"), sep = ";") %>% 
  separate(col = data_year, into = c("dy_1", "dy_2"), sep = ",") %>% 
  mutate(row = row_number()) %>%
  mutate(data_source = ifelse(row %in% c(1:40), ds_1, ds_2)) %>%
  mutate(data_year = ifelse(row %in% c(1:40), dy_1, dy_2)) %>%
  dplyr::select(iso3c, country, species_group, species, data_type, data_type_2, tonnes_per_farm, details, data_source, data_year, geometry)

##### Now add these back into total dataset #####

all_farms_sources_final <- all_farms_sources %>%
  filter(iso3c != "CAN"|species != "Salmonids") %>%
  filter(iso3c != "CAN"|species != "Unfed or algae fed shellfish") %>%
  filter(iso3c != "USA"|species != "Salmonids") %>%
  filter(iso3c != "USA"|species != "Unfed or algae fed shellfish"|data_type != "known location") %>%
  filter(iso3c != "AUS"|species != "Unfed or algae fed shellfish"|data_type != "known location") %>%
  filter(iso3c != "TWN"|species != "General Marine Fish"|data_type != "known location") %>%
  filter(iso3c != "USA"|species != "Shrimp") %>%
  rbind(TWN_finfish_fix, USA_bivalve_fix, USA_salmon_fix, USA_shrimp_fix, CAN_bivalve_fix, CAN_salmon_fix, AUS_bivalve_fix) %>%
  mutate(data_year = as.numeric(data_year)) %>%
  dplyr::select(-species)

summary(all_farms_sources_final)
unique(all_farms_sources_final$data_year)
unique(sort(all_farms_sources$country))
unique(sort(all_farms_sources$iso3c)) ## 73 countries - perfect! 
unique(all_farms_sources_final$species_group)

## Now save this as a csv
st_write(all_farms_sources_final, "all_marine_aquaculture_farms_sources_31km.csv", layer_options = "GEOMETRY=AS_XY", delete_dsn = TRUE)

## check that it wrote correctly

all_farms <- read_csv("all_marine_aquaculture_farms_sources_31km.csv")

test <- all_farms %>%
  filter(iso3c == "IND", str_detect(species_group, "bivalve"))
```