### 2021-07-19 analysis/validation_testing.Rmd

---
title: "Validation Testing"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "07/08/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script tests our validation set of points against our actual known points.

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(ggplot2)
library(fasterize)
library(tabularaster)
library(countrycode)
library(caret)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

cell_rast_df <- raster_df(food_raster) %>%
  dplyr::select(1:2)

## establish CRSs

## food 81km2 at equator 
eez_81 <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/eez_validation_mask.tif"))
plot(area(eez_81))

## 36km2 equal area
eez_36_cea <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/eez_validation_mask_cea_36km.tif"))

new_crs <- "+proj=cea +lat_ts=45 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"
new_raster <- raster(nrows=3000, ncols=4736, xmn=-14207430, xmx=14208570, ymn=-9000182, ymx=8999818, crs = new_crs)

## 25km2 equal area
eez_25_cea <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/eez_validation_mask_cea_25km.tif"))

cea_25_crs <- crs(eez_25_cea)
plot(eez_25_cea)

## 1/4 degree cells 
eez_quart_cell <- raster(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/global_maps/eez_validation_mask_quarter_degree.tif"))

quarter_crs <- crs(eez_quart_cell)

```

Read in validation dataset

```{r}
validation_all <- readRDS(file.path("data/all_validation_farms.rds"))  %>%
  mutate(species_group = case_when(type == "salmon" ~ "Salmonidae fish",
                          type == "bivalve" ~ "Unfed or algae fed bivalve molluscs",
                          type == "shrimp" ~ "Shrimps and prawns", 
                          type == "tuna" ~ "Bluefin tuna", 
                          type == "marine_fish_general" ~ "General marine fish"))

table_3 <- read_csv("analysis/tables/SI_table_3.csv")

iso <-   data.frame(iso3c = countrycode(table_3$country, origin = 'country.name', destination = 'iso3c'))

table_3_iso3c <- table_3 %>%
  cbind(iso)

data_type_val <- validation_all %>%
  left_join(table_3_iso3c, by = c("iso3c", "species_group"))

data_type_A <- data_type_val %>%
  filter(data_type_2 == "A")

mapview(data_type_A, zcol = "source")

table_3_A <- table_3_iso3c %>%
  filter(data_type_2 == "A")

```

Split data

```{r}
validation_farms <- data_type_A %>%
 filter(source == "modeled validation") %>%
  mutate(count_farms = 1)

val_farms_cea_36 <- validation_farms %>%
  st_transform(crs = new_crs)

val_farms_cea_25 <- validation_farms %>%
  st_transform(crs = cea_25_crs)

val_farms_quarter <- validation_farms %>%
  st_transform(crs = quarter_crs)

real_farms <- data_type_A %>%
  filter(source == "real") %>%
  mutate(count_farms = 1)

real_farms_cea_36 <- real_farms %>%
  st_transform(crs = new_crs)

real_farms_cea_25 <- real_farms %>%
  st_transform(crs = cea_25_crs)

real_farms_quart <- real_farms %>%
  st_transform(crs = quarter_crs)
```

Rasterize data to count number of farms per raster cell with 36km2 equal area

```{r}

val_rast <- rasterize(val_farms_cea_36, new_raster, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum")

plot(val_rast, col = "black")
#s <- raster::select(val_rast)
#plot(s)

real_rast <- rasterize(real_farms_cea_36, new_raster, field = "count_farms", fun = "sum")
plot(real_rast)

cellStats(real_rast, "sum") # 8455
plot(real_rast, col = "black")

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8455    8455

```

Use our EEZ mask to create a confusion matrix for CEA 36km2 cell sizes

 - AUC of ~0.61
```{r}

plot(eez_36_cea, col = "black")

world_coast <- as.data.frame(eez_36_cea) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask_cea_36km))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8446

sum(eez_cells_df$real, na.rm = TRUE) # 8377

## still missing some farms.. but nothing egregious. This is probably ok. The results wouldn't change much even if we had the extra ~80 farms

test <- rast_stack*eez_36_cea

cellStats(test,"sum") 
# layer.1 layer.2 
#    8455    7917

## we are losing some farms...

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

confusion_mat

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction       0       1
#          0 1171915    2492
#          1    2762     695
#                                           
#                Accuracy : 0.9955          
#                  95% CI : (0.9954, 0.9957)
#     No Information Rate : 0.9973          
#     P-Value [Acc > NIR] : 1.0000000       
#                                           
#                   Kappa : 0.207           
#                                           
#  Mcnemar's Test P-Value : 0.0002063       
#                                           
#             Sensitivity : 0.9976          
#             Specificity : 0.2181          
#          Pos Pred Value : 0.9979          
#          Neg Pred Value : 0.2010          
#              Prevalence : 0.9973          
#          Detection Rate : 0.9949          
#    Detection Prevalence : 0.9971          
#       Balanced Accuracy : 0.6079          
#                                           
#        'Positive' Class : 0

library(pROC)
roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE) # 0.6019

library(pROC)
library(randomForest)

roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE, percent = TRUE, xlab = "False Positive Percentage", ylab = "True Positive Percentage", col = "#377eb8", lwd = 4, print.auc = TRUE) # 0.6068

roc.info <- roc(eez_cells_df$real_count, eez_cells_df$pred_count, legacy.axes = TRUE)

roc.df <- data.frame(tpp = roc.info$sensitivities*100,
                     fpp = (1-roc.info$specificities)*100,
                     thresholds = roc.info$thresholds)

#### ROC curve try #2
# https://www.datatechnotes.com/2019/03/how-to-create-roc-curve-in-r.html

library(ROCR)

pred = eez_cells_df$pred_count

pred = prediction(pred, eez_cells_df$real_count)

perf = performance(pred, "acc")
plot(perf)

perf_cost <- performance(pred, "cost")
plot(perf_cost)

roc = performance(pred, "tpr", "fpr")

plot(roc, colorize = T, lwd=2)
abline(a=0, b=1)

auc = performance(pred, measure = "auc")
print(auc@y.values)
# 0.6091016

```

Use our EEZ mask to create a confusion matrix for CEA 25km2 cell sizes

 - AUC of ~0.59 - worse than 36km2... makes sense! 
```{r}

val_rast <- rasterize(val_farms_cea_25, eez_25_cea, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum")

plot(val_rast, col = "black")
#s <- raster::select(val_rast)
#plot(s)

real_rast <- rasterize(real_farms_cea_25, eez_25_cea, field = "count_farms", fun = "sum")
plot(real_rast)

cellStats(real_rast, "sum") # 8455
plot(real_rast, col = "black")

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8455    8455

plot(eez_25_cea, col = "black")

world_coast <- as.data.frame(eez_25_cea) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask_cea_25km))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8450

sum(eez_cells_df$real, na.rm = TRUE) # 8386

## still missing some farms.. but nothing egregious. This is probably ok. The results wouldn't change much even if we had the extra ~80 farms. This is likely due to reprojection weirdness

test <- rast_stack*eez_25_cea

cellStats(test,"sum") 
# layer.1 layer.2 
#    8450    8386

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

confusion_mat

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction       0       1
#          0 1689389    2864
#          1    3071     654
#                                           
#                Accuracy : 0.9965          
#                  95% CI : (0.9964, 0.9966)
#     No Information Rate : 0.9979          
#     P-Value [Acc > NIR] : 1.000000        
#                                           
#                   Kappa : 0.1788          
#                                           
#  Mcnemar's Test P-Value : 0.007496        
#                                           
#             Sensitivity : 0.9982          
#             Specificity : 0.1859          
#          Pos Pred Value : 0.9983          
#          Neg Pred Value : 0.1756          
#              Prevalence : 0.9979          
#          Detection Rate : 0.9961          
#    Detection Prevalence : 0.9978          
#       Balanced Accuracy : 0.5920          
#                                           
#        'Positive' Class : 0

library(pROC)
roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE) # 0.6019

library(pROC)
library(randomForest)

roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE, percent = TRUE, xlab = "False Positive Percentage", ylab = "True Positive Percentage", col = "#377eb8", lwd = 4, print.auc = TRUE) # 0.6068

roc.info <- roc(eez_cells_df$real_count, eez_cells_df$pred_count, legacy.axes = TRUE)

roc.df <- data.frame(tpp = roc.info$sensitivities*100,
                     fpp = (1-roc.info$specificities)*100,
                     thresholds = roc.info$thresholds)

#### ROC curve try #2
# https://www.datatechnotes.com/2019/03/how-to-create-roc-curve-in-r.html

library(ROCR)

pred = eez_cells_df$pred_count

pred = prediction(pred, eez_cells_df$real_count)

perf = performance(pred, "acc")
plot(perf)

perf_cost <- performance(pred, "cost")
plot(perf_cost)

roc = performance(pred, "tpr", "fpr")

plot(roc, colorize = T, lwd=2)
abline(a=0, b=1)

auc = performance(pred, measure = "auc")
print(auc@y.values)
# 0.5920433

```

Use our EEZ mask to create a confusion matrix for 0.25 X 0.25 degree cell sizes (about 769.3152km2 at equator). Average cell size is 490.3108 km2

 - AUC of ~0.7241861 ; makes sense that is it better!

```{r}
test <- raster::area(eez_quart_cell)

cellStats(test, "mean") # 490.3108

mean(test@data@values) # 490.3108

val_rast <- rasterize(val_farms_quarter, eez_quart_cell, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum")

plot(val_rast, col = "black")
#s <- raster::select(val_rast)
#plot(s)

real_rast <- rasterize(real_farms_quart, eez_quart_cell, field = "count_farms", fun = "sum")
plot(real_rast)

cellStats(real_rast, "sum") # 8455
plot(real_rast, col = "black")

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8455    8455

plot(eez_quart_cell, col = "black")

world_coast <- as.data.frame(eez_quart_cell) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask_quarter_degree))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8124

sum(eez_cells_df$real, na.rm = TRUE) # 8080

## still missing some farms.. but nothing egregious. This is probably ok. The results wouldn't change much even if we had the extra farms. This is likely due to reprojection weirdness

test <- rast_stack*eez_quart_cell

cellStats(test,"sum") 
# layer.1 layer.2 
#    8124    8080

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

confusion_mat

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction     0     1
#          0 84497   619
#          1   982   527
#                                           
#                Accuracy : 0.9815          
#                  95% CI : (0.9806, 0.9824)
#     No Information Rate : 0.9868          
#     P-Value [Acc > NIR] : 1               
#                                           
#                   Kappa : 0.3878          
#                                           
#  Mcnemar's Test P-Value : <2e-16          
#                                           
#             Sensitivity : 0.9885          
#             Specificity : 0.4599          
#          Pos Pred Value : 0.9927          
#          Neg Pred Value : 0.3492          
#              Prevalence : 0.9868          
#          Detection Rate : 0.9754          
#    Detection Prevalence : 0.9826          
#       Balanced Accuracy : 0.7242          
#                                           
#        'Positive' Class : 0

recall <- 527/(527+619) # 0.46

precision <- 527 / (527 + 982) # 0.3492379

accuracy <- (84497 + 982)/(84497 + 982 + 527 + 619) # 0.9867706

f_measure <- (2*recall*precision)/(recall + precision) # 0.33

specificity = (84497)/(84497 + 982) # 0.9885118

balanced_accuracy <- (recall + specificity)/2 # 0.7242 - this is also the AUC - which is pretty good.

g_measure <- sqrt(precision*recall) # 0.401

library(pROC)
roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE) # 0.7242

library(pROC)
library(randomForest)

roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE, percent = TRUE, xlab = "False Positive Percentage", ylab = "True Positive Percentage", col = "#377eb8", lwd = 4, print.auc = TRUE) # 72.42

roc.info <- roc(eez_cells_df$real_count, eez_cells_df$pred_count, legacy.axes = TRUE)

roc.df <- data.frame(tpp = roc.info$sensitivities*100,
                     fpp = (1-roc.info$specificities)*100,
                     thresholds = roc.info$thresholds)

#### ROC curve try #2
# https://www.datatechnotes.com/2019/03/how-to-create-roc-curve-in-r.html

library(ROCR)

pred = eez_cells_df$pred_count

pred = prediction(pred, eez_cells_df$real_count)

perf = performance(pred, "acc")
plot(perf)

perf_cost <- performance(pred, "cost")
plot(perf_cost)

roc = performance(pred, "tpr", "fpr")

plot(roc, colorize = T, lwd=2)
abline(a=0, b=1)

auc = performance(pred, measure = "auc")
print(auc@y.values)
# 0.7241861

```

Use our EEZ mask to create a confusion matrix for XX degree cell sizes (about 81km2 at equator)

 - AUC of ~ 0.5950; about the same as the 25km2 size. Makes sense, because the average area is somewhere around there (since this isnt equal area). 
```{r}

val_rast <- rasterize(validation_farms, eez_81, field = "count_farms", fun = "sum")
plot(val_rast)

cellStats(val_rast, "sum")

plot(val_rast, col = "black")
#s <- raster::select(val_rast)
#plot(s)

real_rast <- rasterize(real_farms, eez_81, field = "count_farms", fun = "sum")
plot(real_rast)

cellStats(real_rast, "sum") # 8455
plot(real_rast, col = "black")

rast_stack <- stack(val_rast, real_rast)
cellStats(rast_stack, "sum")

# layer.1 layer.2 
#    8455    8455

plot(eez_81, col = "black")

world_coast <- as.data.frame(eez_81) %>%
  mutate(cell = row_number()) %>%
  filter(!is.na(eez_validation_mask))

eez_cells_df <- as.data.frame(rast_stack) %>%
  rename("predicted" = "layer.1", "real" = "layer.2") %>%
  mutate(pred_count = ifelse(predicted >=1, 1, 0),
         real_mod_count = ifelse(predicted >=1&real>=1, 1, 0)) %>%
  mutate(real_count = ifelse(real>=1,1,0)) %>%
  mutate(pred_count = ifelse(is.na(pred_count), 0, pred_count),
         real_count = ifelse(is.na(real_count), 0, real_count), 
         real = ifelse(is.na(real), 0, real),
         predicted = ifelse(is.na(predicted), 0, predicted)) %>%
  mutate(cell = row_number()) %>% 
  filter(cell %in% c(world_coast$cell))

sum(eez_cells_df$predicted, na.rm = TRUE) # 8452

sum(eez_cells_df$real, na.rm = TRUE) # 8350

## still missing some farms.. but nothing egregious. This is probably ok. The results wouldn't change much even if we had the extra farms. This is likely due to reprojection weirdness

test <- rast_stack*eez_81

cellStats(test,"sum") 
# layer.1 layer.2 
#    8124    8080

#### make a confusion matrix

confusion_mat <- caret::confusionMatrix(data = as.factor(eez_cells_df$pred_count), reference = as.factor(eez_cells_df$real_count))

confusion_mat

# Confusion Matrix and Statistics
# 
#           Reference
# Prediction      0      1
#          0 757504   2266
#          1   2137    541
#                                           
#                Accuracy : 0.9942          
#                  95% CI : (0.9941, 0.9944)
#     No Information Rate : 0.9963          
#     P-Value [Acc > NIR] : 1.00000         
#                                           
#                   Kappa : 0.1944          
#                                           
#  Mcnemar's Test P-Value : 0.05373         
#                                           
#             Sensitivity : 0.9972          
#             Specificity : 0.1927          
#          Pos Pred Value : 0.9970          
#          Neg Pred Value : 0.2020          
#              Prevalence : 0.9963          
#          Detection Rate : 0.9935          
#    Detection Prevalence : 0.9965          
#       Balanced Accuracy : 0.5950          
#                                           
#        'Positive' Class : 0

library(pROC)
roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE) # 0.5950

library(pROC)
library(randomForest)

roc(eez_cells_df$real_count, eez_cells_df$pred_count, plot = TRUE, percent = TRUE, xlab = "False Positive Percentage", ylab = "True Positive Percentage", col = "#377eb8", lwd = 4, print.auc = TRUE) # 0.5950

roc.info <- roc(eez_cells_df$real_count, eez_cells_df$pred_count, legacy.axes = TRUE)

roc.df <- data.frame(tpp = roc.info$sensitivities*100,
                     fpp = (1-roc.info$specificities)*100,
                     thresholds = roc.info$thresholds)

#### ROC curve try #2
# https://www.datatechnotes.com/2019/03/how-to-create-roc-curve-in-r.html

library(ROCR)

pred = eez_cells_df$pred_count

pred = prediction(pred, eez_cells_df$real_count)

perf = performance(pred, "acc")
plot(perf)

perf_cost <- performance(pred, "cost")
plot(perf_cost)

roc = performance(pred, "tpr", "fpr")

plot(roc, colorize = T, lwd=2)
abline(a=0, b=1)

auc = performance(pred, measure = "auc")
print(auc@y.values)
# 0.5950

```

Resampling test: 
```{r}

### Manual equal rebalancing
z_ind <- which(eez_cells_df$pred_count == 1)
zz_ind <- which(eez_cells_df$pred_count == 0)

nsamp = 3455
pick_z <- sample(z_ind, nsamp)
pick_zz <- sample(zz_ind, nsamp)

new_data <- eez_cells_df[c(pick_z, pick_zz), ]

confusion_mat <- caret::confusionMatrix(data = as.factor(new_data$pred_count), reference = as.factor(new_data$real_count))

confusion_mat

library(ROCR)

pred = new_data$pred_count

pred = prediction(pred, new_data$real_count)

perf = performance(pred, "acc")
plot(perf)

perf_cost <- performance(pred, "cost")
plot(perf_cost)

roc = performance(pred, "tpr", "fpr")

plot(roc, colorize = T, lwd=2)
abline(a=0, b=1)

auc = performance(pred, measure = "auc")
print(auc@y.values)
# 0.7264348

#### Over/undersampling 
install.packages("ROSE")
library(ROSE)

test <- eez_cells_df %>%
  filter(pred_count == 1)

## not sure what the statistically correct way of picking the sampling number is... 
data_balanced_under <- ovun.sample(pred_count ~ real_count, data = eez_cells_df, method = "under", N = 8455)
table(data_balanced_under$data)

#           real_count
# pred_count    0    1
#          0 4894  106
#          1 2761  694

## still pretty unbalanced, but it is a lot better than it was...

recall <- 694/(694+106) # 0.8675

precision <- 694 / (694 + 2761) # 0.20

accuracy <- (4894 + 2761)/(4894 + 2761 + 106 + 694) # 0.91

f_measure <- (2*recall*precision)/(recall + precision) # 0.33

sensitivity = (4894)/(4894 + 2761) # 0.64

balanced_accuracy <- (recall + sensitivity)/2 # 0.75 - this is also the AUC - which is pretty good.

g_measure <- sqrt(precision*recall) # 0.42

false_pos_rate <- 2761/(2761+4894) # 0.36

```