### 2021-08-12 marine/crustaceans/crustaceans_farms/STEP1_fao_crustacean_production.Rmd

---
title: "FAO crustacean production"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "12/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Clean FAO crustacean data

All data is ultimately adjusted to FAO production data.  This script cleans the fao mariculture data and subsets it to to the relevant years of data and species (crustaceans).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(data.table)
library(countrycode)
library(here)
library(stringr)

path <- "/home/shares/food-systems/Food_footprint/_raw_data"
source(here('marine/salmon/salmon_farm/R/fao_mar_clean.R')) # functions specific to mariculture dealing with compound countries
source('https://raw.githubusercontent.com/OHI-Science/ohiprep_v2019/gh-pages/workflow/R/fao_fxn.R') # function for cleaning FAO files
source(here('workflow/common.R'))

```

Subset the data to include only the species (crustacean) and countries we are interested in. Then summarize by total tonnes of crustacean per country/year. (There are multiple crustacean species)

```{r}

fao <- read.csv(here("data/fao_aquaculture_clean.csv")) %>%
  filter(environment %in% c("Marine", "Brackishwater"))

## Read in fao data and get a species list for crustaceans 
species_list <- read.csv(here("marine/STEP1_species_groups/int/marine_groups.csv"))  %>%
  distinct(species, aq_group)

# the list of species we are including
crustaceans_list <- species_list$species[species_list$aq_group=="crustaceans"]
crustaceans_list

## write out crustacean taxa and production
fao_crustaceans_taxa_prod <- fao %>%
  filter(species %in% crustaceans_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, species, year) %>%
  dplyr::filter(year == 2017) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, species, year) %>%
  ungroup() %>%
  mutate(total = sum(fao_tonnes_production)) %>%
  mutate(perc = fao_tonnes_production/total)

## write out a file with the specific taxa included  
write.csv(fao_crustaceans_taxa_prod, here("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans_taxa_production.csv"), row.names=FALSE)

# include only crustacean
fao_crustaceans <- fao %>%
  filter(species %in% crustaceans_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  filter(value>=500) %>%  # countries with less do not appear to be stable producers at this point
  group_by(country, iso3c, year) %>%
  summarize(fao_tonnes_production = sum(value)) %>%
  arrange(country, iso3c, year)

write.csv(fao_crustaceans, here("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans.csv"), row.names=FALSE)

## explore the data     
 explore <- read_csv(here("marine/crustaceans/crustacean_farms/data/fao_mariculture_crustaceans.csv")) %>%
   filter(year==2017) %>%
   arrange(-fao_tonnes_production) %>%
   mutate(percent = round((fao_tonnes_production/sum(fao_tonnes_production))*100,2))

sum(explore$fao_tonnes_production) # 401719.1
unique(explore$country) # 6 countries

```

```{r}
## look at crustacean taxa and production
fao_crustaceans_taxa_prod <- fao %>%
  filter(species %in% crustaceans_list) %>%  
  filter(!is.na(value)) %>%
  arrange(-value, country) %>%
  group_by(country, year) %>%
  summarise(value = sum(value)) %>%
  filter(value>=500) %>% # countries with less do not appear to be stable producers at this point
  filter(year == 2017)

sum(fao_crustaceans_taxa_prod$value) # 401733.5 - adds 14 tonnes

unique(fao_crustaceans_taxa_prod$country) # 6 countries 
```