### 2021-10-14 marine/united_states/US_mapping.Rmd
started looking into state-by-state US mariculture locations

---
title: "Location of United States Aquaculture Farms"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "10/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script maps all known US mariculture farms

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(raster)
library(tidyverse)
library(stringr)
library(PBSmapping)
library(maptools)
library(rgdal)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(tmap)
library(leaflet)
library(ggplot2)
library(maps)
library(mapdata)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```

# Raster

Standard raster used for this analysis

```{r, include = FALSE}

source(here("_spatial/template_raster.R"))

```

## US Northeast

Source:
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 17, 2020] 
https://www.northeastoceandata.org/data-download/

"This dataset should be considered a work in progress"

States:

 - NY
 - CT
 - ME
 - NH
 - MA
 - Federal waters

Species: 
 - "Shellfish"  "Multi-trophic or other" "Seaweed" "Finfish"

**This data set is not "comprehensive" but provides good sources to look through for individual states (see below). These are polygons.**

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}

us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))

us_ne_active <- us_ne %>%
  filter(status != "Unknown")

us_ne_active$area = st_area(us_ne_active)

# unique(us_ne_active$state) # "Federal waters" "NY"             "CT"             "MA"             "ME"             "NH"   
# unique(us_ne_active$genSpecies) # [1] "Shellfish"  "Multi-trophic or other" "Seaweed" "Finfish"

usa <- map_data('usa')

state <- map_data("state")

ne <- subset(state, region %in% c("new york", "maine", "new hampshire", "massachusetts", "connecticut"))

  ggplot() +
  geom_sf(data = us_ne_active, col = "red") + 
  geom_polygon(data=ne, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "US NE Regional Ocean Council data", caption = "Not comprehensive")

```

## Maine:

 - https://www.maine.gov/dmr/aquaculture/leases/aquaculturemap.html
 - I think we will want to use both "limited purpose" and regular leases.
 - Limited purpose leases are for scientific or smaller/shorter term mariculture projects, that last up to 3 years.
 - I think this can be considered "comprehensive". 
 - Update to date as of September 2021.
 - Has species types specified. 
 - Red polygons for the leases. Blue points for the limited purpose. Limited purpose has length and width measurements, so regardless we will have area for both if needed.

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
maine_aq_lease <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_AQ_Leases/MaineDMR_Aquaculture_-_AQ_Leases.shp"))

# mapview(maine_aq_lease)

maine_aq_lease_active <- maine_aq_lease %>%
  dplyr::filter(STATUS == "A") # 149 leases

```

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
maine_aq_lp <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_Limited_Purpose_Aquaculture_Sites/MaineDMR_Aquaculture_-_Limited_Purpose_Aquaculture_Sites.shp"))

# mapview(maine_aq_lp)

maine_aq_lp_active <- maine_aq_lp %>%
  dplyr::filter(Status == "A") # 749 limited purpose leases active

maine <- subset(state, region %in% c("maine"))

  ggplot() +
  geom_sf(data = maine_aq_lease_active, col = "red") + 
  geom_polygon(data=maine, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Maine data", caption = "Comprehensive, up-to-date (2021)") +
    geom_sf(data = maine_aq_lp_active, col = "blue", alpha = 0.1, size = 0.5)
```

## Alaska:

 - https://adfg.maps.arcgis.com/apps/webappviewer/index.html?id=f3ca95493c1042b39e42a3ecb5dcad6a
 - Will be considered "comprehensive". 
 - Up-to-date as of 2020
 - Specifies species type
 - Csvs with lat lons, but have a "corner" dataset, which we could make into polygons. Has area associated with each site. 
 - Shown in red. 
 - Has 62 active sites.

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
library(usmap)

alaska_details <- read.csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alaska/Active Aquatic Farming Operation Details.csv"))

alaska_corners <- read.csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alaska/Active Aquatic Farming Operation Corners.csv")) ## each group of corners is a polygon representing one farming operation. So, we will need to somehow make those polygons and find the centroid of each group to get points. Save as polygons and as points if possible? Maybe don't even need the exact polygons, since we know the area of each. Almost all are active.

library(tigris)

ak <- counties("Alaska", cb = TRUE) %>%
  filter(NAME %in% c("Aleutians East", "Lake and Peninsula", "Bristol Bay", "Kodiak Island", "Kenai Peninsula", "Anchorage", "Yakutat", "Hoonah-Angoon", "Haines", "Sitka", "Juneau", "Prince of Wales-Hyder", "Ketchikan Gateway", "Wrangell", "Petersburg", "Valdez-Cordova"))

alaska_details_sf <- st_as_sf(alaska_details, coords = c("x", "y"), crs = crs(ak))
alaska_corners_sf <- st_as_sf(alaska_corners, coords = c("x", "y"), crs = crs(ak))

ggplot() +
  geom_sf(data = alaska_corners_sf, col = "red", shape = 5) + 
  geom_sf(data=ak, fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Alaska data", caption = "Comprehensive, up-to-date (2020)")

```

## California:

 - https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentID=187229&inline
 - Report current as of May 2021
 - There is a map with commercial locations and a table with exact farm numbers (pages 7 and 8). We could likely estimate or find the location via Google. 
 - And cross reference with the locations listed here: https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/california_locs.png)

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/california_commercial.png)

## New Hampshire:

 - https://granit.unh.edu/data/search
 - Comprehensive
 - Up-to-date as of 2020
 - Polygons 
 - 40 sites

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
nh_aq <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/New_Hampshire/GRANIT_20211011184728/2020_Aquaculture_Sites.shp")) # 40 sites, assume all active

nh <- counties("New Hampshire", cb = TRUE) %>%
  filter(NAME %in% c("Rockingham"))

  ggplot() +
  geom_sf(data = nh_aq, col = "red") + 
 # geom_polygon(data=nh, aes(x=long, y=lat, group=group), fill = "gray") + 
    geom_sf(data = nh, fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "New Hampshire data", caption = "Comprehensive, up-to-date (2020)")
```

## Florida:

"ESI Human Use and Socio-Economc Resources Points" data

 - https://geodata.myfwc.com/datasets/myfwc::esi-human-use-and-socio-economic-resources-points-florida/about
 - 2015 data
 - 64 sites
 - Point data, no area
 - Not comprehensive (?), only represents points which are at risk from oil spills (which could be relevant to our project).

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}

fl_data <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Florida/ESI_Human-Use_and_Socio-Economic_Resources_Points_Florida/ESI_Human-Use_and_Socio-Economic_Resources_Points_Florida.shp"))

unique(fl_data$TYPE)

fl_aq <- fl_data %>%
  filter(TYPE == "AQUACULTURE")

fl <- subset(state, region %in% c("florida"))

  ggplot() +
  geom_sf(data = fl_aq, col = "red") + 
  geom_polygon(data=fl, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Florida data", caption = "Not Comprehensive, 2015 data")

```

Offshore aquaculture in Florida

 - As far as I can tell, none yet: https://www.flseagrant.org/aquaculture/openocean/

Other than the above, I can't find much info for Florida. Google maps shows a couple shrimp farms.

## Alabama:

 - "Alabama has an emerging oyster industry with 13 farms comprising of 25 acres in 2015. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise." from https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - This has a map... probably our best source for Alabama!: https://alaquaculture.com/state/#list-of-operating-farms

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/alabama_Farms.png)

## Mississippi:

 - Can't find much information for Mississippi.

## Louisana:

 - Pretty much all freshwater from what I can tell.

## Texas:

 - https://www.pewtrusts.org/en/research-and-analysis/articles/2020/05/21/texas-says-shucks-lets-give-oyster-growing-a-try
 - Seems like a very new industry in Texas.

 - Seafood watch 
https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_usfarmedshrimpreport.pdf
 - "Most commercial shrimp aquaculture production in the US is located in Texas, which has seven commercial farms."

http://www.texasaquaculture.org/TxShrimpFms.html
 - They have 9 Texas locations listed here. One is definitely freshwater prawns. Unclear if they are land-based or not (I guess all shrimp farms are techincally land-based since they are ponds). Need to explore these more. 
 - These are current as of 2018.

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/texas_shrimp.png)

## Oregon:

 - Map of general areas where you are allowed to commercially harvest or grow shellfish in Oregon: https://www.oregon.gov/oda/shared/Documents/Publications/FoodSafety/ClassifiedCommercialShellfishGrowingMap.pdf

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/or_areas.png)

 - Shellfish distributors list: https://www.oregon.gov/oda/shared/Documents/Publications/FoodSafety/Shellfishdealers.pdf

 - 2020 shellfish production report: https://www.oregon.gov/oda/shared/Documents/Publications/FoodSafety/AnnualShellfishProductionReport.pdf

## Connecticut:

Shellfish bed managed

 - https://portal.ct.gov/DOAG/Aquaculture1/Aquaculture/Shellfish-Area-Classifications--Maps
 - Only Shellfish
 - Polygons 
 - I believe that these are large areas that have been **approved** for shellfish aquaculture, but not the exact locations of the aquaculture. So this probably isn't helpful.

However, worst case, we could use the data from the Northeast ocean council dataset:

"Connecticut – Data on Connecticut aquaculture was derived from two datasets provided by the State of Connecticut Department of Agriculture/Bureau of Aquaculture and the State of Connecticut Department of Environmental Protection. One dataset contains the location of natural shellfish beds. Natural beds managed by the state, town or used for recreational harvesting were selected from this dataset, resulting in 1855 polygons. The number of beds in use at any given time is subject to change. The second dataset contains 46 polygons representing the approximate geographical location of permitted “Certificate for Aquaculture  Operation” gear sites in the Connecticut coastal waters of Long Island Sound via the State of Connecticut Application for Joint Programmatic General Permit for Aquaculture Department of the Army General Permit. These datasets are up to date as of January 15, 2016. For more information on CT Aquaculture and Shellfish, see the CT Aquaculture Mapping Atlas."

```{r}

ct_shellfish <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/Shellfish_Bed_Managed_shp/SHELLFISH_BED_MANAGED.shp"))

ct <- counties("Connecticut", cb = TRUE)

  ggplot() +
  geom_sf(data = ct_shellfish, col = "red") + 
    geom_sf(data = ct, fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Connecticut data", caption = "Not Comprehensive, General 'approved' areas")
```

Another map, this is probably what we need (seaweed and shellfish): https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

## Washington:

Approved growing areas map

 - https://fortress.wa.gov/doh/oswpviewer/index.html

"DNR leases deep water sites for finfish net pens and floating shellfish rafts and longlines. Salmon aquaculture operations have existed in Washington since the 1970s and are scattered throughout Puget Sound. Current net pen leases include salmon aquaculture, delayed-release salmon, and herring. The use of floating shellfish culture began in 1975 with mussel raft culture in Penn Cove. Current floating shellfish culture on rafts or longlines include oysters, mussels, and scallops."

What we used for global: 
Our Sound, Our Salmon, 2017. [Accessed January 31, 2020] http://www.moldychum.com/violations-prompt-washington-state-cancel-atlantic-salmon-farm-lease/

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/puget_sound.png)

## All states:

 - https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - Current as of 2015 
 - Can definitely use for California to cross reference
 - Has partial data for other states too

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/all_states_partial.png)

## North Carolina:

Shellfish

 - https://deq.nc.gov/about/divisions/marine-fisheries/licenses-permits-and-leases/shellfish-lease-and-franchise#stats-and-figures
 - Comprehensive for shellfish... just need to figure out how download it. 
 - I believe Shellfish is pretty much the only mariculture that occurs in NC. 
 - Has number totals of leases 
 - Will likely need to send an email.

 ![Has lease info for the entire state](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/nc_map.png)

## Virginia:

Shellfish

 - https://mrc.virginia.gov/links.shtm
 - They have oyster lease and application data, which *I think* is the only type of aquaculture there. The map isn't downloadable from what I can tell though, so I bet we will have to email to access.
 - They also have a phone app.

 ![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/va_map.png)

## Maryland:

Shellfish

 - They have individual lease application maps in PDFs, but they would take forever to go through. They must have some sort of internal dataset where they store the approved lease info?: https://dnr.maryland.gov/fisheries/pages/aquaculture/applications.aspx

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/maryland_apps.png)

## Delaware:

"Delaware is the only coastal state with no commercial shellfish aquaculture; a multimillion dollar business in our neighboring states up and down the East Coast."

 - Don't have to worry about Delaware.

## South Carolina:

 - Super small industry here. 
 - https://www.scseagrant.org/aquaculture/
 - https://www.scseagrant.org/wp-content/uploads/Economic-Contribution-of-Oyster-Mariculture-SC.pdf
 - Only 9 farms, all oyster, as of 2021

## Georgia:

 - https://www.gpb.org/news/2020/11/09/will-georgias-fledgling-oyster-industry-sink-it-swims
 - Highly unlikely there are any marine farms here. I think marine aquaculture might be illegal here? And even if it wasn't, they most likely haven't even had a harvest since it was legalized (see above article).

## Massachusetts:

https://www.mass.gov/service-details/aquaculture
  - "Massachusetts has 378 shellfish leases. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise."

https://www.mass.gov/doc/2020-dmf-annual-report/download
 - 386 "private shellfish propogation" permits in 2020

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/mass_permit.png)
 - Also has production data per town or region.

## Rhode Island:

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - No download option.

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/rhode_island_map.png)

## Hawaii

https://hdoa.hawaii.gov/ai/aquaculture-and-livestock-support-services-branch/aquaculture-in-hawaii/

 - Not much info. There is definitely both offshore aquaculture and shrimp aquaculture in Hawaii.

## New Jersey

https://www.nj.gov/agriculture/divisions/anr/sea/

 - "Currently, the program licenses more than 100 finfish, shellfish and aquatic plant producers in the state."

https://www.nj.gov/dep/fgw/pdf/marine/shellfish_leasing_policy_atlantic.pdf

 - this has maps of all shellfish leases in 2020.. but they are all pdf.

## New York

Suffolk County

 - Lease mapper: https://gis.suffolkcountyny.gov/portal/apps/View/index.html?appid=6ae708a346f340ada850544fe25ce0fb
 - Its not totally unreasonable to think that this is the only place in New York that does mariculture? Since its coast is so small.

 - Shellfish map for NY and CT, this seems like what we would need: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/ny_shell.png)