### 2022-01-28 marine/united_states/US_mapping.Rmd

---
title: 'Mariculture farms in the United States'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
pdf_document:
  toc: true
editor_options: 
  chunk_output_type: console
---

Exploring maps and data for all known US mariculture farms

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(raster)
library(tidyverse)
library(stringr)
library(PBSmapping)
library(maptools)
library(rgdal)
library(sf)
library(lwgeom)
library(data.table)
library(here)
library(mapview)
library(tmap)
library(leaflet)
library(ggplot2)
library(maps)
library(mapdata)
library(readr)

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

usa <- map_data('usa')

state <- map_data("state")
```

```{r, include = FALSE}

source(here("_spatial/template_raster.R"))

```

## Overview

There are 23 states with marine coasts.

**There is comprehensive downloadable data for 7 states: **

 - Maine, Alaska, New Hampshire, Connecticut, North Carolina, Rhode Island, Alabama

**There are publicly available map viewers (leases or farm locations) which seem to be comprehensive, but are NOT downloadable for 1 state:**

 - Virginia

**There is general info on lease holders, what they farm, and what municipality or city they are in for 6 states:**

 - California: general areas with acreage for commercial farms
 - Oregon: map of places were commercial harvest/growing is allowed
 - New Jersey: Have pdf maps of all existing shellfish leases and those issued in 2020
 - Massachusetts: Have number of growers per municipality with acreage and species type; also a subset of farms are available in the NE data viewer.
 - South Carolina: Apparently have 9 oyster farms as of 2021
 - Washington: Approximate locations of commercial net pens.

**Subset of farms or outdated downloadable for 2 states:**

 - Florida: A map of mariculture farms which are at risk of oil spills
 - Maryland: Have INDIVIDUAL pdf maps of all lease applications and if they were accepted or not... And was able to download lease data from 2013.

**Couldn't really find any information on 4 states:**

 - Hawaii, Texas, Louisiana, Mississippi

**Probably no mariculture for 2 states:**

 - Delaware: Commercial harvest is illegal; apparently one farm but not production (Halley)
 - Georgia: From what I could tell, oyster farming was illegal here as recent as 2020
 - New York: No production (Halley)

## Comprehensive, downloadable

### Maine

 - Viewer: https://www.maine.gov/dmr/aquaculture/leases/aquaculturemap.html
 - Download: https://dmr-maine.opendata.arcgis.com/datasets/maine::mainedmr-aquaculture-aq-leases/about
 - I think we will want to use regular leases only.
 - Limited purpose leases are for scientific or smaller/shorter term mariculture projects, that last up to 3 years.
 - I think this can be considered "comprehensive". 
 - Update to date as of 2019.
 - Has species types specified.

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
maine_aq_lease <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_AQ_Leases/MaineDMR_Aquaculture_-_AQ_Leases.shp"))

# mapview(maine_aq_lease)

maine_aq_lease_active <- maine_aq_lease %>%
  dplyr::filter(STATUS == "A") # 149 leases

```

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
maine_aq_lp <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maine/MaineDMR_Aquaculture_-_Limited_Purpose_Aquaculture_Sites/MaineDMR_Aquaculture_-_Limited_Purpose_Aquaculture_Sites.shp"))

# mapview(maine_aq_lp)

maine_aq_lp_active <- maine_aq_lp %>%
  dplyr::filter(Status == "A") # 749 limited purpose leases active

maine <- subset(state, region %in% c("maine"))

ggplot() +
  geom_sf(data = maine_aq_lease_active, col = "red") + 
  geom_polygon(data=maine, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Maine data", caption = "Comprehensive, up-to-date (2021)") +
    geom_sf(data = maine_aq_lp_active, col = "blue", alpha = 0.1, size = 0.5)

## leave out limited purpose 
```

### Alaska

 - https://adfg.maps.arcgis.com/apps/webappviewer/index.html?id=f3ca95493c1042b39e42a3ecb5dcad6a
 - Will be considered "comprehensive". 
 - Up-to-date as of 2020
 - Specifies species type
 - Csvs with lat lons, but have a "corner" dataset, which we could make into polygons. Has area associated with each site. 
 - Shown in red. 
 - Has 62 active sites.

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
library(usmap)

alaska_details <- read.csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alaska/Active Aquatic Farming Operation Details.csv"))

alaska_corners <- read.csv(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alaska/Active Aquatic Farming Operation Corners.csv")) ## each group of corners is a polygon representing one farming operation. So, we will need to somehow make those polygons and find the centroid of each group to get points. Save as polygons and as points if possible? Maybe don't even need the exact polygons, since we know the area of each. Almost all are active.

library(tigris)

ak <- counties("Alaska", cb = TRUE) %>%
  filter(NAME %in% c("Aleutians East", "Lake and Peninsula", "Bristol Bay", "Kodiak Island", "Kenai Peninsula", "Anchorage", "Yakutat", "Hoonah-Angoon", "Haines", "Sitka", "Juneau", "Prince of Wales-Hyder", "Ketchikan Gateway", "Wrangell", "Petersburg", "Valdez-Cordova"))

alaska_details_sf <- st_as_sf(alaska_details, coords = c("x", "y"), crs = crs(ak))
alaska_corners_sf <- st_as_sf(alaska_corners, coords = c("x", "y"), crs = crs(ak))

ggplot() +
  geom_sf(data = alaska_corners_sf, col = "red", shape = 5) + 
  geom_sf(data=ak, fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Alaska data", caption = "Comprehensive, up-to-date (2020)")

```

### New Hampshire

 - https://granit.unh.edu/data/search
 - Comprehensive
 - Up-to-date as of 2020
 - Polygons 
 - 40 sites

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
nh_aq <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/New_Hampshire/GRANIT_20211011184728/2020_Aquaculture_Sites.shp")) # 40 sites, assume all active

nh <- counties("New Hampshire", cb = TRUE) %>%
  filter(NAME %in% c("Rockingham"))

  ggplot() +
  geom_sf(data = nh_aq, col = "red") + 
 # geom_polygon(data=nh, aes(x=long, y=lat, group=group), fill = "gray") + 
    geom_sf(data = nh, fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "New Hampshire data", caption = "Comprehensive, up-to-date (2020)")
```

### Connecticut

From the Northeast ocean council dataset:

"Connecticut – Data on Connecticut aquaculture was derived from two datasets provided by the State of Connecticut Department of Agriculture/Bureau of Aquaculture and the State of Connecticut Department of Environmental Protection. One dataset contains the location of natural shellfish beds. Natural beds managed by the state, town or used for recreational harvesting were selected from this dataset, resulting in 1855 polygons. The number of beds in use at any given time is subject to change. The second dataset contains 46 polygons representing the approximate geographical location of permitted “Certificate for Aquaculture  Operation” gear sites in the Connecticut coastal waters of Long Island Sound via the State of Connecticut Application for Joint Programmatic General Permit for Aquaculture Department of the Army General Permit. These datasets are up to date as of January 15, 2016. For more information on CT Aquaculture and Shellfish, see the CT Aquaculture Mapping Atlas."

Map combined with NY, this is what we need (seaweed and shellfish). We will extract the data using the methods described below: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

Current as of 2019

Extract data following these protocols: https://stackoverflow.com/questions/50161492/how-do-i-scrape-data-from-an-arcgis-online-map/50213419#50213419

CT points json: https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/0/query?where=0%3D0&outFields=%2A&f=json

CT polygons json: https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1/query?where=0%3D0&outFields=%2A&f=json

Roadmap for polygons

Viewer: https://www.arcgis.com/apps/webappviewer/index.html?id=f8799cefbb4c4751a209710d14b9ad46

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/f8799cefbb4c4751a209710d14b9ad46/data

Search for layer you want and grab link ("CT Aquaculture Gear): https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1

Run query using that link to output geojson: https://services7.arcgis.com/9fAJJI91yoj2y4Yi/arcgis/rest/services/ConnecticutShellfishOnline2018Draft/FeatureServer/1/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as geojson: https://geojson.io

Check the downloaded geojson file: 
```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/CT_Aquaculture_gear/CT_aquaculture_gear.geojson",  what = "sp") %>%
  st_as_sf()

mapview(spdf)

# library(geojsonio)
# spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Connecticut/CT_aquacultre_points/CT_aquaculture_points.geojson",  what = "sp") %>%
#   st_as_sf()
# 
# 
# plot(spdf)
# 
# mapview(spdf)

```

### North Carolina

Shellfish

 - https://deq.nc.gov/about/divisions/marine-fisheries/licenses-permits-and-leases/shellfish-lease-and-franchise#stats-and-figures
 - Comprehensive for shellfish... just need to figure out how download it. 
 - I believe Shellfish is pretty much the only mariculture that occurs in NC. 
 - Has number totals of leases.
 - Download using protocol described for CT.
 - Current as of 12/21/2021

Roadmap to download

Viewer: https://www.arcgis.com/apps/webappviewer/index.html?id=de86f3bb9e634005b12f69a8a5947367&extent=-8551979.8781%2C4121555.1994%2C-8515290.1046%2C4140072.0696%2C102100

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/de86f3bb9e634005b12f69a8a5947367/data

Search for layer you want and grab link ("Leases" is the layer you want): https://services2.arcgis.com/kCu40SDxsCGcuUWO/arcgis/rest/services/Lease_DB_View/FeatureServer/2

Run query using that link to output geojson: https://services2.arcgis.com/kCu40SDxsCGcuUWO/arcgis/rest/services/Lease_DB_View/FeatureServer/2/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as geojson: https://geojson.io. Once it shows in the map, click "Save" and Save as a geojson file.

Check to see if it worked
```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/North_Carolina/NC_shellfish_leases.geojson",  what = "sp") %>%
  st_as_sf()

mapview(spdf)

```

### Rhode Island

 - Lease map here: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f
 - Julia Livermore: julia.livermore@dem.ri.gov
 - Current as of August 2021
 - Shellfish and kelp

Roadmap to download

Viewer: https://ridemgis.maps.arcgis.com/apps/webappviewer/index.html?id=8beb98d758f14265a84d69758d96742f

I found my way to here:
https://risegis.ri.gov/hosting/rest/services/RIDEM/Marine_Fisheries_Map_Layers/MapServer/0

Then run query using that link to output geojson: https://risegis.ri.gov/hosting/rest/services/RIDEM/Marine_Fisheries_Map_Layers/MapServer/0/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as kml (for some reason, saving as geojson was not working for this one): https://geojson.io. Once it shows in the map, click "Save" and Save as a kml file.

Check to see if it worked

```{r}

rhode_island <- read_sf("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Rhode_Island/ri_aq_sites.kml")

mapview(rhode_island)

```

### Alabama

 - "Alabama has an emerging oyster industry with 13 farms comprising of 25 acres in 2015. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise." from https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - This has a map : https://alaquaculture.com/state/#list-of-operating-farms
 - Here is an even better map with Shellfish leases!: https://aldcnr.maps.arcgis.com/apps/webappviewer/index.html?id=a32dad8dacd249ea86bcb80dd951a424

Roadmap to download

Viewer: https://aldcnr.maps.arcgis.com/apps/webappviewer/index.html?id=a32dad8dacd249ea86bcb80dd951a424

Item data endpoint (just replace id): https://www.arcgis.com/sharing/rest/content/items/a32dad8dacd249ea86bcb80dd951a424/data

Search for layer you want and grab link ("Leases" is the layer you want): https://services7.arcgis.com/jcVUvxpWnajErgCr/arcgis/rest/services/Shellfish_Riparian_Leases_Revised/FeatureServer/0

Run query using that link to output geojson: https://services7.arcgis.com/jcVUvxpWnajErgCr/arcgis/rest/services/Shellfish_Riparian_Leases_Revised/FeatureServer/0/query?where=0%3D0&outFields=%2A&f=geojson

And copy-paste the text in here to save as geojson: https://geojson.io. Once it shows in the map, click "Save" and Save as a geojson file.

Leases current as of: 11/29/2021

```{r}
library(geojsonio)
spdf <- geojson_read("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Alabama/Alabama_Shellfish_Leases.geojson",  what = "sp") %>%
  st_as_sf()

mapview(spdf)

```

## Maps with no download

### Virginia

Shellfish

 - https://mrc.virginia.gov/links.shtm
 - They have oyster lease and application data, which *I think* is the only type of aquaculture there. The map isn't downloadable from what I can tell though, so I bet we will have to email to access.
 - They also have a phone app.
 - web.info@mrc.virginia.gov

 ![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/va_map.png)

## General lease info, areas, etc

### California

 - https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentID=187229&inline
 - Report current as of May 2021
 - There is a map with commercial locations and a table with exact farm numbers (pages 7 and 8). We could likely estimate or find the location via Google. 
 - And cross reference with the locations listed here: https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/california_locs.png)

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/california_commercial.png)

### Oregon

 - Map of general areas where you are allowed to commercially harvest or grow shellfish in Oregon: https://www.oregon.gov/oda/shared/Documents/Publications/FoodSafety/ClassifiedCommercialShellfishGrowingMap.pdf

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/or_areas.png)

 - Shellfish distributors list: https://www.oregon.gov/oda/shared/Documents/Publications/FoodSafety/Shellfishdealers.pdf

 - 2020 shellfish production report: https://www.oregon.gov/oda/shared/Documents/Publications/FoodSafety/AnnualShellfishProductionReport.pdf

### South Carolina

 - Super small industry here. 
 - https://www.scseagrant.org/aquaculture/
 - https://www.scseagrant.org/wp-content/uploads/Economic-Contribution-of-Oyster-Mariculture-SC.pdf
 - Only 9 farms, all oyster, as of 2021

### New Jersey

https://www.nj.gov/agriculture/divisions/anr/sea/

 - "Currently, the program licenses more than 100 finfish, shellfish and aquatic plant producers in the state."

https://www.nj.gov/dep/fgw/pdf/marine/shellfish_leasing_policy_atlantic.pdf

 - this has maps of all shellfish leases in 2020.. but they are all pdf.
 - Amanda Wenczel: amanda.wenczel@ag.nj.gov

### Massachusetts

https://www.mass.gov/service-details/aquaculture
  - "Massachusetts has 378 shellfish leases. Shellfish farms are a source of healthy protein and can contribute to healthy marine ecosystems by cleaning the water column, providing habitat for other sea life, and helping protect shorelines from storm surges and sea level rise.
  - Sean Bowen: Sean.Bowen@mass.gov

https://www.mass.gov/doc/2020-dmf-annual-report/download
 - 386 "private shellfish propogation" permits in 2020

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/mass_permit.png)
 - Also has production data per town or region.

### Washington

Approved growing areas map

 - https://fortress.wa.gov/doh/oswpviewer/index.html

"DNR leases deep water sites for finfish net pens and floating shellfish rafts and longlines. Salmon aquaculture operations have existed in Washington since the 1970s and are scattered throughout Puget Sound. Current net pen leases include salmon aquaculture, delayed-release salmon, and herring. The use of floating shellfish culture began in 1975 with mussel raft culture in Penn Cove. Current floating shellfish culture on rafts or longlines include oysters, mussels, and scallops."

Locations of commercial marine net pens
 - https://ecology.wa.gov/Water-Shorelines/Shoreline-coastal-management/Aquaculture/Net-pens

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/washington_pens.png)

## Subset available

### US Northeast dataset

Source:
Northeast Regional Ocean Council, 2018. Northeast ocean data: aquaculture. [Accessed January 17, 2020] 
https://www.northeastoceandata.org/data-download/

"This dataset should be considered a work in progress"

States:

 - NY
 - CT
 - ME
 - NH
 - MA
 - Federal waters

Species: 
 - "Shellfish"  "Multi-trophic or other" "Seaweed" "Finfish"

**This data set is not "comprehensive" but provides good sources to look through for individual states (see below). These are polygons.**

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}

us_ne <- st_read(file.path(path,"USA/NE/US_NE_Aquaculture.shp"))

us_ne_active <- us_ne %>%
  filter(status != "Unknown")

us_ne_active$area = st_area(us_ne_active)

# unique(us_ne_active$state) # "Federal waters" "NY"             "CT"             "MA"             "ME"             "NH"   
# unique(us_ne_active$genSpecies) # [1] "Shellfish"  "Multi-trophic or other" "Seaweed" "Finfish"

usa <- map_data('usa')

state <- map_data("state")

ne <- subset(state, region %in% c("new york", "maine", "new hampshire", "massachusetts", "connecticut"))

  ggplot() +
  geom_sf(data = us_ne_active, col = "red") + 
  geom_polygon(data=ne, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "US NE Regional Ocean Council data", caption = "Not comprehensive")

```

### Florida

"ESI Human Use and Socio-Economc Resources Points" data

 - https://geodata.myfwc.com/datasets/myfwc::esi-human-use-and-socio-economic-resources-points-florida/about
 - 2015 data
 - 64 sites
 - Point data, no area
 - Not comprehensive (?), only represents points which are at risk from oil spills (which could be relevant to our project).

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}

fl_data <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Florida/ESI_Human-Use_and_Socio-Economic_Resources_Points_Florida/ESI_Human-Use_and_Socio-Economic_Resources_Points_Florida.shp"))

unique(fl_data$TYPE)

fl_aq <- fl_data %>%
  filter(TYPE == "AQUACULTURE")

fl <- subset(state, region %in% c("florida"))

  ggplot() +
  geom_sf(data = fl_aq, col = "red") + 
  geom_polygon(data=fl, aes(x=long, y=lat, group=group), fill = "gray") + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
    labs(title = "Florida data", caption = "Not Comprehensive, 2015 data")

```

Other than the above, I can't find much info for Florida. Google maps shows a couple shrimp farms.

### Maryland

Shellfish

 - They have individual lease application maps in PDFs, but they would take forever to go through. They must have some sort of internal dataset where they store the approved lease info?: https://dnr.maryland.gov/fisheries/pages/aquaculture/applications.aspx
 - Maude Morris: maudee.morris@maryland.gov

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/maryland_apps.png)

List of farms for VA and MD (addresses are on land though): https://oysterrecovery.org/oyster-farms/

Maryland has a viewer for 2013 active leases: https://maryland.maps.arcgis.com/apps/webappviewer/index.html?id=f7db2acc81554601b44c13733a66a60f

 - This viewer has a layer named "ActiveOysterLeases", however, it looks like the last time it was updated is 2013.. I will still download this. Worst case we use this.

I found the layer here: https://gisapps.dnr.state.md.us/arcgis2/rest/services/Biota/ActiveOysterLeases/MapServer/0

Run query using that link to output json (for some reason it won't let me output geojson): https://gisapps.dnr.state.md.us/arcgis2/rest/services/Biota/ActiveOysterLeases/MapServer/0/query?where=0%3D0&outFields=%2A&f=json

```{r include = FALSE}

# test out package to download
devtools::install_github("yonghah/esri2sf")
library("esri2sf")

url <- "https://gisapps.dnr.state.md.us/arcgis2/rest/services/Biota/ActiveOysterLeases/MapServer/0"
df <- esri2sf(url)
mapview(df)

#st_write(df, file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maryland/maryland_2013_oyster_leases.shp"))

```

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'hide'}
md <- st_read(file.path("/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations/USA/Maryland/maryland_2013_oyster_leases.shp"))
```

### All states

 - https://aop.maps.arcgis.com/apps/Cascade/index.html?appid=e97ddd2cd9cf46218cefeed813e26c87
 - another link to map? https://aop.maps.arcgis.com/apps/Shortlist/index.html?appid=32864a80775b4f8abaeb072f0718ee57
 - Current as of 2015 
 - Can definitely use for California to cross reference
 - Has partial data for other states too
 - ask Halley for contact

## Not much info

### Mississippi

 - Can't much info on Mississippi mariculture..

### Louisana

 - Pretty much all freshwater from what I can tell.

### Texas

 - https://www.pewtrusts.org/en/research-and-analysis/articles/2020/05/21/texas-says-shucks-lets-give-oyster-growing-a-try
 - Seems like a very new industry in Texas.

 - Seafood watch 
https://www.seafoodwatch.org/-/m/sfw/pdf/reports/s/mba_seafoodwatch_usfarmedshrimpreport.pdf
 - "Most commercial shrimp aquaculture production in the US is located in Texas, which has seven commercial farms."

http://www.texasaquaculture.org/TxShrimpFms.html
 - They have 9 Texas locations listed here. One is definitely freshwater prawns. Unclear if they are land-based or not (I guess all shrimp farms are techincally land-based since they are ponds). Need to explore these more. 
 - These are current as of 2018.

![](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/texas_shrimp.png)

### Hawaii

https://hdoa.hawaii.gov/ai/aquaculture-and-livestock-support-services-branch/aquaculture-in-hawaii/

 - Not much info. There is definitely both offshore aquaculture and shrimp aquaculture in Hawaii.

## No mariculture?

### Delaware

"Delaware is the only coastal state with no commercial shellfish aquaculture; a multimillion dollar business in our neighboring states up and down the East Coast."

 - Don't have to worry about Delaware.

### Georgia

 - https://www.gpb.org/news/2020/11/09/will-georgias-fledgling-oyster-industry-sink-it-swims
 - Highly unlikely there are any marine farms here. I think marine aquaculture might be illegal here? And even if it wasn't, they most likely haven't even had a harvest since it was legalized (see above article).

### New York

 - According to Halley, there is currently no production in New York state, especially since their coastline is so small.

## Species/production breakdown for US

Take a look at the species in the US that FAO reports

```{r, message=FALSE}
fao_data_all <- read_csv("../../data/fao_aquaculture_clean.csv")

fao_data_us <- fao_data_all %>%
  filter(iso3c == "USA")

fao_data_us_group <- fao_data_us %>%
  filter(year == 2018)

marine <- fao_data_us_group %>%
  filter(environment == "Marine") %>%
  filter(!is.na(value))

# unique(marine$species)

# 16 marine spp
#  [1] "Atlantic salmon"             "Blue mussel"                 "Northern quahog(=Hard clam)"
#  [4] "Sand gaper"                  "American cupped oyster"      "Red drum"                   
#  [7] "Whiteleg shrimp"             "Longfin yellowtail"          "Abalones nei"               
# [10] "Butter clam"                 "Cupped oysters nei"          "Japanese carpet shell"      
# [13] "Marine fishes nei"           "Pacific cupped oyster"       "Pacific geoduck"            
# [16] "Pacific littleneck clam"

## there were 4 species with NA production (so actually 20 marine species total)
marine_na <- fao_data_us_group %>%
  filter(environment == "Marine") %>%
  filter(is.na(value))
#unique(marine_na$species)
# [1] "Cockles nei"          "European flat oyster" "Olympia oyster"       "Pacific horse clam"

# sum(marine$value, na.rm = TRUE)

# 203687 total tonnes for marine.. does this seem correct? 
```

FAO reports 16 different marine species of production in 2018. Additionally, there were 4 species that were previously harvested but had no harvest for 2018. FAO reports a total of 203687 live weight tonnes for 2018.

NOAA Fisheries of the United States 2019 Report: 
https://media.fisheries.noaa.gov/2021-05/FUS2019-FINAL-webready-2.3.pdf?null=

 - NOAA report 44094 metric tonnes of marine aquaculture production for 2018... versus 203687 from FAO. I think this has to do with NOAA reporting meat weight and FAO reporting live weight (includes shells)
 - NOAA also reports that the Gulf region produces th e most marine aquaculture by volume.. which seems problematic for us considering we have very poor farm level data for the Gulf states.

![gulf](/home/sgclawson/github/aquaculture_mapping/marine/united_states/pics/us_mariculture.png)

2018 Census of Aquaculture: https://www.nass.usda.gov/Publications/AgCensus/2017/Online_Resources/Aquaculture/Aqua.pdf

 - Table 1 shows aquaculture farms per state, but it is not split between marine and freshwater from what I can tell. Also split by species group (mollusk, crustaceans, etc.)
 - Table 10 shows saltwater acres and number of farms used for aquaculture production. However, there are landlocked states, so we don't know if it is actually marine or on-land saltwater aquaculture. It does seem reasonable that this would be pretty close to the actual number of farms.. 
 - Table 11 shows sources of water, with "Saltwater" included. This is just where the water comes from. 
 - Table 12 mollusks off bottom seems reasonable to think most of these will be in marine environments. Almost all of the states listed are coastal states.

 - It seems like "Mollusks off bottom", "Mollusks on bottom", "Sea vegetables" are the only ones that I am 80% sure are marine farms... 
 - They also have data at the county level. I know this wouldn't be perfect, but we could filter for coastal counties.

## Climate stressors for marine aquaculture