### 2022-03-02 analysis/paper_stats.Rmd
cleaned code and updated readme

---
title: "Summary stats for paper"
author: "Gage Clawson (UCSB, NCEAS, OHI)"
date: "04/26/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script takes our final farms dataset and calculates summary statistics that will be used in the paper.

```{r}

library(here)
library(dplyr)
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(RCurl)
library(mosaic)
library(googlesheets4)
library(janitor)
library(countrycode)

source(here("_spatial/template_raster.R"))

path <- "/home/shares/food-systems/Food_footprint/_raw_data/Aquaculture_locations"

```

# Read in farms data

```{r}

all_farms <- read_csv("data/all_marine_aquaculture_farms_sources_final.csv")

all_farms_sf <- all_farms %>%
  st_as_sf(coords =c("X", "Y"))

sum(all_farms$tonnes_per_farm) # 29642964

data_type_A <- all_farms %>%
  filter(data_type_2 == "A")

sum(data_type_A$tonnes_per_farm) # 3568135

data_type_no_A <- all_farms %>%
  filter(data_type_2 != "A")
sum(data_type_no_A$tonnes_per_farm) # 26074829

26074829/29642964 # 0.8796296 # ~88% of production is data limited

test <- all_farms %>%
  filter(species_group == "General marine fish")

sum(test$tonnes_per_farm) # 3909255

3909255/29638328 # 0.1318986

```

```{r}
sources <- read_csv("analysis/data/mariculture_mapping_sources.csv") %>%
  clean_names() %>%
  mutate(type = case_when(species == "Salmonids" ~ "salmon",
                          species == "Unfed or algae fed shellfish" ~ "bivalve",
                          species == "Shrimp" ~ "shrimp",
                          species == "Tuna" ~ "tuna",
                          species == "General Marine Fish" ~ "marine_fish_general",
                          species == "Crustaceans" ~ "crustaceans"))

## Numbers by data type
by_data_type <- all_farms %>%
  group_by(data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)

by_data_type
# A tibble: 3 x 4
#   data_type                                    n_obs total_obs   pct
#   <chr>                                        <int>     <int> <dbl>
# 1 known locations                              16032     95331  16.8
# 2 modeled locations, estimated number of farms 69348     95331  72.7
# 3 modeled locations, known number of farms      9951     95331  10.4

# How many countries have spatial locations information?
known_countries <- all_farms %>%
  filter(data_type == "known locations") %>%
  distinct(iso3c)

# China represented 61% of all farms (N=58,166), but only 15 had known locations.

chn_data_estimated <- all_farms %>%
  filter(iso3c == "CHN") 
# 58166 points
58166/95443 # 0.6094318 = ~61%

chn_data_known <- all_farms %>%
  filter(iso3c == "CHN") %>%
  filter(data_type == "known locations") # 5739 locations

# Chile, Norway, and Ireland provided a substantial number of known locations (>1000).
top_known <- all_farms %>%
  filter(data_type == "known locations") %>%
  group_by(iso3c) %>%
    summarise(n_obs = n())

# Only 5 countries (Cambodia, Mauritius, Myanmar, North Korea, and Russia) had no form of farm data at all, however, these countries only account for 0.5% of marine aquaculture production included.

no_info <- all_farms %>%
  group_by(iso3c) %>%
  mutate(check = all(data_type_2 == "F")) %>%
  filter(check == TRUE) 
unique(no_info$country)

# [1] "Cambodia"                              "Myanmar"                              
# [3] "Mauritius"                             "Democratic People’s Republic of Korea"
# [5] "Russian Federation"

mar_production <- read_csv("marine/STEP1_species_groups/int/tonnes_per_country_group.csv")

sum(no_info$tonnes_per_farm) # 156132.3

sum(mar_production$total_tonnes) # 29623515

156132.3/29623515 # 0.005270553

chn_production <- mar_production %>%
  filter(country == "China")
sum(chn_production$total_tonnes)
17696822/29623515 # 0.597391

# Per species group, salmon was the most data rich (98% of points had location information, 1.13% of points were known with no location information, 0.87% estimated), followed by Tuna (34% of points had location information, 66% of points were known with no location information, 0% estimated).

by_species_data_type <- all_farms %>%
  group_by(species_group, data_type) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)

## How much fish and invertebrate mariculture production was included in our assessment?

fao <- read_csv("data/fao_aquaculture_clean.csv")
mar_sp_list <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2020/gh-pages/globalprep/mar/v2020/raw/species_list.csv")

fw <- fao %>%
  dplyr::filter(environment == "Freshwater") %>%
  group_by(species) %>%
  summarize(total = sum(value)) %>%
  ungroup() %>%
  arrange(total)

data.frame(fw)

test <- fao %>%
  filter(year == 2017) %>%
  left_join(mar_sp_list, by = c("species" = "FAO_name")) %>%
  mutate(environment2 = ifelse(Taxon_code == "AL", "seaweed", environment)) %>%
  mutate(environment = ifelse(is.na(environment2), environment, environment2)) %>%
  filter(environment != "Freshwater") %>%
  group_by(environment, year) %>%
  summarise(prod = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(total = sum(prod)) %>%
  mutate(pct = prod/total)

## total fish and invertebrate mariculture production is: 8090542 + 22920871 = 31011413

## we include: 29623515
prod_total <- read_csv("/home/shares/food-systems/Food_footprint/all_food_systems/dataprep/aquaculture/marine/production_extracted_final.csv")

sum(prod_total$total_tonnes) # 29623515

29623515/31011413 # 0.9552456
```

Countries with comprehensive farm locations

However, 17 countries (representing 10% of fish and invertebrate mariculture production) had comprehensive farm locations across all taxa (i.e. no estimation or interpolation required, Table S3)
```{r}

all_comp <- all_farms %>%
  group_by(iso3c) %>%
  mutate(check = all(data_type_2 == "A")) %>%
  filter(check == TRUE)

sort(unique(all_comp$country))

print(unique(all_comp$country))

cat(paste(sort(unique(all_comp$country)), collapse= "', '"))

# Belize, Bulgaria, Canada, Chile, Cyprus, Denmark, Faroe Islands, Finland, French Polynesia, Greece, Hong Kong , Ireland, Jersey, Malta, Norway, United Arab Emirates, United Kingdom

## how much production do they account for?

mar_production <- read_csv("marine/STEP1_species_groups/int/tonnes_per_country_group.csv")

comp_prod <- mar_production %>%
  filter(country %in% c('Belize', 'Bulgaria', 'Canada', 'Chile', 'Cyprus', 'Denmark', 'Faroe Islands', 'Finland', 'French Polynesia', 'Greece', 'China, Hong Kong SAR', 'Ireland', 'Jersey', 'Malta', 'Norway', 'United Arab Emirates', 'United Kingdom'))

sum(comp_prod$total_tonnes) # 3203584

3203584/31011413 # 0.1033034 ~ 10%

```

Countries with comprehensive "number of farm" information

"Thirty-six countries have comprehensive “number of farm” data across all types of mariculture they produce (data types A, B, or C), accounting for 12% of the global fish and invertebrate mariculture production."

```{r}
all_comp <- all_farms %>%
  group_by(iso3c) %>%
  mutate(check = all(data_type_2 %in% c("A","B","C"))) %>%
  filter(check == TRUE)

sort(unique(all_comp$country))

print(unique(all_comp$country))

cat(paste(sort(unique(all_comp$country)), collapse= "', '"))

# 'Albania', 'Belize', 'Bulgaria', 'Canada', 'Chile', 'Colombia', 'Croatia', 'Cyprus', 'Denmark', 'El Salvador ', 'Faroe Islands', 'Finland', 'French Polynesia', 'Germany', 'Greece', 'Guatemala', 'Guernsey', 'Hong Kong ', 'Ireland', 'Italy', 'Jersey', 'Madagascar', 'Malta', 'Netherlands', 'New Zealand', 'Norway', 'Peru', 'Portugal', 'Singapore', 'Slovenia', 'South Africa', 'Sri Lanka', 'Tunisia', 'Turkey', 'United Arab Emirates', 'United Kingdom'

## how much production do they account for?

mar_production <- read_csv("marine/STEP1_species_groups/int/tonnes_per_country_group.csv")

comp_prod <- mar_production %>%
  filter(country %in% c('Albania', 'Belize', 'Bulgaria', 'Canada', 'Chile', 'Colombia', 'Croatia', 'Cyprus', 'Denmark', 'El Salvador ', 'Faroe Islands', 'Finland', 'French Polynesia', 'Germany', 'Greece', 'Guatemala', 'Guernsey', 'Hong Kong ', 'Ireland', 'Italy', 'Jersey', 'Madagascar', 'Malta', 'Netherlands', 'New Zealand', 'Norway', 'Peru', 'Portugal', 'Singapore', 'Slovenia', 'South Africa', 'Sri Lanka', 'Tunisia', 'Turkey', 'United Arab Emirates', 'United Kingdom'))

sum(comp_prod$total_tonnes) # 3788091

3788091/31011413 # 0.1221515 ~ 12%
```

"Using the country/species grouping combinations with presumed comprehensive farm count data (N=74, 17% of production included, data type A/B/C, see Table S3)"
```{r}

test <- all_farms %>%
  filter(data_type_2 %in% c("A", "B", "C")) %>%
  group_by(country, species_group)

cat(paste(sort(unique(test$country)), collapse= "', '"))

length(unique(test$country))

sum(test$tonnes_per_farm)

4963244/29623515 # 0.1675441 ~ 17%

test2 <- test %>%
  group_by(country, species_group) %>%
  summarise(n()) # 74 combos

```

```{r}
by_data_type_spp <- all_farms %>%
  group_by(data_type, species_group) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)

by_spp <- all_farms %>%
  group_by(species_group) %>%
  summarise(n_obs = n()) %>%
  mutate(total_obs = sum(n_obs)) %>% 
  ungroup() %>%
  mutate(pct = n_obs/total_obs*100)
```